### 2019-08-08 data/draft-publications.tex
filter correction

\subsection{Papers in Journals}

\begin{enumerate}
\item 2. S. Aiyappan, A. K. Nandakumaran and Ravi Prakash, Semi-linear optimal control problem on a smooth oscillating domain Communications in Contemporary Mathematics, {\em Communications in Contemporary Mathematics} {\bf } (2019), 26 pages.
\item 1. S. Aiyappan, A. K. Nandakumaran and Abu Sufian, Asymptotic Analysis of a Boundary Optimal Control Problem on a General Branched Structure, to appear in Mathematical Methods in Applied Sciences (MMAS).
\item Benedetti, Carolina and González D’León, Rafael S. and Hanusa, Christopher R.H. and Harris, Pamela E. and Khare, Apoorva and Morales, Alejandro H. and Yip, Martha, A combinatorial model for computing volumes of flow polytopes, {\em Transactions of the American Mathematical Society} {\bf 372} (2019), 3369-3404.
\item Subhajit Ghosh, Total variation cutoff for the transpose top-2 with random shuffle, to appear in Journal of Theoretical Probability.
\item G. K. Basak, M. K. Ghosh and D. Mukherjee, A Stochastic Model with Inflation, Growth and Technology for the Political Business Cycle, {\em Journal of Computational Economics} {\bf 53} (2019), 125-140.
\item Arvind Ayyer and Pooja Singla, Random motion on finite rings, I: commutative rings, {\em Algebras and Representation theory} {\bf } (), .
\item E.K.Narayanan and Pooja Singla, On monomial representations of finitely generated nilpotent groups, {\em Communications in Algebra} {\bf 46} (2018), 2319–2331.
\item Souvik Dhara, Remco van der Hofstad, Johan S.H. van Leeuwaarden, Sanchayan Sen, Heavy-tailed configuration models at criticality, to appear in Annales de l'Institut Henri Poincaré (B) Probabilités et Statistiques.
\item Siddhartha Gadgil, Homogeneous Length Functions on Groups: Intertwined Computer and Human Proofs, to appear in Journal of Automated Reasoning.
\item Srikanth K. Iyer, The Random Connection Model - Connectivity, Edge Lengths and Degree Distributions, {\em Random Structures and Algorithms} {\bf 52} (2018), 283-300.
\item Rahul Vaze, Srikanth K. Iyer, Capacity of Cellular Wireless Networks, {\em IEEE Transactions on Wireless Communications} {\bf 18} (2019), 1490 - 1503.
\item Gokul Nair, Senthil Athmanathan,, Srikanth K. Iyer  and Vishwesha Guttal, Fission-fusion dynamics and group-size-dependent composition in heterogeneous populations, {\em Physical Review E} {\bf 99} (2019), .
\item Subhojoy Gupta, Meromorphic quadratic differentials and measured foliations on a Riemann surface, {\em Mathematische Annalen} {\bf 373} (2019), 73-118.
\item Subhojoy Gupta and Harish Seshadri, On domains biholomorphic to Teichmüller spaces, to appear in International Mathematics Research Notices.
\item Subhojoy Gupta, Harmonic maps and wild Teichmüller spaces, to appear in ournal of Topology and Analysis.
\item Subhojoy Gupta, Limits of harmonic maps and crowned hyperbolic surfaces, to appear in Transactions of the AMS.
\item Luz Roncal and Sundaram Thangavelu, An extension problem and trace Hardy inequality for the sublaplacian on H-type groups, to appear in IMRN.
\item Pradeep Boggarapu, Luz Roncal and Sundaram Thangavelu, On extension problem, trace Hardy and Hardy's inequalities for some fractional Laplacians, to appear in Communications in Pure and Applied Analysis..
\item Indranil Biswas and Vamsi Pritham Pingali, Metric properties of parabolic ample bundles, to appear in International Mathematical Research Notices.
\item Vamsi Pingali and Dror Varolin, Non-uniformly flat affine algebraic hypersurfaces, to appear in Nagoya mathematical journal.
\end{enumerate}

\subsection{Papers in Conference proceedings}

\begin{enumerate}
\item Kriti Goel, Dilip Patil and Jugal Verma, Nullstellens\"atze and Applications {\em } ().
\item Dilip P Patil and Jugal Verma, Rational points and generalized trace forms on a finite algebra over a real closed field {\em } ().
\end{enumerate}

\subsection{Books and book chapters}

\begin{enumerate}
\item Belton, Alexander and Guillot, Dominique and Khare, Apoorva and Putinar, Mihai, A panorama of positivity: I. Dimension free in {\em Birkhauser, Trends in Mathematics}{f } (2019), 117-165.
\item Belton, Alexander and Guillot, Dominique and Khare, Apoorva and Putinar, Mihai, A panorama of positivity: II. Fixed dimension in {\em American Mathematical Society, Contemporary Mathematics}{f } (), .
\item Subhojoy Gupta and Harish Seshadri , Complex geometry of Teichmüller domains in {\em EMS, Handbook of Teichmüller theory}{f 7} (), .
\item Subhojoy Gupta, Holomorphic quadratic differentials in Teichmüller theory in {\em EMS, Handbook of Teichmüller Theory}{f 7} (), .
\end{enumerate}

\subsection{Preprints}

\begin{enumerate}
\item Apoorva Khare, Smooth entrywise positivity preservers, a Horn–Loewner master theorem, and Schur polynomial identities, preprint.
\item Apoorva Khare, The non-compact normed space of norms on a finite-dimensional Banach space, preprint.
\item Projesh Nath Choudhury and Apoorva Khare, Distance matrices of a tree: two more invariants, and in a unified framework, preprint.
\item Arvind Ayyer, Matthieu Josuat-Vergès, Sanjay Ramassamy, Extensions of partial cyclic orders and consecutive coordinate polytopes , preprint.
\item Arvind Ayyer, Pooja Singla, Random motion on finite rings, II: Noncommutative rings , preprint.
\item Arvind Ayyer, Amritanshu Prasad, Steven Spallone, Macdonald trees and determinants of representations for finite Coxeter groups , preprint.
\item Dilip P. Patil, On the connectedness of the fiber product of two connected finte free affine schemes. , preprint.
\item Gautam Bharali and Anwoy Maitra, A weak notion of visibility, a family of examples, and Wolff--Denjoy theorems, submitted for publication.
\item Arvind Ayyer and Pooja Singla, Random motion on finite rings, II: Noncommutative rings, preprint.
\item M. Hassain and Pooja Singla, Representation Growth of Special Compact Linear Groups of Order Two, preprint.
\item Shiv Prakash Patel and Pooja Singla, A multiplicity one theorem for groups of type $A_n$ over discrete valuation rings, preprint.
\item Shifra Reif, R. Venkatesh, On tensor products of irreducible integrable representations, preprint.
\item Deniz Kus, R. Venkatesh, Borel-de Siebenthal theory for affine reflection systems, preprint.
\item Louigi Addario-Berry, Sanchayan Sen, Geometry of the minimal spanning tree of a random 3-regular graph, preprint.
\item Sourav Hait, Sayan Bagchi, Luz Roncal and Sundaram Thangavelu, On the lacunary spherical maximal function on the Heisenberg group, preprint.
\item TIRTHANKAR BHATTACHARYYA, SUBRATA SHYAM ROY, AND TAPESH YADAV, THE AGLER-YOUNG CLASS, submitted for publication.
\item Tirthankar Bhattacharyya, B. Krishna Das, Haripada Sau, Toeplitz operators and pseudo-extensions, submitted for publication.
\item Tirthankar Bhattacharyya, Anindya Biswas, Vikramjeet Singh Chandel, On the Nevanlinna problem - Characterization of all Schur-Agler class solutions affiliated with a given kernel, submitted for publication.
\item Tirthankar Bhattacharyya, B. Krishna Das, Haripada Sau, Toeplitz operators on the symmetrized bidisc, submitted for publication.
\item Vamsi Pritham Pingali, A vector bundle version of the Monge-Ampere equation, preprint.
\item Ved V. Datar and Vamsi Pritham Pingali, On coupled constant scalar curvature Kaehler metrics, preprint.
\item Ved Datar, Adam Jacob, Yuguang Zhang, Adiabatic limits of anti-self-dual connections on collapsed K3 surfaces, preprint.
\item Ved V. Datar, Vamsi Pritham Pingali, On coupled constant scalar curvature K\"ahler metrics, preprint.
\end{enumerate}

% from web page

\subsection{Publications (all kinds) from web page}

\begin{enumerate}
\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi, Generalization of unfolding operator for highly oscillating
smooth boundary domains and homogenization {\em Calc. Var. Partial Differential Equations} {\bf 57} (2018), Art. 86, 30.
\item Akutagawa, Kazuo and Endo, Hisaaki and Seshadri, Harish, A gap theorem for positive {E}instein metrics on the
four-sphere {\em Math. Ann.} {\bf 373} (2019), 1329--1339.
\item Amritkar, R. E. and Rangarajan, G., Spatially synchronous extinction of species under external forcing {\em Physical Review Letters} {\bf 96} (2006), 258102.
\item Anamby, Pramath and Das, Soumya, Distinguishing {H}ermitian cusp forms of degree 2 by a certain
subset of all {F}ourier coefficients {\em Publ. Mat.} {\bf 63} (2019), 307--341.
\item Anamby, Pramath and Das, Soumya, Bounds for the {P}etersson norms of the pullbacks of
{S}aito-{K}urokawa lifts {\em J. Number Theory} {\bf 191} (2018), 289--304.
\item Arunkumar, G. and Kus, Deniz and Venkatesh, R., Root multiplicities for {B}orcherds algebras and graph
coloring {\em J. Algebra} {\bf 499} (2018), 538--569.
\item Ayyer, A. and Finn, C. and Roy, D., Matrix product solution of a left-permeable two-species asymmetric exclusion process {\em Physical Review E} {\bf 97} (2018), 5225.
\item Ayyer, Arvind and Behrend, Roger E., Factorization theorems for classical group characters, with
applications to alternating sign matrices and plane
partitions {\em J. Combin. Theory Ser. A} {\bf 165} (2019), 78--105.
\item Ayyer, Arvind and Bouttier, J\'{e}r\'{e}mie and Corteel, Sylvie and
Linusson, Svante and Nunzi, Fran\c{c}ois, Bumping sequences and multispecies juggling {\em Adv. in Appl. Math.} {\bf 98} (2018), 100--126.
\item Ayyer, Arvind and Finn, Caley and Roy, Dipankar, Matrix product solution of a left-permeable two-species
asymmetric exclusion process {\em Phys. Rev. E} {\bf 97} (2018), 012151, 10.
\item Ayyer, Arvind and Finn, Caley and Roy, Dipankar, The phase diagram for a multispecies left-permeable asymmetric
exclusion process {\em J. Stat. Phys.} {\bf 174} (2019), 605--621.
\item Ayyer, Arvind and Prasad, Amritanshu and Spallone, Steven, Odd partitions in {Y}oung's lattice {\em S\'{e}m. Lothar. Combin.} {\bf 75} (2018), Art. B75g, 13.
\item Ayyer, Arvind and Ramassamy, Sanjay, The {H}ilbert-{G}alton board {\em ALEA Lat. Am. J. Probab. Math. Stat.} {\bf 15} (2018), 755--774.
\item B\"{o}cherer, Siegfried and Das, Soumya, Cuspidality and the growth of {F}ourier coefficients of
modular forms {\em J. Reine Angew. Math.} {\bf 741} (2018), 161--178.
\item B\"{o}cherer, Siegfried and Das, Soumya, Petersson norms of not necessarily cuspidal {J}acobi modular
forms and applications {\em Adv. Math.} {\bf 336} (2018), 335--376.
\item Bagchi, Sayan and Thangavelu, Sundaram, Weighted norm inequalities for {W}eyl multipliers and fourier
multipliers on the {H}eisenberg group {\em J. Anal. Math.} {\bf 136} (2018), 1--29.
\item Baklouti, Ali and Thangavelu, Sundaram, Hardy and {M}iyachi theorems for {H}eisenberg motion groups {\em Nagoya Math. J.} {\bf 229} (2018), 1--20.
\item Balhara, Rakesh, Hardy's inequality for the fractional powers of the {G}rushin
operator {\em Proc. Indian Acad. Sci. Math. Sci.} {\bf 129} (2019), Art. 33, 25.
\item Balodi, Mamta and Palcoux, Sebastien, On {B}oolean intervals of finite groups {\em J. Combin. Theory Ser. A} {\bf 157} (2018), 49--69.
\item Banerjee, Abhishek, On {A}uslander's formula and cohereditary torsion pairs {\em Commun. Contemp. Math.} {\bf 20} (2018), 1750071, 27.
\item Banerjee, Abhishek, Modular {H}ecke algebras over {M}\"{o}bius categories {\em J. Geom. Phys.} {\bf 131} (2018), 23--40.
\item Banerjee, Abhishek, A topological {N}ullstellensatz for tensor-triangulated
categories {\em C. R. Math. Acad. Sci. Paris} {\bf 356} (2018), 365--375.
\item Banerjee, Abhishek, Quasimodular {H}ecke algebras and {H}opf actions {\em J. Noncommut. Geom.} {\bf 12} (2018), 1041--1080.
\item Banerjee, Abhishek, Completions of monoid objects and descent results {\em J. Algebra} {\bf 507} (2018), 362--399.
\item Basu, Arnab and Ghosh, Mrinal K., Nonzero-sum risk-sensitive stochastic games on a countable
state space {\em Math. Oper. Res.} {\bf 43} (2018), 516--532.
\item Belton, Alexander and Guillot, Dominique and Khare, Apoorva
and Putinar, Mihai, Simultaneous kernels of matrix {H}adamard powers {\em Linear Algebra Appl.} {\bf 576} (2019), 142--157.
\item Benedetti, Carolina and Gonz\'{a}lez D'Le\'{o}n, Rafael S. and Hanusa,
Christopher R. H. and Harris, Pamela E. and Khare, Apoorva and
Morales, Alejandro H. and Yip, Martha, The volume of the caracol polytope {\em S\'{e}m. Lothar. Combin.} {\bf 80B} (2018), Art. 87, 12.
\item Bera, Sayani and Verma, Kaushal, Some aspects of shift-like automorphisms of {$\Bbb C^k$} {\em Proc. Indian Acad. Sci. Math. Sci.} {\bf 128} (2018), Art. 10, 48.
\item Bharali, Gautam and Biswas, Indranil and Divakaran, Divakaran
and Janardhanan, Jaikrishnan, Proper holomorphic mappings onto symmetric products of a
{R}iemann surface {\em Doc. Math.} {\bf 23} (2018), 1291--1311.
\item Bhattacharyya, Tirthankar and Sau, Haripada, Holomorphic functions on the symmetrized bidisk---realization,
interpolation and extension {\em J. Funct. Anal.} {\bf 274} (2018), 504--524.
\item Bhimani, Divyang G. and Balhara, Rakesh and Thangavelu,
Sundaram, Hermite multipliers on modulation spaces {\em Analysis and partial differential equations: perspectives from
developing countries} {\bf 275} (2019), 42--64.
\item Bhosle, Usha N., Hitchin pairs on reducible curves {\em Internat. J. Math.} {\bf 29} (2018), 1850015, 49.
\item Biswas, Indranil and D'Mello, Shane and Mukherjee, Ritwik and
Pingali, Vamsi P., Rational cuspidal curves on del-{P}ezzo surfaces {\em J. Singul.} {\bf 17} (2018), 91--107.
\item Biswas, Indranil and Pingali, Vamsi Pritham, A characterization of finite vector bundles on {G}auduchon
astheno-{K}\"{a}hler manifolds {\em \'{E}pijournal Geom. Alg\'{e}brique} {\bf 2} (2018), Art. 6, 13.
\item Biswas, Rahul and Dond, Asha K. and Gudi, Thirupathi, Edge patch-wise local projection stabilized nonconforming
{FEM} for the {O}seen problem {\em Comput. Methods Appl. Math.} {\bf 19} (2019), 189--214.
\item Biswas, Shibananda and Ghosh, Gargi and Misra, Gadadhar and
Shyam Roy, Subrata, On reducing submodules of {H}ilbert modules with
{${S}_n$}-invariant kernels {\em J. Funct. Anal.} {\bf 276} (2019), 751--784.
\item Borah, Diganta and Haridas, Pranav and Verma, Kaushal, Comments on the {G}reen's function of a planar domain {\em Anal. Math. Phys.} {\bf 8} (2018), 383--414.
\item Borkar, V. S and Jain, S and Rangarajan, G., Dynamics of individual specialization and global diversification in   communities {\em Complexity} {\bf 3} (1998), 50.
\item Borkar, V. S. and Jain, S. and Rangarajan, G., Generalized replicator dynamics as a model of specialization and   diversity in societies {\em Advances in Complex Systems} {\bf 1} (1998), 325.
\item Borkar, V. S. and Jain, S. and Rangarajan, G., Collective behaviour and diversity in economic communities: Some insights from an evolutionary game {\em The Application of Econophysics} {\bf } (2003), .
\item Borkar, V. S. and Jain, S. and Rangarajan, G., Evolutionary mechanisms of organization in complex systems {\em Mathematical Models for Bioengineering and Probabilistic Systems} {\bf } (2005), 323.
\item Brenner, Susanne C. and Gudi, Thirupathi and Porwal, Kamana
and Sung, Li-Yeng, A {M}orley finite element method for an elliptic distributed
optimal control problem with pointwise state and control
constraints {\em ESAIM Control Optim. Calc. Var.} {\bf 24} (2018), 1181--1206.
\item Burton, Benjamin A. and Datta, Basudeb and Singh, Nitin and
Spreer, Jonathan, A construction principle for tight and minimal triangulations
of manifolds {\em Exp. Math.} {\bf 27} (2018), 22--36.
\item Carstensen, Carsten and Dond, Asha K. and Rabus, Hella, Quasi-optimality of adaptive mixed {FEM}s for non-selfadjoint
indefinite second-order linear elliptic problems {\em Comput. Methods Appl. Math.} {\bf 19} (2019), 233--250.
\item Chandel, Vikramjeet Singh, The three-point {P}ick-{N}evanlinna interpolation problem on
the polydisc {\em Complex Var. Elliptic Equ.} {\bf 63} (2018), 1341--1352.
\item Chatterjee, Esha and Hassan, Sk. Sarif, On the asymptotic character of a generalized rational
difference equation {\em Discrete Contin. Dyn. Syst.} {\bf 38} (2018), 1707--1718.
\item Chen, Y. and Rangarajan, G. and Ding,. M., General stability analysis of synchronized dynamics in coupled systems {\em Physical Review E} {\bf 67} (2003), 26209.0.
\item Chopra, R. and Murthy, C. R. and Rangarajan, G., Statistical tests for detecting Granger causality {\em IEEE Transactions on Signal Processing} {\bf } (2018), 5803.
\item Choudhury, Projesh Nath and Kannan, M. Rajesh and Sivakumar,
K. C., A note on linear preservers of semipositive and minimally
semipositive matrices {\em Electron. J. Linear Algebra} {\bf 34} (2018), 687--694.
\item Ciaurri, \'{O}scar and Roncal, Luz and Thangavelu, Sundaram, Hardy-type inequalities for fractional powers of the
{D}unkl-{H}ermite operator {\em Proc. Edinb. Math. Soc. (2)} {\bf 61} (2018), 513--544.
\item Das, Soumya and Khan, Rizwanur, The third moment of symmetric square {$L$}-functions {\em Q. J. Math.} {\bf 69} (2018), 1063--1087.
\item Das, Soumya and Kohnen, Winfried, On sign changes of eigenvalues of {S}iegel cusp forms of genus
2 in prime powers {\em Acta Arith.} {\bf 183} (2018), 167--172.
\item Datta, Basudeb and Maity, Dipendu, Semi-equivelar maps on the torus and the {K}lein bottle are
{A}rchimedean {\em Discrete Math.} {\bf 341} (2018), 3296--3309.
\item Daya Sagar B. S. and Rangarajan, G. and Veneziano, D., Fractals in Geophysics {\em Chaos, Solitons and Fractals} {\bf 19} (2004), 237.
\item Dhamala, M. and Rangarajan, G. and Ding, M., Estimating Granger causality from Fourier and wavelet transforms of time series data {\em Physical Review Letters} {\bf 100} (2008), 18701.0.
\item Dhamala, M. and Rangarajan, G. and Ding, M., Analyzing information flow in brain networks with nonparametric Granger causality {\em NeuroImage} {\bf 41} (2008), 354.
\item Ding. M and Rangarajan. G, Parametric spectral analysis {\em Encyclopedia of Computational Neuroscience: Springer Reference} {\bf } (2015), .
\item Dond, Asha K. and Gudi, Thirupathi, Patch-wise local projection stabilized finite element methods
for convection-diffusion-reaction problems {\em Numer. Methods Partial Differential Equations} {\bf 35} (2019), 638--663.
\item Dond, Asha K. and Gudi, Thirupathi and Sau, Ramesh CH., An error analysis of discontinuous finite element methods for
the optimal control problems governed by {S}tokes equation {\em Numer. Funct. Anal. Optim.} {\bf 40} (2019), 421--460.
\item Fritz, Tobias and Gadgil, Siddhartha and Khare, Apoorva and Nielsen, Pace P. and Silberman, Lior and Tao, Terence, Homogeneous length functions on groups {\em Algebra Number Theory} {\bf 12} (2018), 1773--1786.
\item Gaddam, Sharat and Gudi, Thirupathi, Inhomogeneous {D}irichlet boundary condition in the {\it a
posteriori} error control of the obstacle problem {\em Comput. Math. Appl.} {\bf 75} (2018), 2311--2327.
\item Gaddam, Sharat and Gudi, Thirupathi, Bubbles enriched quadratic finite element method for the
3{D}-elliptic obstacle problem {\em Comput. Methods Appl. Math.} {\bf 18} (2018), 223--236.
\item Gade, P. and Rangarajan, G., Frustration induced oscillator death on networks {\em Chaos} {\bf 23} (2013), 13892.
\item Ganapathy, R. and Rangarajan, G. and Sood A. K., Granger causality and cross recurrence plots in rheochaos {\em Physical Review E} {\bf 75} (2007), 7305.
\item Gaussier, Herv\'{e} and Seshadri, Harish, Holomorphic motions and complex geometry {\em Proc. Amer. Math. Soc.} {\bf 147} (2019), 301--313.
\item Gaussier, Herv\'{e} and Seshadri, Harish, On the {G}romov hyperbolicity of convex domains in {$\Bbb
C^n$} {\em Comput. Methods Funct. Theory} {\bf 18} (2018), 617--641.
\item Ghara, Soumitra and Kumar, Surjit, On the sum of two subnormal kernels {\em J. Math. Anal. Appl.} {\bf 469} (2019), 1015--1027.
\item Ghosh, A. and Paria, D. and Rangarajan, G. and Ghosh, A., Velocity fluctuations in helical propulsion: How small can a propeller be {\em Journal of Physical Chemistry Letters} {\bf 5} (2014), 62.
\item Ghosh, Mrinal K. and Pradhan, Somnath, Risk-sensitive stochastic differential games with reflecting
diffusions {\em Stoch. Anal. Appl.} {\bf 36} (2018), 1--27.
\item Ghosh, S. and Rangarajan, G. and Sinha, S., Stability of synchronization in a multi-cellular system {\em European Physics Letters} {\bf 92} (2010), 40012.
\item Gudi, Thirupathi and Majumder, Papri, Convergence analysis of finite element method for a parabolic
obstacle problem {\em J. Comput. Appl. Math.} {\bf 357} (2019), 85--102.
\item Gupta, Rajeev and Kumar, Surjit and Trivedi, Shailesh, Von {N}eumann's inequality for commuting operator-valued
multishifts {\em Proc. Amer. Math. Soc.} {\bf 147} (2019), 2599--2608.
\item Gupta, Rajeev and Ramiz Reza, Md., Operator space structures on {$\ell^1(n)$} {\em Houston J. Math.} {\bf 44} (2018), 1205--1212.
\item Gupta, S. C., The classical {S}tefan problem {\em } {\bf } (2018), xxiii+726.
\item Hazra, Somnath, Homogeneous 2-shifts {\em Complex Anal. Oper. Theory} {\bf 13} (2019), 1729--1763.
\item Iyer, Srikanth K., The random connection model: connectivity, edge lengths, and
degree distributions {\em Random Structures Algorithms} {\bf 52} (2018), 283--300.
\item Jha, Abhash Kumar and Sahu, Brundaban, Rankin-{C}ohen brackets on {J}acobi forms of several variables
and special values of certain {D}irichlet series {\em Int. J. Number Theory} {\bf 15} (2019), 925--933.
\item Joshi, A. U. and Sant, D. A. and Parvez, I. A. and Rangarajan, G. and Limaye, M. A. and Mukherjee, S. and Charola, M. J. and Bhatt, M. N. and Mistry, S. P., Sub-surface profiling of granite pluton using microtremor method: southern Aravalli, Gujarat, India {\em International Journal of Earth Sciences} {\bf 106} (2017), 0.
\item Kanukurthi, Bhavana and Bhavana Obbattu, Sai Lakshmi and
Sekar, Sruthi, Non-malleable randomness encoders and their applications {\em Advances in cryptology---{EUROCRYPT} 2018. {P}art {III}} {\bf 10822} (2018), 589--617.
\item Kashyap, G. A. R. S. R. K. and Bapat, D. and Das, D. and Gowaikar, R. D. and  Amritkar, R. E. and Rangarajan, G. and V. Ravindranath and G. Ambika, Synapse loss and progress of Alzheimer&rsquo;s disease - A network model {\em Scientific Reports} {\bf 9} (2019), .
\item Kataria, Kuldeep Kumar and Vellaisamy, Palaniappan, On the convolution of {M}ittag-{L}effler distributions and its
applications to fractional point processes {\em Stoch. Anal. Appl.} {\bf 37} (2019), 115--122.
\item Khare, Apoorva, Generalized nil-{C}oxeter algebras over discrete complex
reflection groups {\em Trans. Amer. Math. Soc.} {\bf 370} (2018), 2971--2999.
\item Khare, Apoorva, Generalized nil-Coxeter algebras {\em S&eacute;m. Lothar. Combin.} {\bf 80B,} (2018), .
\item Khare, Apoorva and Tao, Terence., Schur polynomials, entrywise positivity preservers, and weak majorization. {\em S&eacute;m. Lothar. Combin.} {\bf 80B,} (2018), .
\item Khare, Apoorva and Tikaradze, Akaki, A {C}arlitz--von {S}taudt type theorem for finite rings {\em Linear Algebra Appl.} {\bf 568} (2019), 106--126.
\item Krakovski, Roi and Onn, Uri and Singla, Pooja, Regular characters of groups of type {$\ssf A_n$} over
discrete valuation rings {\em J. Algebra} {\bf 496} (2018), 116--137.
\item Lin, X. and Kulkarni, P. and Bocci, F. and Schafer, N. and Roy, S. and Tsai, M-Y. and He, Y. and Chen, Y and Rajagopalan, K. and Mooney, S. and Zeng, Y. and Weninger, K. and Grishaev, A. and Onuchic, J. and Levine, H. and Wolynes, P. and Salgia, R. and Rangarajan, G. and Uversky, V. and Orban, J. and Jolly, M. K., Structural and Dynamical Order of a Disordered Protein: Molecular Insights into Conformational Switching of PAGE4 at the Systems Level {\em Biomolecules} {\bf 9} (2019), 77.
\item Lin, X. and Roy, S. and Jolly, M. K. and Bocci, F. and Schafer, N. P. and Tsai, M.-Y. and Chen, Y. and He, Y. and Grishaev, A. and Weninger, K. and Orban, J. and Kulkarni, P. and Rangarajan, G. and Levine, H. and J. N. Onuchic, PAGE4 and conformational switching: Insights from molecular dynamics simulations and implications for prostate cancer {\em Journal of Molecular Biology} {\bf 430} (2018), 2422.
\item Mahajan, Prachi and Verma, Kaushal, A comparison of two biholomorphic invariants {\em Internat. J. Math.} {\bf 30} (2019), 1950012, 16.
\item Mahmoudabadi, G. and Rajagopalan, K. and Getzenberg, R. H. and Hannenhalli, S. and Rangarajan, G. and Kulkarni, P., Intrinsically disordered proteins and conformational noise: Implications in cancer {\em Cell Cycle} {\bf 12} (2013), 26.
\item Maity, Dipendu, Pentagonal maps on the torus and the plane {\em Beitr. Algebra Geom.} {\bf 60} (2019), 17--37.
\item Maity, Dipendu and Upadhyay, Ashish Kumar, On centrally symmetric manifolds {\em J. Ramanujan Math. Soc.} {\bf 34} (2019), 21--27.
\item Mal, Arpita and Sain, Debmalya and Paul, Kallol, On some geometric properties of operator spaces {\em Banach J. Math. Anal.} {\bf 13} (2019), 174--191.
\item Manfio, F. and Turgay, N. C. and Upadhyay, A., Biconservative submanifolds in {$\Bbb{S}^n\times\Bbb{R}$} and
{$\Bbb{H}^n\times\Bbb{R}$} {\em J. Geom. Anal.} {\bf 29} (2019), 283--298.
\item Manikandan, K. and Vishnu Priya, N. and Senthilvelan, M. and
Sankaranarayanan, R., Deformation of dark solitons in a {$\Cal{PT}$}-invariant
variable coefficients nonlocal nonlinear {S}chr\"{o}dinger
equation {\em Chaos} {\bf 28} (2018), 083103, 12.
\item Misra, Gadadhar and Pal, Avijit, Curvature inequalities for operators in the {C}owen-{D}ouglas
class and localization of the {W}allach set {\em J. Anal. Math.} {\bf 136} (2018), 31--54.
\item Mooney, S. M. and Rajagopalan, K. and Rangarajan, G. and Kulkarni, P., Cancer/testis antigens and obligate participation in multiple hallmarks of cancer: an update {\em Asian Journal of Andrology} {\bf 18} (2016), 711.
\item Mukherjee, Sajal Kumar and Bera, Sudip, Combinatorial proofs of the {N}ewton-{G}irard and
{C}hapman-{C}ostas-{S}antos identities {\em Discrete Math.} {\bf 342} (2019), 1577--1580.
\item Nalatore, H. and Ding, M. and Rangarajan, G., Mitigating the effects of measurement noise on Granger causality {\em Physical Review E} {\bf 75} (2007), 12883.
\item Nalatore, H. and Ding, M. and Rangarajan, G., Denoising neural data with state-space smoothing: Method and application {\em Journal of Neuroscience Methods} {\bf 179} (2009), 131.
\item Nalatore, H. and Rangarajan, G., Short-window spectral analysis using AMVAR and multitaper methods: A comparison {\em Biological Cybernetics} {\bf 101} (2009), 71.
\item Nalatore. H. and Sasikumar, N. and Rangarajan, G., Effect of measurement noise on Granger causality {\em Physical Review E} {\bf 90} (2014), 25687.
\item Narayanan, E. K. and Singla, Pooja, On monomial representations of finitely generated nilpotent
groups {\em Comm. Algebra} {\bf 46} (2018), 2319--2331.
\item Pal, Chandan and Pradhan, Somnath, Risk sensitive control of pure jump processes on a general
state space {\em Stochastics} {\bf 91} (2019), 155--174.
\item Pal, Ratna and Verma, Kaushal, Ergodic properties of families of {H}\'{e}non maps {\em Ann. Polon. Math.} {\bf 121} (2018), 45--71.
\item Palaniyandi, P. and Rangarajan, G., Critical lattice size limit for synchronized chaotic state in one- and  two-dimensional diffusively coupled map lattices {\em Physical Review E} {\bf 76} (2007), 11906.
\item Palaniyandi, P. and Rangarajan, G., A mathematical model for storage and recall of images using targeted synchronization of coupled maps {\em Scientific Reports} {\bf 7} (2017), 8921.
\item Paul, Kallol and Sain, Debmalya, Birkhoff-{J}ames orthogonality and its application in the
study of geometry of {B}anach space {\em Advanced topics in mathematical analysis} {\bf } (2019), 245--284.
\item Paul, Kallol and Sain, Debmalya and Mal, Arpita, Approximate {B}irkhoff-{J}ames orthogonality in the space of
bounded linear operators {\em Linear Algebra Appl.} {\bf 537} (2018), 348--357.
\item Paul, Kallol and Sain, Debmalya and Mal, Arpita and Mandal,
Kalidas, Orthogonality of bounded linear operators on complex {B}anach
spaces {\em Adv. Oper. Theory} {\bf 3} (2018), 699--709.
\item Pingali, Vamsi P., A note on the deformed {H}ermitian {Y}ang-{M}ills {PDE} {\em Complex Var. Elliptic Equ.} {\bf 64} (2019), 503--518.
\item Pingali, Vamsi Pritham, A note on higher extremal metrics {\em Trans. Amer. Math. Soc.} {\bf 370} (2018), 6995--7010.
\item Pingali, Vamsi Pritham, Representability of {C}hern-{W}eil forms {\em Math. Z.} {\bf 288} (2018), 629--641.
\item Pritham Pingali, Vamsi, Existence of coupled {K}\"{a}hler-{E}instein metrics using the
continuity method {\em Internat. J. Math.} {\bf 29} (2018), 1850041, 8.
\item Quddus, Safdar, Invariants of the {$\Bbb Z_2$} orbifolds of the {P}odle\'{s} two
spheres {\em J. Noncommut. Geom.} {\bf 13} (2019), 257--267.
\item Rangarajan, G., Symplectic integration of nonlinear Hamiltonian systems {\em Pramana -- Journal of Physics} {\bf 48} (1997), 129.
\item Rangarajan, G., Fractals in geophysics {\em “Application of Fractals in Earth Sciences"} {\bf } (2000), .
\item Rangarajan, G., Applications of Lie algebraic techniques to nonlinear Hamiltoniansystems {\em Nonlinear Phenomena} {\bf } (2000), .
\item Rangarajan, G., Polynomial map factorization of symplectic maps {\em International Journal of Modern Physics C} {\bf 14} (2003), 847.
\item Rangarajan, G. and Ding, M., Integrated approach to the assessment of long range correlation in time series data {\em Physical Review E} {\bf 61} (2000), 4991.
\item Rangarajan, G. and Ding, M., First passage time problem for biased continuous-time random walks {\em Fractals} {\bf 8.0} (2000), 139.
\item Rangarajan, G. and Ding, M., First passage distributions for long memory processes {\em Lecture Notes in Physics} {\bf 621} (2003), 167.
\item Rangarajan, G. and Habib, S. and Ryne, R. D., Lyapunov exponents without rescaling and reorthogonalization {\em Physical Review Letters} {\bf 80} (1998), 3747.
\item Rangarajan, G. and Sachidanand, M., Spherical aberration and its correction using Lie algebraic techniques {\em Pramana - Journal of Physics} {\bf 49} (1997), 635.
\item Rangarajan, G. and Sachidanand, M., Invariant metric for nonlinear symplectic maps {\em Pramana - Journal of Physics} {\bf 58} (2002), 477.
\item Rangarajan, G. and Sant, D. A., A climate predictability index and its applications {\em Geophysical Research Letters} {\bf 24} (1997), 1239.
\item Rangarajan, G. and Sant, D. A., Paleoclimatic data from 74KL and Guliya cores: New insights {\em Geophysical Research Letters} {\bf 27} (2000), 787.
\item Rangarajan, G. and Sant, D. A., Fractal dimensional analysis of Indian climatic dynamics {\em Chaos, Solitons and Fractals} {\bf 19} (2004), 285.
\item Rangarajan, N. and Fox, Z. and Singh, A. and Kulkarni, P. and Rangarajan, G., Disorder, Oscillatory Dynamics and State Switching: The Role of c-Myc {\em Journal of Theoretical Biology} {\bf 386} (2015), 105.
\item Ray, A. and Sehgal, N. and Karunakaran, S. and Rangarajan, G. and  Ravindranath, V., MPTP activates ASK1-p38 MAPK signaling pathway through TNF-dependent Trx1 oxidation in Parkinsonism mouse model {\em Free Radical Biology & Medicine} {\bf 87} (2015), 312.
\item Reddy, Nanda Kishore, Equality of {L}yapunov and stability exponents for products of
isotropic random matrices {\em Int. Math. Res. Not. IMRN} {\bf } (2019), 606--624.
\item Resmi, V. and Ambika, G. and Amritkar, R. E. and Rangarajan, G., Amplitude death in complex networks induced by environment {\em Physical Review E} {\bf 85} (2012), 19593.
\item Reza, Md. Ramiz, Curvature inequalities for operators in the {C}owen-{D}ouglas
class of a planar domain {\em Indiana Univ. Math. J.} {\bf 67} (2018), 1255--1279.
\item Sain, Debmalya, On extreme contractions and the norm attainment set of a
bounded linear operator {\em Ann. Funct. Anal.} {\bf 10} (2019), 135--143.
\item Sain, Debmalya, On the norm attainment set of a bounded linear operator {\em J. Math. Anal. Appl.} {\bf 457} (2018), 67--76.
\item Sain, Debmalya and Paul, Kallol and Bhunia, Pintu and Bag,
Santanu, On the numerical index of polyhedral {B}anach spaces {\em Linear Algebra Appl.} {\bf 577} (2019), 121--133.
\item Sain, Debmalya and Paul, Kallol and Mal, Arpita, A complete characterization of {B}irkhoff-{J}ames
orthogonality in infinite dimensional normed space {\em J. Operator Theory} {\bf 80} (2018), 399--413.
\item Sain, Debmalya and Paul, Kallol and Mal, Arpita, On approximate {B}irkhoff-{J}ames orthogonality and normal
cones in a normed space {\em J. Convex Anal.} {\bf 26} (2019), 341--351.
\item Sampath, D. and Sabitha, K. R. and Hegde, P. and Jayakrishnan, H. R. and Kutty, B.M. and Chattarji, S. and Rangarajan, G. and Laxmi, T.R., A study on fear memory retrieval and REM sleep in maternal separation and isolation stressed rats {\em Behavioural Brain Research} {\bf 273} (2014), 144.
\item Sanki, Bidyut, Filling of closed surfaces {\em J. Topol. Anal.} {\bf 10} (2018), 897--913.
\item Sanki, Bidyut and Gadgil, Siddhartha, Graphs of systoles on hyperbolic surfaces {\em J. Topol. Anal.} {\bf 11} (2019), 1--20.
\item Sant D. A and Parvez, I. A. and Rangarajan, G. and Patel, J. and Bhatt, M. N. and Salam, T. A. S., Subsurface profiling along Banni Plains and bounding faults, Kachchh, Western India using Microtremors method {\em Journal of Asian Earth Sciences} {\bf 146} (2017), 326.
\item Sant, D. A. and Krishnan, K. and Rangarajan, G. and Basavaiah, N. and Pandya and C.,Sharma, M. and Trivedi, Y., Characterization of flood plain and climate change using multi-proxy records {\em Journal of Indian Geophysical Union} {\bf 8.0} (2004), 39.
\item Sant, D. A. and Rangarajan, G., Onset of climate change at last Glacial-Holocene transition: Role of the tropical Pacific {\em Current Science} {\bf 83} (2002), 1398.
\item Sant, D. A. and Rangarajan, G., Study of Indian climate variability using predictability indices {\em Modern Applied Mathematics} {\bf } (2005), 145.
\item Sant, D. A. and Rangarajan, G. and Krishnan, K. and Basavaiah, N. and Pandya, C. and Sharma, M. and Trivedi, Y., “ Flood plain record of southwest Indian monsoon during the last glacial" {\em Man and Environment} {\bf 31} (2006), 9.0.
\item Sant, D.A. and Parvez, I. A. and Rangarajan, G. and Patel, S. J. and Salam, T. A. S. and Bhatt, M. N., Subsurface imaging of brown coal bearing Tertiary sedimentaries - Deccan Trap interface using microtremor method {\em Journal of Applied Geophysics} {\bf 159} (2018), 362.
\item Seshadri, Harish and Verma, Kaushal, Some aspects of the automorphism groups of domains {\em Handbook of group actions. {V}ol. {III}} {\bf 40} (2018), 145--174.
\item Sukumaran, P. and Parvez, I. A. and Sant, D. A. and Rangarajan, G. and Krishnan, K., Profiling of Late Tertiary - Early Quaternary surface in the lower reaches of Narmada Valley using microtremors {\em Journal of Asian Earth Sciences} {\bf 41} (2011), 325.
\item Sukumaran, P. and Sant, D. A. and Krishnan, K. and Rangarajan, G., High resolution facies record on Late Holocene flood plain sediments from lower reaches of Narmada Valley, Western India {\em Journal of the Geological Society of India} {\bf 79} (2012), 41.
\item Thangavelu, Sundaram, An analogue of {P}fannschmidt's theorem for the {H}eisenberg
group {\em J. Anal.} {\bf 26} (2018), 235--244.
\item Thangavelu, Sundaram and Naidu Dogga, Venku, {$L^p$}-{$L^2$} estimates for solutions of the wave equation
associated to the {G}rushin operator {\em Adv. Pure Appl. Math.} {\bf 9} (2018), 85--92.
\item Truccolo, W. A. and Rangarajan, G. and Chen, Y. and Ding, M., Analyzing stability of equilibrium points in neural networks: A general approach {\em Neural Networks} {\bf 16} (2003), 1453.
\item Turgay, Nurettin Cenk and Upadhyay, Abhitosh, On biconservative hypersurfaces in 4-dimensional {R}iemannian
space forms {\em Math. Nachr.} {\bf 292} (2019), 905--921.
\item Verma, Kaushal, Notes on the boundaries of quadrature domains {\em Anal. Math. Phys.} {\bf 9} (2019), 617--638.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan, Govindan
and Lakshmanan, M., On symmetry preserving and symmetry broken bright, dark and
antidark soliton solutions of nonlocal nonlinear {S}chr\"{o}dinger
equation {\em Phys. Lett. A} {\bf 383} (2019), 15--26.
\item Wen. X. and Rangarajan, G. and Ding, M., Is Granger causality a viable technique for analyzing fMRI data? {\em PLoS One} {\bf 8.0} (2013), e67428.
\item “Albo, Z. and Di Prisco, G. V. and Chen, Y. and Rangarajan, G. and Truccolo, W. and Feng, J. and Vertes, R. P. and Ding, M.", Is partial coherence a viable technique for identifying generators of neural oscillations? {\em Biological Cybernetics} {\bf 90} (2004), 318.
\item “Rangarajan, G. and Habib, S. and Ryne, R. D.", A new method for computing Lyapunov exponents {\em Nonlinear Dynamics and Brain Functioning} {\bf } (1999), .
\end{enumerate}