### 2019-08-08 data/draft-sections.tex

\subsection{Summer Students}

\begin{enumerate}
\item K Philip Thomas, IISc-Bangalore did his/her  summer project under the guidance of R. Venkatesh during May 1--31, 2018
\item Rati Ludhani, IIT Bombay did his/her  summer project under the guidance of R. Venkatesh during May 6--31, 2018
\item Nazia V from IISER Pune did his/her  summer project under the guidance of Pooja Singla during 1st June-30th July, 2018
\item      Ms. Meghna, first year BSc,Christ University, Bangalore did his/her  summer project under the guidance of A. K. Nandakumaran during April 11 – May 10, 2018. 
\item Deeparaj Bhat did his/her KVPY summer project under the guidance of Subhojoy Gupta during May - July 2018
\item Arka Ghosh did his/her KVPY summer project under the guidance of Siddhartha Gadgil during May-June 2018
\item Chinmaya Kausik did his/her KVPY summer project under the guidance of Siddhartha Gadgil during May-June 2018
\item Nishit Pandya did his/her KVPY summer project under the guidance of Siddhartha Gadgil during May-June 2018
\item Julian D'Costa did his/her KVPY summer project under the guidance of Siddhartha Gadgil during May-June 2018
\item Vrunda Rathi did his/her KVPY summer project under the guidance of Arvind Ayyer during May-July
\item KARTIK SINGH did his/her KVPY summer project under the guidance of Arvind Ayyer during May-July
\item ADITHYA UPADHYA  did his/her KVPY summer project under the guidance of Arvind Ayyer during May-July
\item MANISH PANDEY did his/her KVPY summer project under the guidance of Arvind Ayyer during May - July
\item Nandan Malhotra, IISER Mohali did his/her  summer project under the guidance of Srikanth K. Iyer during May - July 2018
\end{enumerate}

\section{Resource Generation}

\begin{enumerate}

\item {\bf A. K. Nandakumaran\,:} A. K. Nandakumaran in a DST-SERB (Matrics) project
\begin{description}
  \item[Project Title\,:] Unfolding Operators to Study Homogenization Problems in Domains with Highly Oscillating Structures
  \item[Duration\,:] 2018-2021
  \item[Grant\,:] 6.6 Lakhs
\end{description}

\item {\bf A. K. Nandakumaran\,:} A. K. Nandakumaran in a DST(SERB) project
\begin{description}
  \item[Project Title\,:] Homogenization of Partial Dierential Equations in Certain Complex Domains
  \item[Duration\,:] 2017-2020
  \item[Grant\,:] Rs. 19.316 Lakhs
\end{description}

\item {\bf Apoorva Khare\,:} Sole PI in an Infosys Foundation project (Young Investigator Award) in a Infosys Foundation, India project
\begin{description}
  \item[Project Title\,:] N/A
  \item[Duration\,:] 2017-2019
  \item[Grant\,:] 600,000 per year
\end{description}

\item {\bf Apoorva Khare\,:} Sole PI in a SERB project (Ramanujan Fellowship) in a SERB, DST, Govt. of India project
\begin{description}
  \item[Project Title\,:] N/A
  \item[Duration\,:] 2018-2023
  \item[Grant\,:] 700000 per year
\end{description}

\item {\bf Apoorva Khare\,:} Sole PI an a SERB project (MATRICS grant) in a SERB, DST, Govt. of India project
\begin{description}
  \item[Project Title\,:] Entrywise powers and power series preserving positivity in fixed dimension
  \item[Duration\,:] 2018-2021
  \item[Grant\,:] 200000 per year
\end{description}

\item {\bf Arvind Ayyer\,:} Sole investigator in an SERB project in a SERB project
\begin{description}
  \item[Project Title\,:]  Exactly solvable exclusion processes
  \item[Duration\,:] 2018--2020
  \item[Grant\,:] Rs. 19,30,490/-
\end{description}

\item {\bf Basudeb Datta\,:} Basudeb Datta, Sole investigator in a SERB, DST project
\begin{description}
  \item[Project Title\,:] Minimal and tight triangulations of some pl manifolds
  \item[Duration\,:] May 13, 2018 to May 12, 2021
  \item[Grant\,:] Rs.660,000
\end{description}

\item {\bf Dilip P Patil\,:} Principal Investigator in a SERB Matrics Project in a DST-SERB project
\begin{description}
  \item[Project Title\,:] Burnside Algebras --- Computational Aspects
  \item[Duration\,:] Oct 2018 to Oct 2021
  \item[Grant\,:] 200000 Per year
\end{description}

\item {\bf Gautam Bharali\,:} Sole investigator in a DST project in a DST project
\begin{description}
  \item[Project Title\,:] Swarnajayanti Fellowship
  \item[Duration\,:] 2015--2020
  \item[Grant\,:] Rs. 56.7 lakhs
\end{description}

\item {\bf Govindan Rangarajan\,:} Sole Investigator in a SERB project
\begin{description}
  \item[Project Title\,:] JC Bose Fellowship
  \item[Duration\,:] 2011-present
  \item[Grant\,:] 1,59,85,000
\end{description}

\item {\bf Govindan Rangarajan\,:} Director in a DST-CNRS project
\begin{description}
  \item[Project Title\,:] Indo-French Centre for Applied Matheamtics
  \item[Duration\,:] 2018-present
  \item[Grant\,:] 6,83,00,000 + matching grant from CNRS
\end{description}

\item {\bf Govindan Rangarajan\,:} Principal Investigator in a Broadcom Foundation project
\begin{description}
  \item[Project Title\,:] EMEA Workshops
  \item[Duration\,:] 2017-present
  \item[Grant\,:] 1,35,00,000
\end{description}

\item {\bf Govindan Rangarajan\,:} Co-Principal Investigator in a Tata Trusts project
\begin{description}
  \item[Project Title\,:] Alzheimer's Research
  \item[Duration\,:] 2014-present
  \item[Grant\,:] 75,00,00,000
\end{description}

\item {\bf Manjunath Krishnapur\,:} Manjunath Krishnapur (PI) in a SERB project
\begin{description}
  \item[Project Title\,:] Zeros of random functions and anti-concentration phenomena
  \item[Duration\,:] 2018-2021
  \item[Grant\,:] 6,60,000
\end{description}

\item {\bf Pooja Singla\,:} Pooja Singla and Sole-investigator in a SERB-MATRICS project
\begin{description}
  \item[Project Title\,:] "Representations of Linear Groups over Local Rings : Gelfand Models "
  \item[Duration\,:] March 2019-March 2022
  \item[Grant\,:] 6,00,000
\end{description}

\item {\bf Pooja Singla\,:} Amritanshu Prasad, Uri Onn, James Borger, Pooja Singla and PI, PI, Co-PI, Co-PI  in a MHRD-SPARC  project
\begin{description}
  \item[Project Title\,:] Representation Zeta Functions
  \item[Duration\,:] March 2019 to march 2021
  \item[Grant\,:] 71,46,350
\end{description}

\item {\bf R. Venkatesh\,:} Investigator in a DST project
\begin{description}
  \item[Project Title\,:] Inspire
  \item[Duration\,:] 2016--2020
  \item[Grant\,:] 35 Lakhs
\end{description}

\item {\bf R. Venkatesh\,:} Investigator in a DST-SERB project
\begin{description}
  \item[Project Title\,:] Study on the structure theory of Borcherds-Kac-Moody algebras and their subalgebras
  \item[Duration\,:] 2018-2020
  \item[Grant\,:] 6 Lakhs
\end{description}

\item {\bf Siddhartha Gadgil\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] Homotopy Type theory and Natural langauge processing for Computer-Assisted Mathematics
  \item[Duration\,:] 2018-2021
  \item[Grant\,:] Rs. 19 lakhs
\end{description}

\item {\bf Srikanth K. Iyer\,:} Sole investigator in a SERB project in a SERB project
\begin{description}
  \item[Project Title\,:] Percolation at Criticality and Connectivity in the Random Connection Model
  \item[Duration\,:] 2019-2022
  \item[Grant\,:] six lakhs
\end{description}

\item {\bf Subhojoy Gupta\,:} Sole Investigator in SERB project in a SERB project
\begin{description}
  \item[Project Title\,:] Degenerations of geometric structures and meromorphic differentials on Riemann surfaces
  \item[Duration\,:] 2018-2020
  \item[Grant\,:] 660000
\end{description}

\item {\bf Sundaram Thangavelu\,:} Principal Investigator in a SERB project
\begin{description}
  \item[Project Title\,:] JC Bose National Fellowship
  \item[Duration\,:] 2018-2022
  \item[Grant\,:] 
\end{description}

\item {\bf Vamsi Pritham Pingali\,:} Vamsi Pritham Pingali (Principal Investigator) in a SERB project
\begin{description}
  \item[Project Title\,:] A study of some analytic and computational aspects of Chern-Weil theory
  \item[Duration\,:] 2017-2020
  \item[Grant\,:] 5,42,290 INR
\end{description}

\item {\bf Vamsi Pritham Pingali\,:} Vamsi Pritham Pingali (Principal Investigator) in a Infosys foundation project
\begin{description}
  \item[Project Title\,:] Young Investigator Award (No title)
  \item[Duration\,:] 2017-2019
  \item[Grant\,:] 600000 INR 
\end{description}

\end{enumerate}

\subsection{Awards}

\begin{enumerate}
\item Apoorva Khare was awarded Ramanujan Fellowship (Apr 2018), ICM Satellite Conference invited talk (Aug 2018), Tarun Tejankit Puraskar from Loksatta newspaper (Mar 2019)
\item R. Venkatesh was awarded Associate of the Indian Academy of Sciences 2018-2022
\item Subhojoy Gupta was awarded Simons Visiting Professorship grant to support a visit to Heidelberg University (September 10-17, 2018) following participation at the Oberwolfach workshop
\item Sundaram Thangavelu was awarded V.V. Narlikar Memorial lecture award from INSA
\end{enumerate}

\subsection{Fellowships of Academies}

\begin{enumerate}
\item Apoorva Khare was elected a N/A
\end{enumerate}

\subsection{Editorial boards}
\begin{enumerate}
\item A. K. Nandakumaran, Editorial Member, TMC Bulletin
\item Apoorva Khare, TRIM Series (Hindustan Book Agency); Sao Paulo Journal of Mathematical Sciences
\item Gautam Bharali, Newsletter of the Ramanujan Mathematical Society
\item Sundaram Thangavelu, Proceedings of the Indian Academy of Sciences
\end{enumerate}