### 2020-08-05 data/draft-sections.tex
added \item to activities

\subsection{Summer Students}

\begin{enumerate}
\item Rohit Kumar, Indian Institute of Science  did his/her KVPY summer project under the guidance of R. Venkatesh during May–June 2019
\item Gobinda Debnath, NIT Agartala did his/her IASc-INSA-NASI Focus Area Science Technology Summer Fellowship summer project under the guidance of Ved Datar during May 15, 2019 - July 15, 2020
\item Arnab Roy, Ramakrishna Mission Residential College did his/her Fellowship from the Indian Academy of Sciences summer project under the guidance of Vamsi Pritham Pingali during 26 May - 19 July
\item Ms. Sradha R Bhat, IISER, Thiruvananthapuram did his/her DST Inspire Fellow summer project under the guidance of A. K. Nandakumaran during May 04 – July 04, 2019
\item Ms. Indulekha, IISER, Thiruvananthapuram did his/her May 04 – July 04, 2019 summer project under the guidance of A. K. Nandakumaran during DST Inspire Fellow
\item Ms. Muskan Agarwal, Mount Carmel College, Bangalore did his/her General summer project under the guidance of A. K. Nandakumaran during May 01 – June 09, 2019
\item Ms. Rajashree Dahal, Mount Carmel College, Bangalore did his/her General summer project under the guidance of A. K. Nandakumaran during May 01 – June 09, 2019
\item Mr. Arya Samanta, IISER, Pune did his/her General summer project under the guidance of A. K. Nandakumaran during June-July, 2019
\item Ms Nirmala Aurad, Dept of Mathematics, Gulbarga University did his/her Academy summer project under the guidance of A. K. Nandakumaran during July 03 – August 30, 2019
\item Susheel Shankar, Indian Institute of Science did his/her KVPY summer project under the guidance of Gautam Bharali during May--July, 2019
\item Aritra Das, IISc did his/her KVPY summer project under the guidance of Arvind Ayyer during May-July
\item Mrinmoy Karmakar, IISc did his/her KVPY summer project under the guidance of Arvind Ayyer during May-Jun
\item Omkar Bhalchandra Baraskar, IISc did his/her KVPY summer project under the guidance of Arvind Ayyer during May-Jun
\item Susheel Shankar, IISc did his/her KVPY summer project under the guidance of Arvind Ayyer during May-June
\item Dhanus M Lal, IISc did his/her KVPY summer project under the guidance of Arvind Ayyer during May-June
\item Prathamesh Gajanan Patil, IISc did his/her KVPY summer project under the guidance of Arvind Ayyer during May-June
\item Achal Kumar, IISc did his/her KVPY summer project under the guidance of Arvind Ayyer during May-June
\item Matthew Honnor did his/her  summer project under the guidance of Mahesh Kakde during 1 month
\item Saptarsi Ghosal, CMI did his/her  summer project under the guidance of Manjunath Krishnapur during June-July 2019
\item Deep Ghoshal, ISI Kolkata did his/her  summer project under the guidance of Manjunath Krishnapur during June-July 
\item Arghya Chakraborty, ISI Kolkata did his/her  summer project under the guidance of Manjunath Krishnapur during June-July 
\item Somak Laha, ISI Kolkata did his/her  summer project under the guidance of Manjunath Krishnapur during June-July
\item Sayak Chatterjee did his/her  summer project under the guidance of Manjunath Krishnapur during June-July
\item Anirban Basak, IISER Kolkata did his/her  summer project under the guidance of Manjunath Krishnapur during June-July
\item Nafia Saleem, IIT Madras did his/her SRFP of IAS summer project under the guidance of Manjunath Krishnapur during June-July
\item Rohit Kumar, IISc did his/her KVPY summer project under the guidance of R VENKATESH during May-June 2019
\item Aman Agrawal, IISc did his/her  summer project under the guidance of Apoorva Khare during May-July 2019
\item Sutanay Bhattacharya, IISc did his/her  summer project under the guidance of Apoorva Khare during May-July 2019
\item Samarth Hawaldar, IISc did his/her  summer project under the guidance of Apoorva Khare during May-July 2019
\item Swaroop Hegde, NISER Bhubaneswar did his/her  summer project under the guidance of Apoorva Khare during May-July 2019
\item Anant Mudgal, CMI did his/her  summer project under the guidance of Apoorva Khare during May-July 2019
\item Abhilash Mukherjee, IISc did his/her  summer project under the guidance of Apoorva Khare during May-July 2019
\item Pulkit Sinha, IISc did his/her  summer project under the guidance of Apoorva Khare during May-July 2019
\item Chinmaya Kausik, IISc did his/her KVPY summer project under the guidance of Siddhartha Gadgil during May-June 2019
\item Nilendu Das, IISER Mohali did his/her Academy summer programme summer project under the guidance of Siddhartha Gadgil during May, June 2019
\item Nishit Pandya, IISc did his/her KVPY summer project under the guidance of Siddhartha Gadgil during May, June 2019
\item Julian D'Costa, IISc did his/her KVPY summer project under the guidance of Siddhartha Gadgil during May, June 2019
\end{enumerate}

\section{Resource Generation}

\begin{enumerate}

\item {\bf A. K. Nandakumaran\,:} Sole investigator in a DST (Matrix) project
\begin{description}
  \item[Project Title\,:] Unfolding Operators to Study Homogenization Problems in Domains with Highly Oscillating Structures
  \item[Duration\,:] 2018-2021
  \item[Grant\,:] 6.6 lakhs
\end{description}

\item {\bf A. K. Nandakumaran\,:} Sole investigator in a DST project
\begin{description}
  \item[Project Title\,:] Homogenization of Partial Differential Equations in Certain Complex Domains
  \item[Duration\,:] 2017-2020
  \item[Grant\,:] 19.316 Lakhs
\end{description}

\item {\bf Apoorva Khare\,:} Sole investigator in a Infosys Foundation project
\begin{description}
  \item[Project Title\,:] Young Investigator Award
  \item[Duration\,:] 2017-2019
  \item[Grant\,:] 12 lakhs
\end{description}

\item {\bf Apoorva Khare\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] Ramanujan Fellowship
  \item[Duration\,:] 2018-2023
  \item[Grant\,:] 35 lakhs
\end{description}

\item {\bf Apoorva Khare\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] MATRICS: Entrywise Powers and Power Series Preserving Positivity in Fixed Dimension
  \item[Duration\,:] 2018-2021
  \item[Grant\,:] 6.6 lakhs
\end{description}

\item {\bf Apoorva Khare\,:} Sole investigator in a DST and SERB project
\begin{description}
  \item[Project Title\,:] SwarnaJayanti Fellowship: Transformations preserving positivity in analysis, and connections to algebra and combinatorics
  \item[Duration\,:] 2020-2025
  \item[Grant\,:] Rs. 1,24,88,260
\end{description}

\item {\bf Arvind Ayyer\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:]  Exactly solvable exclusion processes
  \item[Duration\,:] 2017--2020
  \item[Grant\,:] 19,30,490
\end{description}

\item {\bf Arvind Ayyer\,:} Co-investigator in a The Royal Society, UK project
\begin{description}
  \item[Project Title\,:] Determinantal Formulas for Multispecies ASEP  
  \item[Duration\,:] 2019--2021
  \item[Grant\,:] 11,759 pounds
\end{description}

\item {\bf Basudeb Datta\,:} Sole investigator in a SERB, DST project
\begin{description}
  \item[Project Title\,:] Minimal and Tight Triangulations of some PL Manifolds
  \item[Duration\,:] 2018--2021
  \item[Grant\,:] 660,000
\end{description}

\item {\bf Dilip P. Patil\,:} Sole investigator in a DST-SERB project
\begin{description}
  \item[Project Title\,:] Burnside Algebras-Computational Aspects of Commutative Algebra and Algebraic Geometry
  \item[Duration\,:] 2018 to 2021
  \item[Grant\,:] 2 Lakhs per year
\end{description}

\item {\bf E K Narayanan\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] Heckman-Opdam hypergeometric functions and harmonic analysis
  \item[Duration\,:] 2019-2022
  \item[Grant\,:] Six lakhs
\end{description}

\item {\bf Gautam Bharali\,:} Sole investigator in a DST project
\begin{description}
  \item[Project Title\,:] Swarnajayanti Fellowship
  \item[Duration\,:] 2015--2020
  \item[Grant\,:] Rs. 56.7 lakhs
\end{description}

\item {\bf Govindan Rangarajan\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] JC Bose Fellowship
  \item[Duration\,:] 2011-present
  \item[Grant\,:] 80,00,000
\end{description}

\item {\bf Govindan Rangarajan\,:} Sole investigator in a Broadcom Corporation project
\begin{description}
  \item[Project Title\,:] Brain Inspired Computing
  \item[Duration\,:] 2017-present
  \item[Grant\,:] Rs. 1,37,00,000
\end{description}

\item {\bf Govindan Rangarajan\,:} Sole investigator in a DST project
\begin{description}
  \item[Project Title\,:] Indo-French Centre for Applied Mathematics Phase II
  \item[Duration\,:] 2018-present
  \item[Grant\,:] 1,03,77,511
\end{description}

\item {\bf Govindan Rangarajan\,:} Principal investigator in a CNRS, France project
\begin{description}
  \item[Project Title\,:] Indo-French Centre for Applied Mathematics Phase II
  \item[Duration\,:] 2018-present
  \item[Grant\,:] Euros 1,68,000 (Rs. 1,34,40,000)
\end{description}

\item {\bf Manjunath Krishnapur\,:} Principal investigator in a SERB project
\begin{description}
  \item[Project Title\,:] Zeros of random.functions and anti-concentration phenomena
  \item[Duration\,:] 2018-21
  \item[Grant\,:] 200000
\end{description}

\item {\bf R VENKATESH\,:} Sole investigator in a DST-SERB (MATRICS) project
\begin{description}
  \item[Project Title\,:] Study on the structure theory of Borcherds-Kac-Moody algebras and their subalgebras
  \item[Duration\,:] 2019-2021
  \item[Grant\,:] 6,60,000/-
\end{description}

\item {\bf R VENKATESH\,:} Sole investigator in a DST project
\begin{description}
  \item[Project Title\,:] INSPIRE Faculty award
  \item[Duration\,:] 2016-2021
  \item[Grant\,:] 35,00,000/-
\end{description}

\item {\bf R VENKATESH\,:} Sole investigator in a Infosys foundation project
\begin{description}
  \item[Project Title\,:] Infosys Young Investigator award 
  \item[Duration\,:] 2020-2021
  \item[Grant\,:] 6,00,000/-
\end{description}

\item {\bf R. Venkatesh\,:} Sole investigator in a  SERB-MATRICS project
\begin{description}
  \item[Project Title\,:] Study on the structure theory of Borcherds-Kac-Moody algebras and their subalgebras.
  \item[Duration\,:] 2019-2021
  \item[Grant\,:] 6 Lakhs
\end{description}

\item {\bf R. Venkatesh\,:} Sole investigator in a DST-INPIRE project
\begin{description}
  \item[Project Title\,:] Inspire faculty award
  \item[Duration\,:] 2016-2021
  \item[Grant\,:] 35 Lakhs
\end{description}

\item {\bf R. Venkatesh\,:} Sole investigator in a  Infosys foundation project
\begin{description}
  \item[Project Title\,:] Infosys Young Investigator award 
  \item[Duration\,:] 2020-2021
  \item[Grant\,:] 6 Lakhs
\end{description}

\item {\bf RADHIKA GANAPATHY\,:} Sole investigator in a Infosys Foundation project
\begin{description}
  \item[Project Title\,:] Young Investigator
  \item[Duration\,:] 2019-2021
  \item[Grant\,:] Rs. 3,00,000 p.a.+ Rs. 25,000 monthly salary top-up
\end{description}

\item {\bf SANCHAYAN SEN\,:} Sole investigator in a Science & Engineering Research Board (SERB) project
\begin{description}
  \item[Project Title\,:] Geometry of random discrete structures
  \item[Duration\,:] 2020--2023
  \item[Grant\,:] 220,000 per annum
\end{description}

\item {\bf SANCHAYAN SEN\,:} Sole investigator in a Infosys Foundation, Bangalore project
\begin{description}
  \item[Project Title\,:] 
  \item[Duration\,:] 2019--2021
  \item[Grant\,:] 600,000 per annum
\end{description}

\item {\bf Siddhartha Gadgil\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] Homotopy Type Theory and Natural Language Processing for Computer-Assisted mathematics
  \item[Duration\,:] 2018-2021
  \item[Grant\,:] 1898700
\end{description}

\item {\bf Srikanth Krishnan Iyer\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] Percolation at Criticality and Connectivity in the Random Connection Model
  \item[Duration\,:] March 22, 2019 - March 22, 2022
  \item[Grant\,:] 6,60,000
\end{description}

\item {\bf Subhojoy Gupta\,:} Sole investigator in a SERB, DST project
\begin{description}
  \item[Project Title\,:] Degenerations of geometric structures and meromorphic differentials on Riemann surfaces
  \item[Duration\,:] 2018-2021
  \item[Grant\,:] 6 lakhs
\end{description}

\item {\bf Subhojoy Gupta\,:} Co-investigator in a Université Côte-D'azur and IISc  project
\begin{description}
  \item[Project Title\,:] Branched projective structures
  \item[Duration\,:] 2019-2021
  \item[Grant\,:] 6 lakhs
\end{description}

\item {\bf Sundaram Thangavelu\,:} Sole investigator in a DST project
\begin{description}
  \item[Project Title\,:] J.C.Bose Fellowship
  \item[Duration\,:] 2008-2022
  \item[Grant\,:] 
\end{description}

\item {\bf Thirupathi Gudi\,:} Sole investigator in a DST project
\begin{description}
  \item[Project Title\,:] Approximation Theory for PDEconstrained optimization and Variational Inequalities
  \item[Duration\,:] 2019-2022
  \item[Grant\,:] 600000
\end{description}

\item {\bf Vamsi Pritham Pingali\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] A study of some analytic and computational aspects of Chern-Weil theory
  \item[Duration\,:] 2017-2020
  \item[Grant\,:] 542290
\end{description}

\item {\bf Vamsi Pritham Pingali\,:} Sole investigator in a Infosys Foundation project
\begin{description}
  \item[Project Title\,:] Infosys Young Investigator Award
  \item[Duration\,:] 2017-2019
  \item[Grant\,:] 600000
\end{description}

\end{enumerate}

\subsection{Awards}

\begin{enumerate}
\item Apoorva Khare was awarded (1) Delivered a Plenary talk at the 2019 ILAS (International Linear Algebra Society) meeting in Rio de Janeiro, Brazil. Also selected as the LAMA Lecturer. (2) Delivered the 20th Professor Wazir Hasan Abdi Memorial Lecture, in Cochin University of Science and Technology (CUSAT). (3) Awarded the SwarnaJayanti Fellowship for 2018-19, in Mathematics.
\item Arvind Ayyer was awarded My student, Surjadipta De Sarkar received a fellowship and travel grant to attend a program on Algebraic and Enumerative Combinatorics at the Mittag-Leffler Institute from 13 January - 30 April 2020.
\item R VENKATESH was awarded   Infosys Young Investigator award from the Infosys foundation, 2020.
\item R. Venkatesh was awarded Associate of the Indian Academy of Sciences (2018-2022), Infosys Young Investigator award from the Infosys foundation, 2020.
\item Ved Datar was awarded Infosys Young Investigator Award
\end{enumerate}

\subsection{Fellowships of Academies}

\begin{enumerate}

\end{enumerate}

\subsection{Editorial boards}
\begin{enumerate}
\item Apoorva Khare, Journals: (1) Indian Journal of Pure and Applied Mathematics (2) Linear and Multilinear Algebra (3) Sao Paulo Journal of Mathematical Sciences. Book Series: (4) TRIM (Texts and Readings in Mathematics)
\item Arvind Ayyer, Indian Journal of Pure and Applied Mathematics, Algebraic Combinatorics
\item E K Narayanan, Editorial board of the Indian Journal of Pure and Applied Mathematics
\item Gautam Bharali, Newsletter of the Ramanujan Mathematical Society 
\item Mrinal Kanti Ghosh, Associate Editor, Stochastic Analysis and Applications
\item Sundaram Thangavelu, Proceedings of the Indian Academy of Sciences (Math.), Journal of Ramanujan Mathematical Society, Journal of Analysis, Tunisian Journal of Mathematics, Journal of Fourier Analysis and Applications.
\end{enumerate}

\subsection{Other Activities: 2018-19}
\begin{enumerate}
\item A. K. Nandakumaran was Member, Core Committee, KVPY
\item A. K. Nandakumaran was Member, Library Committee
\item Apoorva Khare was Talks: (1) 2019-04-12: Eigenfunctions Seminar, titled "Schoenberg: from metric geometry to matrix positivity" (2) 2019-11-06: 1st Math Symposium of Infosys Young Investigators: Talk, titled "Groups with norms; distance matrices and Zariski density"
\item Apoorva Khare was Organizer of Eigenfunctions,  APRG Seminar, Algebra-Combinatorics Seminar, (co-organiser)
\item Apoorva Khare was Contributed in preparing the PhD Qualifiers in Algebra
\item Apoorva Khare was Student thesis/project supervision: (1) Swarnalipa Datta, MS Thesis (2) Soumik Dutta, BS Project
\item Arvind Ayyer was Organised ICTS workshop on "Universality in random structures: Interfaces, Matrices, Sandpiles" from 14/01 -08/02/2019 jointly with Riddhipratim Basu and Manjunath Krishnapur.
\item Basudeb Datta was B. Datta was an member of the Committees for the following: (i)  Promotion of Professor in Mathematics in HAG scale in NIT, Rourkela in November 2018, (ii) Faculty selection for the post of Associate and Assistant Professors in the Department of Mathematics of IIT, Delhi in May, 2019, (iii) Faculty selection in mathematics at NISER, Bhubaneswar in June, 2019.
\item Basudeb Datta was B. Datta was  examiner for the following PhD theses:  (i) ``Dynamics of maps possessing the almost average shadowing property and variants of transitivity" of Mukta Garg, Univ of Delhi in 2018, (ii) ``On some problems in Algebraic Combinatorics" of Sajal K. Mukherjee, Visva-Bharati University, Santiniketan in 2018 (also the external examiner for the viva-voce examination  in June 2019), (iii) ``Group actions on Dold and Milnor manifolds" of Pinka Dey, IISER Mohali in 2019 (also the external examiner for the viva-voce examination in December 2019), (iv) ``Variants of transitivity, sensitivity and expansivity in non-autonomous discrete systems" of Radhika Vashisht,  Univ. of Delhi in 2020. 
\item Dilip P. Patil was Taught the course: MA 312 Commutative Algebra course during Jan-April 2020
\item Dilip P. Patil was Taught  (12 Week long)  NPTEL Course during Jan-April 2020
\item Gautam Bharali was Member of the Faculty Selection Committee at IIIT Guwahati in June 2018
\item Gautam Bharali was Member of the executive committee of IWM (Indian Women and Mathematics)  
\item R VENKATESH was We hosted 70 students and 5 faculty members from Madras Christian College as a part of their Educational visit to IISc, Bangalore on 16th October 2018. Two lectures were arranged for them on 16th October 2018. 
\item R. Venkatesh was Algebra Symposium, Department of Mathematics, IISc, Bangalore, India (February 12, 2020).
\item R. Venkatesh was Discussion Meeting on Representation Theory 2019, Department of Mathematics, IISc, Bangalore (December 14 - 16, 2019). Partially funded by the UGC-CAS grant.  Two lecture series was delivered by Professor Vyjayanthi Chari (UC Riverside, USA) and Professor Anne Schilling (UC Davis, USA) as part of this discussion meeting.
\item R. Venkatesh was NCM - Instructional School for Teachers on Analysis and PDE, Department of Mathematics, IISc, Bangalore (May 6 - 18, 2019). 
\item Srikanth Krishnan Iyer was Gave a series of ten lectures in the Advanced Instructional School on Stochastic Processes, June 2019 organized by NICER, Bhubaneshwar on Probability Theory.
\item Srikanth Krishnan Iyer was Gave a series of eight lectures in the annual Lecture Series in Probability XIV held at ISI-Delhi between December 2 - 6, 2019 on Limit Theorems in Stochastic Geometry.
\item Srikanth Krishnan Iyer was Gave a talk titled :Random Connection Models" at the RBI, Mumbai on the occasion of Statistics day, June 26, 2019.
\item Srikanth Krishnan Iyer was Gave a talk titled "Phase Transitions in Random Networks" at the FLAIR 4 Big Data Summer School on July 6, 2019.
\item Subhojoy Gupta was Co-organized a discussion meeting at ICTS Bangalore in December 2018, on “Surface-Group Representations and Projective Structures". Website: https://www.icts.res.in/discussion-meeting/SGPS2018.
\item Subhojoy Gupta was Referee for “Memoirs of the American Mathematical Society",  “Indian Journal of Pure and Applied Mathematics"
\item Subhojoy Gupta was M.Sc. thesis referee for a student of Chennai Mathematical Institute,  Satyajit Maity
\item Subhojoy Gupta was Member of Research Advisory Committee for Ramya Nair, a PhD student at IISER Pune. 
\item Thirupathi Gudi was Organized Workshop on Finite Elements for Nonlinear and Multiscale Problems during Feb 28 – Mar 03, 2020.
\item Vamsi Pritham Pingali was Guided the Master's Thesis ("Kahler geometry with a view towards the Calabi conjecture") of Vishnu Nandakumaran of IISER Pune (academic year of 2019-2020)
\item Vamsi Pritham Pingali was Visited ICTP as a junior associate during May 2019 and delivered a talk ("A vector bundle version of the Monge-Ampere equation")
\end{enumerate}

\subsection{Other Activities: 2019-20}
\begin{enumerate}
\item A. K. Nandakumaran was Convener, KVPY
\item A. K. Nandakumaran was Member, Library Committee
\item Apoorva Khare was Member of two UG Project Presentation Evaluation Committees
\item Apoorva Khare was Pre-RMO (Regional Mathematics Olympiad) - helped in setting the question paper
\item Apoorva Khare was Senate nominee for Comprehensive Exam of Vinayak Ramkumar (student of P. Vijay Kumar, ECE)
\item Apoorva Khare was Advising: (1) Prateek Kumar Vishwakarma, Int-PhD (passed his Comprehensive Exam) (2) Babhrubahan Bose, Int-PhD (passed his Comprehensive Exam) (3) G.V.K. Teja, PhD (passed his Comprehensive Exam) (4) Swarnalipa Datta, Int-PhD (downgraded to MS) Mentoring: (5) Dr. Projesh Nath Choudhury, NBHM postdoc
\item Arvind Ayyer was Referee for the thesis of Majil Pratim Saikia at University of Vienna.
\item Arvind Ayyer was Referee for the thesis of Mr. Jyotirmoy Ganguly at IISER Pune.
\item Arvind Ayyer was Thesis examiner for Mr. Sudip Bera, Visva-Bharati University.
\item Gautam Bharali was Member of the executive committee of IWM (Indian Women and Mathematics)
\item Gautam Bharali was Served as an independent expert for the Swiss National Science Foundation for the review of project proposals in pure mathematics in 2019.
\item R VENKATESH was Algebra Symposium, Department of Mathematics, IISc, Bangalore, India (February 12, 2020).
\item R VENKATESH was Discussion Meeting on Representation Theory 2019, Department of Mathematics, IISc, Bangalore (December 14 - 16, 2019). Partially funded by the UGC-CAS grant.  Two lecture series was delivered by Professor Vyjayanthi Chari (UC Riverside, USA) and Professor Anne Schilling (UC Davis, USA) as part of this discussion meeting. 
\item R VENKATESH was NCM - Instructional School for Teachers on Analysis and PDE, Department of Mathematics, IISc, Bangalore (May 6 - 18, 2019). 
\item Siddhartha Gadgil was thesis examiner for a thesis at IIT, Kanpur
\item Siddhartha Gadgil was thesis examiner for a thesis at Delhi University
\item Siddhartha Gadgil was thesis examiner for a thesis for TIFR, Mumbai
\item Siddhartha Gadgil was member of selection committee at BHU
\item Subhojoy Gupta was Member of Research Advisory Committee for Ramya Nair, a PhD student at IISER Pune. 
\item Subhojoy Gupta was Co-organizer of an Advanced Instructional School on Riemannian Geometry, sponsored by the National Centre for Mathematics, held in July 2019.
\item Subhojoy Gupta was Referee for "Proceedings of the London Mathematical Society", "Journal of Topology", "Manuscripta Mathematica" and reviewer for AMS and Zentralblatt Math Reviews
\item Subhojoy Gupta was Supervised UG final year project of Shabarish Chenakkod
\end{enumerate}