### 2020-08-06 data/draft-publications.tex
clean generation

\subsection{Papers in Journals}

\begin{enumerate}
\item S Aiyappan, A. K. Nandakumaran and Abu Sufian, \emph{Asymptotic analysis of Boundary Optimal Control Problem on a General Branched Structure},  Mathematical Methods in the Applied Sciences {\bf 42} (2019. DOI: 10.1002/mma.5748), 6407-6434 , DOI 10.1002/mma.5748.
\item K. Murali, A. K. Nandakumaran, Turgurt Durduran, Hari M Varma, \emph{Recovery of the diffuse correlation spectroscopy data-type from speckle contrast measurements: towards low-cost, deep-tissue blood flow measurements}, Biomedical Optics Express 5395 {\bf 10} (2019), .
\item S Aiyappan, A. K. Nandakumaran and Ravi Prakash, \emph{Locally Periodic Unfolding Operator for Highly Oscillating Rough Domains}, Annali di Matematica Pura ed Applicata {\bf 198} (2019), 1931-1954, DOI 10.1007/s10231-019-00848-7.
\item S Aiyappan, A. K. Nandakumaran and Ravi Prakash, \emph{Semi-linear optimal control problem on a smooth oscillating domain}, Communications in Contemporary Mathematics {\bf } (2019), 1-26, DOI 10.1142/S0219199719500299.
\item S Aiyappan, Editha C. Jose, Ivy Carol B. Lomerio and A. K. Nandakumaran, \emph{ Control Problem on a Rough Circular Domain and Homogenization}, Asymptotic Analysis {\bf 115} (2019),  19-46, DOI 10.3233/ASY-191526.
\item Khare, Apoorva and Rajaratnam, Bala, \emph{Probability inequalities and tail estimates for metric semigroups}, Advances in Operator Theory (Special issue in honor of Rajendra Bhatia) {\bf 5} (2020), 779-795, DOI 10.1007/s43036-020-00048-8, \url{https://link.springer.com/article/10.1007%2Fs43036-020-00048-8}.
\item Ghosh, Subhajit, \emph{Total Variation Cutoff for the Transpose Top-2 with Random Shuffle}, to appear in Journal of Theoretical Probability, DOI 10.1007/s10959-019-00945-6, \url{https://link.springer.com/article/10.1007%2Fs10959-019-00945-6}.
\item Datta, B. and Gupta, S., \emph{Semi-regular tilings of the hyperbolic plane \label{bdsg_2019} }, to appear in Discrete Comput Geom, DOI https://doi.org/10.1007/s00454-019-00156-0, \url{https://doi.org/10.1007/s00454-019-00156-0}.
\item Datta, B. and Sarkar, S., \emph{Equilibrium triangulations of some quasitoric 4-manifolds}, Southeast Asian Bull. Math. {\bf 44} (2020), 57--78.
\item Datta, B. and Maity, D., \emph{Correction to: Semi-equivelar and vertex-transitive maps on the torus}, Beitr Algebra Geom {\bf 61} (2020), 187--188.
\item Martin Kreuzer and Dilip P. Patil , \emph{COMPUTATIONAL ASPECTS OF BURNSIDE RINGS PART II: IMPORTANT MAPS}, to appear in Beiträge zur Algebra und Geometrie / Contributions to Algebra and Geometry.
\item Bharali, Gautam and Maitra, Anwoy, \emph{A weak notion of visibility, a family of examples, and Wolff--Denjoy theorems}, to appear in Annali della Scuola Normale Superiore di Pisa, DOI 10.2422/2036-2145.201906\_007, \url{http://annaliscienze.sns.it/index.php?page=ForthcomingArticles}.
\item Maitra, Anwoy, \emph{On the continuous extension of Kobayashi isometries}, Proc. Amer. Math. Soc. {\bf 148} (2020), 3437--3451, \url{https://www.ams.org/journals/proc/2020-148-08/S0002-9939-2020-15038-1/}.
\item Vishnu Priya, N. and Senthilvelan, M. and  Rangarajan, G. and Lakshmanan, M, \emph{On symmetry preserving and symmetry broken bright, dark and antidark soliton solutions of nonlocal nonlinear Schrödinger equation}, Physics Letters A {\bf 383} (2019), 15-26.
\item Kashyap, G. A. R. S. R. K. and Bapat, D. and Das, D. and Gowaikar, R. D. and Amritkar, R. E. and Rangarajan, G. and Ravindranath, V. and Ambika, G, \emph{Synapse loss and progress of Alzheimer’s disease - A network model}, Scientific Reports {\bf 9} (2019), 6555.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan, G., \emph{On the role of four-wave mixing effect in the interactions between nonlinear modes of coupled generalized nonlinear Schrodinger equation}, Chaos {\bf 29} (2019), 123135.
\item Sarkar, Amar Deep and Verma, Kaushal, \emph{A submultiplicative property of the Carath\'{e}odory metric on planar domains}, Proc. Indian Acad. Sci. (Math. Sci.) {\bf } (), , \url{https://www.ias.ac.in/article/fulltext/pmsc/130/0035}.
\item Frauke Bleher, Ted Chinburg, Ralph Greenberg, Mahesh Kakde, George Pappas, Romyar Sharifi, Martin Taylor, \emph{Higher Chern Classes in Iwasawa Theory}, American Journal of Mathematics {\bf 142} (2020), 627-682, \url{https://muse.jhu.edu/issue/42062}.
\item Basu, R. and Ganguly, S. and Hegde, M. and Krishnapur, M., \emph{Lower Deviations in β-ensembles and Law of Iterated Logarithm in Last Passage Percolation}, to appear in Israel Journal of Mathematics.
\item M. K. Ghosh and S. Pradhan, \emph{Zero-Sum Risk-Sensitive Stochastic Differential Games with Reflecting Diffusions in the Orthant}, to appear in ESIAM: Control, Optimization and Calculus of Variations , DOI Doi:10.105/cocv/2020029.
\item M. K. Ghosh, K. Suresh Kumar, C. Pal and S. Pradhan, \emph{Nonzero-sum risk-sensitive stochastic differential games with discounted costs, Stochastic Analysis and Applications}, to appear in Stochastic Analysis and Applications, DOI Doi: 10.1080/07362994.2020.1796707.
\item Iyer, S. K. and Yogeshwaran, D., \emph{Thresholds for vanishing of isolated faces in random Cech and Vietoris-Rips Complexes}, Annals de l'Institut Henri Poincar$\acute{e}$ - Probabiliti$\acute{e}$s et Statistiques {\bf 56} (2020), 1869-1897, \url{https://doi.org/10.1214/19-AIHP1020}.
\item Datta, Basudeb and Gupta, Subhojoy, \emph{Semi-regular tilings of the hyperbolic plane}, to appear in Discrete and Computational Geometry, DOI https://doi.org/10.1007/s00454-019-00156-0.
\item Thirupathi Gudi, Ramesh Ch. Sau, \emph{Finite Element Analysis of the Constrained Dirichlet Boundary Control Problem Governed by the Diffusion Problem}, to appear in ESAIM: Control, Optim. Calc. Var, DOI doi: 10.1051/cocv/2019068..
\item Bhattacharyya, Tirthankar and Biswas, Anindya and Chandel, Vikramjeet Singh, \emph{On the Nevanlinna problem: characterization of all Schur–Agler class solutions affiliated with a given kernel},  Studia Math. {\bf 255} (2020), 83–107., DOI 10.4064/sm190505-8-10.
\item Bhattacharyya, Tirthankar and Das, Bata Krishna and Sau, Haripada, \emph{Toeplitz operators on the symmetrized bidisc}, International Mathematics Research Notices (IMRN) {\bf } (), , DOI 10.1093/imrn/rnz333.
\item Pingali, Vamsi Pritham and Varolin, Dror, \emph{Nonuniformly flat affine algebraic hypersurfaces}, to appear in Nagoya Mathematical Journal, DOI https://doi.org/10.1017/nmj.2019.2, \url{https://www.cambridge.org/core/journals/nagoya-mathematical-journal/article/nonuniformly-flat-affine-algebraic-hypersurfaces/3704C1133CB22A05D7FB13765156FC4E}.
\item Alvarez-Consul, Luis and Garcia-Fernandez, Mario  and García-Prada, Oscar and Pingali, Vamsi Pritham , \emph{Gravitating vortices and the Einstein–Bogomol’nyi equations}, to appear in Mathematische Annalen, DOI https://doi.org/10.1007/s00208-020-01964-z, \url{https://link.springer.com/content/pdf/10.1007/s00208-020-01964-z.pdf}.
\item Datar, Ved and Jacob,Adam and Zhang, Yuguang, \emph{Adiabatic limits of anti-self-dual connections on collapsed K3 surfaces}, to appear in Journal of differential geometry, \url{https://arxiv.org/pdf/1809.08583.pdf}.
\item Datar, Ved and Pingali, Vamsi Pritham, \emph{On coupled constant scalar curvature Kaehler metrics}, to appear in Journal of symplectic geometry, \url{https://arxiv.org/pdf/1901.10454v2.pdf}.
\item Pradhan, P. and Sahu, B., \emph{A characterization of the family of secant lines to a hyperbolic quadric in PG(3,q), q odd}, Discrete Mathematics {\bf 343} (2020), , DOI 10.1016/j.disc.2020.112044, \url{https://doi.org/10.1016/j.disc.2020.112044}.
\item Arun Maiti, \emph{Quasi-vertex-transitive maps on the plane}, Discrete Mathematics {\bf 343} (2020), 111911-111917, \url{https://www.sciencedirect.com/science/article/pii/S0012365X20301035}.
\item Datt, Gopal and Kumar, Sanjay, \emph{A criterion for quasinormality in $\mathbb{C}^n$},  Proc. Indian Acad. Sci. Math. Sci. {\bf 129} (2019), Paper No. 11, 13 pp., DOI https://doi.org/10.1007/s12044-018-0449-5     (This article was published before the review period i.e., April 1, 2019, BUT this is not yet included on the page  https://siddhartha-gadgil.github.io/DeptWeb/pubs.html), \url{https://link.springer.com/article/10.1007/s12044-018-0449-5?shared-article-renderer}.
\item Hatui, Sumana, \emph{Schur multipliers of special p-groups of rank 2}, Journal of Group Theory {\bf 23} (2020), 85-95, DOI https://doi.org/10.1515/jgth-2019-0045, \url{https://www.degruyter.com/view/journals/jgth/23/1/article-p85.xml}.
\item Choudhury, Projesh Nath and Tsatsomeros Michael J., \emph{Algorithmic Detection and Construction of N-matrices}, Linear Algebra and Its Applications {\bf 602} (2020), 46-56, DOI 10.1016/j.laa.2020.04.028, \url{https://doi.org/10.1016/j.laa.2020.04.028}.
\item Choudhury, Projesh Nath and Sivakumar, K.C., \emph{Matrices with Positive Semidefinite Real Part}, to appear in Linear and Multilinear Algebra, DOI 10.1080/03081087.2019.1602588, \url{https://doi.org/10.1080/03081087.2019.1602588}.
\item Chim, Kwok Chi; Shorey, Tarlok Nath; Sinha, Sneh Bala , \emph{On Baker's explicit abc-conjecture},  Publ. Math. Debrecen {\bf 94} (2019), 435–453.
\item Carstensen, C. and Mallik, G. and Nataraj, N., \emph{Nonconforming Finite Element Discretization for Semilinear Problems with Trilinear Nonlinearity}, IMA Journal of Numerical Analysis {\bf } (2020), , DOI https://doi.org/10.1093/imanum/drz071, \url{https://doi.org/10.1093/imanum/drz071}.
\item Mal, Arpita and Sain, Debmalya and Paul, Kallol, \emph{On some geometric properties of operator spaces}, Banach J. Math. Anal. {\bf 13} (2019), 174-191, \url{https://projecteuclid.org/euclid.bjma/1543914019}.
\end{enumerate}

\subsection{Papers in Conference proceedings}

\subsection{Books}

\subsection{Book Chapters}

\begin{enumerate}
\item Gupta, Subhojoy and Seshadri, Harish, \emph{Complex geometry of Teichmüller domains} in Papadopoulos, Athanase (ed.) (ed.), {\em Handbook of Teichmüller theory} , IRMA Lect. Math. Theor. Phys , Eur. Math. Soc., Zürich (2020), 26.
\item Gupta, Subhojoy, \emph{Holomorphic quadratic differentials in Teichmüller theory} in Papadopoulos, Athanase (ed.) (ed.), {\em Handbook of Teichmüller theory} , IRMA Lect. Math. Theor. Phys. , Eur. Math. Soc., Zürich (2020), 36.
\end{enumerate}

\subsection{Preprints}

\begin{enumerate}
\item Khare, Apoorva, \emph{Sharp nonzero lower bounds for the Schur product theorem}, submitted for publication, \url{https://arxiv.org/abs/1910.03537}.
\item Lübeck, Frank and Prasad, Dipendra and Ayyer, Arvind, \emph{A character relationship between symmetric group and hyperoctahedral group }, preprint, \url{https://arxiv.org/abs/1912.08576}.
\item Ayyer, Arvind and Sinha, Shubham , \emph{The size of t-cores and hook lengths of random cells in random partitions }, preprint, \url{https://arxiv.org/abs/1911.03135}.
\item Ayyer, Arvind, \emph{ A simple symmetric exclusion process driven by an asymmetric tracer particle }, preprint, \url{https://arxiv.org/abs/2001.02425}.
\item David Burns, Mahesh Kakde, \emph{On Weil-Etale Cohomology and Artin L-series}, preprint, \url{https://nms.kcl.ac.uk/david.burns/web_page/bk.pdf}.
\item Gupta, Purvi and Wawrzyniak, Chloe Urbanski, \emph{STABILITY OF THE HULL(S) OF AN $n$-SPHERE IN $\mathbb{C}^n$}, submitted for publication, \url{https://arxiv.org/pdf/2002.08699.pdf}.
\item Radhika Ganapathy, \emph{A Hecke algebra isomorphism over close local fields}, preprint.
\item Miermont, Grégory and Sen, Sanchayan , \emph{On breadth-first constructions of scaling limits of random graphs and random unicellular maps}, submitted for publication, \url{https://arxiv.org/abs/1908.04403}.
\item Gupta, Subhojoy and Mj, Mahan, \emph{Meromorphic projective structures, grafting and the monodromy map}, preprint, \url{http://arxiv.org/abs/1904.03804}.
\item Gupta, Subhojoy, \emph{Monodromy groups of $CP^1$-structures on punctured surfaces}, preprint, \url{https://arxiv.org/abs/1909.10771}.
\item Biswas, Indranil and Gupta, Subhojoy and Mj, Mahan and Whang, Junho Peter, \emph{Surface group representations in SL(2,C) with finite mapping class orbits}, preprint, \url{https://arxiv.org/abs/1707.00071}.
\item Ganguly, Pritam and Thangavelu,Sundaram, \emph{On lacunary spherical maximal function on Heisenberg groups}, preprint.
\item Bhattacharyya, Tirthankar and Kumar, Poornendu and Sau, Haripada, \emph{Distinguished Varieties Through the Berger--Coburn--Lebow Theorem}, submitted for publication, \url{https://arxiv.org/pdf/2001.01410.pdf}.
\item Bhattacharyya, T. and  Biswas, A. and Maitra, A., \emph{An unbounded realization of the symmetrized bidisc}, submitted for publication, \url{https://arxiv.org/pdf/2005.00289.pdf }.
\item Pingali, Vamsi Pritham, \emph{The deformed Hermitian Yang-Mills equation on three-folds}, preprint, \url{https://arxiv.org/abs/1910.01870}.
\item Garcia-Fernandez, Mario and Pingali, Vamsi Pritham and Yao, Chengjian , \emph{Gravitating vortices with positive curvature}, preprint, \url{https://arxiv.org/abs/1911.09616}.
\item Quddus, Safdar, \emph{Equivariant Cartan Homotopy Formula for $DG$-algebra.}, preprint, \url{https://arxiv.org/abs/2002.02192}.
\item Arun Maiti, \emph{OPTIMAL INDEX GROWTH OF CLOSED GEODESICS}, submitted for publication.
\item  Poria, Anirudha, \emph{Uncertainty Principles for the Fourier and the Short-Time Fourier Transforms}, submitted for publication, \url{https://arxiv.org/abs/2004.04184}.
\item Poria, Anirudha, \emph{Uncertainty principles for the Opdam-Cherednik transform on modulation spaces}, submitted for publication, \url{https://arxiv.org/abs/2005.14274}.
\item Bera, Sudip, \emph{ENUMERATION OF WEIGHTED PATHS ON A DIGRAPH AND BLOCK HOOK DETERMINANT}, submitted for publication.
\item Behera, Kiran Kumar, \emph{Self-inversive polynomials from rational transformation of linear functionals}, submitted for publication, \url{https://arxiv.org/abs/1909.12548}.
\item Hatui, Sumana and Singla Pooja, \emph{On Schur multiplier and Projective representations of Heisenberg groups}, submitted for publication, \url{https://arxiv.org/abs/1909.06589}.
\item Chattopahyay, Arup and Sarkar, Jaydeb and Sarkar, Srijan, \emph{Multiplicities, invariant subspaces and an additive formula.}, preprint, \url{https://arxiv.org/abs/1812.05435}.
\item Ghara, Soumitra and Kumar, Surjit and Pramanick, Paramita, \emph{$K$-homogeneous tuple of operators on bounded symmetric domains}, preprint.
\end{enumerate}

% from web page

\subsection{Publications (all kinds) from web page}

\begin{enumerate}
\item Aas, Erik and Ayyer, Arvind and Linusson, Svante and Potka,
Samu, {\em The exact phase diagram for a semipermeable {TASEP} with
nonlocal boundary jumps}, J. Phys. A {\bf 52} (2019), 355001, 19, DOI 10.1088/1751-8121/ab2e96, \url{https://doi.org/10.1088/1751-8121/ab2e96}.
\item Aiyappan, S. and Jose, Editha C. and Lomerio, Ivy Carol B. and
Nandakumaran, A. K., {\em Control problem on a rough circular domain and homogenization}, Asymptot. Anal. {\bf 115} (2019), 19--46, DOI 10.3233/asy-191526, \url{https://doi.org/10.3233/asy-191526}.
\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi, {\em Locally periodic unfolding operator for highly oscillating
rough domains}, Ann. Mat. Pura Appl. (4) {\bf 198} (2019), 1931--1954, DOI 10.1007/s10231-019-00848-7, \url{https://doi.org/10.1007/s10231-019-00848-7}.
\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi, {\em Generalization of unfolding operator for highly oscillating
smooth boundary domains and homogenization}, Calc. Var. Partial Differential Equations {\bf 57} (2018), Art. 86, 30, DOI 10.1007/s00526-018-1354-6, \url{https://doi.org/10.1007/s00526-018-1354-6}.
\item Aiyappan, S. and Nandakumaran, A. K. and Sufian, Abu, {\em Asymptotic analysis of a boundary optimal control problem on a
general branched structure}, Math. Methods Appl. Sci. {\bf 42} (2019), 6407--6434, DOI 10.1002/mma.5748, \url{https://doi.org/10.1002/mma.5748}.
\item Aiyappan, S. and Sardar, Bidhan Chandra, {\em Biharmonic equation in a highly oscillating domain and
homogenization of an associated control problem}, Appl. Anal. {\bf 98} (2019), 2783--2801, DOI 10.1080/00036811.2018.1471207, \url{https://doi.org/10.1080/00036811.2018.1471207}.
\item Akutagawa, Kazuo and Endo, Hisaaki and Seshadri, Harish, {\em A gap theorem for positive {E}instein metrics on the
four-sphere}, Math. Ann. {\bf 373} (2019), 1329--1339, DOI 10.1007/s00208-018-1749-x, \url{https://doi.org/10.1007/s00208-018-1749-x}.
\item Anamby, Pramath and Das, Soumya, {\em Distinguishing {H}ermitian cusp forms of degree 2 by a certain
subset of all {F}ourier coefficients}, Publ. Mat. {\bf 63} (2019), 307--341, DOI 10.5565/PUBLMAT6311911, \url{https://doi.org/10.5565/PUBLMAT6311911}.
\item Anamby, Pramath and Das, Soumya, {\em Bounds for the {P}etersson norms of the pullbacks of
{S}aito-{K}urokawa lifts}, J. Number Theory {\bf 191} (2018), 289--304, DOI 10.1016/j.jnt.2018.03.011, \url{https://doi.org/10.1016/j.jnt.2018.03.011}.
\item Arunkumar, G. and Kus, Deniz and Venkatesh, R., {\em Root multiplicities for {B}orcherds algebras and graph
coloring}, J. Algebra {\bf 499} (2018), 538--569, DOI 10.1016/j.jalgebra.2017.11.050, \url{https://doi.org/10.1016/j.jalgebra.2017.11.050}.
\item Ayyer, Arvind and Behrend, Roger E., {\em Factorization theorems for classical group characters, with
applications to alternating sign matrices and plane
partitions}, J. Combin. Theory Ser. A {\bf 165} (2019), 78--105, DOI 10.1016/j.jcta.2019.01.001, \url{https://doi.org/10.1016/j.jcta.2019.01.001}.
\item Ayyer, Arvind and Bouttier, J\'{e}r\'{e}mie and Corteel, Sylvie and
Linusson, Svante and Nunzi, Fran\c{c}ois, {\em Bumping sequences and multispecies juggling}, Adv. in Appl. Math. {\bf 98} (2018), 100--126, DOI 10.1016/j.aam.2018.03.001, \url{https://doi.org/10.1016/j.aam.2018.03.001}.
\item Ayyer, Arvind and Finn, Caley and Roy, Dipankar, {\em Matrix product solution of a left-permeable two-species
asymmetric exclusion process}, Phys. Rev. E {\bf 97} (2018), 012151, 10, DOI 10.1103/physreve.97.012151, \url{https://doi.org/10.1103/physreve.97.012151}.
\item Ayyer, Arvind and Finn, Caley and Roy, Dipankar, {\em The phase diagram for a multispecies left-permeable asymmetric
exclusion process}, J. Stat. Phys. {\bf 174} (2019), 605--621, DOI 10.1007/s10955-018-2183-x, \url{https://doi.org/10.1007/s10955-018-2183-x}.
\item Ayyer, Arvind and Linusson, Svante, {\em Reverse juggling processes}, Random Structures Algorithms {\bf 55} (2019), 56--72, DOI 10.1002/rsa.20825, \url{https://doi.org/10.1002/rsa.20825}.
\item Ayyer, Arvind and Prasad, Amritanshu and Spallone, Steven, {\em Odd partitions in {Y}oung's lattice}, S\'{e}m. Lothar. Combin. {\bf 75} (2018), Art. B75g, 13.
\item Ayyer, Arvind and Ramassamy, Sanjay, {\em The {H}ilbert-{G}alton board}, ALEA Lat. Am. J. Probab. Math. Stat. {\bf 15} (2018), 755--774, DOI 10.30757/alea.v15-28, \url{https://doi.org/10.30757/alea.v15-28}.
\item B\"{o}cherer, Siegfried and Das, Soumya, {\em Cuspidality and the growth of {F}ourier coefficients of
modular forms}, J. Reine Angew. Math. {\bf 741} (2018), 161--178, DOI 10.1515/crelle-2015-0075, \url{https://doi.org/10.1515/crelle-2015-0075}.
\item B\"{o}cherer, Siegfried and Das, Soumya, {\em Petersson norms of not necessarily cuspidal {J}acobi modular
forms and applications}, Adv. Math. {\bf 336} (2018), 335--376, DOI 10.1016/j.aim.2018.07.033, \url{https://doi.org/10.1016/j.aim.2018.07.033}.
\item Bagchi, Sayan and Thangavelu, Sundaram, {\em Weighted norm inequalities for {W}eyl multipliers and fourier
multipliers on the {H}eisenberg group}, J. Anal. Math. {\bf 136} (2018), 1--29, DOI 10.1007/s11854-018-0053-8, \url{https://doi.org/10.1007/s11854-018-0053-8}.
\item Baklouti, Ali and Thangavelu, Sundaram, {\em Hardy and {M}iyachi theorems for {H}eisenberg motion groups}, Nagoya Math. J. {\bf 229} (2018), 1--20, DOI 10.1017/nmj.2016.58, \url{https://doi.org/10.1017/nmj.2016.58}.
\item Balakumar, G. P. and Borah, Diganta and Mahajan, Prachi and
Verma, Kaushal, {\em Remarks on the higher dimensional {S}uita conjecture}, Proc. Amer. Math. Soc. {\bf 147} (2019), 3401--3411, DOI 10.1090/proc/14421, \url{https://doi.org/10.1090/proc/14421}.
\item Balhara, Rakesh, {\em Hardy's inequality for the fractional powers of the {G}rushin
operator}, Proc. Indian Acad. Sci. Math. Sci. {\bf 129} (2019), Art. 33, 25, DOI 10.1007/s12044-019-0471-2, \url{https://doi.org/10.1007/s12044-019-0471-2}.
\item Balodi, Mamta and Palcoux, Sebastien, {\em On {B}oolean intervals of finite groups}, J. Combin. Theory Ser. A {\bf 157} (2018), 49--69, DOI 10.1016/j.jcta.2018.02.004, \url{https://doi.org/10.1016/j.jcta.2018.02.004}.
\item Banerjee, Abhishek, {\em On {A}uslander's formula and cohereditary torsion pairs}, Commun. Contemp. Math. {\bf 20} (2018), 1750071, 27, DOI 10.1142/S0219199717500717, \url{https://doi.org/10.1142/S0219199717500717}.
\item Banerjee, Abhishek, {\em Modular {H}ecke algebras over {M}\"{o}bius categories}, J. Geom. Phys. {\bf 131} (2018), 23--40, DOI 10.1016/j.geomphys.2018.04.008, \url{https://doi.org/10.1016/j.geomphys.2018.04.008}.
\item Banerjee, Abhishek, {\em On differential torsion theories and rings with several
objects}, Canad. Math. Bull. {\bf 62} (2019), 703--714, DOI 10.4153/s0008439518000656, \url{https://doi.org/10.4153/s0008439518000656}.
\item Banerjee, Abhishek, {\em A topological {N}ullstellensatz for tensor-triangulated
categories}, C. R. Math. Acad. Sci. Paris {\bf 356} (2018), 365--375, DOI 10.1016/j.crma.2018.02.012, \url{https://doi.org/10.1016/j.crma.2018.02.012}.
\item Banerjee, Abhishek, {\em Quasimodular {H}ecke algebras and {H}opf actions}, J. Noncommut. Geom. {\bf 12} (2018), 1041--1080, DOI 10.4171/JNCG/297, \url{https://doi.org/10.4171/JNCG/297}.
\item Banerjee, Abhishek, {\em Completions of monoid objects and descent results}, J. Algebra {\bf 507} (2018), 362--399, DOI 10.1016/j.jalgebra.2018.04.022, \url{https://doi.org/10.1016/j.jalgebra.2018.04.022}.
\item Banerjee, Abhishek and Kour, Surjeet, {\em {$(A,\delta)$}-modules, {H}ochschild homology and higher
derivations}, Ann. Mat. Pura Appl. (4) {\bf 198} (2019), 1781--1802, DOI 10.1007/s10231-019-00844-x, \url{https://doi.org/10.1007/s10231-019-00844-x}.
\item Basu, Arnab and Ghosh, Mrinal K., {\em Nonzero-sum risk-sensitive stochastic games on a countable
state space}, Math. Oper. Res. {\bf 43} (2018), 516--532, DOI 10.1287/moor.2017.0870, \url{https://doi.org/10.1287/moor.2017.0870}.
\item Belton, Alexander and Guillot, Dominique and Khare, Apoorva
and Putinar, Mihai, {\em Simultaneous kernels of matrix {H}adamard powers}, Linear Algebra Appl. {\bf 576} (2019), 142--157, DOI 10.1016/j.laa.2018.03.035, \url{https://doi.org/10.1016/j.laa.2018.03.035}.
\item Benedetti, Carolina and Gonz\'{a}lez D'Le\'{o}n, Rafael S. and Hanusa,
Christopher R. H. and Harris, Pamela E. and Khare, Apoorva and
Morales, Alejandro H. and Yip, Martha, {\em The volume of the caracol polytope}, S\'{e}m. Lothar. Combin. {\bf 80B} (2018), Art. 87, 12.
\item Benedetti, Carolina and Gonz\'{a}lez D'Le\'{o}n, Rafael S. and Hanusa,
Christopher R. H. and Harris, Pamela E. and Khare, Apoorva and
Morales, Alejandro H. and Yip, Martha, {\em A combinatorial model for computing volumes of flow polytopes}, Trans. Amer. Math. Soc. {\bf 372} (2019), 3369--3404, DOI 10.1090/tran/7743, \url{https://doi.org/10.1090/tran/7743}.
\item Bera, Sayani and Verma, Kaushal, {\em Some aspects of shift-like automorphisms of {$\Bbb C^k$}}, Proc. Indian Acad. Sci. Math. Sci. {\bf 128} (2018), Art. 10, 48, DOI 10.1007/s12044-018-0388-1, \url{https://doi.org/10.1007/s12044-018-0388-1}.
\item Bharali, Gautam and Biswas, Indranil and Divakaran, Divakaran
and Janardhanan, Jaikrishnan, {\em Proper holomorphic mappings onto symmetric products of a
{R}iemann surface}, Doc. Math. {\bf 23} (2018), 1291--1311.
\item Bharali, Gautam and Chandel, Vikramjeet Singh, {\em Pick interpolation on the polydisc: small families of
sufficient kernels}, Complex Anal. Oper. Theory {\bf 13} (2019), 2069--2093, DOI 10.1007/s11785-017-0701-5, \url{https://doi.org/10.1007/s11785-017-0701-5}.
\item Bhattacharyya, Tirthankar and Rao, T. S. S. R. K., {\em Survey of last ten years of work done in {I}ndia in some
selected areas of functional analysis and operator theory}, Indian J. Pure Appl. Math. {\bf 50} (2019), 599--617, DOI 10.1007/s13226-019-0345-4, \url{https://doi.org/10.1007/s13226-019-0345-4}.
\item Bhattacharyya, Tirthankar and Sau, Haripada, {\em Holomorphic functions on the symmetrized bidisk---realization,
interpolation and extension}, J. Funct. Anal. {\bf 274} (2018), 504--524, DOI 10.1016/j.jfa.2017.09.013, \url{https://doi.org/10.1016/j.jfa.2017.09.013}.
\item Bhimani, Divyang G. and Balhara, Rakesh and Thangavelu,
Sundaram, {\em Hermite multipliers on modulation spaces}, Analysis and partial differential equations: perspectives from
developing countries {\bf 275} (2019), 42--64.
\item Bhosle, Usha N., {\em Hitchin pairs on reducible curves}, Internat. J. Math. {\bf 29} (2018), 1850015, 49, DOI 10.1142/S0129167X18500155, \url{https://doi.org/10.1142/S0129167X18500155}.
\item Biswas, Indranil and D'Mello, Shane and Mukherjee, Ritwik and
Pingali, Vamsi P., {\em Rational cuspidal curves on del-{P}ezzo surfaces}, J. Singul. {\bf 17} (2018), 91--107.
\item Biswas, Indranil and Dumitrescu, Sorin and Gupta, Subhojoy, {\em Branched projective structures on a {R}iemann surface and
logarithmic connections}, Doc. Math. {\bf 24} (2019), 2299--2337.
\item Biswas, Indranil and Pingali, Vamsi Pritham, {\em A characterization of finite vector bundles on {G}auduchon
astheno-{K}\"{a}hler manifolds}, \'{E}pijournal Geom. Alg\'{e}brique {\bf 2} (2018), Art. 6, 13.
\item Biswas, Rahul and Dond, Asha K. and Gudi, Thirupathi, {\em Edge patch-wise local projection stabilized nonconforming
{FEM} for the {O}seen problem}, Comput. Methods Appl. Math. {\bf 19} (2019), 189--214, DOI 10.1515/cmam-2018-0020, \url{https://doi.org/10.1515/cmam-2018-0020}.
\item Biswas, Shibananda and Ghosh, Gargi and Misra, Gadadhar and
Shyam Roy, Subrata, {\em On reducing submodules of {H}ilbert modules with
{${S}_n$}-invariant kernels}, J. Funct. Anal. {\bf 276} (2019), 751--784, DOI 10.1016/j.jfa.2018.10.025, \url{https://doi.org/10.1016/j.jfa.2018.10.025}.
\item Boggarapu, Pradeep and Roncal, Luz and Thangavelu, Sundaram, {\em On extension problem, trace {H}ardy and {H}ardy's inequalities
for some fractional {L}aplacians}, Commun. Pure Appl. Anal. {\bf 18} (2019), 2575--2605, DOI 10.3934/cpaa.2019116, \url{https://doi.org/10.3934/cpaa.2019116}.
\item Borah, Diganta and Haridas, Pranav and Verma, Kaushal, {\em Comments on the {G}reen's function of a planar domain}, Anal. Math. Phys. {\bf 8} (2018), 383--414, DOI 10.1007/s13324-017-0177-5, \url{https://doi.org/10.1007/s13324-017-0177-5}.
\item Brenner, Susanne C. and Gudi, Thirupathi and Porwal, Kamana
and Sung, Li-Yeng, {\em A {M}orley finite element method for an elliptic distributed
optimal control problem with pointwise state and control
constraints}, ESAIM Control Optim. Calc. Var. {\bf 24} (2018), 1181--1206, DOI 10.1051/cocv/2017031, \url{https://doi.org/10.1051/cocv/2017031}.
\item Burton, Benjamin A. and Datta, Basudeb and Singh, Nitin and
Spreer, Jonathan, {\em A construction principle for tight and minimal triangulations
of manifolds}, Exp. Math. {\bf 27} (2018), 22--36, DOI 10.1080/10586458.2016.1212747, \url{https://doi.org/10.1080/10586458.2016.1212747}.
\item Carstensen, Carsten and Dond, Asha K. and Rabus, Hella, {\em Quasi-optimality of adaptive mixed {FEM}s for non-selfadjoint
indefinite second-order linear elliptic problems}, Comput. Methods Appl. Math. {\bf 19} (2019), 233--250, DOI 10.1515/cmam-2019-0034, \url{https://doi.org/10.1515/cmam-2019-0034}.
\item Chandel, Vikramjeet Singh, {\em The three-point {P}ick-{N}evanlinna interpolation problem on
the polydisc}, Complex Var. Elliptic Equ. {\bf 63} (2018), 1341--1352, DOI 10.1080/17476933.2017.1370461, \url{https://doi.org/10.1080/17476933.2017.1370461}.
\item Chatterjee, Esha and Hassan, Sk. Sarif, {\em On the asymptotic character of a generalized rational
difference equation}, Discrete Contin. Dyn. Syst. {\bf 38} (2018), 1707--1718, DOI 10.3934/dcds.2018070, \url{https://doi.org/10.3934/dcds.2018070}.
\item Choudhury, Projesh Nath and Kannan, M. Rajesh and Sivakumar,
K. C., {\em A note on linear preservers of semipositive and minimally
semipositive matrices}, Electron. J. Linear Algebra {\bf 34} (2018), 687--694, DOI 10.13001/1081-3810.3864, \url{https://doi.org/10.13001/1081-3810.3864}.
\item Ciaurri, \'{O}scar and Roncal, Luz and Thangavelu, Sundaram, {\em Hardy-type inequalities for fractional powers of the
{D}unkl-{H}ermite operator}, Proc. Edinb. Math. Soc. (2) {\bf 61} (2018), 513--544, DOI 10.1017/s0013091517000311, \url{https://doi.org/10.1017/s0013091517000311}.
\item Das, Soumya and Jha, Abhash Kumar, {\em Analytic properties of twisted real-analytic {H}ermitian
{K}lingen type {E}isenstein series and applications}, Abh. Math. Semin. Univ. Hambg. {\bf 89} (2019), 105--116, DOI 10.1007/s12188-019-00206-7, \url{https://doi.org/10.1007/s12188-019-00206-7}.
\item Das, Soumya and Khan, Rizwanur, {\em The third moment of symmetric square {$L$}-functions}, Q. J. Math. {\bf 69} (2018), 1063--1087, DOI 10.1093/qmath/hay012, \url{https://doi.org/10.1093/qmath/hay012}.
\item Das, Soumya and Kohnen, Winfried, {\em On sign changes of eigenvalues of {S}iegel cusp forms of genus
2 in prime powers}, Acta Arith. {\bf 183} (2018), 167--172, DOI 10.4064/aa170419-4-11, \url{https://doi.org/10.4064/aa170419-4-11}.
\item Das, Soumya and Pal, Ritwik, {\em The first negative eigenvalue of {Y}oshida lifts}, Res. Number Theory {\bf 5} (2019), Paper No. 20, 9, DOI 10.1007/s40993-019-0158-x, \url{https://doi.org/10.1007/s40993-019-0158-x}.
\item Datta, Basudeb and Maity, Dipendu, {\em Semi-equivelar maps on the torus and the {K}lein bottle are
{A}rchimedean}, Discrete Math. {\bf 341} (2018), 3296--3309, DOI 10.1016/j.disc.2018.08.016, \url{https://doi.org/10.1016/j.disc.2018.08.016}.
\item Dond, Asha K. and Gudi, Thirupathi, {\em Patch-wise local projection stabilized finite element methods
for convection-diffusion-reaction problems}, Numer. Methods Partial Differential Equations {\bf 35} (2019), 638--663, DOI 10.1002/num.22317, \url{https://doi.org/10.1002/num.22317}.
\item Dond, Asha K. and Gudi, Thirupathi and Sau, Ramesh CH., {\em An error analysis of discontinuous finite element methods for
the optimal control problems governed by {S}tokes equation}, Numer. Funct. Anal. Optim. {\bf 40} (2019), 421--460, DOI 10.1080/01630563.2018.1538158, \url{https://doi.org/10.1080/01630563.2018.1538158}.
\item Fritz, Tobias and Gadgil, Siddhartha and Khare, Apoorva and Nielsen, Pace P. and Silberman, Lior and Tao, Terence, {\em Homogeneous length functions on groups}, Algebra Number Theory {\bf 12} (2018), 1773--1786, DOI 10.2140/ant.2018.12.1773, \url{https://doi.org/10.2140/ant.2018.12.1773}.
\item Gaddam, Sharat and Gudi, Thirupathi, {\em Inhomogeneous {D}irichlet boundary condition in the {\it a
posteriori} error control of the obstacle problem}, Comput. Math. Appl. {\bf 75} (2018), 2311--2327, DOI 10.1016/j.camwa.2017.12.010, \url{https://doi.org/10.1016/j.camwa.2017.12.010}.
\item Gaddam, Sharat and Gudi, Thirupathi, {\em Bubbles enriched quadratic finite element method for the
3{D}-elliptic obstacle problem}, Comput. Methods Appl. Math. {\bf 18} (2018), 223--236, DOI 10.1515/cmam-2017-0018, \url{https://doi.org/10.1515/cmam-2017-0018}.
\item Garani, Shayan Srinivasa and Seshadri, Harish, {\em An algorithmic approach to {S}outh {I}ndian classical music}, J. Math. Music {\bf 13} (2019), 107--134, DOI 10.1080/17459737.2019.1604845, \url{https://doi.org/10.1080/17459737.2019.1604845}.
\item Gaussier, Herv\'{e} and Seshadri, Harish, {\em Holomorphic motions and complex geometry}, Proc. Amer. Math. Soc. {\bf 147} (2019), 301--313, DOI 10.1090/proc/14217, \url{https://doi.org/10.1090/proc/14217}.
\item Gaussier, Herv\'{e} and Seshadri, Harish, {\em On the {G}romov hyperbolicity of convex domains in {$\Bbb
C^n$}}, Comput. Methods Funct. Theory {\bf 18} (2018), 617--641, DOI 10.1007/s40315-018-0243-5, \url{https://doi.org/10.1007/s40315-018-0243-5}.
\item Ghara, Soumitra and Kumar, Surjit, {\em On the sum of two subnormal kernels}, J. Math. Anal. Appl. {\bf 469} (2019), 1015--1027, DOI 10.1016/j.jmaa.2018.09.048, \url{https://doi.org/10.1016/j.jmaa.2018.09.048}.
\item Ghosh, Mrinal K. and Pradhan, Somnath, {\em Risk-sensitive stochastic differential games with reflecting
diffusions}, Stoch. Anal. Appl. {\bf 36} (2018), 1--27, DOI 10.1080/07362994.2017.1356732, \url{https://doi.org/10.1080/07362994.2017.1356732}.
\item Gudi, Thirupathi and Majumder, Papri, {\em Convergence analysis of finite element method for a parabolic
obstacle problem}, J. Comput. Appl. Math. {\bf 357} (2019), 85--102, DOI 10.1016/j.cam.2019.02.026, \url{https://doi.org/10.1016/j.cam.2019.02.026}.
\item Gudi, Thirupathi and Majumder, Papri, {\em Conforming and discontinuous {G}alerkin {FEM} in space for
solving parabolic obstacle problem}, Comput. Math. Appl. {\bf 78} (2019), 3896--3915, DOI 10.1016/j.camwa.2019.06.022, \url{https://doi.org/10.1016/j.camwa.2019.06.022}.
\item Gupta, Rajeev and Kumar, Surjit and Trivedi, Shailesh, {\em Von {N}eumann's inequality for commuting operator-valued
multishifts}, Proc. Amer. Math. Soc. {\bf 147} (2019), 2599--2608, DOI 10.1090/proc/14410, \url{https://doi.org/10.1090/proc/14410}.
\item Gupta, Rajeev and Ramiz Reza, Md., {\em Operator space structures on {$\ell^1(n)$}}, Houston J. Math. {\bf 44} (2018), 1205--1212.
\item Gupta, S. C., {\em The classical {S}tefan problem},  {\bf } (2018), xxiii+726.
\item Gupta, Subhojoy, {\em Limits of harmonic maps and crowned hyperbolic surfaces}, Trans. Amer. Math. Soc. {\bf 372} (2019), 7573--7596, DOI 10.1090/tran/7777, \url{https://doi.org/10.1090/tran/7777}.
\item Gupta, Subhojoy and Wolf, Michael, {\em Meromorphic quadratic differentials and measured foliations on
a {R}iemann surface}, Math. Ann. {\bf 373} (2019), 73--118, DOI 10.1007/s00208-018-1674-z, \url{https://doi.org/10.1007/s00208-018-1674-z}.
\item Hazra, Somnath, {\em Homogeneous 2-shifts}, Complex Anal. Oper. Theory {\bf 13} (2019), 1729--1763, DOI 10.1007/s11785-018-0822-5, \url{https://doi.org/10.1007/s11785-018-0822-5}.
\item Iyer, Srikanth K., {\em The random connection model: connectivity, edge lengths, and
degree distributions}, Random Structures Algorithms {\bf 52} (2018), 283--300, DOI 10.1002/rsa.20741, \url{https://doi.org/10.1002/rsa.20741}.
\item Jha, Abhash Kumar and Juyal, Abhishek and Pandey, Manish
Kumar, {\em On simultaneous non-vanishing of twisted {$L$}-functions
associated to newforms on {$\Gamma_0(N)$}}, J. Ramanujan Math. Soc. {\bf 34} (2019), 245--252.
\item Jha, Abhash Kumar and Sahu, Brundaban, {\em Rankin-{C}ohen brackets on {J}acobi forms of several variables
and special values of certain {D}irichlet series}, Int. J. Number Theory {\bf 15} (2019), 925--933, DOI 10.1142/S1793042119500490, \url{https://doi.org/10.1142/S1793042119500490}.
\item Jha, Abhash Kumar and Sahu, Brundaban, {\em Differential operators on {J}acobi forms and special values of
certain {D}irichlet series}, Automorphic forms and related topics {\bf 732} (2019), 91--99, DOI 10.1090/conm/732/14793, \url{https://doi.org/10.1090/conm/732/14793}.
\item Kanukurthi, Bhavana and Bhavana Obbattu, Sai Lakshmi and
Sekar, Sruthi, {\em Non-malleable randomness encoders and their applications}, Advances in cryptology---{EUROCRYPT} 2018. {P}art {III} {\bf 10822} (2018), 589--617.
\item Kashyap, G. A. R. S. R. K. and Bapat, D. and Das, D. and Gowaikar, R. D. and  Amritkar, R. E. and Rangarajan, G. and V. Ravindranath and G. Ambika, {\em Synapse loss and progress of Alzheimers disease - A network model}, Scientific Reports {\bf 9} (2019), .
\item Kataria, Kuldeep Kumar and Vellaisamy, Palaniappan, {\em On the convolution of {M}ittag-{L}effler distributions and its
applications to fractional point processes}, Stoch. Anal. Appl. {\bf 37} (2019), 115--122, DOI 10.1080/07362994.2018.1538803, \url{https://doi.org/10.1080/07362994.2018.1538803}.
\item Khan, Mohammad Yasir Feroz and Biswas, Rahul and Sau, Ramesh
Chandra, {\em Instability of {L}agrange interpolation on {$H^1(\Omega)\cap
C(\overline\Omega)$}}, Math. Student {\bf 88} (2019), 153--158.
\item Khare, Apoorva, {\em Generalized nil-{C}oxeter algebras over discrete complex
reflection groups}, Trans. Amer. Math. Soc. {\bf 370} (2018), 2971--2999, DOI 10.1090/tran/7304, \url{https://doi.org/10.1090/tran/7304}.
\item Khare, Apoorva and Tikaradze, Akaki, {\em A {C}arlitz--von {S}taudt type theorem for finite rings}, Linear Algebra Appl. {\bf 568} (2019), 106--126, DOI 10.1016/j.laa.2018.05.035, \url{https://doi.org/10.1016/j.laa.2018.05.035}.
\item Krakovski, Roi and Onn, Uri and Singla, Pooja, {\em Regular characters of groups of type {$\ssf A_n$} over
discrete valuation rings}, J. Algebra {\bf 496} (2018), 116--137, DOI 10.1016/j.jalgebra.2017.10.018, \url{https://doi.org/10.1016/j.jalgebra.2017.10.018}.
\item Lin, X. and Kulkarni, P. and Bocci, F. and Schafer, N. and Roy, S. and Tsai, M-Y. and He, Y. and Chen, Y and Rajagopalan, K. and Mooney, S. and Zeng, Y. and Weninger, K. and Grishaev, A. and Onuchic, J. and Levine, H. and Wolynes, P. and Salgia, R. and Rangarajan, G. and Uversky, V. and Orban, J. and Jolly, M. K., {\em Structural and Dynamical Order of a Disordered Protein: Molecular Insights into Conformational Switching of PAGE4 at the Systems Level}, Biomolecules {\bf 9} (2019), 77.
\item Mahajan, Prachi and Verma, Kaushal, {\em A comparison of two biholomorphic invariants}, Internat. J. Math. {\bf 30} (2019), 1950012, 16, DOI 10.1142/S0129167X19500125, \url{https://doi.org/10.1142/S0129167X19500125}.
\item Maity, Dipendu, {\em Pentagonal maps on the torus and the plane}, Beitr. Algebra Geom. {\bf 60} (2019), 17--37, DOI 10.1007/s13366-018-0405-7, \url{https://doi.org/10.1007/s13366-018-0405-7}.
\item Maity, Dipendu and Upadhyay, Ashish Kumar, {\em On centrally symmetric manifolds}, J. Ramanujan Math. Soc. {\bf 34} (2019), 21--27.
\item Mal, Arpita and Paul, Kallol and Rao, T. S. S. R. K. and Sain,
Debmalya, {\em Approximate {B}irkhoff-{J}ames orthogonality and smoothness in
the space of bounded linear operators}, Monatsh. Math. {\bf 190} (2019), 549--558, DOI 10.1007/s00605-019-01289-3, \url{https://doi.org/10.1007/s00605-019-01289-3}.
\item Mal, Arpita and Sain, Debmalya and Paul, Kallol, {\em On some geometric properties of operator spaces}, Banach J. Math. Anal. {\bf 13} (2019), 174--191, DOI 10.1215/17358787-2018-0021, \url{https://doi.org/10.1215/17358787-2018-0021}.
\item Manfio, F. and Turgay, N. C. and Upadhyay, A., {\em Biconservative submanifolds in {$\Bbb{S}^n\times\Bbb{R}$} and
{$\Bbb{H}^n\times\Bbb{R}$}}, J. Geom. Anal. {\bf 29} (2019), 283--298, DOI 10.1007/s12220-018-9990-9, \url{https://doi.org/10.1007/s12220-018-9990-9}.
\item Manikandan, K. and Vishnu Priya, N. and Senthilvelan, M. and
Sankaranarayanan, R., {\em Deformation of dark solitons in a {$\Cal{PT}$}-invariant
variable coefficients nonlocal nonlinear {S}chr\"{o}dinger
equation}, Chaos {\bf 28} (2018), 083103, 12, DOI 10.1063/1.5039901, \url{https://doi.org/10.1063/1.5039901}.
\item Misra, Gadadhar, {\em Operators in the {C}owen-{D}ouglas class and related topics}, Handbook of analytic operator theory {\bf } (2019), 87--137.
\item Misra, Gadadhar and Pal, Avijit, {\em Curvature inequalities for operators in the {C}owen-{D}ouglas
class and localization of the {W}allach set}, J. Anal. Math. {\bf 136} (2018), 31--54, DOI 10.1007/s11854-018-0054-7, \url{https://doi.org/10.1007/s11854-018-0054-7}.
\item Misra, Gadadhar and Pal, Avijit and Varughese, Cherian, {\em Contractivity and complete contractivity for finite
dimensional {B}anach spaces}, J. Operator Theory {\bf 82} (2019), 23--47, DOI 10.7900/jot.2018jun13.2225, \url{https://doi.org/10.7900/jot.2018jun13.2225}.
\item Misra, Gadadhar and Reza, Md. Ramiz, {\em Curvature inequalities and extremal operators}, Illinois J. Math. {\bf 63} (2019), 193--217, DOI 10.1215/00192082-7768711, \url{https://doi.org/10.1215/00192082-7768711}.
\item Mukherjee, Sajal Kumar and Bera, Sudip, {\em Combinatorial proofs of the {N}ewton-{G}irard and
{C}hapman-{C}ostas-{S}antos identities}, Discrete Math. {\bf 342} (2019), 1577--1580, DOI 10.1016/j.disc.2019.02.013, \url{https://doi.org/10.1016/j.disc.2019.02.013}.
\item Narayanan, E. K. and Singla, Pooja, {\em On monomial representations of finitely generated nilpotent
groups}, Comm. Algebra {\bf 46} (2018), 2319--2331, DOI 10.1080/00927872.2017.1378894, \url{https://doi.org/10.1080/00927872.2017.1378894}.
\item Pal, Chandan and Pradhan, Somnath, {\em Risk sensitive control of pure jump processes on a general
state space}, Stochastics {\bf 91} (2019), 155--174, DOI 10.1080/17442508.2018.1521413, \url{https://doi.org/10.1080/17442508.2018.1521413}.
\item Pal, Ratna and Verma, Kaushal, {\em Ergodic properties of families of {H}\'{e}non maps}, Ann. Polon. Math. {\bf 121} (2018), 45--71, DOI 10.4064/ap170928-2-4, \url{https://doi.org/10.4064/ap170928-2-4}.
\item Paul, Kallol and Sain, Debmalya, {\em Birkhoff-{J}ames orthogonality and its application in the
study of geometry of {B}anach space}, Advanced topics in mathematical analysis {\bf } (2019), 245--284.
\item Paul, Kallol and Sain, Debmalya and Ghosh, Puja, {\em Symmetry of {B}irkhoff-{J}ames orthogonality of bounded linear
operators}, Ulam type stability {\bf } (2019), 331--344.
\item Paul, Kallol and Sain, Debmalya and Mal, Arpita, {\em Approximate {B}irkhoff-{J}ames orthogonality in the space of
bounded linear operators}, Linear Algebra Appl. {\bf 537} (2018), 348--357, DOI 10.1016/j.laa.2017.10.008, \url{https://doi.org/10.1016/j.laa.2017.10.008}.
\item Paul, Kallol and Sain, Debmalya and Mal, Arpita and Mandal,
Kalidas, {\em Orthogonality of bounded linear operators on complex {B}anach
spaces}, Adv. Oper. Theory {\bf 3} (2018), 699--709, DOI 10.15352/aot.1712-1268, \url{https://doi.org/10.15352/aot.1712-1268}.
\item Pingali, Vamsi P., {\em A note on the deformed {H}ermitian {Y}ang-{M}ills {PDE}}, Complex Var. Elliptic Equ. {\bf 64} (2019), 503--518, DOI 10.1080/17476933.2018.1454914, \url{https://doi.org/10.1080/17476933.2018.1454914}.
\item Pingali, Vamsi Pritham, {\em A note on higher extremal metrics}, Trans. Amer. Math. Soc. {\bf 370} (2018), 6995--7010, DOI 10.1090/tran/7416, \url{https://doi.org/10.1090/tran/7416}.
\item Pingali, Vamsi Pritham, {\em Representability of {C}hern-{W}eil forms}, Math. Z. {\bf 288} (2018), 629--641, DOI 10.1007/s00209-017-1903-2, \url{https://doi.org/10.1007/s00209-017-1903-2}.
\item Pritham Pingali, Vamsi, {\em Existence of coupled {K}\"{a}hler-{E}instein metrics using the
continuity method}, Internat. J. Math. {\bf 29} (2018), 1850041, 8, DOI 10.1142/S0129167X18500416, \url{https://doi.org/10.1142/S0129167X18500416}.
\item Quddus, Safdar, {\em Invariants of the {$\Bbb Z_2$} orbifolds of the {P}odle\'{s} two
spheres}, J. Noncommut. Geom. {\bf 13} (2019), 257--267, DOI 10.4171/JNCG/320, \url{https://doi.org/10.4171/JNCG/320}.
\item Reddy, Nanda Kishore, {\em Equality of {L}yapunov and stability exponents for products of
isotropic random matrices}, Int. Math. Res. Not. IMRN {\bf } (2019), 606--624, DOI 10.1093/imrn/rnx134, \url{https://doi.org/10.1093/imrn/rnx134}.
\item Reza, Md. Ramiz, {\em Curvature inequalities for operators in the {C}owen-{D}ouglas
class of a planar domain}, Indiana Univ. Math. J. {\bf 67} (2018), 1255--1279, DOI 10.1512/iumj.2018.67.7320, \url{https://doi.org/10.1512/iumj.2018.67.7320}.
\item Roth, Julien and Upadhyay, Abhitosh, {\em On compact anisotropic {W}eingarten hypersurfaces in
{E}uclidean space}, Arch. Math. (Basel) {\bf 113} (2019), 213--224, DOI 10.1007/s00013-019-01315-8, \url{https://doi.org/10.1007/s00013-019-01315-8}.
\item Roy, Krishanu and Venkatesh, R., {\em Maximal closed subroot systems of real affine root systems}, Transform. Groups {\bf 24} (2019), 1261--1308, DOI 10.1007/s00031-018-9510-9, \url{https://doi.org/10.1007/s00031-018-9510-9}.
\item Sain, Debmalya, {\em On extreme contractions and the norm attainment set of a
bounded linear operator}, Ann. Funct. Anal. {\bf 10} (2019), 135--143, DOI 10.1215/20088752-2018-0014, \url{https://doi.org/10.1215/20088752-2018-0014}.
\item Sain, Debmalya, {\em On the norm attainment set of a bounded linear operator}, J. Math. Anal. Appl. {\bf 457} (2018), 67--76, DOI 10.1016/j.jmaa.2017.07.070, \url{https://doi.org/10.1016/j.jmaa.2017.07.070}.
\item Sain, Debmalya, {\em Smooth points in operator spaces and some
{B}ishop-{P}helps-{B}ollob\'{a}s type theorems in {B}anach spaces}, Oper. Matrices {\bf 13} (2019), 433--445, DOI 10.7153/oam-2019-13-32, \url{https://doi.org/10.7153/oam-2019-13-32}.
\item Sain, Debmalya and Paul, Kallol and Bhunia, Pintu and Bag,
Santanu, {\em On the numerical index of polyhedral {B}anach spaces}, Linear Algebra Appl. {\bf 577} (2019), 121--133, DOI 10.1016/j.laa.2019.04.024, \url{https://doi.org/10.1016/j.laa.2019.04.024}.
\item Sain, Debmalya and Paul, Kallol and Mal, Arpita, {\em A complete characterization of {B}irkhoff-{J}ames
orthogonality in infinite dimensional normed space}, J. Operator Theory {\bf 80} (2018), 399--413.
\item Sain, Debmalya and Paul, Kallol and Mal, Arpita, {\em On approximate {B}irkhoff-{J}ames orthogonality and normal
cones in a normed space}, J. Convex Anal. {\bf 26} (2019), 341--351.
\item Sain, Debmalya and Paul, Kallol and Mandal, Kalidas, {\em On two extremum problems related to the norm of a bounded
linear operator}, Oper. Matrices {\bf 13} (2019), 421--432, DOI 10.7153/oam-2019-13-31, \url{https://doi.org/10.7153/oam-2019-13-31}.
\item Sain, Debmalya and Ray, Anubhab and Paul, Kallol, {\em Extreme contractions on finite-dimensional polygonal {B}anach
spaces}, J. Convex Anal. {\bf 26} (2019), 877--885.
\item Sanki, Bidyut, {\em Filling of closed surfaces}, J. Topol. Anal. {\bf 10} (2018), 897--913, DOI 10.1142/S1793525318500309, \url{https://doi.org/10.1142/S1793525318500309}.
\item Sanki, Bidyut and Gadgil, Siddhartha, {\em Graphs of systoles on hyperbolic surfaces}, J. Topol. Anal. {\bf 11} (2019), 1--20, DOI 10.1142/S1793525319500018, \url{https://doi.org/10.1142/S1793525319500018}.
\item Seshadri, Harish, {\em Differential geometry in {I}ndia}, Indian J. Pure Appl. Math. {\bf 50} (2019), 795--799, DOI 10.1007/s13226-019-0355-2, \url{https://doi.org/10.1007/s13226-019-0355-2}.
\item Seshadri, Harish and Verma, Kaushal, {\em Some aspects of the automorphism groups of domains}, Handbook of group actions. {V}ol. {III} {\bf 40} (2018), 145--174.
\item Thangavelu, Sundaram, {\em An analogue of {P}fannschmidt's theorem for the {H}eisenberg
group}, J. Anal. {\bf 26} (2018), 235--244, DOI 10.1007/s41478-018-0147-9, \url{https://doi.org/10.1007/s41478-018-0147-9}.
\item Thangavelu, Sundaram and Naidu Dogga, Venku, {\em {$L^p$}-{$L^2$} estimates for solutions of the wave equation
associated to the {G}rushin operator}, Adv. Pure Appl. Math. {\bf 9} (2018), 85--92, DOI 10.1515/apam-2017-0042, \url{https://doi.org/10.1515/apam-2017-0042}.
\item Turgay, Nurettin Cenk and Upadhyay, Abhitosh, {\em On biconservative hypersurfaces in 4-dimensional {R}iemannian
space forms}, Math. Nachr. {\bf 292} (2019), 905--921, DOI 10.1002/mana.201700328, \url{https://doi.org/10.1002/mana.201700328}.
\item Urbina-Romero, Wilfredo, {\em Gaussian harmonic analysis},  {\bf } (2019), xix+477, DOI 10.1007/978-3-030-05597-4, \url{https://doi.org/10.1007/978-3-030-05597-4}.
\item Verma, Kaushal, {\em Notes on the boundaries of quadrature domains}, Anal. Math. Phys. {\bf 9} (2019), 617--638, DOI 10.1007/s13324-018-0221-0, \url{https://doi.org/10.1007/s13324-018-0221-0}.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan,
Govindan, {\em On the role of four-wave mixing effect in the interactions
between nonlinear modes of coupled generalized nonlinear
{S}chr\"{o}dinger equation}, Chaos {\bf 29} (2019), 123135, 14, DOI 10.1063/1.5121245, \url{https://doi.org/10.1063/1.5121245}.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan, Govindan
and Lakshmanan, M., {\em On symmetry preserving and symmetry broken bright, dark and
antidark soliton solutions of nonlocal nonlinear {S}chr\"{o}dinger
equation}, Phys. Lett. A {\bf 383} (2019), 15--26, DOI 10.1016/j.physleta.2018.10.011, \url{https://doi.org/10.1016/j.physleta.2018.10.011}.
\end{enumerate}