### 2020-08-10 data/draft-publications.tex
hyperlinks with title

\subsection{Papers in Journals}

\begin{enumerate}
\item S Aiyappan, A. K. Nandakumaran and Abu Sufian: Asymptotic analysis of Boundary Optimal Control Problem on a General Branched Structure, \emph{Mathematical Methods in the Applied Sciences} {\bf 42} (2019. DOI: 10.1002/mma.5748), 6407-6434.
\item K. Murali, A. K. Nandakumaran, Turgurt Durduran, Hari M Varma: Recovery of the diffuse correlation spectroscopy data-type from speckle contrast measurements: towards low-cost, deep-tissue blood flow measurements, \emph{Biomedical Optics Express 5395} {\bf 10} (2019).
\item S Aiyappan, A. K. Nandakumaran and Ravi Prakash: Locally Periodic Unfolding Operator for Highly Oscillating Rough Domains, \emph{Annali di Matematica Pura ed Applicata} {\bf 198} (2019), 1931-1954.
\item S Aiyappan, A. K. Nandakumaran and Ravi Prakash: Semi-linear optimal control problem on a smooth oscillating domain, \emph{Communications in Contemporary Mathematics} {\bf } (2019), 1-26.
\item S Aiyappan, Editha C. Jose, Ivy Carol B. Lomerio and A. K. Nandakumaran: Control Problem on a Rough Circular Domain and Homogenization, \emph{Asymptotic Analysis} {\bf 115} (2019), 19-46.
\item M. Balodi, A. Banerjee, S. Ray: Cohomology of modules over H-categories and co-H-categories, to appear in \emph{Canadian Journal of Mathematics}.
\item M. Balodi, A. Banerjee, S. Ray: Entwined modules over linear categories and Galois extensions, to appear in \emph{Israel Journal of Mathematics (Accepted)}.
\item A. Banerjee: On Noetherian schemes over $(\mathcal C,\otimes,1)$ and the category of quasi-coherent sheaves, to appear in \emph{Indiana University Mathematics Journal}.
\item A. Banerjee: On differential torsion theories and rings with several objects, \emph{Canadian Math Bulletin} {\bf 62} (2019), 703--714.
\item Khare, Apoorva and Rajaratnam, Bala: Probability inequalities and tail estimates for metric semigroups, \emph{Advances in Operator Theory (Special issue in honor of Rajendra Bhatia)} {\bf 5} (2020), 779-795.
\item Ghosh, Subhajit: Total Variation Cutoff for the Transpose Top-2 with Random Shuffle, to appear in \emph{Journal of Theoretical Probability}.
\item Datta, B. and Gupta, S.: Semi-regular tilings of the hyperbolic plane \label{bdsg_2019}, to appear in \emph{Discrete Comput Geom}.
\item Datta, B. and Sarkar, S.: Equilibrium triangulations of some quasitoric 4-manifolds, \emph{Southeast Asian Bull. Math.} {\bf 44} (2020), 57--78.
\item Datta, B. and Maity, D.: Correction to: Semi-equivelar and vertex-transitive maps on the torus, \emph{Beitr Algebra Geom} {\bf 61} (2020), 187--188.
\item Martin Kreuzer and Dilip P. Patil: COMPUTATIONAL ASPECTS OF BURNSIDE RINGS PART II: IMPORTANT MAPS, to appear in \emph{Beiträge zur Algebra und Geometrie / Contributions to Algebra and Geometry}.
\item Gadadhar Misra: Professor, \emph{Adv. Math.} {\bf 351} (2019), 1105 - 1138.
\item Bharali, Gautam and Maitra, Anwoy: A weak notion of visibility, a family of examples, and Wolff--Denjoy theorems, to appear in \emph{Annali della Scuola Normale Superiore di Pisa}.
\item Maitra, Anwoy: On the continuous extension of Kobayashi isometries, \emph{Proc. Amer. Math. Soc.} {\bf 148} (2020), 3437--3451.
\item Vishnu Priya, N. and Senthilvelan, M. and  Rangarajan, G. and Lakshmanan, M: On symmetry preserving and symmetry broken bright, dark and antidark soliton solutions of nonlocal nonlinear Schrödinger equation, \emph{Physics Letters A} {\bf 383} (2019), 15-26.
\item Kashyap, G. A. R. S. R. K. and Bapat, D. and Das, D. and Gowaikar, R. D. and Amritkar, R. E. and Rangarajan, G. and Ravindranath, V. and Ambika, G: Synapse loss and progress of Alzheimer’s disease - A network model, \emph{Scientific Reports} {\bf 9} (2019), 6555.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan, G.: On the role of four-wave mixing effect in the interactions between nonlinear modes of coupled generalized nonlinear Schrodinger equation, \emph{Chaos} {\bf 29} (2019), 123135.
\item Sarkar, Amar Deep and Verma, Kaushal: A submultiplicative property of the Carath\'{e}odory metric on planar domains, \emph{Proc. Indian Acad. Sci. (Math. Sci.)} {\bf } ().
\item Frauke Bleher, Ted Chinburg, Ralph Greenberg, Mahesh Kakde, George Pappas, Romyar Sharifi, Martin Taylor: Higher Chern Classes in Iwasawa Theory, \emph{American Journal of Mathematics} {\bf 142} (2020), 627-682.
\item Basu, R. and Ganguly, S. and Hegde, M. and Krishnapur, M.: Lower Deviations in β-ensembles and Law of Iterated Logarithm in Last Passage Percolation, to appear in \emph{Israel Journal of Mathematics}.
\item M. K. Ghosh and S. Pradhan: Zero-Sum Risk-Sensitive Stochastic Differential Games with Reflecting Diffusions in the Orthant, to appear in \emph{ESIAM: Control, Optimization and Calculus of Variations}.
\item M. K. Ghosh, K. Suresh Kumar, C. Pal and S. Pradhan: Nonzero-sum risk-sensitive stochastic differential games with discounted costs, Stochastic Analysis and Applications, to appear in \emph{Stochastic Analysis and Applications}.
\item Iyer, S. K. and Yogeshwaran, D.: Thresholds for vanishing of isolated faces in random Cech and Vietoris-Rips Complexes, \emph{Annals de l'Institut Henri Poincar$\acute{e}$ - Probabiliti$\acute{e}$s et Statistiques} {\bf 56} (2020), 1869-1897.
\item Datta, Basudeb and Gupta, Subhojoy: Semi-regular tilings of the hyperbolic plane, to appear in \emph{Discrete and Computational Geometry}.
\item Thirupathi Gudi, Ramesh Ch. Sau: Finite Element Analysis of the Constrained Dirichlet Boundary Control Problem Governed by the Diffusion Problem, to appear in \emph{ESAIM: Control, Optim. Calc. Var}.
\item Bhattacharyya, Tirthankar and Biswas, Anindya and Chandel, Vikramjeet Singh: On the Nevanlinna problem: characterization of all Schur–Agler class solutions affiliated with a given kernel, \emph{Studia Math.} {\bf 255} (2020), 83–107..
\item Bhattacharyya, Tirthankar and Das, Bata Krishna and Sau, Haripada: Toeplitz operators on the symmetrized bidisc, \emph{International Mathematics Research Notices (IMRN)} {\bf } ().
\item Pingali, Vamsi Pritham and Varolin, Dror: Nonuniformly flat affine algebraic hypersurfaces, to appear in \emph{Nagoya Mathematical Journal}.
\item Alvarez-Consul, Luis and Garcia-Fernandez, Mario  and García-Prada, Oscar and Pingali, Vamsi Pritham: Gravitating vortices and the Einstein–Bogomol’nyi equations, to appear in \emph{Mathematische Annalen}.
\item Datar, Ved and Jacob,Adam and Zhang, Yuguang: Adiabatic limits of anti-self-dual connections on collapsed K3 surfaces, to appear in \emph{Journal of differential geometry}.
\item Datar, Ved and Pingali, Vamsi Pritham: On coupled constant scalar curvature Kaehler metrics, to appear in \emph{Journal of symplectic geometry}.
\item Pradhan, P. and Sahu, B.: A characterization of the family of secant lines to a hyperbolic quadric in PG(3,q), q odd, \emph{Discrete Mathematics} {\bf 343} (2020).
\item Arun Maiti: Quasi-vertex-transitive maps on the plane, \emph{Discrete Mathematics} {\bf 343} (2020), 111911-111917.
\item Datt, Gopal and Kumar, Sanjay: A criterion for quasinormality in $\mathbb{C}^n$, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 129} (2019), Paper No. 11, 13 pp..
\item Hatui, Sumana: Schur multipliers of special p-groups of rank 2, \emph{Journal of Group Theory} {\bf 23} (2020), 85-95.
\item Choudhury, Projesh Nath and Tsatsomeros Michael J.: Algorithmic Detection and Construction of N-matrices, \emph{Linear Algebra and Its Applications} {\bf 602} (2020), 46-56.
\item Choudhury, Projesh Nath and Sivakumar, K.C.: Matrices with Positive Semidefinite Real Part, to appear in \emph{Linear and Multilinear Algebra}.
\item Chim, Kwok Chi; Shorey, Tarlok Nath; Sinha, Sneh Bala: On Baker's explicit abc-conjecture, \emph{Publ. Math. Debrecen} {\bf 94} (2019), 435–453.
\item Carstensen, C. and Mallik, G. and Nataraj, N.: Nonconforming Finite Element Discretization for Semilinear Problems with Trilinear Nonlinearity, \emph{IMA Journal of Numerical Analysis} {\bf } (2020).
\item Mal, Arpita and Sain, Debmalya and Paul, Kallol: On some geometric properties of operator spaces, \emph{Banach J. Math. Anal.} {\bf 13} (2019), 174-191.
\item Roth, Julien and Upadhyay, Abhitosh: On almost stable CMC hypersurfaces in manifolds of bounded sectional curvature, \emph{Bulletin of the Australian Mathematical Society} {\bf 101} (2020), 333-338.
\item Roth, Julien and Upadhyay, Abhitosh: f-Biharmonic Submanifolds of Generalized Space Forms, \emph{Results in Mathematics} {\bf 75} (2020).
\item Barbosa, Ezequiel and Santana, Farley and Upadhyay, Abhitosh: On stable CMC free-boundary surfaces in a strictly convex domain of a bi-invariant Lie group, to appear in \emph{International Journal of Mathematics}.
\end{enumerate}

\subsection{Papers in Journals with links}

\begin{enumerate}
\item S Aiyappan, A. K. Nandakumaran and Abu Sufian: Asymptotic analysis of Boundary Optimal Control Problem on a General Branched Structure, \emph{Mathematical Methods in the Applied Sciences} {\bf 42} (2019. DOI: 10.1002/mma.5748), 6407-6434.
\item K. Murali, A. K. Nandakumaran, Turgurt Durduran, Hari M Varma: Recovery of the diffuse correlation spectroscopy data-type from speckle contrast measurements: towards low-cost, deep-tissue blood flow measurements, \emph{Biomedical Optics Express 5395} {\bf 10} (2019).
\item S Aiyappan, A. K. Nandakumaran and Ravi Prakash: Locally Periodic Unfolding Operator for Highly Oscillating Rough Domains, \emph{Annali di Matematica Pura ed Applicata} {\bf 198} (2019), 1931-1954.
\item S Aiyappan, A. K. Nandakumaran and Ravi Prakash: Semi-linear optimal control problem on a smooth oscillating domain, \emph{Communications in Contemporary Mathematics} {\bf } (2019), 1-26.
\item S Aiyappan, Editha C. Jose, Ivy Carol B. Lomerio and A. K. Nandakumaran: Control Problem on a Rough Circular Domain and Homogenization, \emph{Asymptotic Analysis} {\bf 115} (2019), 19-46.
\item M. Balodi, A. Banerjee, S. Ray: Cohomology of modules over H-categories and co-H-categories, to appear in \emph{Canadian Journal of Mathematics}.
\item M. Balodi, A. Banerjee, S. Ray: Entwined modules over linear categories and Galois extensions, to appear in \emph{Israel Journal of Mathematics (Accepted)}.
\item A. Banerjee: On Noetherian schemes over $(\mathcal C,\otimes,1)$ and the category of quasi-coherent sheaves, to appear in \emph{Indiana University Mathematics Journal}.
\item A. Banerjee: On differential torsion theories and rings with several objects, \emph{Canadian Math Bulletin} {\bf 62} (2019), 703--714.
\item Khare, Apoorva and Rajaratnam, Bala: Probability inequalities and tail estimates for metric semigroups, \emph{Advances in Operator Theory (Special issue in honor of Rajendra Bhatia)} {\bf 5} (2020), 779-795.
\item Ghosh, Subhajit: Total Variation Cutoff for the Transpose Top-2 with Random Shuffle, to appear in \emph{Journal of Theoretical Probability}.
\item Datta, B. and Gupta, S.: Semi-regular tilings of the hyperbolic plane \label{bdsg_2019}, to appear in \emph{Discrete Comput Geom}.
\item Datta, B. and Sarkar, S.: Equilibrium triangulations of some quasitoric 4-manifolds, \emph{Southeast Asian Bull. Math.} {\bf 44} (2020), 57--78.
\item Datta, B. and Maity, D.: Correction to: Semi-equivelar and vertex-transitive maps on the torus, \emph{Beitr Algebra Geom} {\bf 61} (2020), 187--188.
\item Martin Kreuzer and Dilip P. Patil: COMPUTATIONAL ASPECTS OF BURNSIDE RINGS PART II: IMPORTANT MAPS, to appear in \emph{Beiträge zur Algebra und Geometrie / Contributions to Algebra and Geometry}.
\item Gadadhar Misra: Professor, \emph{Adv. Math.} {\bf 351} (2019), 1105 - 1138.
\item Bharali, Gautam and Maitra, Anwoy: A weak notion of visibility, a family of examples, and Wolff--Denjoy theorems, to appear in \emph{Annali della Scuola Normale Superiore di Pisa}.
\item Maitra, Anwoy: On the continuous extension of Kobayashi isometries, \emph{Proc. Amer. Math. Soc.} {\bf 148} (2020), 3437--3451.
\item Vishnu Priya, N. and Senthilvelan, M. and  Rangarajan, G. and Lakshmanan, M: On symmetry preserving and symmetry broken bright, dark and antidark soliton solutions of nonlocal nonlinear Schrödinger equation, \emph{Physics Letters A} {\bf 383} (2019), 15-26.
\item Kashyap, G. A. R. S. R. K. and Bapat, D. and Das, D. and Gowaikar, R. D. and Amritkar, R. E. and Rangarajan, G. and Ravindranath, V. and Ambika, G: Synapse loss and progress of Alzheimer’s disease - A network model, \emph{Scientific Reports} {\bf 9} (2019), 6555.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan, G.: On the role of four-wave mixing effect in the interactions between nonlinear modes of coupled generalized nonlinear Schrodinger equation, \emph{Chaos} {\bf 29} (2019), 123135.
\item Sarkar, Amar Deep and Verma, Kaushal: A submultiplicative property of the Carath\'{e}odory metric on planar domains, \emph{Proc. Indian Acad. Sci. (Math. Sci.)} {\bf } ().
\item Frauke Bleher, Ted Chinburg, Ralph Greenberg, Mahesh Kakde, George Pappas, Romyar Sharifi, Martin Taylor: Higher Chern Classes in Iwasawa Theory, \emph{American Journal of Mathematics} {\bf 142} (2020), 627-682.
\item Basu, R. and Ganguly, S. and Hegde, M. and Krishnapur, M.: Lower Deviations in β-ensembles and Law of Iterated Logarithm in Last Passage Percolation, to appear in \emph{Israel Journal of Mathematics}.
\item M. K. Ghosh and S. Pradhan: Zero-Sum Risk-Sensitive Stochastic Differential Games with Reflecting Diffusions in the Orthant, to appear in \emph{ESIAM: Control, Optimization and Calculus of Variations}.
\item M. K. Ghosh, K. Suresh Kumar, C. Pal and S. Pradhan: Nonzero-sum risk-sensitive stochastic differential games with discounted costs, Stochastic Analysis and Applications, to appear in \emph{Stochastic Analysis and Applications}.
\item Iyer, S. K. and Yogeshwaran, D.: Thresholds for vanishing of isolated faces in random Cech and Vietoris-Rips Complexes, \emph{Annals de l'Institut Henri Poincar$\acute{e}$ - Probabiliti$\acute{e}$s et Statistiques} {\bf 56} (2020), 1869-1897.
\item Datta, Basudeb and Gupta, Subhojoy: Semi-regular tilings of the hyperbolic plane, to appear in \emph{Discrete and Computational Geometry}.
\item Thirupathi Gudi, Ramesh Ch. Sau: Finite Element Analysis of the Constrained Dirichlet Boundary Control Problem Governed by the Diffusion Problem, to appear in \emph{ESAIM: Control, Optim. Calc. Var}.
\item Bhattacharyya, Tirthankar and Biswas, Anindya and Chandel, Vikramjeet Singh: On the Nevanlinna problem: characterization of all Schur–Agler class solutions affiliated with a given kernel, \emph{Studia Math.} {\bf 255} (2020), 83–107..
\item Bhattacharyya, Tirthankar and Das, Bata Krishna and Sau, Haripada: Toeplitz operators on the symmetrized bidisc, \emph{International Mathematics Research Notices (IMRN)} {\bf } ().
\item Pingali, Vamsi Pritham and Varolin, Dror: Nonuniformly flat affine algebraic hypersurfaces, to appear in \emph{Nagoya Mathematical Journal}.
\item Alvarez-Consul, Luis and Garcia-Fernandez, Mario  and García-Prada, Oscar and Pingali, Vamsi Pritham: Gravitating vortices and the Einstein–Bogomol’nyi equations, to appear in \emph{Mathematische Annalen}.
\item Datar, Ved and Jacob,Adam and Zhang, Yuguang: Adiabatic limits of anti-self-dual connections on collapsed K3 surfaces, to appear in \emph{Journal of differential geometry}.
\item Datar, Ved and Pingali, Vamsi Pritham: On coupled constant scalar curvature Kaehler metrics, to appear in \emph{Journal of symplectic geometry}.
\item Pradhan, P. and Sahu, B.: \href{https://doi.org/10.1016/j.disc.2020.112044}{A characterization of the family of secant lines to a hyperbolic quadric in PG(3,q), q odd}, \emph{Discrete Mathematics} {\bf 343} (2020).
\item Arun Maiti: \href{https://www.sciencedirect.com/science/article/pii/S0012365X20301035}{Quasi-vertex-transitive maps on the plane}, \emph{Discrete Mathematics} {\bf 343} (2020), 111911-111917.
\item Datt, Gopal and Kumar, Sanjay: \href{https://link.springer.com/article/10.1007/s12044-018-0449-5?shared-article-renderer}{A criterion for quasinormality in $\mathbb{C}^n$}, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 129} (2019), Paper No. 11, 13 pp..
\item Hatui, Sumana: \href{https://www.degruyter.com/view/journals/jgth/23/1/article-p85.xml}{Schur multipliers of special p-groups of rank 2}, \emph{Journal of Group Theory} {\bf 23} (2020), 85-95.
\item Choudhury, Projesh Nath and Tsatsomeros Michael J.: \href{https://doi.org/10.1016/j.laa.2020.04.028}{Algorithmic Detection and Construction of N-matrices}, \emph{Linear Algebra and Its Applications} {\bf 602} (2020), 46-56.
\item Choudhury, Projesh Nath and Sivakumar, K.C.: \href{https://doi.org/10.1080/03081087.2019.1602588}{Matrices with Positive Semidefinite Real Part}, to appear in \emph{Linear and Multilinear Algebra}.
\item Chim, Kwok Chi; Shorey, Tarlok Nath; Sinha, Sneh Bala: On Baker's explicit abc-conjecture, \emph{Publ. Math. Debrecen} {\bf 94} (2019), 435–453.
\item Carstensen, C. and Mallik, G. and Nataraj, N.: \href{https://doi.org/10.1093/imanum/drz071}{Nonconforming Finite Element Discretization for Semilinear Problems with Trilinear Nonlinearity}, \emph{IMA Journal of Numerical Analysis} {\bf } (2020).
\item Mal, Arpita and Sain, Debmalya and Paul, Kallol: \href{https://projecteuclid.org/euclid.bjma/1543914019}{On some geometric properties of operator spaces}, \emph{Banach J. Math. Anal.} {\bf 13} (2019), 174-191.
\item Roth, Julien and Upadhyay, Abhitosh: \href{https://www.cambridge.org/core/journals/bulletin-of-the-australian-mathematical-society/article/on-almost-stable-cmc-hypersurfaces-in-manifolds-of-bounded-sectional-curvature/E0BF84D4D46C82F6D75CDCEAAF5DEE94}{On almost stable CMC hypersurfaces in manifolds of bounded sectional curvature}, \emph{Bulletin of the Australian Mathematical Society} {\bf 101} (2020), 333-338.
\item Roth, Julien and Upadhyay, Abhitosh: \href{https://link.springer.com/article/10.1007/s00025-019-1142-4?shared-article-renderer#citeas}{f-Biharmonic Submanifolds of Generalized Space Forms}, \emph{Results in Mathematics} {\bf 75} (2020).
\item Barbosa, Ezequiel and Santana, Farley and Upadhyay, Abhitosh: On stable CMC free-boundary surfaces in a strictly convex domain of a bi-invariant Lie group, to appear in \emph{International Journal of Mathematics}.
\end{enumerate}

\subsection{Papers in Conference proceedings}

\begin{enumerate}
\item A. Banerjee: Spectra as universal objects in categories of supports in {\em Banach Center Publications(Quantum Dynamics Proceedings)} Ludwik Dabrowski, Ryszard Nest, Adam Skalski (ed.) , Banach Center Publications , IMPAN, Warsaw, Poland (2020), (accepted).
\item IMPAN, Warsaw, Poland: 2020 in {\em (accepted)} No (ed.)   ().
\item S. Ghara and G. Misra: Decomposition of the tensor product of two Hilbert modules in {\em Operator Theory, Operator Algebras and Their Interactions with Geometry and Topology Ronald G. Douglas Memorial Volume} Curto, R.E., Helton, W., Lin, H., Tang, X., Yang, R., Yu, G. (ed.) , Operator Theory: Advances and Applications , Birkhauser (2020).
\item Birkhauser: 2020 in {\em } Yes (ed.) , Gadadhar Misra , Decomposition of the tensor product of two Hilbert modules (Operator Theory, Operator Algebras and Their Interactions with Geometry and Topology Ronald G. Douglas Memorial Volume), R. Curto, W. Helton, H. Lin, X. Tang, R. Yang, G. Yu.
\item Decomposition of the tensor product of two Hilbert modules: Operator Theory, Operator Algebras and Their Interactions with Geometry and Topology Ronald G. Douglas Memorial Volume in {\em R. Curto, W. Helton, H. Lin, X. Tang, R. Yang, G. Yu} Operator Theory: Advances and Applications (ed.) , Birkhauser , 2020 (), No.
\item 2020:  in {\em No}  (ed.)   ().
\end{enumerate}

\subsection{Books}

\subsection{Book Chapters}

\begin{enumerate}
\item Gupta, Subhojoy and Seshadri, Harish: Complex geometry of Teichmüller domains in {\em Handbook of Teichmüller theory} Papadopoulos, Athanase (ed.), IRMA Lect. Math. Theor. Phys, Eur. Math. Soc., Zürich (2020), 26.
\item Gupta, Subhojoy: Holomorphic quadratic differentials in Teichmüller theory in {\em Handbook of Teichmüller theory} Papadopoulos, Athanase (ed.), IRMA Lect. Math. Theor. Phys., Eur. Math. Soc., Zürich (2020), 36.
\end{enumerate}

\subsection{Preprints}

\begin{enumerate}
\item Khare, Apoorva: Sharp nonzero lower bounds for the Schur product theorem, submitted for publication.
\item Lübeck, Frank and Prasad, Dipendra and Ayyer, Arvind: A character relationship between symmetric group and hyperoctahedral group, preprint.
\item Ayyer, Arvind and Sinha, Shubham: The size of t-cores and hook lengths of random cells in random partitions, preprint.
\item Ayyer, Arvind: A simple symmetric exclusion process driven by an asymmetric tracer particle, preprint.
\item David Burns, Mahesh Kakde: On Weil-Etale Cohomology and Artin L-series, preprint.
\item Gupta, Purvi and Wawrzyniak, Chloe Urbanski: STABILITY OF THE HULL(S) OF AN $n$-SPHERE IN $\mathbb{C}^n$, submitted for publication.
\item Radhika Ganapathy: A Hecke algebra isomorphism over close local fields, preprint.
\item Miermont, Grégory and Sen, Sanchayan: On breadth-first constructions of scaling limits of random graphs and random unicellular maps, submitted for publication.
\item Gupta, Subhojoy and Mj, Mahan: Meromorphic projective structures, grafting and the monodromy map, preprint.
\item Gupta, Subhojoy: Monodromy groups of $CP^1$-structures on punctured surfaces, preprint.
\item Biswas, Indranil and Gupta, Subhojoy and Mj, Mahan and Whang, Junho Peter: Surface group representations in SL(2,C) with finite mapping class orbits, preprint.
\item Ganguly, Pritam and Thangavelu,Sundaram: On lacunary spherical maximal function on Heisenberg groups, preprint.
\item Bhattacharyya, Tirthankar and Kumar, Poornendu and Sau, Haripada: Distinguished Varieties Through the Berger--Coburn--Lebow Theorem, submitted for publication.
\item Bhattacharyya, T. and  Biswas, A. and Maitra, A.: An unbounded realization of the symmetrized bidisc, submitted for publication.
\item Pingali, Vamsi Pritham: The deformed Hermitian Yang-Mills equation on three-folds, preprint.
\item Garcia-Fernandez, Mario and Pingali, Vamsi Pritham and Yao, Chengjian: Gravitating vortices with positive curvature, preprint.
\item Quddus, Safdar: Equivariant Cartan Homotopy Formula for $DG$-algebra., preprint.
\item Arun Maiti: OPTIMAL INDEX GROWTH OF CLOSED GEODESICS, submitted for publication.
\item Poria, Anirudha: Uncertainty Principles for the Fourier and the Short-Time Fourier Transforms, submitted for publication.
\item Poria, Anirudha: Uncertainty principles for the Opdam-Cherednik transform on modulation spaces, submitted for publication.
\item Bera, Sudip: ENUMERATION OF WEIGHTED PATHS ON A DIGRAPH AND BLOCK HOOK DETERMINANT, submitted for publication.
\item Behera, Kiran Kumar: Self-inversive polynomials from rational transformation of linear functionals, submitted for publication.
\item Hatui, Sumana and Singla Pooja: On Schur multiplier and Projective representations of Heisenberg groups, submitted for publication.
\item Chattopahyay, Arup and Sarkar, Jaydeb and Sarkar, Srijan: Multiplicities, invariant subspaces and an additive formula., preprint.
\item Ghara, Soumitra and Kumar, Surjit and Pramanick, Paramita: $K$-homogeneous tuple of operators on bounded symmetric domains, preprint.
\item Toft, J. and Gumber, A. and Manna, R. and Ratnakumar, P. K.: Translation and modulation invariant Hilbert spaces, submitted for publication.
\end{enumerate}

\subsection{Preprints with links}

\begin{enumerate}
\item Khare, Apoorva: Sharp nonzero lower bounds for the Schur product theorem, submitted for publication.
\item Lübeck, Frank and Prasad, Dipendra and Ayyer, Arvind: A character relationship between symmetric group and hyperoctahedral group, preprint.
\item Ayyer, Arvind and Sinha, Shubham: The size of t-cores and hook lengths of random cells in random partitions, preprint.
\item Ayyer, Arvind: A simple symmetric exclusion process driven by an asymmetric tracer particle, preprint.
\item David Burns, Mahesh Kakde: On Weil-Etale Cohomology and Artin L-series, preprint.
\item Gupta, Purvi and Wawrzyniak, Chloe Urbanski: STABILITY OF THE HULL(S) OF AN $n$-SPHERE IN $\mathbb{C}^n$, submitted for publication.
\item Radhika Ganapathy: A Hecke algebra isomorphism over close local fields, preprint.
\item Miermont, Grégory and Sen, Sanchayan: On breadth-first constructions of scaling limits of random graphs and random unicellular maps, submitted for publication.
\item Gupta, Subhojoy and Mj, Mahan: Meromorphic projective structures, grafting and the monodromy map, preprint.
\item Gupta, Subhojoy: Monodromy groups of $CP^1$-structures on punctured surfaces, preprint.
\item Biswas, Indranil and Gupta, Subhojoy and Mj, Mahan and Whang, Junho Peter: Surface group representations in SL(2,C) with finite mapping class orbits, preprint.
\item Ganguly, Pritam and Thangavelu,Sundaram: On lacunary spherical maximal function on Heisenberg groups, preprint.
\item Bhattacharyya, Tirthankar and Kumar, Poornendu and Sau, Haripada: Distinguished Varieties Through the Berger--Coburn--Lebow Theorem, submitted for publication.
\item Bhattacharyya, T. and  Biswas, A. and Maitra, A.: An unbounded realization of the symmetrized bidisc, submitted for publication.
\item Pingali, Vamsi Pritham: The deformed Hermitian Yang-Mills equation on three-folds, preprint.
\item Garcia-Fernandez, Mario and Pingali, Vamsi Pritham and Yao, Chengjian: Gravitating vortices with positive curvature, preprint.
\item Quddus, Safdar: \href{https://arxiv.org/abs/2002.02192}{Equivariant Cartan Homotopy Formula for $DG$-algebra.}, preprint.
\item Arun Maiti: OPTIMAL INDEX GROWTH OF CLOSED GEODESICS, submitted for publication.
\item Poria, Anirudha: \href{https://arxiv.org/abs/2004.04184}{Uncertainty Principles for the Fourier and the Short-Time Fourier Transforms}, submitted for publication.
\item Poria, Anirudha: \href{https://arxiv.org/abs/2005.14274}{Uncertainty principles for the Opdam-Cherednik transform on modulation spaces}, submitted for publication.
\item Bera, Sudip: ENUMERATION OF WEIGHTED PATHS ON A DIGRAPH AND BLOCK HOOK DETERMINANT, submitted for publication.
\item Behera, Kiran Kumar: \href{https://arxiv.org/abs/1909.12548}{Self-inversive polynomials from rational transformation of linear functionals}, submitted for publication.
\item Hatui, Sumana and Singla Pooja: \href{https://arxiv.org/abs/1909.06589}{On Schur multiplier and Projective representations of Heisenberg groups}, submitted for publication.
\item Chattopahyay, Arup and Sarkar, Jaydeb and Sarkar, Srijan: \href{https://arxiv.org/abs/1812.05435}{Multiplicities, invariant subspaces and an additive formula.}, preprint.
\item Ghara, Soumitra and Kumar, Surjit and Pramanick, Paramita: $K$-homogeneous tuple of operators on bounded symmetric domains, preprint.
\item Toft, J. and Gumber, A. and Manna, R. and Ratnakumar, P. K.: \href{https://arxiv.org/abs/2004.02430v1}{Translation and modulation invariant Hilbert spaces}, submitted for publication.
\end{enumerate}

% from web page

\subsection{Publications (all kinds) from web page}

\begin{enumerate}
\item Aas, Erik and Ayyer, Arvind and Linusson, Svante and Potka,
Samu: The exact phase diagram for a semipermeable {TASEP} with
nonlocal boundary jumps, \emph{J. Phys. A} {\bf 52} (2019), 355001, 19.
\item Aiyappan, S. and Jose, Editha C. and Lomerio, Ivy Carol B. and
Nandakumaran, A. K.: Control problem on a rough circular domain and homogenization, \emph{Asymptot. Anal.} {\bf 115} (2019), 19--46.
\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi: Semi-linear optimal control problem on a smooth oscillating
domain, \emph{Commun. Contemp. Math.} {\bf 22} (2020), 1950029, 26.
\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi: Locally periodic unfolding operator for highly oscillating
rough domains, \emph{Ann. Mat. Pura Appl. (4)} {\bf 198} (2019), 1931--1954.
\item Aiyappan, S. and Nandakumaran, A. K. and Sufian, Abu: Asymptotic analysis of a boundary optimal control problem on a
general branched structure, \emph{Math. Methods Appl. Sci.} {\bf 42} (2019), 6407--6434.
\item Aiyappan, S. and Sardar, Bidhan Chandra: Biharmonic equation in a highly oscillating domain and
homogenization of an associated control problem, \emph{Appl. Anal.} {\bf 98} (2019), 2783--2801.
\item Akutagawa, Kazuo and Endo, Hisaaki and Seshadri, Harish: A gap theorem for positive {E}instein metrics on the
four-sphere, \emph{Math. Ann.} {\bf 373} (2019), 1329--1339.
\item Anamby, Pramath and Das, Soumya: Distinguishing {H}ermitian cusp forms of degree 2 by a certain
subset of all {F}ourier coefficients, \emph{Publ. Mat.} {\bf 63} (2019), 307--341.
\item Anand, Jatin and Bhattacharyya, Tirthankar and Srivastava,
Sachi: Some thoughts on composition operators on subspaces of the
{H}ardy space, \emph{Arch. Math. (Basel)} {\bf 114} (2020), 431--444.
\item Ayyer, Arvind: Squareness for the {M}onopole-{D}imer model, \emph{Ann. Comb.} {\bf 24} (2020), 237--255.
\item Ayyer, Arvind and Behrend, Roger E.: Factorization theorems for classical group characters, with
applications to alternating sign matrices and plane
partitions, \emph{J. Combin. Theory Ser. A} {\bf 165} (2019), 78--105.
\item Ayyer, Arvind and Behrend, Roger E. and Fischer, Ilse: Extreme diagonally and antidiagonally symmetric alternating
sign matrices of odd order, \emph{Adv. Math.} {\bf 367} (2020), 107125, 56.
\item Ayyer, Arvind and Finn, Caley and Roy, Dipankar: The phase diagram for a multispecies left-permeable asymmetric
exclusion process, \emph{J. Stat. Phys.} {\bf 174} (2019), 605--621.
\item Ayyer, Arvind and Fischer, Ilse: Bijective proofs of skew {S}chur polynomial factorizations, \emph{J. Combin. Theory Ser. A} {\bf 174} (2020), 105241, 40.
\item Ayyer, Arvind and Linusson, Svante: Reverse juggling processes, \emph{Random Structures Algorithms} {\bf 55} (2019), 56--72.
\item Ayyer, Arvind and Singla, Pooja: Random motion on finite rings, {I}: commutative rings, \emph{Algebr. Represent. Theory} {\bf 23} (2020), 583--604.
\item Balakumar, G. P. and Borah, Diganta and Mahajan, Prachi and
Verma, Kaushal: Remarks on the higher dimensional {S}uita conjecture, \emph{Proc. Amer. Math. Soc.} {\bf 147} (2019), 3401--3411.
\item Balhara, Rakesh: Hardy's inequality for the fractional powers of the {G}rushin
operator, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 129} (2019), Art. 33, 25.
\item Banerjee, Abhishek: On differential torsion theories and rings with several
objects, \emph{Canad. Math. Bull.} {\bf 62} (2019), 703--714.
\item Banerjee, Abhishek and Kour, Surjeet: {$(A,\delta)$}-modules, {H}ochschild homology and higher
derivations, \emph{Ann. Mat. Pura Appl. (4)} {\bf 198} (2019), 1781--1802.
\item Belton, Alexander and Guillot, Dominique and Khare, Apoorva
and Putinar, Mihai: Simultaneous kernels of matrix {H}adamard powers, \emph{Linear Algebra Appl.} {\bf 576} (2019), 142--157.
\item Benedetti, Carolina and Gonz\'{a}lez D'Le\'{o}n, Rafael S. and Hanusa,
Christopher R. H. and Harris, Pamela E. and Khare, Apoorva and
Morales, Alejandro H. and Yip, Martha: A combinatorial model for computing volumes of flow polytopes, \emph{Trans. Amer. Math. Soc.} {\bf 372} (2019), 3369--3404.
\item Bera, Sayani and Pal, Ratna and Verma, Kaushal: A rigidity theorem for {H}\'{e}non maps, \emph{Eur. J. Math.} {\bf 6} (2020), 508--532.
\item Bharali, Gautam: On the growth of the {B}ergman metric near a point of infinite
type, \emph{J. Geom. Anal.} {\bf 30} (2020), 1238--1258.
\item Bharali, Gautam and Chandel, Vikramjeet Singh: Pick interpolation on the polydisc: small families of
sufficient kernels, \emph{Complex Anal. Oper. Theory} {\bf 13} (2019), 2069--2093.
\item Bhattacharyya, Tirthankar and Rao, T. S. S. R. K.: Survey of last ten years of work done in {I}ndia in some
selected areas of functional analysis and operator theory, \emph{Indian J. Pure Appl. Math.} {\bf 50} (2019), 599--617.
\item Bhimani, Divyang G. and Balhara, Rakesh and Thangavelu,
Sundaram: Hermite multipliers on modulation spaces, \emph{Analysis and partial differential equations: perspectives from
developing countries} {\bf 275} (2019), 42--64.
\item Biswas, Indranil and Dumitrescu, Sorin and Gupta, Subhojoy: Branched projective structures on a {R}iemann surface and
logarithmic connections, \emph{Doc. Math.} {\bf 24} (2019), 2299--2337.
\item Biswas, Rahul and Dond, Asha K. and Gudi, Thirupathi: Edge patch-wise local projection stabilized nonconforming
{FEM} for the {O}seen problem, \emph{Comput. Methods Appl. Math.} {\bf 19} (2019), 189--214.
\item Biswas, Shibananda and Ghosh, Gargi and Misra, Gadadhar and
Shyam Roy, Subrata: On reducing submodules of {H}ilbert modules with
{${S}_n$}-invariant kernels, \emph{J. Funct. Anal.} {\bf 276} (2019), 751--784.
\item Boggarapu, Pradeep and Roncal, Luz and Thangavelu, Sundaram: On extension problem, trace {H}ardy and {H}ardy's inequalities
for some fractional {L}aplacians, \emph{Commun. Pure Appl. Anal.} {\bf 18} (2019), 2575--2605.
\item Carstensen, Carsten and Dond, Asha K. and Rabus, Hella: Quasi-optimality of adaptive mixed {FEM}s for non-selfadjoint
indefinite second-order linear elliptic problems, \emph{Comput. Methods Appl. Math.} {\bf 19} (2019), 233--250.
\item Chakraborty, Atlanta and Chandru, Vijay and Rao, M. R.: A linear programming primer: from {F}ourier to {K}armarkar, \emph{Ann. Oper. Res.} {\bf 287} (2020), 593--616.
\item Chandel, Vikramjeet Singh: The 3-point spectral pick interpolation problem and an
application to holomorphic correspondences, \emph{J. Geom. Anal.} {\bf 30} (2020), 551--572.
\item Choudhury, Projesh Nath and Tsatsomeros, Michael J.: Algorithmic detection and construction of {N}-matrices, \emph{Linear Algebra Appl.} {\bf 602} (2020), 46--56.
\item Cicuttin, Matteo and Ern, Alexandre and Gudi, Thirupathi: Hybrid high-order methods for the elliptic obstacle problem, \emph{J. Sci. Comput.} {\bf 83} (2020), Paper No. 8, 18.
\item Das, Soumya and Jha, Abhash Kumar: Analytic properties of twisted real-analytic {H}ermitian
{K}lingen type {E}isenstein series and applications, \emph{Abh. Math. Semin. Univ. Hambg.} {\bf 89} (2019), 105--116.
\item Das, Soumya and Pal, Ritwik: The first negative eigenvalue of {Y}oshida lifts, \emph{Res. Number Theory} {\bf 5} (2019), Paper No. 20, 9.
\item Dond, Asha K. and Gudi, Thirupathi: Patch-wise local projection stabilized finite element methods
for convection-diffusion-reaction problems, \emph{Numer. Methods Partial Differential Equations} {\bf 35} (2019), 638--663.
\item Dond, Asha K. and Gudi, Thirupathi and Sau, Ramesh CH.: An error analysis of discontinuous finite element methods for
the optimal control problems governed by {S}tokes equation, \emph{Numer. Funct. Anal. Optim.} {\bf 40} (2019), 421--460.
\item Gadgil, Siddhartha: Homogeneous {L}ength {F}unctions on {G}roups: {I}ntertwined
{C}omputer and {H}uman {P}roofs, \emph{J. Automat. Reason.} {\bf 64} (2020), 677--688.
\item Garani, Shayan Srinivasa and Seshadri, Harish: An algorithmic approach to {S}outh {I}ndian classical music, \emph{J. Math. Music} {\bf 13} (2019), 107--134.
\item Gaussier, Herv\'{e} and Seshadri, Harish: Holomorphic motions and complex geometry, \emph{Proc. Amer. Math. Soc.} {\bf 147} (2019), 301--313.
\item Ghara, Soumitra and Kumar, Surjit: On the sum of two subnormal kernels, \emph{J. Math. Anal. Appl.} {\bf 469} (2019), 1015--1027.
\item Gudi, Thirupathi and Majumder, Papri: Crouzeix-{R}aviart finite element approximation for the
parabolic obstacle problem, \emph{Comput. Methods Appl. Math.} {\bf 20} (2020), 273--292.
\item Gudi, Thirupathi and Majumder, Papri: Convergence analysis of finite element method for a parabolic
obstacle problem, \emph{J. Comput. Appl. Math.} {\bf 357} (2019), 85--102.
\item Gudi, Thirupathi and Majumder, Papri: Conforming and discontinuous {G}alerkin {FEM} in space for
solving parabolic obstacle problem, \emph{Comput. Math. Appl.} {\bf 78} (2019), 3896--3915.
\item Gupta, Rajeev and Kumar, Surjit and Trivedi, Shailesh: Unitary equivalence of operator-valued multishifts, \emph{J. Math. Anal. Appl.} {\bf 487} (2020), 124032, 23.
\item Gupta, Rajeev and Kumar, Surjit and Trivedi, Shailesh: Von {N}eumann's inequality for commuting operator-valued
multishifts, \emph{Proc. Amer. Math. Soc.} {\bf 147} (2019), 2599--2608.
\item Gupta, Subhojoy: Limits of harmonic maps and crowned hyperbolic surfaces, \emph{Trans. Amer. Math. Soc.} {\bf 372} (2019), 7573--7596.
\item Gupta, Subhojoy and Mj, Mahan: Monodromy representations of meromorphic projective
structures, \emph{Proc. Amer. Math. Soc.} {\bf 148} (2020), 2069--2078.
\item Gupta, Subhojoy and Seshadri, Harish: On domains biholomorphic to {T}eichm\"{u}ller spaces, \emph{Int. Math. Res. Not. IMRN} {\bf } (2020), 2542--2560.
\item Gupta, Subhojoy and Wolf, Michael: Meromorphic quadratic differentials and measured foliations on
a {R}iemann surface, \emph{Math. Ann.} {\bf 373} (2019), 73--118.
\item Hatui, Sumana: Schur multipliers of special {$p$}-groups of rank 2, \emph{J. Group Theory} {\bf 23} (2020), 85--95.
\item Hazra, Somnath: Homogeneous 2-shifts, \emph{Complex Anal. Oper. Theory} {\bf 13} (2019), 1729--1763.
\item Jha, Abhash Kumar and Juyal, Abhishek and Pandey, Manish
Kumar: On simultaneous non-vanishing of twisted {$L$}-functions
associated to newforms on {$\Gamma_0(N)$}, \emph{J. Ramanujan Math. Soc.} {\bf 34} (2019), 245--252.
\item Jha, Abhash Kumar and Sahu, Brundaban: Rankin-{C}ohen brackets on {J}acobi forms of several variables
and special values of certain {D}irichlet series, \emph{Int. J. Number Theory} {\bf 15} (2019), 925--933.
\item Jha, Abhash Kumar and Sahu, Brundaban: Differential operators on {J}acobi forms and special values of
certain {D}irichlet series, \emph{Automorphic forms and related topics} {\bf 732} (2019), 91--99.
\item Kashyap, G. A. R. S. R. K. and Bapat, D. and Das, D. and Gowaikar, R. D. and  Amritkar, R. E. and Rangarajan, G. and V. Ravindranath and G. Ambika: Synapse loss and progress of Alzheimers disease - A network model, \emph{Scientific Reports} {\bf 9} (2019).
\item Kataria, Kuldeep Kumar and Vellaisamy, Palaniappan: On the convolution of {M}ittag-{L}effler distributions and its
applications to fractional point processes, \emph{Stoch. Anal. Appl.} {\bf 37} (2019), 115--122.
\item Khan, Mohammad Yasir Feroz and Biswas, Rahul and Sau, Ramesh
Chandra: Instability of {L}agrange interpolation on {$H^1(\Omega)\cap
C(\overline\Omega)$}, \emph{Math. Student} {\bf 88} (2019), 153--158.
\item Khare, Apoorva and Tikaradze, Akaki: A {C}arlitz--von {S}taudt type theorem for finite rings, \emph{Linear Algebra Appl.} {\bf 568} (2019), 106--126.
\item Lin, X. and Kulkarni, P. and Bocci, F. and Schafer, N. and Roy, S. and Tsai, M-Y. and He, Y. and Chen, Y and Rajagopalan, K. and Mooney, S. and Zeng, Y. and Weninger, K. and Grishaev, A. and Onuchic, J. and Levine, H. and Wolynes, P. and Salgia, R. and Rangarajan, G. and Uversky, V. and Orban, J. and Jolly, M. K.: Structural and Dynamical Order of a Disordered Protein: Molecular Insights into Conformational Switching of PAGE4 at the Systems Level, \emph{Biomolecules} {\bf 9} (2019), 77.
\item Mahajan, Prachi and Verma, Kaushal: A comparison of two biholomorphic invariants, \emph{Internat. J. Math.} {\bf 30} (2019), 1950012, 16.
\item Maiti, Arun: Quasi-vertex-transitive maps on the plane, \emph{Discrete Math.} {\bf 343} (2020), 111911, 7.
\item Maitra, Anwoy: On the continuous extension of {K}obayashi isometries, \emph{Proc. Amer. Math. Soc.} {\bf 148} (2020), 3437--3451.
\item Maitra, Anwoy: A form of {S}chwarz's lemma and a bound for the {K}obayashi
metric on convex domains, \emph{J. Math. Anal. Appl.} {\bf 484} (2020), 123694, 16.
\item Maity, Dipendu: Pentagonal maps on the torus and the plane, \emph{Beitr. Algebra Geom.} {\bf 60} (2019), 17--37.
\item Maity, Dipendu and Upadhyay, Ashish Kumar: On centrally symmetric manifolds, \emph{J. Ramanujan Math. Soc.} {\bf 34} (2019), 21--27.
\item Mal, Arpita and Paul, Kallol and Rao, T. S. S. R. K. and Sain,
Debmalya: Approximate {B}irkhoff-{J}ames orthogonality and smoothness in
the space of bounded linear operators, \emph{Monatsh. Math.} {\bf 190} (2019), 549--558.
\item Mal, Arpita and Sain, Debmalya and Paul, Kallol: On some geometric properties of operator spaces, \emph{Banach J. Math. Anal.} {\bf 13} (2019), 174--191.
\item Manfio, F. and Turgay, N. C. and Upadhyay, A.: Biconservative submanifolds in {$\Bbb{S}^n\times\Bbb{R}$} and
{$\Bbb{H}^n\times\Bbb{R}$}, \emph{J. Geom. Anal.} {\bf 29} (2019), 283--298.
\item Misra, Gadadhar: Operators in the {C}owen-{D}ouglas class and related topics, \emph{Handbook of analytic operator theory} {\bf } (2019), 87--137.
\item Misra, Gadadhar and Pal, Avijit and Varughese, Cherian: Contractivity and complete contractivity for finite
dimensional {B}anach spaces, \emph{J. Operator Theory} {\bf 82} (2019), 23--47.
\item Misra, Gadadhar and Reza, Md. Ramiz: Curvature inequalities and extremal operators, \emph{Illinois J. Math.} {\bf 63} (2019), 193--217.
\item Mukherjee, Sajal Kumar and Bera, Sudip: Combinatorial proofs of the {N}ewton-{G}irard and
{C}hapman-{C}ostas-{S}antos identities, \emph{Discrete Math.} {\bf 342} (2019), 1577--1580.
\item Nikolov, Nikolai and Verma, Kaushal: On the squeezing function and {F}ridman invariants, \emph{J. Geom. Anal.} {\bf 30} (2020), 1218--1225.
\item Pal, Chandan and Pradhan, Somnath: Risk sensitive control of pure jump processes on a general
state space, \emph{Stochastics} {\bf 91} (2019), 155--174.
\item Paul, Kallol and Sain, Debmalya: Birkhoff-{J}ames orthogonality and its application in the
study of geometry of {B}anach space, \emph{Advanced topics in mathematical analysis} {\bf } (2019), 245--284.
\item Paul, Kallol and Sain, Debmalya and Ghosh, Puja: Symmetry of {B}irkhoff-{J}ames orthogonality of bounded linear
operators, \emph{Ulam type stability} {\bf } (2019), 331--344.
\item Pingali, Vamsi P.: A note on the deformed {H}ermitian {Y}ang-{M}ills {PDE}, \emph{Complex Var. Elliptic Equ.} {\bf 64} (2019), 503--518.
\item Pingali, Vamsi Pritham: A vector bundle version of the {M}onge-{A}mp\`ere equation, \emph{Adv. Math.} {\bf 360} (2020), 106921, 40.
\item Pingali, Vamsi Pritham: Quillen metrics and perturbed equations, \emph{Lett. Math. Phys.} {\bf 110} (2020), 1861--1875.
\item Poria, Anirudha: Positive weight function and classification of {$g$}-frames, \emph{Numer. Funct. Anal. Optim.} {\bf 41} (2020), 950--968.
\item Quddus, Safdar: Invariants of the {$\Bbb Z_2$} orbifolds of the {P}odle\'{s} two
spheres, \emph{J. Noncommut. Geom.} {\bf 13} (2019), 257--267.
\item Reddy, Nanda Kishore: Equality of {L}yapunov and stability exponents for products of
isotropic random matrices, \emph{Int. Math. Res. Not. IMRN} {\bf } (2019), 606--624.
\item Roth, Julien and Upadhyay, Abhitosh: On compact anisotropic {W}eingarten hypersurfaces in
{E}uclidean space, \emph{Arch. Math. (Basel)} {\bf 113} (2019), 213--224.
\item Roth, Julien and Upadhyay, Abhitosh: On almost stable {CMC} hypersurfaces in manifolds of bounded
sectional curvature, \emph{Bull. Aust. Math. Soc.} {\bf 101} (2020), 333--338.
\item Roy, Krishanu and Venkatesh, R.: Maximal closed subroot systems of real affine root systems, \emph{Transform. Groups} {\bf 24} (2019), 1261--1308.
\item Sain, Debmalya: On extreme contractions and the norm attainment set of a
bounded linear operator, \emph{Ann. Funct. Anal.} {\bf 10} (2019), 135--143.
\item Sain, Debmalya: On the norm attainment set of a bounded linear operator and
semi-inner-products in normed spaces, \emph{Indian J. Pure Appl. Math.} {\bf 51} (2020), 179--186.
\item Sain, Debmalya: Smooth points in operator spaces and some
{B}ishop-{P}helps-{B}ollob\'{a}s type theorems in {B}anach spaces, \emph{Oper. Matrices} {\bf 13} (2019), 433--445.
\item Sain, Debmalya and Mal, Arpita and Paul, Kallol: Some remarks on {B}irkhoff-{J}ames orthogonality of linear
operators, \emph{Expo. Math.} {\bf 38} (2020), 138--147.
\item Sain, Debmalya and Paul, Kallol and Bhunia, Pintu and Bag,
Santanu: On the numerical index of polyhedral {B}anach spaces, \emph{Linear Algebra Appl.} {\bf 577} (2019), 121--133.
\item Sain, Debmalya and Paul, Kallol and Mal, Arpita: On approximate {B}irkhoff-{J}ames orthogonality and normal
cones in a normed space, \emph{J. Convex Anal.} {\bf 26} (2019), 341--351.
\item Sain, Debmalya and Paul, Kallol and Mandal, Kalidas: On two extremum problems related to the norm of a bounded
linear operator, \emph{Oper. Matrices} {\bf 13} (2019), 421--432.
\item Sain, Debmalya and Ray, Anubhab and Paul, Kallol: Extreme contractions on finite-dimensional polygonal {B}anach
spaces, \emph{J. Convex Anal.} {\bf 26} (2019), 877--885.
\item Sanki, Bidyut and Gadgil, Siddhartha: Graphs of systoles on hyperbolic surfaces, \emph{J. Topol. Anal.} {\bf 11} (2019), 1--20.
\item Sarkar, Amar Deep and Verma, Kaushal: Boundary behaviour of some conformal invariants on planar
domains, \emph{Comput. Methods Funct. Theory} {\bf 20} (2020), 145--158.
\item Seshadri, Harish: Differential geometry in {I}ndia, \emph{Indian J. Pure Appl. Math.} {\bf 50} (2019), 795--799.
\item Turgay, Nurettin Cenk and Upadhyay, Abhitosh: On biconservative hypersurfaces in 4-dimensional {R}iemannian
space forms, \emph{Math. Nachr.} {\bf 292} (2019), 905--921.
\item Urbina-Romero, Wilfredo: Gaussian harmonic analysis, \emph{} {\bf } (2019), xix+477.
\item Verma, Kaushal: Notes on the boundaries of quadrature domains, \emph{Anal. Math. Phys.} {\bf 9} (2019), 617--638.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan,
Govindan: On the role of four-wave mixing effect in the interactions
between nonlinear modes of coupled generalized nonlinear
{S}chr\"{o}dinger equation, \emph{Chaos} {\bf 29} (2019), 123135, 14.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan, Govindan
and Lakshmanan, M.: On symmetry preserving and symmetry broken bright, dark and
antidark soliton solutions of nonlocal nonlinear {S}chr\"{o}dinger
equation, \emph{Phys. Lett. A} {\bf 383} (2019), 15--26.
\end{enumerate}

\subsection{Publications (all kinds) from web page with hyperlinks}

\begin{enumerate}
\item Aas, Erik and Ayyer, Arvind and Linusson, Svante and Potka,
Samu: \href{https://doi.org/10.1088/1751-8121/ab2e96}{The exact phase diagram for a semipermeable {TASEP} with
nonlocal boundary jumps}, \emph{J. Phys. A} {\bf 52} (2019), 355001, 19.
\item Aiyappan, S. and Jose, Editha C. and Lomerio, Ivy Carol B. and
Nandakumaran, A. K.: \href{https://doi.org/10.3233/asy-191526}{Control problem on a rough circular domain and homogenization}, \emph{Asymptot. Anal.} {\bf 115} (2019), 19--46.
\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi: \href{https://doi.org/10.1142/S0219199719500299}{Semi-linear optimal control problem on a smooth oscillating
domain}, \emph{Commun. Contemp. Math.} {\bf 22} (2020), 1950029, 26.
\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi: \href{https://doi.org/10.1007/s10231-019-00848-7}{Locally periodic unfolding operator for highly oscillating
rough domains}, \emph{Ann. Mat. Pura Appl. (4)} {\bf 198} (2019), 1931--1954.
\item Aiyappan, S. and Nandakumaran, A. K. and Sufian, Abu: \href{https://doi.org/10.1002/mma.5748}{Asymptotic analysis of a boundary optimal control problem on a
general branched structure}, \emph{Math. Methods Appl. Sci.} {\bf 42} (2019), 6407--6434.
\item Aiyappan, S. and Sardar, Bidhan Chandra: \href{https://doi.org/10.1080/00036811.2018.1471207}{Biharmonic equation in a highly oscillating domain and
homogenization of an associated control problem}, \emph{Appl. Anal.} {\bf 98} (2019), 2783--2801.
\item Akutagawa, Kazuo and Endo, Hisaaki and Seshadri, Harish: \href{https://doi.org/10.1007/s00208-018-1749-x}{A gap theorem for positive {E}instein metrics on the
four-sphere}, \emph{Math. Ann.} {\bf 373} (2019), 1329--1339.
\item Anamby, Pramath and Das, Soumya: \href{https://doi.org/10.5565/PUBLMAT6311911}{Distinguishing {H}ermitian cusp forms of degree 2 by a certain
subset of all {F}ourier coefficients}, \emph{Publ. Mat.} {\bf 63} (2019), 307--341.
\item Anand, Jatin and Bhattacharyya, Tirthankar and Srivastava,
Sachi: \href{https://doi.org/10.1007/s00013-019-01406-6}{Some thoughts on composition operators on subspaces of the
{H}ardy space}, \emph{Arch. Math. (Basel)} {\bf 114} (2020), 431--444.
\item Ayyer, Arvind: \href{https://doi.org/10.1007/s00026-019-00480-5}{Squareness for the {M}onopole-{D}imer model}, \emph{Ann. Comb.} {\bf 24} (2020), 237--255.
\item Ayyer, Arvind and Behrend, Roger E.: \href{https://doi.org/10.1016/j.jcta.2019.01.001}{Factorization theorems for classical group characters, with
applications to alternating sign matrices and plane
partitions}, \emph{J. Combin. Theory Ser. A} {\bf 165} (2019), 78--105.
\item Ayyer, Arvind and Behrend, Roger E. and Fischer, Ilse: \href{https://doi.org/10.1016/j.aim.2020.107125}{Extreme diagonally and antidiagonally symmetric alternating
sign matrices of odd order}, \emph{Adv. Math.} {\bf 367} (2020), 107125, 56.
\item Ayyer, Arvind and Finn, Caley and Roy, Dipankar: \href{https://doi.org/10.1007/s10955-018-2183-x}{The phase diagram for a multispecies left-permeable asymmetric
exclusion process}, \emph{J. Stat. Phys.} {\bf 174} (2019), 605--621.
\item Ayyer, Arvind and Fischer, Ilse: \href{https://doi.org/10.1016/j.jcta.2020.105241}{Bijective proofs of skew {S}chur polynomial factorizations}, \emph{J. Combin. Theory Ser. A} {\bf 174} (2020), 105241, 40.
\item Ayyer, Arvind and Linusson, Svante: \href{https://doi.org/10.1002/rsa.20825}{Reverse juggling processes}, \emph{Random Structures Algorithms} {\bf 55} (2019), 56--72.
\item Ayyer, Arvind and Singla, Pooja: \href{https://doi.org/10.1007/s10468-019-09864-w}{Random motion on finite rings, {I}: commutative rings}, \emph{Algebr. Represent. Theory} {\bf 23} (2020), 583--604.
\item Balakumar, G. P. and Borah, Diganta and Mahajan, Prachi and
Verma, Kaushal: \href{https://doi.org/10.1090/proc/14421}{Remarks on the higher dimensional {S}uita conjecture}, \emph{Proc. Amer. Math. Soc.} {\bf 147} (2019), 3401--3411.
\item Balhara, Rakesh: \href{https://doi.org/10.1007/s12044-019-0471-2}{Hardy's inequality for the fractional powers of the {G}rushin
operator}, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 129} (2019), Art. 33, 25.
\item Banerjee, Abhishek: \href{https://doi.org/10.4153/s0008439518000656}{On differential torsion theories and rings with several
objects}, \emph{Canad. Math. Bull.} {\bf 62} (2019), 703--714.
\item Banerjee, Abhishek and Kour, Surjeet: \href{https://doi.org/10.1007/s10231-019-00844-x}{{$(A,\delta)$}-modules, {H}ochschild homology and higher
derivations}, \emph{Ann. Mat. Pura Appl. (4)} {\bf 198} (2019), 1781--1802.
\item Belton, Alexander and Guillot, Dominique and Khare, Apoorva
and Putinar, Mihai: \href{https://doi.org/10.1016/j.laa.2018.03.035}{Simultaneous kernels of matrix {H}adamard powers}, \emph{Linear Algebra Appl.} {\bf 576} (2019), 142--157.
\item Benedetti, Carolina and Gonz\'{a}lez D'Le\'{o}n, Rafael S. and Hanusa,
Christopher R. H. and Harris, Pamela E. and Khare, Apoorva and
Morales, Alejandro H. and Yip, Martha: \href{https://doi.org/10.1090/tran/7743}{A combinatorial model for computing volumes of flow polytopes}, \emph{Trans. Amer. Math. Soc.} {\bf 372} (2019), 3369--3404.
\item Bera, Sayani and Pal, Ratna and Verma, Kaushal: \href{https://doi.org/10.1007/s40879-019-00326-7}{A rigidity theorem for {H}\'{e}non maps}, \emph{Eur. J. Math.} {\bf 6} (2020), 508--532.
\item Bharali, Gautam: \href{https://doi.org/10.1007/s12220-019-00342-9}{On the growth of the {B}ergman metric near a point of infinite
type}, \emph{J. Geom. Anal.} {\bf 30} (2020), 1238--1258.
\item Bharali, Gautam and Chandel, Vikramjeet Singh: \href{https://doi.org/10.1007/s11785-017-0701-5}{Pick interpolation on the polydisc: small families of
sufficient kernels}, \emph{Complex Anal. Oper. Theory} {\bf 13} (2019), 2069--2093.
\item Bhattacharyya, Tirthankar and Rao, T. S. S. R. K.: \href{https://doi.org/10.1007/s13226-019-0345-4}{Survey of last ten years of work done in {I}ndia in some
selected areas of functional analysis and operator theory}, \emph{Indian J. Pure Appl. Math.} {\bf 50} (2019), 599--617.
\item Bhimani, Divyang G. and Balhara, Rakesh and Thangavelu,
Sundaram: Hermite multipliers on modulation spaces, \emph{Analysis and partial differential equations: perspectives from
developing countries} {\bf 275} (2019), 42--64.
\item Biswas, Indranil and Dumitrescu, Sorin and Gupta, Subhojoy: Branched projective structures on a {R}iemann surface and
logarithmic connections, \emph{Doc. Math.} {\bf 24} (2019), 2299--2337.
\item Biswas, Rahul and Dond, Asha K. and Gudi, Thirupathi: \href{https://doi.org/10.1515/cmam-2018-0020}{Edge patch-wise local projection stabilized nonconforming
{FEM} for the {O}seen problem}, \emph{Comput. Methods Appl. Math.} {\bf 19} (2019), 189--214.
\item Biswas, Shibananda and Ghosh, Gargi and Misra, Gadadhar and
Shyam Roy, Subrata: \href{https://doi.org/10.1016/j.jfa.2018.10.025}{On reducing submodules of {H}ilbert modules with
{${S}_n$}-invariant kernels}, \emph{J. Funct. Anal.} {\bf 276} (2019), 751--784.
\item Boggarapu, Pradeep and Roncal, Luz and Thangavelu, Sundaram: \href{https://doi.org/10.3934/cpaa.2019116}{On extension problem, trace {H}ardy and {H}ardy's inequalities
for some fractional {L}aplacians}, \emph{Commun. Pure Appl. Anal.} {\bf 18} (2019), 2575--2605.
\item Carstensen, Carsten and Dond, Asha K. and Rabus, Hella: \href{https://doi.org/10.1515/cmam-2019-0034}{Quasi-optimality of adaptive mixed {FEM}s for non-selfadjoint
indefinite second-order linear elliptic problems}, \emph{Comput. Methods Appl. Math.} {\bf 19} (2019), 233--250.
\item Chakraborty, Atlanta and Chandru, Vijay and Rao, M. R.: \href{https://doi.org/10.1007/s10479-019-03186-2}{A linear programming primer: from {F}ourier to {K}armarkar}, \emph{Ann. Oper. Res.} {\bf 287} (2020), 593--616.
\item Chandel, Vikramjeet Singh: \href{https://doi.org/10.1007/s12220-018-00123-w}{The 3-point spectral pick interpolation problem and an
application to holomorphic correspondences}, \emph{J. Geom. Anal.} {\bf 30} (2020), 551--572.
\item Choudhury, Projesh Nath and Tsatsomeros, Michael J.: \href{https://doi.org/10.1016/j.laa.2020.04.028}{Algorithmic detection and construction of {N}-matrices}, \emph{Linear Algebra Appl.} {\bf 602} (2020), 46--56.
\item Cicuttin, Matteo and Ern, Alexandre and Gudi, Thirupathi: \href{https://doi.org/10.1007/s10915-020-01195-z}{Hybrid high-order methods for the elliptic obstacle problem}, \emph{J. Sci. Comput.} {\bf 83} (2020), Paper No. 8, 18.
\item Das, Soumya and Jha, Abhash Kumar: \href{https://doi.org/10.1007/s12188-019-00206-7}{Analytic properties of twisted real-analytic {H}ermitian
{K}lingen type {E}isenstein series and applications}, \emph{Abh. Math. Semin. Univ. Hambg.} {\bf 89} (2019), 105--116.
\item Das, Soumya and Pal, Ritwik: \href{https://doi.org/10.1007/s40993-019-0158-x}{The first negative eigenvalue of {Y}oshida lifts}, \emph{Res. Number Theory} {\bf 5} (2019), Paper No. 20, 9.
\item Dond, Asha K. and Gudi, Thirupathi: \href{https://doi.org/10.1002/num.22317}{Patch-wise local projection stabilized finite element methods
for convection-diffusion-reaction problems}, \emph{Numer. Methods Partial Differential Equations} {\bf 35} (2019), 638--663.
\item Dond, Asha K. and Gudi, Thirupathi and Sau, Ramesh CH.: \href{https://doi.org/10.1080/01630563.2018.1538158}{An error analysis of discontinuous finite element methods for
the optimal control problems governed by {S}tokes equation}, \emph{Numer. Funct. Anal. Optim.} {\bf 40} (2019), 421--460.
\item Gadgil, Siddhartha: \href{https://doi.org/10.1007/s10817-019-09523-1}{Homogeneous {L}ength {F}unctions on {G}roups: {I}ntertwined
{C}omputer and {H}uman {P}roofs}, \emph{J. Automat. Reason.} {\bf 64} (2020), 677--688.
\item Garani, Shayan Srinivasa and Seshadri, Harish: \href{https://doi.org/10.1080/17459737.2019.1604845}{An algorithmic approach to {S}outh {I}ndian classical music}, \emph{J. Math. Music} {\bf 13} (2019), 107--134.
\item Gaussier, Herv\'{e} and Seshadri, Harish: \href{https://doi.org/10.1090/proc/14217}{Holomorphic motions and complex geometry}, \emph{Proc. Amer. Math. Soc.} {\bf 147} (2019), 301--313.
\item Ghara, Soumitra and Kumar, Surjit: \href{https://doi.org/10.1016/j.jmaa.2018.09.048}{On the sum of two subnormal kernels}, \emph{J. Math. Anal. Appl.} {\bf 469} (2019), 1015--1027.
\item Gudi, Thirupathi and Majumder, Papri: \href{https://doi.org/10.1515/cmam-2019-0057}{Crouzeix-{R}aviart finite element approximation for the
parabolic obstacle problem}, \emph{Comput. Methods Appl. Math.} {\bf 20} (2020), 273--292.
\item Gudi, Thirupathi and Majumder, Papri: \href{https://doi.org/10.1016/j.cam.2019.02.026}{Convergence analysis of finite element method for a parabolic
obstacle problem}, \emph{J. Comput. Appl. Math.} {\bf 357} (2019), 85--102.
\item Gudi, Thirupathi and Majumder, Papri: \href{https://doi.org/10.1016/j.camwa.2019.06.022}{Conforming and discontinuous {G}alerkin {FEM} in space for
solving parabolic obstacle problem}, \emph{Comput. Math. Appl.} {\bf 78} (2019), 3896--3915.
\item Gupta, Rajeev and Kumar, Surjit and Trivedi, Shailesh: \href{https://doi.org/10.1016/j.jmaa.2020.124032}{Unitary equivalence of operator-valued multishifts}, \emph{J. Math. Anal. Appl.} {\bf 487} (2020), 124032, 23.
\item Gupta, Rajeev and Kumar, Surjit and Trivedi, Shailesh: \href{https://doi.org/10.1090/proc/14410}{Von {N}eumann's inequality for commuting operator-valued
multishifts}, \emph{Proc. Amer. Math. Soc.} {\bf 147} (2019), 2599--2608.
\item Gupta, Subhojoy: \href{https://doi.org/10.1090/tran/7777}{Limits of harmonic maps and crowned hyperbolic surfaces}, \emph{Trans. Amer. Math. Soc.} {\bf 372} (2019), 7573--7596.
\item Gupta, Subhojoy and Mj, Mahan: \href{https://doi.org/10.1090/proc/14866}{Monodromy representations of meromorphic projective
structures}, \emph{Proc. Amer. Math. Soc.} {\bf 148} (2020), 2069--2078.
\item Gupta, Subhojoy and Seshadri, Harish: \href{https://doi.org/10.1093/imrn/rny204}{On domains biholomorphic to {T}eichm\"{u}ller spaces}, \emph{Int. Math. Res. Not. IMRN} {\bf } (2020), 2542--2560.
\item Gupta, Subhojoy and Wolf, Michael: \href{https://doi.org/10.1007/s00208-018-1674-z}{Meromorphic quadratic differentials and measured foliations on
a {R}iemann surface}, \emph{Math. Ann.} {\bf 373} (2019), 73--118.
\item Hatui, Sumana: \href{https://doi.org/10.1515/jgth-2019-0045}{Schur multipliers of special {$p$}-groups of rank 2}, \emph{J. Group Theory} {\bf 23} (2020), 85--95.
\item Hazra, Somnath: \href{https://doi.org/10.1007/s11785-018-0822-5}{Homogeneous 2-shifts}, \emph{Complex Anal. Oper. Theory} {\bf 13} (2019), 1729--1763.
\item Jha, Abhash Kumar and Juyal, Abhishek and Pandey, Manish
Kumar: On simultaneous non-vanishing of twisted {$L$}-functions
associated to newforms on {$\Gamma_0(N)$}, \emph{J. Ramanujan Math. Soc.} {\bf 34} (2019), 245--252.
\item Jha, Abhash Kumar and Sahu, Brundaban: \href{https://doi.org/10.1142/S1793042119500490}{Rankin-{C}ohen brackets on {J}acobi forms of several variables
and special values of certain {D}irichlet series}, \emph{Int. J. Number Theory} {\bf 15} (2019), 925--933.
\item Jha, Abhash Kumar and Sahu, Brundaban: \href{https://doi.org/10.1090/conm/732/14793}{Differential operators on {J}acobi forms and special values of
certain {D}irichlet series}, \emph{Automorphic forms and related topics} {\bf 732} (2019), 91--99.
\item Kashyap, G. A. R. S. R. K. and Bapat, D. and Das, D. and Gowaikar, R. D. and  Amritkar, R. E. and Rangarajan, G. and V. Ravindranath and G. Ambika: Synapse loss and progress of Alzheimers disease - A network model, \emph{Scientific Reports} {\bf 9} (2019).
\item Kataria, Kuldeep Kumar and Vellaisamy, Palaniappan: \href{https://doi.org/10.1080/07362994.2018.1538803}{On the convolution of {M}ittag-{L}effler distributions and its
applications to fractional point processes}, \emph{Stoch. Anal. Appl.} {\bf 37} (2019), 115--122.
\item Khan, Mohammad Yasir Feroz and Biswas, Rahul and Sau, Ramesh
Chandra: Instability of {L}agrange interpolation on {$H^1(\Omega)\cap
C(\overline\Omega)$}, \emph{Math. Student} {\bf 88} (2019), 153--158.
\item Khare, Apoorva and Tikaradze, Akaki: \href{https://doi.org/10.1016/j.laa.2018.05.035}{A {C}arlitz--von {S}taudt type theorem for finite rings}, \emph{Linear Algebra Appl.} {\bf 568} (2019), 106--126.
\item Lin, X. and Kulkarni, P. and Bocci, F. and Schafer, N. and Roy, S. and Tsai, M-Y. and He, Y. and Chen, Y and Rajagopalan, K. and Mooney, S. and Zeng, Y. and Weninger, K. and Grishaev, A. and Onuchic, J. and Levine, H. and Wolynes, P. and Salgia, R. and Rangarajan, G. and Uversky, V. and Orban, J. and Jolly, M. K.: Structural and Dynamical Order of a Disordered Protein: Molecular Insights into Conformational Switching of PAGE4 at the Systems Level, \emph{Biomolecules} {\bf 9} (2019), 77.
\item Mahajan, Prachi and Verma, Kaushal: \href{https://doi.org/10.1142/S0129167X19500125}{A comparison of two biholomorphic invariants}, \emph{Internat. J. Math.} {\bf 30} (2019), 1950012, 16.
\item Maiti, Arun: \href{https://doi.org/10.1016/j.disc.2020.111911}{Quasi-vertex-transitive maps on the plane}, \emph{Discrete Math.} {\bf 343} (2020), 111911, 7.
\item Maitra, Anwoy: \href{https://doi.org/10.1090/proc/15038}{On the continuous extension of {K}obayashi isometries}, \emph{Proc. Amer. Math. Soc.} {\bf 148} (2020), 3437--3451.
\item Maitra, Anwoy: \href{https://doi.org/10.1016/j.jmaa.2019.123694}{A form of {S}chwarz's lemma and a bound for the {K}obayashi
metric on convex domains}, \emph{J. Math. Anal. Appl.} {\bf 484} (2020), 123694, 16.
\item Maity, Dipendu: \href{https://doi.org/10.1007/s13366-018-0405-7}{Pentagonal maps on the torus and the plane}, \emph{Beitr. Algebra Geom.} {\bf 60} (2019), 17--37.
\item Maity, Dipendu and Upadhyay, Ashish Kumar: On centrally symmetric manifolds, \emph{J. Ramanujan Math. Soc.} {\bf 34} (2019), 21--27.
\item Mal, Arpita and Paul, Kallol and Rao, T. S. S. R. K. and Sain,
Debmalya: \href{https://doi.org/10.1007/s00605-019-01289-3}{Approximate {B}irkhoff-{J}ames orthogonality and smoothness in
the space of bounded linear operators}, \emph{Monatsh. Math.} {\bf 190} (2019), 549--558.
\item Mal, Arpita and Sain, Debmalya and Paul, Kallol: \href{https://doi.org/10.1215/17358787-2018-0021}{On some geometric properties of operator spaces}, \emph{Banach J. Math. Anal.} {\bf 13} (2019), 174--191.
\item Manfio, F. and Turgay, N. C. and Upadhyay, A.: \href{https://doi.org/10.1007/s12220-018-9990-9}{Biconservative submanifolds in {$\Bbb{S}^n\times\Bbb{R}$} and
{$\Bbb{H}^n\times\Bbb{R}$}}, \emph{J. Geom. Anal.} {\bf 29} (2019), 283--298.
\item Misra, Gadadhar: Operators in the {C}owen-{D}ouglas class and related topics, \emph{Handbook of analytic operator theory} {\bf } (2019), 87--137.
\item Misra, Gadadhar and Pal, Avijit and Varughese, Cherian: \href{https://doi.org/10.7900/jot.2018jun13.2225}{Contractivity and complete contractivity for finite
dimensional {B}anach spaces}, \emph{J. Operator Theory} {\bf 82} (2019), 23--47.
\item Misra, Gadadhar and Reza, Md. Ramiz: \href{https://doi.org/10.1215/00192082-7768711}{Curvature inequalities and extremal operators}, \emph{Illinois J. Math.} {\bf 63} (2019), 193--217.
\item Mukherjee, Sajal Kumar and Bera, Sudip: \href{https://doi.org/10.1016/j.disc.2019.02.013}{Combinatorial proofs of the {N}ewton-{G}irard and
{C}hapman-{C}ostas-{S}antos identities}, \emph{Discrete Math.} {\bf 342} (2019), 1577--1580.
\item Nikolov, Nikolai and Verma, Kaushal: \href{https://doi.org/10.1007/s12220-019-00237-9}{On the squeezing function and {F}ridman invariants}, \emph{J. Geom. Anal.} {\bf 30} (2020), 1218--1225.
\item Pal, Chandan and Pradhan, Somnath: \href{https://doi.org/10.1080/17442508.2018.1521413}{Risk sensitive control of pure jump processes on a general
state space}, \emph{Stochastics} {\bf 91} (2019), 155--174.
\item Paul, Kallol and Sain, Debmalya: Birkhoff-{J}ames orthogonality and its application in the
study of geometry of {B}anach space, \emph{Advanced topics in mathematical analysis} {\bf } (2019), 245--284.
\item Paul, Kallol and Sain, Debmalya and Ghosh, Puja: Symmetry of {B}irkhoff-{J}ames orthogonality of bounded linear
operators, \emph{Ulam type stability} {\bf } (2019), 331--344.
\item Pingali, Vamsi P.: \href{https://doi.org/10.1080/17476933.2018.1454914}{A note on the deformed {H}ermitian {Y}ang-{M}ills {PDE}}, \emph{Complex Var. Elliptic Equ.} {\bf 64} (2019), 503--518.
\item Pingali, Vamsi Pritham: \href{https://doi.org/10.1016/j.aim.2019.106921}{A vector bundle version of the {M}onge-{A}mp\`ere equation}, \emph{Adv. Math.} {\bf 360} (2020), 106921, 40.
\item Pingali, Vamsi Pritham: \href{https://doi.org/10.1007/s11005-020-01279-9}{Quillen metrics and perturbed equations}, \emph{Lett. Math. Phys.} {\bf 110} (2020), 1861--1875.
\item Poria, Anirudha: \href{https://doi.org/10.1080/01630563.2020.1728771}{Positive weight function and classification of {$g$}-frames}, \emph{Numer. Funct. Anal. Optim.} {\bf 41} (2020), 950--968.
\item Quddus, Safdar: \href{https://doi.org/10.4171/JNCG/320}{Invariants of the {$\Bbb Z_2$} orbifolds of the {P}odle\'{s} two
spheres}, \emph{J. Noncommut. Geom.} {\bf 13} (2019), 257--267.
\item Reddy, Nanda Kishore: \href{https://doi.org/10.1093/imrn/rnx134}{Equality of {L}yapunov and stability exponents for products of
isotropic random matrices}, \emph{Int. Math. Res. Not. IMRN} {\bf } (2019), 606--624.
\item Roth, Julien and Upadhyay, Abhitosh: \href{https://doi.org/10.1007/s00013-019-01315-8}{On compact anisotropic {W}eingarten hypersurfaces in
{E}uclidean space}, \emph{Arch. Math. (Basel)} {\bf 113} (2019), 213--224.
\item Roth, Julien and Upadhyay, Abhitosh: \href{https://doi.org/10.1017/s0004972719000935}{On almost stable {CMC} hypersurfaces in manifolds of bounded
sectional curvature}, \emph{Bull. Aust. Math. Soc.} {\bf 101} (2020), 333--338.
\item Roy, Krishanu and Venkatesh, R.: \href{https://doi.org/10.1007/s00031-018-9510-9}{Maximal closed subroot systems of real affine root systems}, \emph{Transform. Groups} {\bf 24} (2019), 1261--1308.
\item Sain, Debmalya: \href{https://doi.org/10.1215/20088752-2018-0014}{On extreme contractions and the norm attainment set of a
bounded linear operator}, \emph{Ann. Funct. Anal.} {\bf 10} (2019), 135--143.
\item Sain, Debmalya: \href{https://doi.org/10.1007/s13226-020-0393-9}{On the norm attainment set of a bounded linear operator and
semi-inner-products in normed spaces}, \emph{Indian J. Pure Appl. Math.} {\bf 51} (2020), 179--186.
\item Sain, Debmalya: \href{https://doi.org/10.7153/oam-2019-13-32}{Smooth points in operator spaces and some
{B}ishop-{P}helps-{B}ollob\'{a}s type theorems in {B}anach spaces}, \emph{Oper. Matrices} {\bf 13} (2019), 433--445.
\item Sain, Debmalya and Mal, Arpita and Paul, Kallol: \href{https://doi.org/10.1016/j.exmath.2019.01.001}{Some remarks on {B}irkhoff-{J}ames orthogonality of linear
operators}, \emph{Expo. Math.} {\bf 38} (2020), 138--147.
\item Sain, Debmalya and Paul, Kallol and Bhunia, Pintu and Bag,
Santanu: \href{https://doi.org/10.1016/j.laa.2019.04.024}{On the numerical index of polyhedral {B}anach spaces}, \emph{Linear Algebra Appl.} {\bf 577} (2019), 121--133.
\item Sain, Debmalya and Paul, Kallol and Mal, Arpita: On approximate {B}irkhoff-{J}ames orthogonality and normal
cones in a normed space, \emph{J. Convex Anal.} {\bf 26} (2019), 341--351.
\item Sain, Debmalya and Paul, Kallol and Mandal, Kalidas: \href{https://doi.org/10.7153/oam-2019-13-31}{On two extremum problems related to the norm of a bounded
linear operator}, \emph{Oper. Matrices} {\bf 13} (2019), 421--432.
\item Sain, Debmalya and Ray, Anubhab and Paul, Kallol: Extreme contractions on finite-dimensional polygonal {B}anach
spaces, \emph{J. Convex Anal.} {\bf 26} (2019), 877--885.
\item Sanki, Bidyut and Gadgil, Siddhartha: \href{https://doi.org/10.1142/S1793525319500018}{Graphs of systoles on hyperbolic surfaces}, \emph{J. Topol. Anal.} {\bf 11} (2019), 1--20.
\item Sarkar, Amar Deep and Verma, Kaushal: \href{https://doi.org/10.1007/s40315-020-00303-2}{Boundary behaviour of some conformal invariants on planar
domains}, \emph{Comput. Methods Funct. Theory} {\bf 20} (2020), 145--158.
\item Seshadri, Harish: \href{https://doi.org/10.1007/s13226-019-0355-2}{Differential geometry in {I}ndia}, \emph{Indian J. Pure Appl. Math.} {\bf 50} (2019), 795--799.
\item Turgay, Nurettin Cenk and Upadhyay, Abhitosh: \href{https://doi.org/10.1002/mana.201700328}{On biconservative hypersurfaces in 4-dimensional {R}iemannian
space forms}, \emph{Math. Nachr.} {\bf 292} (2019), 905--921.
\item Urbina-Romero, Wilfredo: \href{https://doi.org/10.1007/978-3-030-05597-4}{Gaussian harmonic analysis}, \emph{} {\bf } (2019), xix+477.
\item Verma, Kaushal: \href{https://doi.org/10.1007/s13324-018-0221-0}{Notes on the boundaries of quadrature domains}, \emph{Anal. Math. Phys.} {\bf 9} (2019), 617--638.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan,
Govindan: \href{https://doi.org/10.1063/1.5121245}{On the role of four-wave mixing effect in the interactions
between nonlinear modes of coupled generalized nonlinear
{S}chr\"{o}dinger equation}, \emph{Chaos} {\bf 29} (2019), 123135, 14.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan, Govindan
and Lakshmanan, M.: \href{https://doi.org/10.1016/j.physleta.2018.10.011}{On symmetry preserving and symmetry broken bright, dark and
antidark soliton solutions of nonlocal nonlinear {S}chr\"{o}dinger
equation}, \emph{Phys. Lett. A} {\bf 383} (2019), 15--26.
\end{enumerate}