### 2020-08-10 data/draft-sections.tex

\subsection{Summer Students}

\begin{enumerate}
\item Rohit Kumar, Indian Institute of Science did his/her KVPY summer project under the guidance of R. Venkatesh during May–June 2019
\item Gobinda Debnath, NIT Agartala did his/her IASc-INSA-NASI Focus Area Science Technology Summer Fellowship summer project under the guidance of Ved Datar during May 15, 2019 - July 15, 2020
\item Arnab Roy, Ramakrishna Mission Residential College did his/her Fellowship from the Indian Academy of Sciences summer project under the guidance of Vamsi Pritham Pingali during 26 May - 19 July
\item Ms. Sradha R Bhat, IISER, Thiruvananthapuram did his/her DST Inspire Fellow summer project under the guidance of A. K. Nandakumaran during May 04 – July 04, 2019
\item Ms. Indulekha, IISER, Thiruvananthapuram did his/her May 04 – July 04, 2019 summer project under the guidance of A. K. Nandakumaran during DST Inspire Fellow
\item Ms. Muskan Agarwal, Mount Carmel College, Bangalore did his/her General summer project under the guidance of A. K. Nandakumaran during May 01 – June 09, 2019
\item Ms. Rajashree Dahal, Mount Carmel College, Bangalore did his/her General summer project under the guidance of A. K. Nandakumaran during May 01 – June 09, 2019
\item Mr. Arya Samanta, IISER, Pune did his/her General summer project under the guidance of A. K. Nandakumaran during June-July, 2019
\item Ms Nirmala Aurad, Dept of Mathematics, Gulbarga University did his/her Academy summer project under the guidance of A. K. Nandakumaran during July 03 – August 30, 2019
\item Susheel Shankar, Indian Institute of Science did his/her KVPY summer project under the guidance of Gautam Bharali during May--July, 2019
\item Aritra Das, IISc did his/her KVPY summer project under the guidance of Arvind Ayyer during May-July
\item Mrinmoy Karmakar, IISc did his/her KVPY summer project under the guidance of Arvind Ayyer during May-Jun
\item Omkar Bhalchandra Baraskar, IISc did his/her KVPY summer project under the guidance of Arvind Ayyer during May-Jun
\item Susheel Shankar, IISc did his/her KVPY summer project under the guidance of Arvind Ayyer during May-June
\item Dhanus M Lal, IISc did his/her KVPY summer project under the guidance of Arvind Ayyer during May-June
\item Prathamesh Gajanan Patil, IISc did his/her KVPY summer project under the guidance of Arvind Ayyer during May-June
\item Achal Kumar, IISc did his/her KVPY summer project under the guidance of Arvind Ayyer during May-June
\item Matthew Honnor did his/her  summer project under the guidance of Mahesh Kakde during 1 month
\item Saptarsi Ghosal, CMI did his/her  summer project under the guidance of Manjunath Krishnapur during June-July 2019
\item Deep Ghoshal, ISI Kolkata did his/her  summer project under the guidance of Manjunath Krishnapur during June-July
\item Arghya Chakraborty, ISI Kolkata did his/her  summer project under the guidance of Manjunath Krishnapur during June-July
\item Somak Laha, ISI Kolkata did his/her  summer project under the guidance of Manjunath Krishnapur during June-July
\item Sayak Chatterjee did his/her  summer project under the guidance of Manjunath Krishnapur during June-July
\item Anirban Basak, IISER Kolkata did his/her  summer project under the guidance of Manjunath Krishnapur during June-July
\item Nafia Saleem, IIT Madras did his/her SRFP of IAS summer project under the guidance of Manjunath Krishnapur during June-July
\item Rohit Kumar, IISc did his/her KVPY summer project under the guidance of R VENKATESH during May-June 2019
\item Aman Agrawal, IISc did his/her  summer project under the guidance of Apoorva Khare during May-July 2019
\item Sutanay Bhattacharya, IISc did his/her  summer project under the guidance of Apoorva Khare during May-July 2019
\item Samarth Hawaldar, IISc did his/her  summer project under the guidance of Apoorva Khare during May-July 2019
\item Swaroop Hegde, NISER Bhubaneswar did his/her  summer project under the guidance of Apoorva Khare during May-July 2019
\item Anant Mudgal, CMI did his/her  summer project under the guidance of Apoorva Khare during May-July 2019
\item Abhilash Mukherjee, IISc did his/her  summer project under the guidance of Apoorva Khare during May-July 2019
\item Pulkit Sinha, IISc did his/her  summer project under the guidance of Apoorva Khare during May-July 2019
\item Chinmaya Kausik, IISc did his/her KVPY summer project under the guidance of Siddhartha Gadgil during May-June 2019
\item Nilendu Das, IISER Mohali did his/her Academy summer programme summer project under the guidance of Siddhartha Gadgil during May, June 2019
\item Nishit Pandya, IISc did his/her KVPY summer project under the guidance of Siddhartha Gadgil during May, June 2019
\item Julian D'Costa, IISc did his/her KVPY summer project under the guidance of Siddhartha Gadgil during May, June 2019
\end{enumerate}

\subsection{List of Visitors to the Department}

\begin{enumerate}
\item Dr.~Malabika Pramanik, University of British Columbia, Vancouver, Canada, August 2, 2017--August 2, 2017.
\item Dr.~Jugal Verma, IIT Bombay, September 5, 2017--September 5, 2017.
\item Dr.~Anirban Basak, Weizmann Institute, Rehovot, Israel, October 5, 2017--October 8, 2017.
\item Dr.~Neeraj Kayal, Microsoft Research, Bangalore, October 6, 2017--October 6, 2017.
\item Dr.~Nishant Chandgotia, Tel Aviv University, Israel, October 7, 2017--October 10, 2017.
\item Dr.~Krishanu Roy, IMSc, Chennai, November 20, 2017--November 24, 2017.
\item Dr.~Patrice Philippon, CNRS, Paris, France, November 22, 2017--November 22, 2017.
\item Dr.~Kyoung-Seog Lee, Center for Geometry and Physics, Institute for Basic Science (IBS) Pohang, Republic of Korea, December 4, 2017--December 4, 2017.
\item Dr.~Ved Datar, University of California, Berkeley, USA, January 3, 2018--January 3, 2018.
\item Dr.~Moumanti Podder, Georgia Institute of Technology, Atlanta, USA, January 7, 2018--January 12, 2018.
\item Dr.~Augusto Visintin, Universita' degli Studi di Trento, Italy, January 10, 2018--January 13, 2018.
\item Dr.~Alexander Bufetov, CNRS, Marseilles, France; and Steklov Mathematical Institute, Moscow, Russia, January 11, 2018--January 20, 2018.
\item Dr.~Shreedevi K. Masuti, IMSc, Chennai, January 11, 2018--January 11, 2018.
\item Dr.~Jayadev Athreya, University of Washington, Seattle, USA, January 17, 2018--January 17, 2018.
\item Dr.~Krishna B. Athreya, Iowa State University, Ames, USA, January 19, 2018--January 19, 2018.
\item Dr.~Pranav Pandit, IHES, Bures-sur-Yvette, France, January 31, 2018--January 31, 2018.
\item Dr.~Svante Linusson, KTH, Stockholm, Sweden, February 3, 2018--February 11, 2018.
\item Dr.~Samu Potka, KTH, Stockholm, Sweden, February 3, 2018--February 11, 2018.
\item Dr.~Dominique Guillot, University of Delaware, Newark, USA, February 4, 2018--February 16, 2018.
\item Dr.~Aprameyan Parthasarathy, Universität Paderborn, Germany, March 1, 2018--April 27, 2018.
\item Dr.~Mahesh Kakde, King's College, London, UK, March 2, 2018--March 2, 2018.
\item Dr.~Arnab Saha, Australian National University, Canberra, Australia, March 4, 2018--March 10, 2018.
\item Dr.~Koushik Ramachandran, Oklahoma State University, Stillwater, USA, March 6, 2018--March 7, 2018.
\item Dr.~Vikas Krishnamurthy, Schrodinger Institute, Vienna, Austria, March 6, 2018--March 6, 2018.
\item Dr.~Subhajit Goswami, IHES, Bures-sur-Yvette, France, March 19, 2018--March 19, 2018.
\item Dr.~Sarang Sane, IIT Madras, March 23, 2018--March 23, 2018.
\item Dr.~Siegfried Böcherer, Universität Mannheim, Germany, March 24, 2018--April 8, 2018.
\item Dr.~Ananthnarayan Hariharan, IIT Bombay, April 6, 2018--April 7, 2018.
\item Dr.~C. S. Aravinda, TIFR-CAM, Bangalore, April 6, 2018--April 6, 2018.
\item Dr.~Bidyut Sanki, IMSc, Chennai, April 24, 2018--April 29, 2018.
\item Dr.~Satyajit Guin, IISER Mohali, April 25, 2018--April 25, 2018.
\item Dr.~Prashant Jha, Louisiana State University, Baton Rouge, USA, May 1, 2018--May 1, 2018.
\item Dr.~Ramesh Kasilingam, ISI Bangalore, May 4, 2018--May 4, 2018.
\item Dr.~Radhika Ganapathy, TIFR, Mumbai, May 24, 2018--May 24, 2018.
\item Dr.~Luz Roncal, BCAM, Bilbao, Spain, June 1, 2018--August 31, 2018.
\item Dr.~Terrence George, Brown University, Providence, USA, June 2, 2018--June 30, 2018.
\item Dr.~Aprameyan Parthasarathy, Universität Paderborn, Germany, June 3, 2018--June 23, 2018.
\item Dr.~Sudhir Ghorpade, IIT Bombay, June 4, 2018--June 4, 2018.
\item Dr.~Yifei Chen, Chinese Academy of Sciences, Beijing, China, June 5, 2018--June 12, 2018.
\item Dr.~Rajeeva L. Karandikar, CMI, Chennai, June 27, 2018--June 28, 2018.
\item Dr.~Soumyashant Nayak, University of Pennsylvania, Philadelphia, USA, June 27, 2018--June 27, 2018.
\item Dr.~M. S. Raghunathan, CBS, Mumbai, June 28, 2018--June 29, 2018.
\item Dr.~Shubham Sinha, UCSD, San Diego, USA, July 1, 2018--July 30, 2018.
\item Dr.~Terrence George, Brown University, Providence, USA, July 2, 2018--August 8, 2018.
\item Dr.~Anand Deopurkar, Australian National University, Canberra, Australia, July 2, 2018--July 4, 2018.
\item Dr.~Asilata Bapat, Australian National University, Canberra, Australia, July 2, 2018--July 4, 2018.
\item Dr.~Mizanur Rahaman, University of Waterloo, Canada, July 9, 2018--July 11, 2018.
\item Dr.~Kunal Datta, INRIA, Sophia-Antipolis, France, July 23, 2018--July 25, 2018.
\item Dr.~Sumana Hatui, HRI, Allahabad, July 23, 2018--July 23, 2018.
\item Dr.~Neha Prabhu, Queen's University, Kingston, Canada, July 23, 2018--July 23, 2018.
\item Dr.~Bhimarthi Ravinder, CMI, Chennai, July 26, 2018--July 27, 2018.
\item Dr.~Tulasi Ram Reddy, NYU Abu Dhabi, UAE, July 31, 2018--August 12, 2018.
\item Dr.~Ravi Raghunathan, IIT Bombay, August 3, 2018--August 3, 2018.
\item Dr.~Jayadev Athreya, University of Washington, Seattle, USA, August 10, 2018--August 10, 2018.
\item Dr.~Stephan Baier, RKMVU, Belur, August 26, 2018--August 27, 2018.
\item Dr.~A. Raghuram, IISER Pune, September 24, 2018--September 26, 2018.
\item Dr.~Bharatram Rangarajan, Hebrew University, Jerusalem, Israel, September 26, 2018--September 26, 2018.
\item Dr.~Amritanshu Prasad, IMSc, Chennai, September 27, 2018--September 28, 2018.
\item Dr.~Jean Baptiste Gatsinzi, Botswana University of Science and Technology, Botswana, October 1, 2018--October 31, 2018.
\item Dr.~Dootika Vats, University of Warwick, UK, October 3, 2018--October 4, 2018.
\item Dr.~Ashutosh Kumar, NUS, Singapore, October 9, 2018--October 12, 2018.
\item Dr.~Dhruv Ranganathan, IAS Princeton, MIT Boston (USA); CMI (Chennai); Cambridge (UK), October 29, 2018--November 2, 2018.
\item Dr.~K. N. Raghavan, IMSc, Chennai, November 9, 2018--November 9, 2018.
\item Dr.~Daniel Seco, ICMAT, Madrid, Spain, November 13, 2018--November 14, 2018.
\item Dr.~Purvi Gupta, Rutgers University, New Brunswick, USA, November 18, 2018--November 23, 2018.
\item Dr.~Anuradha Sharma, IIIT Delhi, December 3, 2018--December 3, 2018.
\item Dr.~Harald Upmeier, Universität Marburg, Germany, December 4, 2018--February 28, 2019.
\item Dr.~Sameer Chavan, IIT Kanpur, December 10, 2018--December 13, 2018.
\item Dr.~Katharina Jochemko, KTH, Stockholm, Sweden, December 11, 2018--December 14, 2018.
\item Dr.~Alexander Belton, University of Lancaster, UK, December 12, 2018--December 21, 2018.
\item Dr.~Gopal Srinivasan, IIT Bombay, December 13, 2018--December 13, 2018.
\item Dr.~Thorsten Heidersdorf, Max Planck Institut, Bonn, Germany, December 18, 2018--December 20, 2018.
\item Dr.~Rongwei Yang, SUNY Albany, USA, December 19, 2018--December 21, 2018.
\item Dr.~Prasad Tetali, Georgia Institute of Technology, Atlanta, USA, January 1, 2019--January 1, 2019.
\item Dr.~T. V. H. Prathamesh, University of Innsbruck, Austria, January 3, 2019--January 4, 2019.
\item Dr.~Oscar Ciaurri, Universidad de la Rioja, Logronyo, Spain, January 7, 2019--January 25, 2019.
\item Dr.~Dominique Guillot, University of Delaware, Newark, USA, January 10, 2019--January 17, 2019.
\item Dr.~Ramesh Gangolli, University of Washington, Seattle, USA, January 13, 2019--January 26, 2019.
\item Dr.~Sanjay Ramassamy, École Normale Supériure, Paris, France, January 13, 2019--January 26, 2019.
\item Dr.~Hiroshige Shiga, Tokyo Institute of Technology, Japan, January 17, 2019--January 26, 2019.
\item Dr.~Pranav Pandit, ICTS Bangalore, January 21, 2019--January 21, 2019.
\item Dr.~Niranjan Balachandran, IIT Bombay, January 29, 2019--January 30, 2019.
\item Dr.~Siegfried Böcherer, Universität Mannheim, Germany, January 31, 2019--February 10, 2019.
\item Dr.~Mohan R., ISI Bangalore, February 8, 2019--February 8, 2019.
\item Dr.~Nishant Chandgotia, Hebrew University, Jerusalem, Israel, February 11, 2019--February 12, 2019.
\item Dr.~Arpan Kabiraj, CMI, Chennai, February 11, 2019--February 11, 2019.
\item Dr.~Nishad Kothari, University of Campinas, São Paulo, Brazil, February 13, 2019--February 13, 2019.
\item Dr.~James Cann, University College London, UK, February 15, 2019--March 10, 2019.
\item Dr.~Klas Pettersson, University of Tromsø – The Arctic University of Norway, February 26, 2019--February 26, 2019.
\item Dr.~Parthanil Roy, ISI Bangalore, March 1, 2019--March 1, 2019.
\item Dr.~Xavier Viennot, CNRS and LaBRI, France, March 6, 2019--March 9, 2019.
\item Dr.~Agnid Banerjee, TIFR-CAM, Bangalore, March 8, 2019--March 8, 2019.
\item Dr.~Ravi A. Rao, TIFR, Mumbai, March 9, 2019--March 15, 2019.
\item Dr.~Soma Maity, IISER Kolkata, March 10, 2019--March 12, 2019.
\item Dr.~Agnid Banerjee, TIFR-CAM, Bangalore, March 15, 2019--March 15, 2019.
\item Dr.~Soumya Bhattachrarya, IISER Kolkata, March 17, 2019--March 23, 2019.
\item Dr.~Sayan Chakraborty, ISI Kolkata, April 8, 2019--April 8, 2019.
\item Dr.~Jyotirmoy Ganguly, IISER Pune, April 14, 2019--April 19, 2019.
\item Dr.~Siegfried Böcherer, Universität Mannheim, Germany, April 22, 2019--April 22, 2019.
\item Dr.~Indrava Roy, IMSc, Chennai, April 28, 2019--April 30, 2019.
\item Dr.~Vineeth Chintala, IIT Bombay, May 1, 2019--May 1, 2019.
\item Dr.~Sankaran Viswanath, IMSc, Chennai, May 3, 2019--May 7, 2019.
\item Dr.~Amritanshu Prasad, IMSc, Chennai, May 8, 2019--May 8, 2019.
\item Dr.~Sudipto Choudhury, IIT Bombay, May 20, 2019--May 22, 2019.
\item Dr.~Siddhartha Sahi, Rutgers University, New Brunswick, USA, June 2, 2019--June 6, 2019.
\item Dr.~Sanjay P. K., NIT Calicut, June 5, 2019--June 30, 2019.
\item Dr.~Sanjay Parui, NISER, Bhubaneswar, June 9, 2019--June 24, 2019.
\item Dr.~Sasmita Patnaik, IIT Kanpur, June 23, 2019--June 28, 2019.
\item Dr.~Shubham Sinha, UCSD, San Diego, USA, June 24, 2019--July 2, 2019.
\item Dr.~Subhajit Goswami, IHES, Bures-sur-Yvette, France, July 24, 2019--July 27, 2019.
\item Dr.~Marc Hubert-Nicole, University of Marseilles, France, July 28, 2019--September 2, 2019.
\item Dr.~Rakesh, University of Delaware, Newark, USA, August 2, 2019--August 2, 2019.
\item Dr.~Terrence George, Brown University, Providence, USA, August 7, 2019--August 7, 2019.
\item Dr.~Akashdeep Dey, Princeton University, Princeton, USA, August 12, 2019--August 16, 2019.
\item Dr.~Hariharan Narayanan, TIFR, Mumbai, August 14, 2019--August 14, 2019.
\item Dr.~Piyush Srivastava, TIFR, Mumbai, August 14, 2019--August 14, 2019.
\item Dr.~Nikhil Srivastava, University of California, Berkeley, USA, August 16, 2019--August 16, 2019.
\item Dr.~Mahan Mj, TIFR, Mumbai, August 25, 2019--August 31, 2019.
\item Dr.~T. V. H. Prathamesh, KREA University, Sricity, August 28, 2019--August 30, 2019.
\item Dr.~Soumyashant Nayak, University of Pennsylvania, Philadelphia, USA, August 28, 2019--August 29, 2019.
\item Dr.~Soumik Pal, University of Washington, Seattle, USA, August 30, 2019--August 30, 2019.
\item Dr.~Chandan Dalawat, HRI, Allahabad, September 2, 2019--September 6, 2019.
\item Dr.~Guhan Venkat, Université Laval, Quebec, Canada; Morningside Center of Mathematics, Beijing, China, September 4, 2019--September 7, 2019.
\item Dr.~Senthil Raani, IISER Berhampur, September 6, 2019--September 11, 2019.
\item Dr.~Jean-Marie de Koninck, Université Laval, Quebec, Canada, September 11, 2019--September 13, 2019.
\item Dr.~Rekha Biswal, Max Planck Institut, Bonn, Germany, September 13, 2019--September 17, 2019.
\item Dr.~Prasad Tetali, Georgia Institute of Technology, Atlanta, USA, September 13, 2019--September 13, 2019.
\item Dr.~Kartick Adhikari, Technion, Haifa, Israel, September 15, 2019--September 20, 2019.
\item Dr.~Martin Kreuzer, Universität Passau, Germany, September 16, 2019--October 1, 2019.
\item Dr.~Anantha Padmanabha, IMSc, Chennai, September 18, 2019--September 20, 2019.
\item Dr.~Santosh Nadimpalli, Radboud University, Nijmegen, Netherlands, September 19, 2019--September 21, 2019.
\item Dr.~Divyang Bhimani, TIFR-CAM, Bangalore, September 20, 2019--September 20, 2019.
\item Dr.~Sushmita Venugopalan, IMSc, Chennai, September 21, 2019--September 29, 2019.
\item Dr.~Rajeeva L. Karandikar, CMI, Chennai, September 27, 2019--September 27, 2019.
\item Dr.~Gianluca Faraco, TIFR, Mumbai, October 7, 2019--October 10, 2019.
\item Dr.~Atul Dixit, IIT Gandhinagar, October 9, 2019--October 11, 2019.
\item Dr.~Aswin Balasubramanian, Rutgers University, New Brunswick, USA, October 14, 2019--October 14, 2019.
\item Dr.~Uri Onn, Australian National University, Canberra, Australia, October 23, 2019--October 30, 2019.
\item Dr.~Souvik Pal, HRI, Allahabad, October 24, 2019--November 2, 2019.
\item Dr.~Sorin Dumitrescu, Université Côte d'Azur, Nice, October 26, 2019--October 30, 2019.
\item Dr.~Indranil Biswas, TIFR Mumbai, October 26, 2019--October 30, 2019.
\item Dr.~Ramesh Manna, TIFR-CAM, Bangalore, October 30, 2019--October 30, 2019.
\item Dr.~Tathagata Basak, Iowa State University, Ames, USA, November 4, 2019--November 4, 2019.
\item Dr.~Apratim Chakraborty, ISI Bangalore, November 4, 2019--November 4, 2019.
\item Dr.~Chandrasheel Bhagwat, IISER Pune, November 6, 2019--November 6, 2019.
\item Dr.~Klas Pettersson, UiT – The Arctic University of Norway, Narvik, Norway, November 6, 2019--November 6, 2019.
\item Dr.~Harald Upmeier, Universität Marburg, Germany, November 7, 2019--March 15, 2020.
\item Dr.~Uri Onn, Australian National University, Canberra, Australia, November 8, 2019--November 9, 2019.
\item Dr.~Bernhard Heim, German University of Technology, Oman, November 14, 2019--November 14, 2019.
\item Dr.~Luz Roncal, BCAM, Bilbao, Spain, November 21, 2019--December 7, 2019.
\item Dr.~Sugata Mondal, TIFR, Mumbai, November 21, 2019--November 21, 2019.
\item Dr.~Pavol Hell, Simon Fraser University, Canada, November 22, 2019--November 22, 2019.
\item Dr.~Gianluca Faraco, TIFR, Mumbai, November 26, 2019--December 4, 2019.
\item Dr.~Indranath Sengupta, IIT Gandhinagar, November 27, 2019--November 29, 2019.
\item Dr.~Vyjayanthi Chari, University of California, Riverside, USA, December 1, 2019--December 31, 2019.
\item Dr.~John Meakin, University of Nebraska, Lincoln, USA, December 3, 2019--December 4, 2019.
\item Dr.~Anne Schilling, University of California, Davis, USA, December 14, 2019--December 16, 2019.
\item Dr.~Amritanshu Prasad, IMSc, Chennai, December 14, 2019--December 16, 2019.
\item Dr.~Sankaran Viswanath, IMSc, Chennai, December 14, 2019--December 16, 2019.
\item Dr.~K. N. Raghavan, IMSc, Chennai, December 14, 2019--December 16, 2019.
\item Dr.~Tanushree Khandai, IISER Mohali, December 14, 2019--December 16, 2019.
\item Dr.~Bhimarthi Ravinder, IIT Tirupati, December 14, 2019--December 16, 2019.
\item Dr.~Sachin Sharma, IIT Kanpur, December 14, 2019--December 16, 2019.
\item Dr.~Sunil Chhita, Durham University, December 16, 2019--December 19, 2019.
\item Dr.~Ken Ono, University of Virginia, Charlottesville, USA, December 18, 2019--December 18, 2019.
\item Dr.~Sujatha Ramdorai, University of British Columbia, Vancouver, Canada, December 20, 2019--December 20, 2019.
\item Dr.~Akshaa Vatwani, IIT, Gandhinagar, February 26, 2020--February 26, 2020.
\item Dr.~Arijit Chakrabarty, ISI, Kolkata, February 26, 2020--.
\item Dr.~Debapratim Banerjee, ISI, Kolkata, March 9, 2020--March 11, 2020.
\end{enumerate}

\section{Resource Generation}

\begin{enumerate}

\item {\bf A. K. Nandakumaran\,:} Sole investigator in a DST (Matrix) project
\begin{description}
  \item[Project Title\,:] Unfolding Operators to Study Homogenization Problems in Domains with Highly Oscillating Structures
  \item[Duration\,:] 2018-2021
  \item[Grant\,:] 6.6 lakhs
\end{description}

\item {\bf A. K. Nandakumaran\,:} Sole investigator in a DST project
\begin{description}
  \item[Project Title\,:] Homogenization of Partial Differential Equations in Certain Complex Domains
  \item[Duration\,:] 2017-2020
  \item[Grant\,:] 19.316 Lakhs
\end{description}

\item {\bf Abhishek Banerjee\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] MATRICS Grant : MTR/2017/000112("Algebraic Geometry over symmetric monoidal categories and tensor triangular geometry")
  \item[Duration\,:] 2018-2021
  \item[Grant\,:] Rs 6.6 lakh
\end{description}

\item {\bf Apoorva Khare\,:} Sole investigator in a Infosys Foundation project
\begin{description}
  \item[Project Title\,:] Young Investigator Award
  \item[Duration\,:] 2017-2019
  \item[Grant\,:] 12 lakhs
\end{description}

\item {\bf Apoorva Khare\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] Ramanujan Fellowship
  \item[Duration\,:] 2018-2023
  \item[Grant\,:] 35 lakhs
\end{description}

\item {\bf Apoorva Khare\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] MATRICS: Entrywise Powers and Power Series Preserving Positivity in Fixed Dimension
  \item[Duration\,:] 2018-2021
  \item[Grant\,:] 6.6 lakhs
\end{description}

\item {\bf Apoorva Khare\,:} Sole investigator in a DST and SERB project
\begin{description}
  \item[Project Title\,:] SwarnaJayanti Fellowship: Transformations preserving positivity in analysis, and connections to algebra and combinatorics
  \item[Duration\,:] 2020-2025
  \item[Grant\,:] Rs. 1,24,88,260
\end{description}

\item {\bf Arvind Ayyer\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] Exactly solvable exclusion processes
  \item[Duration\,:] 2017--2020
  \item[Grant\,:] 19,30,490
\end{description}

\item {\bf Arvind Ayyer\,:} Co-investigator in a The Royal Society, UK project
\begin{description}
  \item[Project Title\,:] Determinantal Formulas for Multispecies ASEP
  \item[Duration\,:] 2019--2021
  \item[Grant\,:] 11,759 pounds
\end{description}

\item {\bf Basudeb Datta\,:} Sole investigator in a SERB, DST project
\begin{description}
  \item[Project Title\,:] Minimal and Tight Triangulations of some PL Manifolds
  \item[Duration\,:] 2018--2021
  \item[Grant\,:] 660,000
\end{description}

\item {\bf Dilip P. Patil\,:} Sole investigator in a DST-SERB project
\begin{description}
  \item[Project Title\,:] Burnside Algebras-Computational Aspects of Commutative Algebra and Algebraic Geometry
  \item[Duration\,:] 2018 to 2021
  \item[Grant\,:] 2 Lakhs per year
\end{description}

\item {\bf E K Narayanan\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] Heckman-Opdam hypergeometric functions and harmonic analysis
  \item[Duration\,:] 2019-2022
  \item[Grant\,:] Six lakhs
\end{description}

\item {\bf Gadadhar Misra\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] The role of curvature in Operator Theory
  \item[Duration\,:] 2018 - 2021
  \item[Grant\,:] 660000
\end{description}

\item {\bf Gadadhar Misra\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] J C Bose Fellowship
  \item[Duration\,:] 2018 - 2021
  \item[Grant\,:] 
\end{description}

\item {\bf Gautam Bharali\,:} Sole investigator in a DST project
\begin{description}
  \item[Project Title\,:] Swarnajayanti Fellowship
  \item[Duration\,:] 2015--2020
  \item[Grant\,:] Rs. 56.7 lakhs
\end{description}

\item {\bf Govindan Rangarajan\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] JC Bose Fellowship
  \item[Duration\,:] 2011-present
  \item[Grant\,:] 80,00,000
\end{description}

\item {\bf Govindan Rangarajan\,:} Sole investigator in a Broadcom Corporation project
\begin{description}
  \item[Project Title\,:] Brain Inspired Computing
  \item[Duration\,:] 2017-present
  \item[Grant\,:] Rs. 1,37,00,000
\end{description}

\item {\bf Govindan Rangarajan\,:} Sole investigator in a DST project
\begin{description}
  \item[Project Title\,:] Indo-French Centre for Applied Mathematics Phase II
  \item[Duration\,:] 2018-present
  \item[Grant\,:] 1,03,77,511
\end{description}

\item {\bf Govindan Rangarajan\,:} Principal investigator in a CNRS, France project
\begin{description}
  \item[Project Title\,:] Indo-French Centre for Applied Mathematics Phase II
  \item[Duration\,:] 2018-present
  \item[Grant\,:] Euros 1,68,000 (Rs. 1,34,40,000)
\end{description}

\item {\bf Manjunath Krishnapur\,:} Principal investigator in a SERB project
\begin{description}
  \item[Project Title\,:] Zeros of random.functions and anti-concentration phenomena
  \item[Duration\,:] 2018-21
  \item[Grant\,:] 200000
\end{description}

\item {\bf R VENKATESH\,:} Sole investigator in a DST-SERB (MATRICS) project
\begin{description}
  \item[Project Title\,:] Study on the structure theory of Borcherds-Kac-Moody algebras and their subalgebras
  \item[Duration\,:] 2019-2021
  \item[Grant\,:] 6,60,000/-
\end{description}

\item {\bf R VENKATESH\,:} Sole investigator in a DST project
\begin{description}
  \item[Project Title\,:] INSPIRE Faculty award
  \item[Duration\,:] 2016-2021
  \item[Grant\,:] 35,00,000/-
\end{description}

\item {\bf R VENKATESH\,:} Sole investigator in a Infosys foundation project
\begin{description}
  \item[Project Title\,:] Infosys Young Investigator award
  \item[Duration\,:] 2020-2021
  \item[Grant\,:] 6,00,000/-
\end{description}

\item {\bf R. Venkatesh\,:} Sole investigator in a SERB-MATRICS project
\begin{description}
  \item[Project Title\,:] Study on the structure theory of Borcherds-Kac-Moody algebras and their subalgebras.
  \item[Duration\,:] 2019-2021
  \item[Grant\,:] 6 Lakhs
\end{description}

\item {\bf R. Venkatesh\,:} Sole investigator in a DST-INPIRE project
\begin{description}
  \item[Project Title\,:] Inspire faculty award
  \item[Duration\,:] 2016-2021
  \item[Grant\,:] 35 Lakhs
\end{description}

\item {\bf R. Venkatesh\,:} Sole investigator in a Infosys foundation project
\begin{description}
  \item[Project Title\,:] Infosys Young Investigator award
  \item[Duration\,:] 2020-2021
  \item[Grant\,:] 6 Lakhs
\end{description}

\item {\bf RADHIKA GANAPATHY\,:} Sole investigator in a Infosys Foundation project
\begin{description}
  \item[Project Title\,:] Young Investigator
  \item[Duration\,:] 2019-2021
  \item[Grant\,:] Rs. 3,00,000 p.a.+ Rs. 25,000 monthly salary top-up
\end{description}

\item {\bf SANCHAYAN SEN\,:} Sole investigator in a Science \& Engineering Research Board (SERB) project
\begin{description}
  \item[Project Title\,:] Geometry of random discrete structures
  \item[Duration\,:] 2020--2023
  \item[Grant\,:] 220,000 per annum
\end{description}

\item {\bf SANCHAYAN SEN\,:} Sole investigator in a Infosys Foundation, Bangalore project
\begin{description}
  \item[Project Title\,:] 
  \item[Duration\,:] 2019--2021
  \item[Grant\,:] 600,000 per annum
\end{description}

\item {\bf Siddhartha Gadgil\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] Homotopy Type Theory and Natural Language Processing for Computer-Assisted mathematics
  \item[Duration\,:] 2018-2021
  \item[Grant\,:] 1898700
\end{description}

\item {\bf Srikanth Krishnan Iyer\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] Percolation at Criticality and Connectivity in the Random Connection Model
  \item[Duration\,:] March 22, 2019 - March 22, 2022
  \item[Grant\,:] 6,60,000
\end{description}

\item {\bf Subhojoy Gupta\,:} Sole investigator in a SERB, DST project
\begin{description}
  \item[Project Title\,:] Degenerations of geometric structures and meromorphic differentials on Riemann surfaces
  \item[Duration\,:] 2018-2021
  \item[Grant\,:] 6 lakhs
\end{description}

\item {\bf Subhojoy Gupta\,:} Co-investigator in a Université Côte-D'azur and IISc project
\begin{description}
  \item[Project Title\,:] Branched projective structures
  \item[Duration\,:] 2019-2021
  \item[Grant\,:] 6 lakhs
\end{description}

\item {\bf Sundaram Thangavelu\,:} Sole investigator in a DST project
\begin{description}
  \item[Project Title\,:] J.C.Bose Fellowship
  \item[Duration\,:] 2008-2022
  \item[Grant\,:] 
\end{description}

\item {\bf Thirupathi Gudi\,:} Sole investigator in a DST project
\begin{description}
  \item[Project Title\,:] Approximation Theory for PDEconstrained optimization and Variational Inequalities
  \item[Duration\,:] 2019-2022
  \item[Grant\,:] 600000
\end{description}

\item {\bf Vamsi Pritham Pingali\,:} Sole investigator in a SERB project
\begin{description}
  \item[Project Title\,:] A study of some analytic and computational aspects of Chern-Weil theory
  \item[Duration\,:] 2017-2020
  \item[Grant\,:] 542290
\end{description}

\item {\bf Vamsi Pritham Pingali\,:} Sole investigator in a Infosys Foundation project
\begin{description}
  \item[Project Title\,:] Infosys Young Investigator Award
  \item[Duration\,:] 2017-2019
  \item[Grant\,:] 600000
\end{description}

\end{enumerate}

\subsection{Awards}

\begin{enumerate}
\item Apoorva Khare was awarded (1) Delivered a Plenary talk at the 2019 ILAS (International Linear Algebra Society) meeting in Rio de Janeiro, Brazil. Also selected as the LAMA Lecturer. (2) Delivered the 20th Professor Wazir Hasan Abdi Memorial Lecture, in Cochin University of Science and Technology (CUSAT). (3) Awarded the SwarnaJayanti Fellowship for 2018-19, in Mathematics.
\item Arvind Ayyer was awarded My student, Surjadipta De Sarkar received a fellowship and travel grant to attend a program on Algebraic and Enumerative Combinatorics at the Mittag-Leffler Institute from 13 January - 30 April 2020.
\item R VENKATESH was awarded Infosys Young Investigator award from the Infosys foundation, 2020.
\item R. Venkatesh was awarded Associate of the Indian Academy of Sciences (2018-2022), Infosys Young Investigator award from the Infosys foundation, 2020.
\item Ved Datar was awarded Infosys Young Investigator Award
\end{enumerate}

\subsection{Fellowships of Academies}

\subsection{Editorial boards}
\begin{enumerate}
\item Apoorva Khare, Journals: (1) Indian Journal of Pure and Applied Mathematics (2) Linear and Multilinear Algebra (3) Sao Paulo Journal of Mathematical Sciences. Book Series: (4) TRIM (Texts and Readings in Mathematics)
\item Arvind Ayyer, Indian Journal of Pure and Applied Mathematics, Algebraic Combinatorics
\item E K Narayanan, Editorial board of the Indian Journal of Pure and Applied Mathematics
\item Gautam Bharali, Newsletter of the Ramanujan Mathematical Society
\item Mrinal Kanti Ghosh, Associate Editor, Stochastic Analysis and Applications
\item Sundaram Thangavelu, Proceedings of the Indian Academy of Sciences (Math.), Journal of Ramanujan Mathematical Society, Journal of Analysis, Tunisian Journal of Mathematics, Journal of Fourier Analysis and Applications.
\end{enumerate}

\subsection{Other Activities: 2018-19}
\begin{enumerate}
\item A. K. Nandakumaran was Member, Core Committee, KVPY
\item A. K. Nandakumaran was Member, Library Committee
\item Apoorva Khare was Talks: (1) 2019-04-12: Eigenfunctions Seminar, titled "Schoenberg: from metric geometry to matrix positivity" (2) 2019-11-06: 1st Math Symposium of Infosys Young Investigators: Talk, titled "Groups with norms; distance matrices and Zariski density"
\item Apoorva Khare was Organizer of Eigenfunctions,  APRG Seminar, Algebra-Combinatorics Seminar, (co-organiser)
\item Apoorva Khare was Contributed in preparing the PhD Qualifiers in Algebra
\item Apoorva Khare was Student thesis/project supervision: (1) Swarnalipa Datta, MS Thesis (2) Soumik Dutta, BS Project
\item Arvind Ayyer was Organised ICTS workshop on "Universality in random structures: Interfaces, Matrices, Sandpiles" from 14/01 -08/02/2019 jointly with Riddhipratim Basu and Manjunath Krishnapur.
\item Basudeb Datta was B. Datta was an member of the Committees for the following: (i)  Promotion of Professor in Mathematics in HAG scale in NIT, Rourkela in November 2018, (ii) Faculty selection for the post of Associate and Assistant Professors in the Department of Mathematics of IIT, Delhi in May, 2019, (iii) Faculty selection in mathematics at NISER, Bhubaneswar in June, 2019.
\item Basudeb Datta was B. Datta was  examiner for the following PhD theses:  (i) ``Dynamics of maps possessing the almost average shadowing property and variants of transitivity" of Mukta Garg, Univ of Delhi in 2018, (ii) ``On some problems in Algebraic Combinatorics" of Sajal K. Mukherjee, Visva-Bharati University, Santiniketan in 2018 (also the external examiner for the viva-voce examination  in June 2019), (iii) ``Group actions on Dold and Milnor manifolds" of Pinka Dey, IISER Mohali in 2019 (also the external examiner for the viva-voce examination in December 2019), (iv) ``Variants of transitivity, sensitivity and expansivity in non-autonomous discrete systems" of Radhika Vashisht,  Univ. of Delhi in 2020.
\item Dilip P. Patil was Taught the course: MA 312 Commutative Algebra course during Jan-April 2020
\item Dilip P. Patil was Taught  (12 Week long)  NPTEL Course during Jan-April 2020
\item Gadadhar Misra was Convener, Symposium: Mathematics and Society, INSA Anniversary Conference, Dec 2019
\item Gautam Bharali was Member of the Faculty Selection Committee at IIIT Guwahati in June 2018
\item Gautam Bharali was Member of the executive committee of IWM (Indian Women and Mathematics)
\item R VENKATESH was We hosted 70 students and 5 faculty members from Madras Christian College as a part of their Educational visit to IISc, Bangalore on 16th October 2018. Two lectures were arranged for them on 16th October 2018.
\item R. Venkatesh was Algebra Symposium, Department of Mathematics, IISc, Bangalore, India (February 12, 2020).
\item R. Venkatesh was Discussion Meeting on Representation Theory 2019, Department of Mathematics, IISc, Bangalore (December 14 - 16, 2019). Partially funded by the UGC-CAS grant.  Two lecture series was delivered by Professor Vyjayanthi Chari (UC Riverside, USA) and Professor Anne Schilling (UC Davis, USA) as part of this discussion meeting.
\item R. Venkatesh was NCM - Instructional School for Teachers on Analysis and PDE, Department of Mathematics, IISc, Bangalore (May 6 - 18, 2019).
\item Srikanth Krishnan Iyer was Gave a series of ten lectures in the Advanced Instructional School on Stochastic Processes, June 2019 organized by NICER, Bhubaneshwar on Probability Theory.
\item Srikanth Krishnan Iyer was Gave a series of eight lectures in the annual Lecture Series in Probability XIV held at ISI-Delhi between December 2 - 6, 2019 on Limit Theorems in Stochastic Geometry.
\item Srikanth Krishnan Iyer was Gave a talk titled :Random Connection Models" at the RBI, Mumbai on the occasion of Statistics day, June 26, 2019.
\item Srikanth Krishnan Iyer was Gave a talk titled "Phase Transitions in Random Networks" at the FLAIR 4 Big Data Summer School on July 6, 2019.
\item Subhojoy Gupta was Co-organized a discussion meeting at ICTS Bangalore in December 2018, on “Surface-Group Representations and Projective Structures". Website: https://www.icts.res.in/discussion-meeting/SGPS2018.
\item Subhojoy Gupta was Referee for “Memoirs of the American Mathematical Society",  “Indian Journal of Pure and Applied Mathematics"
\item Subhojoy Gupta was M.Sc. thesis referee for a student of Chennai Mathematical Institute,  Satyajit Maity
\item Subhojoy Gupta was Member of Research Advisory Committee for Ramya Nair, a PhD student at IISER Pune.
\item Thirupathi Gudi was Organized Workshop on Finite Elements for Nonlinear and Multiscale Problems during Feb 28 – Mar 03, 2020.
\item Vamsi Pritham Pingali was Guided the Master's Thesis ("Kahler geometry with a view towards the Calabi conjecture") of Vishnu Nandakumaran of IISER Pune (academic year of 2019-2020)
\item Vamsi Pritham Pingali was Visited ICTP as a junior associate during May 2019 and delivered a talk ("A vector bundle version of the Monge-Ampere equation")
\end{enumerate}

\subsection{Other Activities: 2019-20}
\begin{enumerate}
\item A. K. Nandakumaran was Convener, KVPY
\item A. K. Nandakumaran was Member, Library Committee
\item Apoorva Khare was Member of two UG Project Presentation Evaluation Committees
\item Apoorva Khare was Pre-RMO (Regional Mathematics Olympiad) - helped in setting the question paper
\item Apoorva Khare was Senate nominee for Comprehensive Exam of Vinayak Ramkumar (student of P. Vijay Kumar, ECE)
\item Apoorva Khare was Advising: (1) Prateek Kumar Vishwakarma, Int-PhD (passed his Comprehensive Exam) (2) Babhrubahan Bose, Int-PhD (passed his Comprehensive Exam) (3) G.V.K. Teja, PhD (passed his Comprehensive Exam) (4) Swarnalipa Datta, Int-PhD (downgraded to MS) Mentoring: (5) Dr. Projesh Nath Choudhury, NBHM postdoc
\item Arvind Ayyer was Referee for the thesis of Majil Pratim Saikia at University of Vienna.
\item Arvind Ayyer was Referee for the thesis of Mr. Jyotirmoy Ganguly at IISER Pune.
\item Arvind Ayyer was Thesis examiner for Mr. Sudip Bera, Visva-Bharati University.
\item Gadadhar Misra was Member, Scientific Committee, Hilbert Modules in Analytic Function Spaces workshop,     Tsinghua Sanya International Mathematics Forum (TSIMF), Sanya, China, Dec 30, 2019 - Jan 3, 2020
\item Gautam Bharali was Member of the executive committee of IWM (Indian Women and Mathematics)
\item Gautam Bharali was Served as an independent expert for the Swiss National Science Foundation for the review of project proposals in pure mathematics in 2019.
\item R VENKATESH was Algebra Symposium, Department of Mathematics, IISc, Bangalore, India (February 12, 2020).
\item R VENKATESH was Discussion Meeting on Representation Theory 2019, Department of Mathematics, IISc, Bangalore (December 14 - 16, 2019). Partially funded by the UGC-CAS grant.  Two lecture series was delivered by Professor Vyjayanthi Chari (UC Riverside, USA) and Professor Anne Schilling (UC Davis, USA) as part of this discussion meeting.
\item R VENKATESH was NCM - Instructional School for Teachers on Analysis and PDE, Department of Mathematics, IISc, Bangalore (May 6 - 18, 2019). 
\item Siddhartha Gadgil was thesis examiner for a thesis at IIT, Kanpur
\item Siddhartha Gadgil was thesis examiner for a thesis at Delhi University
\item Siddhartha Gadgil was thesis examiner for a thesis for TIFR, Mumbai
\item Siddhartha Gadgil was member of selection committee at BHU
\item Subhojoy Gupta was Member of Research Advisory Committee for Ramya Nair, a PhD student at IISER Pune.
\item Subhojoy Gupta was Co-organizer of an Advanced Instructional School on Riemannian Geometry, sponsored by the National Centre for Mathematics, held in July 2019.
\item Subhojoy Gupta was Referee for "Proceedings of the London Mathematical Society", "Journal of Topology", "Manuscripta Mathematica" and reviewer for AMS and Zentralblatt Math Reviews
\item Subhojoy Gupta was Supervised UG final year project of Shabarish Chenakkod
\end{enumerate}