### 2020-08-10 data/publications_2018-20.tex
improvements for next year

\subsubsection{\!\!\!{\sl Publications}}

\subsubsection*{{\small $(i)$ {\sc Research Papers in Refereed Journals}}}
\medskip

\begin{center}
	\begin{longtable}{|c|c|c|c|}

		\hline
		                    &                   &                   &                            \\[-4mm]
		{\bf Year}          & {\bf Thrust Area} & {\bf Other Areas} & {\bf Total}                \\[.4mm]\hline
		\endfirsthead

		\hline
		                    &                   &                   &                            \\[-4mm]
		{\bf Year}          & {\bf Thrust Area} & {\bf Other Areas} & {\bf Total} \hspace{1.1cm} \\[.4mm]\hline
		\endhead

		                    &                   &                   &                            \\[-4mm]
		{\em 2018--19}      & 72                & 34                & 106                        \\[.4mm]\hline
		                    &                   &                   &                            \\[-4mm]
		{\em Accepted for}  & 13                & 7                 & 20                         \\
		{\em publication}   &                   &                   &                            \\[0.4mm]\hline
		                    &                   &                   &                            \\[-4mm]
		{\em Submitted} for & 17                & 12                & 29                         \\
		{\em Publication}   &                   &                   &                            \\[0.4mm]\hline
		                    &                   &                   &                            \\[-4mm]
		{\bf Total}         & {\bf 102}         & {\bf 53}          & {\bf 155}                  \\[0.4mm]\hline
	\end{longtable}
\end{center}

\noindent{{\small $(i\text{-}a)$ {\sc Published in the period under review}}}
\medskip

%% YEAR 1 LISTING %%%%%
\noindent{{\small {\bf April 2018--March 2019:}}}
\medskip

\noindent{{\small {\sl In the identified thrust area:}}}\vspace{-2.5mm}
\begin{enumerate}[leftmargin=27pt]

	\item Aiyappan, S. and Nandakumaran, A.K. and Prakash, Ravi: Generalization of unfolding operator for highly oscillating
	      smooth boundary domains and homogenization, {\em Calc. Var. Partial Differential Equations} {\bf 57} (2018), Art.~86, 30.

	\item Akutagawa, Kazuo and Endo, Hisaaki and Seshadri, Harish: A gap theorem for positive {E}instein metrics on the
	      four-sphere, {\em Math. Ann.} {\bf 373} (2019), 1329--1339.

	\item\label{ayyer:exclusion-process} Ayyer, A. and Finn, C. and Roy, D.: Matrix product solution of a left-permeable two-species asymmetric exclusion process, {\em Physical Review E} {\bf 97} (2018), 5225.

	\item\label{ayyer:multispecies-juggling} Ayyer, Arvind and Bouttier, J\'{e}r\'{e}mie and Corteel, Sylvie and
	      Linusson, Svante and Nunzi, Fran\c{c}ois: Bumping sequences and multispecies juggling, {\em Adv. in Appl. Math.} {\bf 98} (2018), 100--126.

	\item   Ayyer, Arvind and Finn, Caley and Roy, Dipankar: The phase diagram for a multispecies left-permeable asymmetric
	      exclusion process, {\em J. Stat. Phys.} {\bf 174} (2019), 605--621.

	\item\label{ayyer:Hilbert-Galton} Ayyer, Arvind and Ramassamy, Sanjay: The {H}ilbert--{G}alton board, {\em ALEA Lat. Am. J. Probab. Math. Stat.} {\bf 15} (2018), 755--774.

	\item Bagchi, Sayan and  Thangavelu, Sundaram: Weighted norm inequalities for {W}eyl multipliers and Fourier multipliers on the {H}eisenberg group, {\em J. Anal. Math.} {\bf 136} (2018), 1--29.

	\item Baklouti, Ali and Thangavelu, Sundaram: Hardy and {M}iyachi theorems for {H}eisenberg motion groups, {\em Nagoya Math. J.} {\bf 229} (2018), 1--20.

	\item Balhara, Rakesh: Hardy's inequality for the fractional powers of the {G}rushin
	      operator, {\em Proc. Indian Acad. Sci. Math. Sci.} {\bf 129} (2019), Art.~33, 25.

	\item\label{ghosh:political-business-cycle} Basak, G. K. and Ghosh, M. K. and Mukherjee, D.: A stochastic model with inflation,
	      growth and technology for the political business cycle, {\em Journal of Computational Economics} {\bf 53} (2019), 125-140.

	\item Basu, Arnab and Ghosh, Mrinal K.: Nonzero-sum risk-sensitive stochastic games on a countable
	      state space, {\em Math. Oper. Res.} {\bf 43} (2018), 516--532.

	\item Belton, Alexander and Guillot, Dominique and Khare, Apoorva
	      and Putinar, Mihai: Simultaneous kernels of matrix {H}adamard powers, {\em Linear Algebra Appl.} {\bf 576} (2019), 142--157.

	\item Bera, Sayani and Verma, Kaushal: Some aspects of shift-like automorphisms of {$\mathbb C^k$},
	      {\em Proc. Indian Acad. Sci. Math. Sci.} {\bf 128} (2018), Art.~10, 48.

	\item Bhattacharyya, Tirthankar and Sau, Haripada: Holomorphic functions on the symmetrized
	      bidisk\,---\,realization, interpolation and extension, {\em J. Funct. Anal.} {\bf 274} (2018), 504--524.

	\item Bhimani, Divyang G. and Balhara, Rakesh and Thangavelu,
	      Sundaram: Hermite multipliers on modulation spaces, {\em Analysis and partial differential equations:
			      perspectives from developing countries} {\bf 275} (2019), 42--64.

	\item Biswas, Indranil and D'Mello, Shane and Mukherjee, Ritwik and Pingali, Vamsi P.: Rational cuspidal
	      curves on del-{P}ezzo surfaces, {\em J. Singul.} {\bf 17} (2018), 91--107.

	\item Biswas, Indranil and Pingali, Vamsi Pritham: A characterization of finite vector bundles on
		      {G}auduchon astheno-{K}\"{a}hler manifolds, {\em \'{E}pijournal Geom. Alg\'{e}brique} {\bf 2} (2018), Art.~6, 13.

	\item Biswas, Rahul and Dond, Asha K. and Gudi, Thirupathi: Edge patch-wise local projection stabilized nonconforming
		      {FEM} for the {O}seen problem, {\em Comput. Methods Appl. Math.} {\bf 19} (2019), 189--214.

	\item\label{Misra:symm-polydisc} Biswas, Shibananda and Ghosh, Gargi and Misra, Gadadhar and
	      Shyam Roy, Subrata: On reducing submodules
	      of {H}ilbert modules with {${S}_n$}-invariant kernels, {\em J. Funct. Anal.} {\bf 276} (2019), 751--784.

	\item Borah, Diganta and Haridas, Pranav and Verma, Kaushal: Comments on the {G}reen's function of a
	      planar domain, {\em Anal. Math. Phys.} {\bf 8} (2018), 383--414.

	\item Brenner, Susanne C. and Gudi, Thirupathi and Porwal, Kamana
	      and Sung, Li-Yeng: A {M}orley finite element method for an elliptic distributed
	      optimal control problem with pointwise state and control
	      constraints, {\em ESAIM Control Optim. Calc. Var.} {\bf 24} (2018), 1181--1206.

	\item Carstensen, Carsten and Dond, Asha K. and Rabus, Hella: Quasi-optimality of adaptive mixed {FEM}s for non-selfadjoint
	      indefinite second-order linear elliptic problems, {\em Comput. Methods Appl. Math.} {\bf 19} (2019), 233--250.

	\item Chandel, Vikramjeet Singh: The three-point {P}ick--{N}evanlinna interpolation problem on
	      the polydisc, {\em Complex Var. Elliptic Equ.} {\bf 63} (2018), 1341--1352.

	\item\label{rangarajan:Granger-causality} Chopra, R. and Murthy, C. R. and Rangarajan, G.: Statistical tests for detecting Granger causality, {\em IEEE Transactions on Signal Processing} {\bf } (2018), 5803--5816.

	\item Ciaurri, \'{O}scar and Roncal, Luz and Thangavelu, Sundaram: Hardy-type inequalities for fractional powers
	      of the {D}unkl-{H}ermite operator, {\em Proc. Edinb. Math. Soc. (2)} {\bf 61} (2018), 513--544.

	\item Dond, Asha K. and Gudi, Thirupathi: Patch-wise local projection stabilized finite element methods
	      for convection-diffusion-reaction problems, {\em Numer. Methods Partial Differential Equations}
		      {\bf 35} (2019), 638--663.

	\item Dond, Asha K. and Gudi, Thirupathi and Sau, Ramesh Ch.: An error analysis of discontinuous finite element
	      methods for the optimal control problems governed by {S}tokes equation, {\em Numer. Funct. Anal. Optim.}
		      {\bf 40} (2019), 421--460.

	\item Gaddam, Sharat and Gudi, Thirupathi: Inhomogeneous {D}irichlet boundary condition in the {\it a
			      posteriori} error control of the obstacle problem, {\em Comput. Math. Appl.} {\bf 75} (2018), 2311--2327.

	\item Gaddam, Sharat and Gudi, Thirupathi: Bubbles enriched quadratic finite element method for the
	      3{D}-elliptic obstacle problem, {\em Comput. Methods Appl. Math.} {\bf 18} (2018), 223--236.

	\item Gaussier, Herv\'{e} and Seshadri, Harish: Holomorphic motions and complex geometry,
	      {\em Proc. Amer. Math. Soc.} {\bf 147} (2019), 301--313.

	\item Gaussier, Herv\'{e} and Seshadri, Harish, On the {G}romov hyperbolicity of convex domains in {$\mathbb
				      C^n$}, {\em Comput. Methods Funct. Theory} {\bf 18} (2018), 617--641.

	\item Ghara, Soumitra and Kumar, Surjit: On the sum of two subnormal kernels, {\em J. Math. Anal. Appl.}
		      {\bf 469} (2019), 1015--1027.

	\item Ghara, Soumitra: The orbit of a bounded operator under the M{\"o}bius group modulo similarity equivalence.

	\item Ghosh, Mrinal K. and Pradhan, Somnath: Risk-sensitive stochastic differential games with reflecting
	      diffusions, {\em Stoch. Anal. Appl.} {\bf 36} (2018), 1--27.

	\item Gudi, Thirupathi and Majumder, Papri: Convergence analysis of finite element method for a parabolic
	      obstacle problem, {\em J. Comput. Appl. Math.} {\bf 357} (2019), 85--102.

	\item Gupta, Rajeev and Kumar, Surjit and Trivedi, Shailesh: Von {N}eumann's inequality for commuting
	      operator-valued multishifts, {\em Proc. Amer. Math. Soc.} {\bf 147} (2019), 2599--2608.

	\item Gupta, Rajeev and Ramiz Reza, Md.: Operator space structures on {$\ell^1(n)$}, {\em Houston J. Math.}
		      {\bf 44} (2018), 1205--1212.

	\item Hazra, Somnath: Homogeneous 2-shifts, {\em Complex Anal. Oper. Theory} {\bf 13} (2019), 1729--1763.

	\item\label{iyer:connection-model} Iyer, Srikanth K.: The random connection model: connectivity, edge lengths,
	      and degree distributions, {\em Random Structures Algorithms} {\bf 52} (2018), 283--300.

	\item Kashyap, G. and Bapat, D. and Das, D. and Gowaikar, R.D. and  Amritkar, R.E. and Rangarajan, G. and V. Ravindranath and G. Ambika: Synapse loss and progress of Alzheimer's disease\,--\,A network model,
	      {\em Scientific Reports} {\bf 9} (2019).

	\item Khare, Apoorva and Tao, Terence: Schur polynomials, entrywise positivity preservers, and weak
	      majorization, {\em S\'em. Lothar. Combin.} {\bf 80B,} (2018).

	\item Lin, X. and Kulkarni, P. and Bocci, F. and Schafer, N. and Roy, S. and Tsai, M-Y. and He, Y. and Chen, Y and Rajagopalan, K. and Mooney, S. and Zeng, Y. and Weninger, K. and Grishaev, A. and Onuchic, J. and Levine, H. and Wolynes, P. and Salgia, R. and Rangarajan, G. and Uversky, V. and Orban, J. and Jolly, M. K.: Structural and
	      dynamical order of a disordered protein: molecular insights into conformational
	      switching of PAGE4 at the systems level, {\em Biomolecules} {\bf 9} (2019), 77.

	\item\label{rangarajan:prostate-cancer} Lin, X. and Roy, S. and Jolly, M. K. and Bocci, F. and Schafer, N. P. and Tsai, M.-Y. and Chen, Y. and He, Y. and Grishaev, A. and Weninger, K. and Orban, J. and Kulkarni, P. and Rangarajan, G. and Levine, H. and J. N. Onuchic: PAGE4 and conformational switching: Insights from molecular dynamics simulations and implications for
	      prostate cancer, {\em Journal of Molecular Biology} {\bf 430} (2018), 2422--2438.

	\item Mahajan, Prachi and Verma, Kaushal: A comparison of two biholomorphic invariants, {\em Internat. J. Math.} {\bf 30} (2019), 1950012, 16 pp.

	\item Mal, Arpita and Sain, Debmalya and Paul, Kallol: On some geometric properties of operator spaces, {\em Banach J. Math. Anal.} {\bf 13} (2019), 174--191.

	\item Manfio, F. and Turgay, N. C. and Upadhyay, A.: Biconservative submanifolds in
		      {$\mathbb{S}^n\times\mathbb{R}$} and
		      {$\mathbb{H}^n\times\mathbb{R}$}, {\em J. Geom. Anal.} {\bf 29} (2019), 283--298.

	\item Manikandan, K. and Vishnu Priya, N. and Senthilvelan, M. and
	      Sankaranarayanan, R.: Deformation of dark solitons in a {$\mathcal{PT}$}-invariant
	      variable coefficients nonlocal nonlinear {S}chr\"{o}dinger
	      equation, {\em Chaos} {\bf 28} (2018), 083103, 12.

	\item Misra, Gadadhar and Pal, Avijit: Curvature inequalities for operators in the {C}owen--{D}ouglas
	      class and localization of the {W}allach set, {\em J. Anal. Math.} {\bf 136} (2018), 31--54.

	\item Misra, Gadadhar and Pal, Avijit and Varughese, Cherian: Contractivity and complete contractivity for finite dimensional Banach spaces, {\em J. Operator Theory} {\bf 82} (2019), 23--47.

	\item Pal, Chandan and Pradhan, Somnath: Risk sensitive control of pure jump processes on a general
	      state space, {\em Stochastics} {\bf 91} (2019), 155--174.

	\item Pal, Ratna and Verma, Kaushal: Ergodic properties of families of {H}\'{e}non maps, {\em Ann. Polon. Math.}
		      {\bf 121} (2018), 45--71.

	\item Paul, Kallol and Sain, Debmalya: Birkhoff--{J}ames orthogonality and its application in the
	      study of geometry of {B}anach space, {\em Advanced topics in mathematical analysis} {\bf } (2019), 245--284.

	\item Paul, Kallol and Sain, Debmalya and Mal, Arpita: Approximate {B}irkhoff--{J}ames orthogonality in the space of
	      bounded linear operators, {\em Linear Algebra Appl.} {\bf 537} (2018), 348--357.

	\item Paul, Kallol and Sain, Debmalya and Mal, Arpita and Mandal,
	      Kalidas: Orthogonality of bounded linear operators on complex {B}anach
	      spaces, {\em Adv. Oper. Theory} {\bf 3} (2018), 699--709.

	\item Pingali, Vamsi P.: A note on the deformed {H}ermitian {Y}ang--{M}ills {PDE}, {\em Complex Var. Elliptic Equ.}
		      {\bf 64} (2019), 503--518.

	\item Pingali, Vamsi Pritham: A note on higher extremal metrics, {\em Trans. Amer. Math. Soc.} {\bf 370} (2018), 6995--7010.

	\item Pingali, Vamsi Pritham: Representability of {C}hern--{W}eil forms, {\em Math. Z.} {\bf 288} (2018), 629--641.

	\item Pritham Pingali, Vamsi: Existence of coupled {K}\"{a}hler--{E}instein metrics using the
	      continuity method, {\em Internat. J. Math.} {\bf 29} (2018), 1850041, 8.

	\item Reddy, Nanda Kishore: Equality of {L}yapunov and stability exponents for products of
	      isotropic random matrices, {\em Int. Math. Res. Not. IMRN} {\bf } (2019), 606--624.

	\item Reza, Md. Ramiz: Curvature inequalities for operators in the {C}owen--{D}ouglas
	      class of a planar domain, {\em Indiana Univ. Math. J.} {\bf 67} (2018), 1255--1279.

	\item Sain, Debmalya: On extreme contractions and the norm attainment set of a
	      bounded linear operator, {\em Ann. Funct. Anal.} {\bf 10} (2019), 135--143.

	\item Sain, Debmalya: On the norm attainment set of a bounded linear operator, {\em J. Math. Anal. Appl.} {\bf 457} (2018), 67--76.

	\item Sain, Debmalya and Paul, Kallol and Bhunia, Pintu and Bag,
	      Santanu: On the numerical index of polyhedral {B}anach spaces, {\em Linear Algebra Appl.} {\bf 577} (2019), 121--133.

	\item Sain, Debmalya and Paul, Kallol and Mal, Arpita: A complete characterization of {B}irkhoff--{J}ames
	      orthogonality in infinite dimensional normed space, {\em J. Operator Theory} {\bf 80} (2018), 399--413.

	\item Sain, Debmalya and Paul, Kallol and Mal, Arpita: On approximate {B}irkhoff--{J}ames orthogonality and normal
	      cones in a normed space, {\em J. Convex Anal.} {\bf 26} (2019), 341--351.

	\item\label{rangarajan:microtremor-method} Sant, D.A. and Parvez, I. A. and Rangarajan, G. and Patel, S. J. and Salam, T. A. S. and Bhatt, M. N.: Subsurface imaging of brown coal bearing Tertiary sedimentaries\,---\,Deccan Trap
	      interface using microtremor method, {\em Journal of Applied Geophysics} {\bf 159} (2018), 362--373.

	\item Seshadri, Harish and Verma, Kaushal: Some aspects of the automorphism groups of domains {\em Handbook of group actions, {V}ol.~{III}} {\bf 40} (2018), 145--174.

	\item Thangavelu, Sundaram: An analogue of {P}fannschmidt's theorem for the {H}eisenberg
	      group, {\em J. Anal.} {\bf 26} (2018), 235--244.

	\item Thangavelu, Sundaram and Naidu Dogga, Venku: {$L^p$}-{$L^2$} estimates for solutions of the wave equation
	      associated to the {G}rushin operator, {\em Adv. Pure Appl. Math.} {\bf 9} (2018), 85--92.

	\item Turgay, Nurettin Cenk and Upadhyay, Abhitosh: On biconservative hypersurfaces in 4-dimensional {R}iemannian
	      space forms, {\em Math. Nachr.} {\bf 292} (2019), 905--921.

	\item Verma, Kaushal: Notes on the boundaries of quadrature domains, {\em Anal. Math. Phys.} {\bf 9} (2019), 617--638.

	\item Vishnu Priya, N. and Senthilvelan, M. and  Rangarajan, Govindan
	      and Lakshmanan, M.: On symmetry preserving and symmetry broken bright, dark and
	      antidark soliton solutions of nonlocal nonlinear {S}chr\"{o}dinger
	      equation, {\em Phys. Lett. A} {\bf 383} (2019), 15--26.
\end{enumerate}

\noindent{{\small {\sl In other areas:}}}\vspace{-2.5mm}

\begin{enumerate}[resume, leftmargin=27pt]
	\item Anamby, Pramath and Das, Soumya: Distinguishing {H}ermitian cusp forms of degree 2 by
	      a certain subset of all {F}ourier coefficients, {\em Publ. Mat.} {\bf 63} (2019), 307--341.

	\item Anamby, Pramath and Das, Soumya: Bounds for the {P}etersson norms of the pullbacks of
		      {S}aito-{K}urokawa lifts, {\em J. Number Theory} {\bf 191} (2018), 289--304.

	\item Arunkumar, G. and Kus, Deniz and Venkatesh, R.: Root multiplicities for {B}orcherds
	      algebras and graph coloring, {\em J. Algebra} {\bf 499} (2018), 538--569.

	\item\label{ayyer:plane-partitions} Ayyer, Arvind and Behrend, Roger E.: Factorization theorems for classical group characters,
	      with applications to alternating sign matrices and plane
	      partitions, {\em J. Combin. Theory Ser. A} {\bf 165} (2019), 78--105.

	\item Ayyer, Arvind and Prasad, Amritanshu and Spallone, Steven: Odd partitions in {Y}oung's lattice,
	      {\em S\'{e}m. Lothar. Combin.} {\bf 75} (2018), Art. B75g, 13.

	\item Balodi, Mamta and Palcoux, Sebastien: On {B}oolean intervals of finite groups, {\em J. Combin. Theory Ser. A} {\bf 157}
	      (2018), 49--69.

	\item Benedetti, Carolina and Gonz\'{a}lez D'Le\'{o}n, Rafael S. and Hanusa,
	      Christopher R.H. and Harris, Pamela E. and Khare, Apoorva and
	      Morales, Alejandro H. and Yip, Martha: The volume of the caracol polytope, {\em S\'{e}m. Lothar. Combin.}
	      {\bf 80B} (2018), Art.~87, 12.

	\item Bhosle, Usha N.: Hitchin pairs on reducible curves, {\em Internat. J. Math.} {\bf 29}
	      (2018), 1850015, 49.

	\item B\"{o}cherer, Siegfried and Das, Soumya: Cuspidality and the growth of {F}ourier coefficients of
	      modular forms, {\em J. Reine Angew. Math.} {\bf 741} (2018), 161--178.

	\item B\"{o}cherer, Siegfried and Das, Soumya: Petersson norms of not necessarily cuspidal {J}acobi
	      modular forms and applications, {\em Adv. Math.} {\bf 336} (2018), 335--376.

	\item Banerjee, Abhishek: On {A}uslander's formula and cohereditary torsion pairs, {\em Commun. Contemp. Math.} {\bf 20} (2018),
	      1750071, 27.

	\item Banerjee, Abhishek: Modular {H}ecke algebras over {M}\"{o}bius categories, {\em J. Geom. Phys.} {\bf 131} (2018), 23--40.

	\item Banerjee, Abhishek: A topological {N}ullstellensatz for tensor-triangulated
	      categories, {\em C. R. Math. Acad. Sci. Paris} {\bf 356} (2018), 365--375.

	\item Banerjee, Abhishek: Quasimodular {H}ecke algebras and {H}opf actions, {\em J. Noncommut. Geom.} {\bf 12} (2018), 1041--1080.

	\item Banerjee, Abhishek: Completions of monoid objects and descent results, {\em J. Algebra} {\bf 507} (2018), 362--399.

	\item Burton, Benjamin A. and Datta, Basudeb and Singh, Nitin and
	      Spreer, Jonathan: A construction principle for tight and minimal triangulations
	      of manifolds, {\em Exp. Math.} {\bf 27} (2018), 22--36.

	\item Choudhury, Projesh Nath and Kannan, M. Rajesh and Sivakumar,
	      K. C.: A note on linear preservers of semipositive and minimally
	      semipositive matrices, {\em Electron. J. Linear Algebra} {\bf 34} (2018), 687--694.

	\item Das, Soumya and Khan, Rizwanur: The third moment of symmetric square {$L$}-functions, {\em Q. J. Math.} {\bf 69} (2018),
	      1063--1087.

	\item Das, Soumya and Kohnen, Winfried: On sign changes of eigenvalues of {S}iegel cusp forms of genus
	      2 in prime powers, {\em Acta Arith.} {\bf 183} (2018), 167--172.

	\item Datta, Basudeb and Maity, Dipendu: Semi-equivelar maps on the torus and the {K}lein bottle are
		      {A}rchimedean, {\em Discrete Math.} {\bf 341} (2018), 3296--3309.

	\item Fritz, Tobias and Gadgil, Siddhartha and Khare, Apoorva and Nielsen, Pace P. and Silberman, Lior and Tao, Terence: Homogeneous length functions on groups, {\em Algebra Number Theory} {\bf 12} (2018), 1773--1786.

	\item\label{gupta:measured-foliations} Gupta, Subhojoy and Wolf, Michael: Meromorphic quadratic differentials and measured foliations on a Riemann surface, {\em Math Ann.} {\bf 373} (2019), 73--118.

	\item Jha, Abhash Kumar and Sahu, Brundaban: Rankin--{C}ohen brackets on {J}acobi forms of several variables
	      and special values of certain {D}irichlet series, {\em Int. J. Number Theory} {\bf 15} (2019), 925--933.

	\item Khare, Apoorva: Generalized nil-{C}oxeter algebras over discrete complex
	      reflection groups, {\em Trans. Amer. Math. Soc.} {\bf 370} (2018), 2971--2999.

	\item Khare, Apoorva: Generalized nil-Coxeter algebras, {\em S\'em Lothar. Combin.} {\bf 80B,} (2018), No.~29.

	\item Khare, Apoorva and Tikaradze, Akaki: A {C}arlitz--von {S}taudt type theorem for finite rings,
	      {\em Linear Algebra Appl.} {\bf 568} (2019), 106--126.

	\item Krakovski, Roi and Onn, Uri and Singla, Pooja: Regular characters of groups of type {$A_n$} over
	      discrete valuation rings, {\em J. Algebra} {\bf 496} (2018), 116--137.

	\item Maity, Dipendu: Pentagonal maps on the torus and the plane, {\em Beitr. Algebra Geom.} {\bf 60} (2019), 17--37.

	\item Maity, Dipendu and Upadhyay, Ashish Kumar: On centrally symmetric manifolds, {\em J. Ramanujan Math. Soc.} {\bf 34}
	      (2019), 21--27.

	\item Mukherjee, Sajal Kumar and Bera, Sudip: Combinatorial proofs of the {N}ewton--{G}irard and
		      {C}hapman--{C}ostas--{S}antos identities, {\em Discrete Math.} {\bf 342} (2019), 1577--1580.

	\item Narayanan, E. K. and Singla, Pooja: On monomial representations of finitely generated nilpotent
	      groups, {\em Comm. Algebra} {\bf 46} (2018), 2319--2331.

	\item Quddus, Safdar: Invariants of the {$\mathbb Z_2$} orbifolds of the {P}odle\'{s} two
	      spheres, {\em J. Noncommut. Geom.} {\bf 13} (2019), 257--267.

	\item Sanki, Bidyut: Filling of closed surfaces, {\em J. Topol. Anal.} {\bf 10} (2018), 897--913.

	\item Sanki, Bidyut and Gadgil, Siddhartha: Graphs of systoles on hyperbolic surfaces, {\em J. Topol. Anal.}
		      {\bf 11} (2019), 1--20.

\end{enumerate}
\vspace{0.5mm}

%% YEAR 2 LISTING %%%%%
\noindent{{\small {\bf April 2019--March 2020:}}}
\medskip

\noindent{{\small {\sl In the identified thrust area:}}}\vspace{-2.5mm}

\begin{enumerate}[resume, leftmargin=27pt]
	\item Aas, Erik and Ayyer, Arvind and Linusson, Svante and Potka,
	      Samu: \href{https://doi.org/10.1088/1751-8121/ab2e96}{The exact phase diagram for a
		      semipermeable {TASEP} with nonlocal boundary jumps}, \emph{J. Phys. A} {\bf 52} (2019),
	      355001, 19.

	\item Aiyappan, S. and Jose, Editha C. and Lomerio, Ivy Carol B. and
	      Nandakumaran, A. K.: \href{https://doi.org/10.3233/asy-191526}{Control problem on a rough
		      circular domain and homogenization}, \emph{Asymptot. Anal.} {\bf 115} (2019), 19--46.

	\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi: \href{https://doi.org/10.1007/s10231-019-00848-7}{Locally periodic unfolding operator for highly oscillating
		      rough domains}, \emph{Ann. Mat. Pura Appl. (4)} {\bf 198} (2019), 1931--1954.

	\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi: \href{https://doi.org/10.1142/S0219199719500299}{Semi-linear optimal control problem on a smooth oscillating
		      domain}, \emph{Commun. Contemp. Math.} {\bf 22} (2020), 1950029, 26.

	\item Aiyappan, S. and Nandakumaran, A. K. and Sufian, Abu: \href{https://doi.org/10.1002/mma.5748}{Asymptotic analysis of a boundary optimal control problem on a general branched structure}, \emph{Math. Methods Appl. Sci.} {\bf 42} (2019), 6407--6434.

	\item Aiyappan, S. and Sardar, Bidhan Chandra: \href{https://doi.org/10.1080/00036811.2018.1471207}{Biharmonic equation in a highly oscillating domain and
		      homogenization of an associated control problem}, \emph{Appl. Anal.} {\bf 98} (2019), 2783--2801.

	\item Anand, Jatin and Bhattacharyya, Tirthankar and Srivastava,
	      Sachi: \href{https://doi.org/10.1007/s00013-019-01406-6}{Some thoughts on composition operators on subspaces of the
			      {H}ardy space}, \emph{Arch. Math. (Basel)} {\bf 114} (2020), 431--444.

	\item Ayyer, Arvind and Behrend, Roger E. and Fischer, Ilse: \href{https://doi.org/10.1016/j.aim.2020.107125}{Extreme diagonally and antidiagonally symmetric alternating
		      sign matrices of odd order}, \emph{Adv. Math.} {\bf 367} (2020), 107125, 56.

	\item Ayyer, Arvind and Behrend, Roger E.: \href{https://doi.org/10.1016/j.jcta.2019.01.001}{Factorization theorems for classical group characters, with
		      applications to alternating sign matrices and plane
		      partitions}, \emph{J. Combin. Theory Ser. A} {\bf 165} (2019), 78--105.

	\item Ayyer, Arvind and Fischer, Ilse: \href{https://doi.org/10.1016/j.jcta.2020.105241}{Bijective proofs of skew {S}chur polynomial factorizations}, \emph{J. Combin. Theory Ser. A} {\bf 174} (2020), 105241, 40.

	\item Ayyer, Arvind and Linusson, Svante: \href{https://doi.org/10.1002/rsa.20825}{Reverse juggling processes}, \emph{Random Structures Algorithms} {\bf 55} (2019), 56--72.

	\item\label{ayyer-singla:commutative-rings} Ayyer, Arvind and Singla, Pooja:
	      \href{https://doi.org/10.1007/s10468-019-09864-w}{Random motion on finite rings, {I}: commutative
		      rings}, \emph{Algebr. Represent. Theory} {\bf 23} (2020), 583--604.

	\item Ayyer, Arvind: \href{https://doi.org/10.1007/s00026-019-00480-5}{Squareness for the {M}onopole-{D}imer model}, \emph{Ann. Comb.} {\bf 24} (2020), 237--255.

	\item Balakumar, G. P. and Borah, Diganta and Mahajan, Prachi and
	      Verma, Kaushal: \href{https://doi.org/10.1090/proc/14421}{Remarks on the higher dimensional {S}uita conjecture}, \emph{Proc. Amer. Math. Soc.} {\bf 147} (2019), 3401--3411.

	\item Bera, Sayani and Pal, Ratna and Verma, Kaushal: \href{https://doi.org/10.1007/s40879-019-00326-7}{A rigidity theorem for {H}\'{e}non maps}, \emph{Eur. J. Math.} {\bf 6} (2020), 508--532.

	\item Bharali, Gautam and Chandel, Vikramjeet Singh: \href{https://doi.org/10.1007/s11785-017-0701-5}{Pick interpolation on the polydisc: small families of
		      sufficient kernels}, \emph{Complex Anal. Oper. Theory} {\bf 13} (2019), 2069--2093.

	\item Bharali, Gautam: \href{https://doi.org/10.1007/s12220-019-00342-9}{On the growth of the {B}ergman metric near a point of infinite
		      type}, \emph{J. Geom. Anal.} {\bf 30} (2020), 1238--1258.

	\item Bhattacharyya, Tirthankar and Biswas, Anindya and Chandel, Vikramjeet Singh: On the Nevanlinna problem: characterization of all Schur–Agler class solutions affiliated with a given kernel, \emph{Studia Math.} {\bf 255} (2020), 83–107.

	\item\label{Bhattacharyya:symm-bidisc} Bhattacharyya, Tirthankar and Das, Bata Krishna and Sau, Haripada: Toeplitz
	      operators on the symmetrized bidisc, \emph{International Mathematics Research Notices (IMRN)} (2020).

	\item Bhattacharyya, Tirthankar and Rao, T. S. S. R. K.: \href{https://doi.org/10.1007/s13226-019-0345-4}{Survey of last ten years of work done in {I}ndia in some
		      selected areas of functional analysis and operator theory}, \emph{Indian J. Pure Appl. Math.} {\bf 50} (2019), 599--617.

	\item Boggarapu, Pradeep and Roncal, Luz and Thangavelu, Sundaram: \href{https://doi.org/10.3934/cpaa.2019116}{On extension problem, trace {H}ardy and {H}ardy's inequalities
		      for some fractional {L}aplacians}, \emph{Commun. Pure Appl. Anal.} {\bf 18} (2019), 2575--2605.

	\item Chakraborty, Atlanta and Chandru, Vijay and Rao, M. R.: \href{https://doi.org/10.1007/s10479-019-03186-2}{A linear programming primer: from {F}ourier to {K}armarkar}, \emph{Ann. Oper. Res.} {\bf 287} (2020), 593--616.

	\item Chandel, Vikramjeet Singh: \href{https://doi.org/10.1007/s12220-018-00123-w}{The 3-point spectral pick interpolation problem and an
		      application to holomorphic correspondences}, \emph{J. Geom. Anal.} {\bf 30} (2020), 551--572.

	\item Cicuttin, Matteo and Ern, Alexandre and Gudi, Thirupathi: \href{https://doi.org/10.1007/s10915-020-01195-z}{Hybrid high-order methods for the elliptic obstacle problem}, \emph{J. Sci. Comput.} {\bf 83} (2020), Paper No. 8, 18.

	\item Gudi, Thirupathi and Majumder, Papri: \href{https://doi.org/10.1016/j.camwa.2019.06.022}{Conforming and discontinuous {G}alerkin {FEM} in space for
	      solving parabolic obstacle problem}, \emph{Comput. Math. Appl.} {\bf 78} (2019), 3896--3915.

	\item Gudi, Thirupathi and Majumder, Papri: \href{https://doi.org/10.1515/cmam-2019-0057}{Crouzeix-{R}aviart finite element approximation for the
	      parabolic obstacle problem}, \emph{Comput. Methods Appl. Math.} {\bf 20} (2020), 273--292.

	\item Gupta, Rajeev and Kumar, Surjit and Trivedi, Shailesh: \href{https://doi.org/10.1090/proc/14410}{Von {N}eumann's inequality for commuting operator-valued
		      multishifts}, \emph{Proc. Amer. Math. Soc.} {\bf 147} (2019), 2599--2608.

	\item Iyer, S. K. and Yogeshwaran, D.: Thresholds for vanishing of isolated faces in random Cech and Vietoris-Rips Complexes, \emph{Annals de l'Institut Henri Poincar$\acute{e}$ - Probabiliti$\acute{e}$s et Statistiques} {\bf 56} (2020), 1869-1897.

	\item K. Murali, A. K. Nandakumaran, Turgurt Durduran, Hari M Varma: Recovery of the diffuse correlation spectroscopy data-type from speckle contrast measurements: towards low-cost, deep-tissue blood flow measurements, \emph{Biomedical Optics Express 5395} {\bf 10} (2019).

	\item Kashyap, G. A. R. S. R. K. and Bapat, D. and Das, D. and Gowaikar, R. D. and  Amritkar, R. E. and Rangarajan, G. and V. Ravindranath and G. Ambika: Synapse loss and progress of Alzheimers disease - A network model, \emph{Scientific Reports} {\bf 9} (2019).

	\item Kataria, Kuldeep Kumar and Vellaisamy, Palaniappan: \href{https://doi.org/10.1080/07362994.2018.1538803}{On the convolution of {M}ittag-{L}effler distributions and its
	      applications to fractional point processes}, \emph{Stoch. Anal. Appl.} {\bf 37} (2019), 115--122.

	\item Khan, Mohammad Yasir Feroz and Biswas, Rahul and Sau, Ramesh
	      Chandra: Instability of {L}agrange interpolation on {$H^1(\Omega)\cap
				      C(\overline\Omega)$}, \emph{Math. Student} {\bf 88} (2019), 153--158.

	\item Khare, Apoorva and Tikaradze, Akaki: \href{https://doi.org/10.1016/j.laa.2018.05.035}{A {C}arlitz--von {S}taudt type theorem for finite rings}, \emph{Linear Algebra Appl.} {\bf 568} (2019), 106--126.
	\item Khare, Apoorva and Rajaratnam, Bala: Probability inequalities and tail estimates for metric semigroups, \emph{Advances in Operator Theory (Special issue in honor of Rajendra Bhatia)} {\bf 5} (2020), 779-795.

	\item Kor\'{a}nyi, Adam and Misra, Gadadhar: Homogeneous Hermitian holomorphic vector bundles and the Cowen-Douglas class over bounded symmetric domains, \emph{Adv. Math.} {\bf 351} (2019), 1105 - 1138.

	\item Maitra, Anwoy: On the continuous extension of Kobayashi isometries, \emph{Proc. Amer. Math. Soc.} {\bf 148} (2020), 3437--3451.

	\item Maitra, Anwoy: \href{https://doi.org/10.1016/j.jmaa.2019.123694}{A form of {S}chwarz's lemma and a bound for the {K}obayashi
		      metric on convex domains}, \emph{J. Math. Anal. Appl.} {\bf 484} (2020), 123694, 16.
	\item Mal, Arpita and Paul, Kallol and Rao, T. S. S. R. K. and Sain,
	      Debmalya: \href{https://doi.org/10.1007/s00605-019-01289-3}{Approximate {B}irkhoff-{J}ames orthogonality and smoothness in
	      the space of bounded linear operators}, \emph{Monatsh. Math.} {\bf 190} (2019), 549--558.

	\item Maitra, Anwoy: \href{https://doi.org/10.1090/proc/15038}{On the continuous extension of {K}obayashi isometries}, \emph{Proc. Amer. Math. Soc.} {\bf 148} (2020), 3437--3451.

	\item Mal, Arpita and Sain, Debmalya and Paul, Kallol: \href{https://projecteuclid.org/euclid.bjma/1543914019}{On some geometric properties of operator spaces}, \emph{Banach J. Math. Anal.} {\bf 13} (2019), 174-191.

	\item Misra, Gadadhar and Reza, Md. Ramiz: \href{https://doi.org/10.1215/00192082-7768711}{Curvature inequalities and extremal operators}, \emph{Illinois J. Math.} {\bf 63} (2019), 193--217.

	\item Nikolov, Nikolai and Verma, Kaushal: \href{https://doi.org/10.1007/s12220-019-00237-9}{On the squeezing function and {F}ridman invariants}, \emph{J. Geom. Anal.} {\bf 30} (2020), 1218--1225.

	\item Paul, Kallol and Sain, Debmalya and Ghosh, Puja: Symmetry of {B}irkhoff-{J}ames orthogonality of bounded linear
	      operators, \emph{Ulam type stability} {\bf } (2019), 331--344.

	\item Poria, Anirudha: \href{https://doi.org/10.1080/01630563.2020.1728771}{Positive weight function and classification of {$g$}-frames}, \emph{Numer. Funct. Anal. Optim.} {\bf 41} (2020), 950--968.

	\item Roth, Julien and Upadhyay, Abhitosh: \href{https://doi.org/10.1007/s00013-019-01315-8}{On compact anisotropic {W}eingarten hypersurfaces in
			      {E}uclidean space}, \emph{Arch. Math. (Basel)} {\bf 113} (2019), 213--224.

	\item Roth, Julien and Upadhyay, Abhitosh: \href{https://doi.org/10.1017/s0004972719000935}{On almost stable {CMC} hypersurfaces in manifolds of bounded
		      sectional curvature}, \emph{Bull. Aust. Math. Soc.} {\bf 101} (2020), 333--338.

	\item Roth, Julien and Upadhyay, Abhitosh: \href{https://link.springer.com/article/10.1007/s00025-019-1142-4?shared-article-renderer#citeas}{f-Biharmonic Submanifolds of Generalized Space Forms}, \emph{Results in Mathematics} {\bf 75} (2020).

	\item Roth, Julien and Upadhyay, Abhitosh: \href{https://www.cambridge.org/core/journals/bulletin-of-the-australian-mathematical-society/article/on-almost-stable-cmc-hypersurfaces-in-manifolds-of-bounded-sectional-curvature/E0BF84D4D46C82F6D75CDCEAAF5DEE94}{On almost stable CMC hypersurfaces in manifolds of bounded sectional curvature}, \emph{Bulletin of the Australian Mathematical Society} {\bf 101} (2020), 333-338.

	\item Sain, Debmalya and Mal, Arpita and Paul, Kallol: \href{https://doi.org/10.1016/j.exmath.2019.01.001}{Some remarks on {B}irkhoff-{J}ames orthogonality of linear
	      operators}, \emph{Expo. Math.} {\bf 38} (2020), 138--147.

	\item Sain, Debmalya and Paul, Kallol and Mandal, Kalidas: \href{https://doi.org/10.7153/oam-2019-13-31}{On two extremum problems related to the norm of a bounded
		      linear operator}, \emph{Oper. Matrices} {\bf 13} (2019), 421--432.

	\item Sain, Debmalya and Ray, Anubhab and Paul, Kallol: Extreme contractions on finite-dimensional polygonal {B}anach
	      spaces, \emph{J. Convex Anal.} {\bf 26} (2019), 877--885.

	\item Sain, Debmalya: \href{https://doi.org/10.1007/s13226-020-0393-9}{On the norm attainment set of a bounded linear
		      operator and
		      semi-inner-products in normed spaces}, \emph{Indian J. Pure Appl. Math.} {\bf 51} (2020), 179--186.

	\item Sain, Debmalya: \href{https://doi.org/10.7153/oam-2019-13-32}{Smooth points in operator spaces and some
		      {B}ishop-{P}helps-{B}ollob\'{a}s type theorems in {B}anach spaces}, \emph{Oper. Matrices} {\bf 13} (2019), 433--445.

	\item Sarkar, Amar Deep and Verma, Kaushal: A submultiplicative property of the Carath\'{e}odory metric on planar domains, \emph{Proc. Indian Acad. Sci. (Math. Sci.)} {\bf 130} (2020).

	\item Sarkar, Amar Deep and Verma, Kaushal: \href{https://doi.org/10.1007/s40315-020-00303-2}{Boundary behaviour of some conformal invariants on planar
		      domains}, \emph{Comput. Methods Funct. Theory} {\bf 20} (2020), 145--158.

	\item Vishnu Priya, N. and Senthilvelan, M. and  Rangarajan, G. and Lakshmanan, M: On symmetry preserving and symmetry broken bright, dark and antidark soliton solutions of nonlocal nonlinear Schrödinger equation, \emph{Physics Letters A} {\bf 383} (2019), 15-26.

	\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan,
	      Govindan: \href{https://doi.org/10.1063/1.5121245}{On the role of four-wave mixing effect in the interactions
	      between nonlinear modes of coupled generalized nonlinear
		      {S}chr\"{o}dinger equation}, \emph{Chaos} {\bf 29} (2019), 123135, 14.

	\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan, G.: On the role of four-wave mixing effect in the interactions between nonlinear modes of coupled generalized nonlinear Schrodinger equation, \emph{Chaos} {\bf 29} (2019), 123135.\end{enumerate}

\noindent{{\small {\sl In other areas:}}}\vspace{-2.5mm}

\begin{enumerate}[resume, leftmargin=27pt]

	\item Anamby, Pramath and Das, Soumya: \href{https://doi.org/10.5565/PUBLMAT6311911}{Distinguishing
			      {H}ermitian cusp forms of degree 2 by a certain subset of all {F}ourier coefficients},
	      \emph{Publ. Mat.} {\bf 63} (2019), 307--341.

	\item Banerjee, Abhishek and Kour, Surjeet: \href{https://doi.org/10.1007/s10231-019-00844-x}{{$(A,\delta)$}-modules, {H}ochschild homology and higher
		      derivations}, \emph{Ann. Mat. Pura Appl. (4)} {\bf 198} (2019), 1781--1802.

	\item Banerjee, Abhishek: \href{https://doi.org/10.4153/s0008439518000656}{On differential torsion theories and rings with
		      several objects}, \emph{Canad. Math. Bull.} {\bf 62} (2019), 703--714.

	\item Benedetti, Carolina and Gonz\'{a}lez D'Le\'{o}n, Rafael S. and Hanusa,
	      Christopher R. H. and Harris, Pamela E. and Khare, Apoorva and
	      Morales, Alejandro H. and Yip, Martha: \href{https://doi.org/10.1090/tran/7743}{A combinatorial model for computing volumes of flow polytopes}, \emph{Trans. Amer. Math. Soc.} {\bf 372} (2019), 3369--3404.

	\item Biswas, Indranil and Dumitrescu, Sorin and Gupta, Subhojoy: Branched projective structures on a {R}iemann surface and
	      logarithmic connections, \emph{Doc. Math.} {\bf 24} (2019), 2299--2337.

	\item Carstensen, C. and Mallik, G. and Nataraj, N.: \href{https://doi.org/10.1093/imanum/drz071}{Nonconforming Finite Element Discretization for Semilinear Problems with Trilinear Nonlinearity}, \emph{IMA Journal of Numerical Analysis} {\bf } (2020).

	\item Chim, Kwok Chi; Shorey, Tarlok Nath; Sinha, Sneh Bala: On Baker's explicit abc-conjecture, \emph{Publ. Math. Debrecen} {\bf 94} (2019), 435–453.

	\item Choudhury, Projesh Nath and Tsatsomeros Michael J.: \href{https://doi.org/10.1016/j.laa.2020.04.028}{Algorithmic Detection and Construction of N-matrices}, \emph{Linear Algebra and Its Applications} {\bf 602} (2020), 46-56.

	\item Choudhury, Projesh Nath and Tsatsomeros, Michael J.: \href{https://doi.org/10.1016/j.laa.2020.04.028}{Algorithmic detection and construction of {N}-matrices}, \emph{Linear Algebra Appl.} {\bf 602} (2020), 46--56.

	\item Das, Soumya and Jha, Abhash Kumar: \href{https://doi.org/10.1007/s12188-019-00206-7}{Analytic properties of twisted real-analytic {H}ermitian
	      {K}lingen type {E}isenstein series and applications}, \emph{Abh. Math. Semin. Univ. Hambg.} {\bf 89} (2019), 105--116.

	\item Das, Soumya and Pal, Ritwik: \href{https://doi.org/10.1007/s40993-019-0158-x}{The first negative eigenvalue of {Y}oshida lifts}, \emph{Res. Number Theory} {\bf 5} (2019), Paper No. 20, 9.

	\item Datt, Gopal and Kumar, Sanjay: \href{https://link.springer.com/article/10.1007/s12044-018-0449-5?shared-article-renderer}{A criterion for quasinormality in $\mathbb{C}^n$}, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 129} (2019), Paper No. 11, 13 pp.

	\item Datta, B. and Maity, D.: Correction to: Semi-equivelar and vertex-transitive maps on the torus, \emph{Beitr Algebra Geom} {\bf 61} (2020), 187--188.

	\item Datta, B. and Sarkar, S.: Equilibrium triangulations of some quasitoric 4-manifolds, \emph{Southeast Asian Bull. Math.} {\bf 44} (2020), 57--78.

	\item Frauke Bleher, Ted Chinburg, Ralph Greenberg, Mahesh Kakde, George Pappas, Romyar Sharifi, Martin Taylor: Higher Chern Classes in Iwasawa Theory, \emph{American Journal of Mathematics} {\bf 142} (2020), 627--682.

	\item Gadgil, Siddhartha: \href{https://doi.org/10.1007/s10817-019-09523-1}{Homogeneous {L}ength {F}unctions on {G}roups: {I}ntertwined
	      {C}omputer and {H}uman {P}roofs}, \emph{J. Automat. Reason.} {\bf 64} (2020), 677--688.

	\item Garani, Shayan Srinivasa and Seshadri, Harish: \href{https://doi.org/10.1080/17459737.2019.1604845}{An algorithmic approach to {S}outh {I}ndian classical music}, \emph{J. Math. Music} {\bf 13} (2019), 107--134.

	\item Gupta, Subhojoy and Mj, Mahan: \href{https://doi.org/10.1090/proc/14866}{Monodromy representations of meromorphic projective
		      structures}, \emph{Proc. Amer. Math. Soc.} {\bf 148} (2020), 2069--2078.

	\item Gupta, Subhojoy and Seshadri, Harish: \href{https://doi.org/10.1093/imrn/rny204}{On domains biholomorphic to {T}eichm\"{u}ller spaces}, \emph{Int. Math. Res. Not. IMRN} {\bf } (2020), 2542--2560.

	\item Gupta, Subhojoy and Wolf, Michael: \href{https://doi.org/10.1007/s00208-018-1674-z}{Meromorphic quadratic differentials and measured foliations on
		      a {R}iemann surface}, \emph{Math. Ann.} {\bf 373} (2019), 73--118.

	\item Gupta, Subhojoy: \href{https://doi.org/10.1090/tran/7777}{Limits of harmonic maps and crowned hyperbolic surfaces}, \emph{Trans. Amer. Math. Soc.} {\bf 372} (2019), 7573--7596.

	\item Hatui, Sumana: \href{https://doi.org/10.1515/jgth-2019-0045}{Schur multipliers of special {$p$}-groups of rank 2}, \emph{J. Group Theory} {\bf 23} (2020), 85--95.

	\item Hatui, Sumana: \href{https://www.degruyter.com/view/journals/jgth/23/1/article-p85.xml}{Schur multipliers of special p-groups of rank 2}, \emph{Journal of Group Theory} {\bf 23} (2020), 85-95.

	\item Jha, Abhash Kumar and Juyal, Abhishek and Pandey, Manish
	      Kumar: On simultaneous non-vanishing of twisted {$L$}-functions
	      associated to newforms on {$\Gamma_0(N)$}, \emph{J. Ramanujan Math. Soc.} {\bf 34} (2019), 245--252.

	\item Jha, Abhash Kumar and Sahu, Brundaban: \href{https://doi.org/10.1090/conm/732/14793}{Differential operators on {J}acobi forms and special values of
		      certain {D}irichlet series}, \emph{Automorphic forms and related topics} {\bf 732} (2019), 91--99.

	\item Jha, Abhash Kumar and Sahu, Brundaban: \href{https://doi.org/10.1142/S1793042119500490}{Rankin-{C}ohen brackets on {J}acobi forms of several variables
	      and special values of certain {D}irichlet series}, \emph{Int. J. Number Theory} {\bf 15} (2019), 925--933.

	\item Maiti, Arun: \href{https://doi.org/10.1016/j.disc.2020.111911}{Quasi-vertex-transitive maps on the plane}, \emph{Discrete Math.} {\bf 343} (2020), 111911, 7.

	\item Maiti, Arun: \href{https://www.sciencedirect.com/science/article/pii/S0012365X20301035}{Quasi-vertex-transitive maps on the plane}, \emph{Discrete Mathematics} {\bf 343} (2020), 111911-111917.

	\item Maity, Dipendu: \href{https://doi.org/10.1007/s13366-018-0405-7}{Pentagonal maps on the torus and the plane}, \emph{Beitr. Algebra Geom.} {\bf 60} (2019), 17--37.
	\item Maity, Dipendu and Upadhyay, Ashish Kumar: On centrally symmetric manifolds, \emph{J. Ramanujan Math. Soc.} {\bf 34} (2019), 21--27.

	\item Mukherjee, Sajal Kumar and Bera, Sudip: \href{https://doi.org/10.1016/j.disc.2019.02.013}{Combinatorial proofs of the {N}ewton-{G}irard and
		      {C}hapman-{C}ostas-{S}antos identities}, \emph{Discrete Math.} {\bf 342} (2019), 1577--1580.

	\item Pingali, Vamsi Pritham: \href{https://doi.org/10.1007/s11005-020-01279-9}{Quillen metrics and perturbed equations}, \emph{Lett. Math. Phys.} {\bf 110} (2020), 1861--1875.

	\item Pingali, Vamsi Pritham: \href{https://doi.org/10.1016/j.aim.2019.106921}{A vector bundle version of the {M}onge-{A}mp\`ere equation}, \emph{Adv. Math.} {\bf 360} (2020), 106921, 40.

	\item Pradhan, P. and Sahu, B.: \href{https://doi.org/10.1016/j.disc.2020.112044}{A characterization of the family of secant lines to a hyperbolic quadric in PG(3,q), q odd}, \emph{Discrete Mathematics} {\bf 343} (2020).

	\item Quddus, Safdar: \href{https://doi.org/10.4171/JNCG/320}{Invariants of the {$\mathbb{Z}_2$} orbifolds of the {P}odle\'{s} two
	      spheres}, \emph{J. Noncommut. Geom.} {\bf 13} (2019), 257--267.

	\item Roy, Krishanu and Venkatesh, R.: \href{https://doi.org/10.1007/s00031-018-9510-9}{Maximal closed subroot systems of real affine root systems}, \emph{Transform. Groups} {\bf 24} (2019), 1261--1308.

	\item Seshadri, Harish: \href{https://doi.org/10.1007/s13226-019-0355-2}{Differential geometry in {I}ndia}, \emph{Indian J. Pure Appl. Math.} {\bf 50} (2019), 795--799.

\end{enumerate}
\vspace{0.5mm}

\noindent{{\small $(i\text{-}b)$ {\sc Accepted for publication at the time of reporting}}}
\medskip

%%THIS LIST IS NOT ORGANISED BY YEAR %%%%%
%\noindent{{\small {\bf  April 2018--March 2019:}}}
%\medskip

\noindent{{\small {\sl In the identified thrust area:}}}\vspace{-2.5mm}

\begin{enumerate}[resume, leftmargin=27pt]
	\item Basu, R. and Ganguly, S. and Hegde, M. and Krishnapur, M.: Lower Deviations in β-ensembles and Law of Iterated Logarithm in Last Passage Percolation, to appear in \emph{Israel Journal of Mathematics}.

	\item\label{bharali:Wolff-Denjoy} Bharali, Gautam and Maitra, Anwoy: A weak notion of visibility,
	      a family of examples, and Wolff--Denjoy theorems, to appear in \emph{Annali della Scuola Normale Superiore di Pisa}.

	\item\label{sen:models-at-criticality} Dhara, Souvik and van der Hofstad, Remco and van Leeuwaarden, Johan S.H. and Sen, Sanchayan: Heavy-tailed configuration models at criticality, to appear in {\em Annales de l'Institut Henri
			      Poincar\'e (B) Probabilit\'es et Statistiques}.

	\item Ghosh, M. K. and K. Suresh Kumar and C. Pal and S. Pradhan: Nonzero-sum risk-sensitive stochastic differential games with discounted costs, Stochastic Analysis and Applications, to appear in \emph{Stochastic Analysis and Applications}.

	\item Ghosh, M. K. and Pradhan, S.: Zero-Sum Risk-Sensitive Stochastic Differential Games with Reflecting Diffusions in the Orthant, to appear in \emph{ESIAM: Control, Optimization and Calculus of Variations}.

	\item Ghosh, Subhajit: Total variation cutoff for the transpose top-2 with random shuffle, to appear in {\em Journal of Theoretical Probability}.

	\item Gudi, Thirupathi and  Sau, Ramesh Ch.: Finite Element Analysis of the Constrained Dirichlet Boundary Control Problem Governed by the Diffusion Problem, to appear in \emph{ESAIM: Control, Optim. Calc. Var}.

	\item\label{iyer:heterogeneous-populations} Nair, Gokul and Athmanathan, Senthil and Iyer, Srikanth K. and Guttal, Vishwesha: Fission-fusion dynamics and group-size-dependent composition in heterogeneous populations,
	      to appear in {\em Physical Review E}.

	\item\label{thangavelu:H-type-groups} Roncal, Luz and Thangavelu, Sundaram: An extension problem and trace Hardy
	      inequality for the sublaplacian on H-type groups, to appear in {\em International Mathematical Research Notices}.

	\item\label{iyer:wireless-networks} Vaze, Rahul and Iyer, Srikanth K., Capacity of cellular wireless networks,
	      to appear in {\em IEEE Transactions on Wireless Communications}.

\end{enumerate}

\noindent{{\small {\sl In other areas:}}}\vspace{-2.5mm}

\begin{enumerate}[resume, leftmargin=27pt]
	\item Alvarez-Consul, Luis and Garcia-Fernandez, Mario  and García-Prada, Oscar and Pingali, Vamsi Pritham: Gravitating vortices and the Einstein-Bogomol’nyi equations, to appear in \emph{Mathematische Annalen}.

	\item Balodi, M. and Banerjee, A. and Ray, S.: Cohomology of modules over H-categories and co-H-categories, to appear in \emph{Canadian Journal of Mathematics}.

	\item Balodi, M. and Banerjee, A. and Ray, S.: Entwined modules over linear categories and Galois extensions, to appear in \emph{Israel Journal of Mathematics}.

	\item Banerjee, A.: On Noetherian schemes over $(\mathcal C,\otimes,1)$ and the category of quasi-coherent sheaves, to appear in \emph{Indiana University Mathematics Journal}.

	\item Barbosa, Ezequiel and Santana, Farley and Upadhyay, Abhitosh: On stable CMC free-boundary surfaces in a strictly convex domain of a bi-invariant Lie group, to appear in \emph{International Journal of Mathematics}.

	\item Choudhury, Projesh Nath and Sivakumar, K.C.: \href{https://doi.org/10.1080/03081087.2019.1602588}{Matrices with Positive Semidefinite Real Part}, to appear in \emph{Linear and Multilinear Algebra}.

	\item\label{datta-gupta:hyperbolic-plane} Datta, Basudeb and Gupta, Subhojoy: Semi-regular tilings of the hyperbolic plane,
	      to appear in \emph{Discrete and Computational Geometry}.

	\item Gupta, Subhojoy: Harmonic maps and wild Teichm{\"u}ller spaces, to appear in {\em Journal of Topology and
			      Analysis}.

	\item Kreuzer, M. and Patil, D. Patil: Computational aspects of Burnside rings part II: important maps, to appear in \emph{Beitr\"age zur Algebra und Geometrie / Contributions to Algebra and Geometry}.

	\item\label{pingali:algebraic-hypersurfaces} Pingali, Vamsi Pritham and Varolin, Dror: Non-uniformly flat affine algebraic hypersurfaces, to appear in {\em Nagoya Mathematical Journal}.

	\item Ray, Samarpita: Closure operations, Continuous valuations on monoids and Spectral spaces, to appear in {\em Journal of algebra and its applications}.
\end{enumerate}
\vspace{0.5mm}

\noindent{{\small $(i\text{-}c)$ {\sc Submitted/preprints at the time of reporting}}}
\medskip

%% THIS LIST IS NOT ORGANISED BY YEAR%%%%%
%\noindent{{\small {\bf April 2018--March 2019:}}}
%\medskip

\noindent{{\small {\sl In the identified thrust area:}}}\vspace{-2.5mm}

\begin{enumerate}[resume, leftmargin=27pt]
	\item\label{sen:random-3-regular-graph} Addario-Berry, Louigi and Sen, Sanchayan: Geometry of the minimal spanning tree of a random 3-regular graph.

	\item\label{ayyer:partial-cyclic-orders} Ayyer, Arvind and Josuat-Verg\`es, Matthieu and Ramassamy, Sanjay: Extensions of partial cyclic orders and consecutive coordinate polytopes.

	\item\label{ayyer:Coxeter-groups} Ayyer, Arvind and Prasad, Amritanshu and Spallone, Steven: Macdonald trees and
	      determinants of representations for finite Coxeter groups.

	\item Ayyer, Arvind and Sinha, Shubham: The size of t-cores and hook lengths of random cells in random partitions.

	\item Ayyer, Arvind: A simple symmetric exclusion process driven by an asymmetric tracer particle.

	\item\label{ayyer-singla:noncommutative-rings} Ayyer, Arvind and Singla, Pooja: Random motion on finite rings, II:
	      Noncommutative rings.

	\item\label{thangavelu:Heisenberg-group} Bagchi, Sayan and  Hait, Sourav and  Roncal, Luz and Thangavelu, Sundaram: On the lacunary spherical maximal function on the Heisenberg group.

	\item Bera, Sudip: Enumeration of weighted paths on a digraph and block hook determinant.

	\item Behera, Kiran Kumar: \href{https://arxiv.org/abs/1909.12548}{Self-inversive polynomials from rational transformation of linear functionals}.

	\item\label{bhattacharyya:pseudo-extensions} Bhattacharyya, Tirthankar and Das, B. Krishna  and Sau, Haripada: Toeplitz operators
	      and pseudo-extensions.

	\item Bhattacharyya, Tirthankar and Kumar, Poornendu and Sau, Haripada: Distinguished Varieties Through the Berger--Coburn--Lebow Theorem.

	\item Bhattacharyya, T. and  Biswas, A. and Maitra, A.: An unbounded realization of the symmetrized bidisc.

	\item Bhattacharyya, Tirthankar and  Roy, Subrata Shyam and Yadav, Tapesh: The Agler--Young Class.

	\item B{\"o}cherer, Siegfried and Das, Soumya: On fundamental Fourier coefficients of Siegel modular forms.

	\item Chattopahyay, Arup and Sarkar, Jaydeb and Sarkar, Srijan: \href{https://arxiv.org/abs/1812.05435}{Multiplicities, invariant subspaces and an additive formula.}.

	\item\label{datar:K3-surfaces} Datar, Ved and Jacob,Adam and Zhang, Yuguang: Adiabatic limits of anti-self-dual
	      connections on collapsed K3 surfaces, to appear in \emph{Journal of differential geometry}.

	\item\label{datar-pingali:Kahler-metrics} Datar, Ved and Pingali, Vamsi Pritham: On coupled constant scalar curvature Kaehler
	      metrics, to appear in \emph{Journal of symplectic geometry}.

	\item Ghara, Soumitra and Kumar, Surjit and Pramanick, Paramita: $K$-homogeneous tuple of operators on bounded symmetric domains.

	\item Gupta, Purvi and Wawrzyniak, Chloe Urbanski: STABILITY OF THE HULL(S) OF AN $n$-SPHERE IN $\mathbb{C}^n$.

	\item\label{khare:schur-polynomial-identities} Khare, Apoorva: Smooth entrywise positivity preservers, a
	      Horn-Loewner master theorem, and Schur polynomial identities.

	\item Ganguly, Pritam and Thangavelu,Sundaram: On lacunary spherical maximal function on Heisenberg groups.

	\item Khare, Apoorva: The non-compact normed space of norms on a finite-dimensional Banach space.

	\item Khare, Apoorva: Sharp nonzero lower bounds for the Schur product theorem.

	\item L\"ubeck, Frank and Prasad, Dipendra and Ayyer, Arvind: A character relationship between symmetric group and hyperoctahedral group.

	\item Maiti, Arun: Loop products on level homology and closed geodesics.

	\item Miermont, Gr\'egory and Sen, Sanchayan: On breadth-first constructions of scaling limits of random graphs and random unicellular maps.

	\item\label{Misra:Jordan-Keppler} Misra, Gadadhar and Upmeier, Harald: Singular Hilbert modules on
	      Jordan--Keppler varieties.

	\item Poria, Anirudha: \href{https://arxiv.org/abs/2004.04184}{Uncertainty Principles for the Fourier and the Short-Time Fourier Transforms}.

	\item Poria, Anirudha: \href{https://arxiv.org/abs/2005.14274}{Uncertainty principles for the Opdam-Cherednik transform on modulation spaces}.

	\item Toft, J. and Gumber, A. and Manna, R. and Ratnakumar, P. K.: \href{https://arxiv.org/abs/2004.02430v1}{Translation and modulation invariant Hilbert spaces}.
\end{enumerate}

\noindent{{\small {\sl In other areas:}}}\vspace{-2.5mm}

\begin{enumerate}[resume, leftmargin=27pt]

	\item Balodi, Mamta and Banerjee, Abhishek: Cycles over DGH-semicategories and pairings in categorical Hopf-cyclic cohomology.

	\item Balodi, Mamta: Morita invariance of equivariant and Hopf-cyclic cohomology of module algebras over Hopf algebroids.

	\item Biswas, Indranil and Gupta, Subhojoy and Mj, Mahan and Whang, Junho Peter: Surface group representations in $SL(2,\mathbb{C})$ with finite mapping class orbits.

	\item David Burns, Mahesh Kakde: On Weil-Etale Cohomology and Artin L-series.

	\item\label{khare:matrices-tree}  Choudhury, Projesh Nath and Khare, Apoorva: Distance matrices of a tree: two
	      more invariants, and in a unified framework.
	\item Garcia-Fernandez, Mario and Pingali, Vamsi Pritham and Yao, Chengjian: Gravitating vortices with positive curvature.

	\item Gupta, Subhojoy and Mj, Mahan: Meromorphic projective structures, grafting and the monodromy map.

	\item Gupta, Subhojoy: Monodromy groups of $CP^1$-structures on punctured surfaces.

	\item\label{singla:compact-linear-groups} Hassain, M. and Singla, Pooja; Representation Growth of Special Compact Linear Groups of Order Two.

	\item Hatui, Sumana and Singla Pooja: \href{https://arxiv.org/abs/1909.06589}{On Schur multiplier and Projective representations of Heisenberg groups}.

	\item Kus, Deniz and Venkatesh, R: Borel--de Siebenthal theory for affine reflection systems.

	\item Maiti, Arun: Optimal index growth of closed geodesics.

	\item\label{singla:discrete-valuation-rings} Patel, Shiv Prakash and Singla, Pooja: A multiplicity one theorem for groups of type $A_n$ over discrete valuation rings.

	\item Patil, Dilip P.: On the connectedness of the fiber product of two connected finte free affine schemes.

	\item Pingali, Vamsi Pritham: The deformed Hermitian Yang-Mills equation on three-folds.

	\item\label{pingali:Monge-Ampere-equation} Pingali, Vamsi Pritham: A vector bundle version of the Monge--Ampere equation.

	\item Quddus, Safdar: \href{https://arxiv.org/abs/2002.02192}{Equivariant Cartan Homotopy Formula for $DG$-algebra.}.

	\item Radhika Ganapathy: A Hecke algebra isomorphism over close local fields.

	\item Reif, Shifra and Venkatesh, R.: On tensor products of irreducible integrable representations.

\end{enumerate}

\subsubsection*{{\small $(ii)$ {\sc Research Papers in Conference Proceedings}}}

%% YEAR 1 LISTING %%%%%
\noindent{{\small {\bf April 2018--March 2019:}}}\vspace{-1mm}

\renewcommand{\theenumi}{C.\arabic{enumi}}
\begin{enumerate}%[leftmargin=27pt]

	\item Kanukurthi, Bhavana and Obbattu, Sai Lakshmi Bhavana and
	      Sekar, Sruthi: Non-malleable randomness encoders and their applications, {\em Advances in
	      Cryptology\,---\,\linebreak{EUROCRYPT} 2018: {P}art {III}} {\bf 10822} (2018), 589--617.

	\item S. Ghara and G. Misra: Decomposition of the tensor product of two Hilbert modules in {\em Operator Theory, Operator Algebras and Their Interactions with Geometry and Topology Ronald G. Douglas Memorial Volume} Curto, R.E., Helton, W., Lin, H., Tang, X., Yang, R., Yu, G. (ed.), Operator Theory: Advances and Applications , Birkhauser (2020).

	\item Gadadhar Misra: Decomposition of the tensor product of two Hilbert modules in {\em Operator Theory, Operator Algebras and Their Interactions with Geometry and Topology Ronald G. Douglas Memorial Volume} R. Curto, W. Helton, H. Lin, X. Tang, R. Yang, G. Yu (ed.), Operator Theory: Advances and Applications , Birkhauser (2020).

	\item\label{patil:Nullstellensaetze} Goel, Kriti and Patil, Dilip P. and Verma, Jugal: Nullstellens\"atze and Applications, {\em Proceedings of the International Conference on Mathematics and Centenary Celebrations of Ramjas College}, forthcoming.

	\item\label{patil:generalized-trace-forms} Patil, Dilip P.  and Verma, Jugal: Rational points and generalized trace forms on a finite algebra over a real closed field, {\em Proceedings of the International Conference on Computational Mathematics,  Banasthali Vidyapeeth}.
\end{enumerate}

%% YEAR 2 LISTING %%%%%
\noindent{{\small {\bf April 2019--March 2020:}}}\vspace{-1mm}

\begin{enumerate}[resume]%[leftmargin=27pt]

	\item Banerjee, A.: Spectra as universal objects in categories of supports in {\em Banach Center Publications(Quantum Dynamics Proceedings)} Ludwik Dabrowski, Ryszard Nest, Adam Skalski (ed.), Banach Center Publications , IMPAN, Warsaw, Poland (2020).

\end{enumerate}

\subsubsection*{{\small $(iii)$ {\sc Books \& Monographs}}}\vspace{-1mm}

%% YEAR 1 LISTING %%%%%
\noindent{{\small {\bf April 2018--March 2019:}}}

\renewcommand{\theenumi}{B.\arabic{enumi}}
\begin{enumerate}%[leftmargin=27pt]

	\item Belton, Alexander and Guillot, Dominique and Khare, Apoorva and Putinar, Mihai: A panorama of positivity: I.
	      Dimension, in {\em Analysis of Operators on Function Spaces (The Serguei Shimorin Memorial Volume)}, eds.
	      A.~Aleman, H.~Hedenmalm, D.~Khavinson, M.~Putinar, Trends in Mathematics,  Birkhauser, pages 117--165, 2019.

	\item Belton, Alexander and Guillot, Dominique and Khare, Apoorva and Putinar, Mihai: A panorama of positivity: II.
	      Fixed dimension, in {\em Complex Analysis and Spectral Theory}, eds.  G.~Dales, D.~Khavinson, J.~Mashreghi, Eds.,
	      CRM Proceedings, AMS  Contemporary Mathematics, American Mathematical Society, 2019.

\end{enumerate}

%% YEAR 2 LISTING %%%%%
\noindent{{\small {\bf April 2019--March 2020:}}}

\begin{enumerate}[resume]%[leftmargin=27pt]

	\item Gupta, Subhojoy and Seshadri, Harish: Complex geometry of Teichm{\"u}ller domains in {\em Handbook
			      of Teichm{\"u}ller Theory} Papadopoulos, Athanase (ed.), IRMA Lect. Math. Theor. Phys, Eur. Math. Soc., Z{\"u}rich
	      (2020), 26.

	\item Gupta, Subhojoy: Holomorphic quadratic differentials in Teichm{\"u}ller theory in {\em Handbook
			      of Teichm{\"u}ller Theory} Papadopoulos, Athanase (ed.), IRMA Lect. Math. Theor. Phys., Eur. Math. Soc., Z{\"u}rich
	      (2020), 36.

	\item Misra, Gadadhar: Operators in the {C}owen-{D}ouglas class and related topics, \emph{Handbook of analytic operator theory} {\bf } (2019), 87--137.
\end{enumerate}