### 2020-08-10 sorted.tex

\item Alvarez-Consul, Luis and Garcia-Fernandez, Mario  and García-Prada, Oscar and Pingali, Vamsi Pritham: Gravitating vortices and the Einstein–Bogomol’nyi equations, to appear in \emph{Mathematische Annalen}.

\item Basu, R. and Ganguly, S. and Hegde, M. and Krishnapur, M.: Lower Deviations in β-ensembles and Law of Iterated Logarithm in Last Passage Percolation, to appear in \emph{Israel Journal of Mathematics}.

\item Ghosh, M. K. and K. Suresh Kumar and C. Pal and S. Pradhan: Nonzero-sum risk-sensitive stochastic differential games with discounted costs, Stochastic Analysis and Applications, to appear in \emph{Stochastic Analysis and Applications}.

\item Ghosh, Subhajit: Total Variation Cutoff for the Transpose Top-2 with Random Shuffle, to appear in \emph{Journal of Theoretical Probability}.

\item Ghosh, Subhajit: Total variation cutoff for the transpose top-2 with random shuffle, to appear in {\em Journal of Theoretical Probability}.

\item Gudi, Thirupathi and  Sau, Ramesh Ch.: Finite Element Analysis of the Constrained Dirichlet Boundary Control Problem Governed by the Diffusion Problem, to appear in \emph{ESAIM: Control, Optim. Calc. Var}.

\item \label{bharali:Wolff-Denjoy} Bharali, Gautam and Maitra, Anwoy: A weak notion of visibility,
      a family of examples, and Wolff--Denjoy theorems, to appear in \emph{Annali della Scuola Normale Superiore di Pisa}.

\item \label{iyer:heterogeneous-populations} Nair, Gokul and Athmanathan, Senthil and Iyer, Srikanth K. and Guttal, Vishwesha: Fission-fusion dynamics and group-size-dependent composition in heterogeneous populations,
      to appear in {\em Physical Review E}.

\item \label{iyer:wireless-networks} Vaze, Rahul and Iyer, Srikanth K., Capacity of cellular wireless networks,
      to appear in {\em IEEE Transactions on Wireless Communications}.

\item \label{pingali:algebraic-hypersurfaces} Pingali, Vamsi Pritham and Varolin, Dror: Non-uniformly flat affine algebraic hypersurfaces, to appear in {\em Nagoya Mathematical Journal}.

\item \label{sen:models-at-criticality} Dhara, Souvik and van der Hofstad, Remco and van Leeuwaarden, Johan S.H. and Sen, Sanchayan: Heavy-tailed configuration models at criticality, to appear in {\em Annales de l'Institut Henri
              Poincar\'e (B) Probabilit\'es et Statistiques}.

\item \label{thangavelu:H-type-groups} Roncal, Luz and Thangavelu, Sundaram: An extension problem and trace Hardy
      inequality for the sublaplacian on H-type groups, to appear in {\em International Mathematical Research Notices}.