### 2021-08-19 data/draft-publications.tex
publications for DST-FIST, timetable

\subsection{Papers in Journals (with links)}

\begin{enumerate}
\item S Aiyappan, A. K. Nandakumaran and Abu Sufian: Asymptotic analysis of Boundary Optimal Control Problem on a General Branched Structure, \emph{Mathematical Methods in the Applied Sciences} {\bf 42} (2019. DOI: 10.1002/mma.5748), 6407-6434.
\item K. Murali, A. K. Nandakumaran, Turgurt Durduran, Hari M Varma: Recovery of the diffuse correlation spectroscopy data-type from speckle contrast measurements: towards low-cost, deep-tissue blood flow measurements, \emph{Biomedical Optics Express 5395} {\bf 10} (2019).
\item S Aiyappan, A. K. Nandakumaran and Ravi Prakash: Locally Periodic Unfolding Operator for Highly Oscillating Rough Domains, \emph{Annali di Matematica Pura ed Applicata} {\bf 198} (2019), 1931-1954.
\item S Aiyappan, A. K. Nandakumaran and Ravi Prakash: Semi-linear optimal control problem on a smooth oscillating domain, \emph{Communications in Contemporary Mathematics} {\bf } (2019), 1-26.
\item S Aiyappan, Editha C. Jose, Ivy Carol B. Lomerio and A. K. Nandakumaran: Control Problem on a Rough Circular Domain and Homogenization, \emph{Asymptotic Analysis} {\bf 115} (2019), 19-46.
\item A. Banerjee: On differential torsion theories and rings with several objects, \emph{Canadian Math Bulletin} {\bf 62} (2019), 703--714.
\item Khare, Apoorva and Rajaratnam, Bala: Probability inequalities and tail estimates for metric semigroups, \emph{Advances in Operator Theory (Special issue in honor of Rajendra Bhatia)} {\bf 5} (2020), 779-795.
\item Datta, B. and Sarkar, S.: Equilibrium triangulations of some quasitoric 4-manifolds, \emph{Southeast Asian Bull. Math.} {\bf 44} (2020), 57--78.
\item Datta, B. and Maity, D.: Correction to: Semi-equivelar and vertex-transitive maps on the torus, \emph{Beitr Algebra Geom} {\bf 61} (2020), 187--188.
\item Gadadhar Misra: Professor, \emph{Adv. Math.} {\bf 351} (2019), 1105 - 1138.
\item Maitra, Anwoy: On the continuous extension of Kobayashi isometries, \emph{Proc. Amer. Math. Soc.} {\bf 148} (2020), 3437--3451.
\item Vishnu Priya, N. and Senthilvelan, M. and  Rangarajan, G. and Lakshmanan, M: On symmetry preserving and symmetry broken bright, dark and antidark soliton solutions of nonlocal nonlinear Schrödinger equation, \emph{Physics Letters A} {\bf 383} (2019), 15-26.
\item Kashyap, G. A. R. S. R. K. and Bapat, D. and Das, D. and Gowaikar, R. D. and Amritkar, R. E. and Rangarajan, G. and Ravindranath, V. and Ambika, G: Synapse loss and progress of Alzheimer’s disease - A network model, \emph{Scientific Reports} {\bf 9} (2019), 6555.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan, G.: On the role of four-wave mixing effect in the interactions between nonlinear modes of coupled generalized nonlinear Schrodinger equation, \emph{Chaos} {\bf 29} (2019), 123135.
\item Sarkar, Amar Deep and Verma, Kaushal: A submultiplicative property of the Carath\'{e}odory metric on planar domains, \emph{Proc. Indian Acad. Sci. (Math. Sci.)} {\bf } ().
\item Frauke Bleher, Ted Chinburg, Ralph Greenberg, Mahesh Kakde, George Pappas, Romyar Sharifi, Martin Taylor: Higher Chern Classes in Iwasawa Theory, \emph{American Journal of Mathematics} {\bf 142} (2020), 627-682.
\item Iyer, S. K. and Yogeshwaran, D.: Thresholds for vanishing of isolated faces in random Cech and Vietoris-Rips Complexes, \emph{Annals de l'Institut Henri Poincar$\acute{e}$ - Probabiliti$\acute{e}$s et Statistiques} {\bf 56} (2020), 1869-1897.
\item Bhattacharyya, Tirthankar and Biswas, Anindya and Chandel, Vikramjeet Singh: On the Nevanlinna problem: characterization of all Schur–Agler class solutions affiliated with a given kernel, \emph{Studia Math.} {\bf 255} (2020), 83–107..
\item Bhattacharyya, Tirthankar and Das, Bata Krishna and Sau, Haripada: Toeplitz operators on the symmetrized bidisc, \emph{International Mathematics Research Notices (IMRN)} {\bf } ().
\item Arun Maiti: \href{https://www.sciencedirect.com/science/article/pii/S0012365X20301035}{Quasi-vertex-transitive maps on the plane}, \emph{Discrete Mathematics} {\bf 343} (2020), 111911-111917.
\item Carstensen, C. and Mallik, G. and Nataraj, N.: \href{https://doi.org/10.1093/imanum/drz071}{Nonconforming Finite Element Discretization for Semilinear Problems with Trilinear Nonlinearity}, \emph{IMA Journal of Numerical Analysis} {\bf } (2020).
\item Chim, Kwok Chi; Shorey, Tarlok Nath; Sinha, Sneh Bala: On Baker's explicit abc-conjecture, \emph{Publ. Math. Debrecen} {\bf 94} (2019), 435–453.
\item Choudhury, Projesh Nath and Tsatsomeros Michael J.: \href{https://doi.org/10.1016/j.laa.2020.04.028}{Algorithmic Detection and Construction of N-matrices}, \emph{Linear Algebra and Its Applications} {\bf 602} (2020), 46-56.
\item Datt, Gopal and Kumar, Sanjay: \href{https://link.springer.com/article/10.1007/s12044-018-0449-5?shared-article-renderer}{A criterion for quasinormality in $\mathbb{C}^n$}, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 129} (2019), Paper No. 11, 13 pp..
\item Hatui, Sumana: \href{https://www.degruyter.com/view/journals/jgth/23/1/article-p85.xml}{Schur multipliers of special p-groups of rank 2}, \emph{Journal of Group Theory} {\bf 23} (2020), 85-95.
\item Mal, Arpita and Sain, Debmalya and Paul, Kallol: \href{https://projecteuclid.org/euclid.bjma/1543914019}{On some geometric properties of operator spaces}, \emph{Banach J. Math. Anal.} {\bf 13} (2019), 174-191.
\item Pradhan, P. and Sahu, B.: \href{https://doi.org/10.1016/j.disc.2020.112044}{A characterization of the family of secant lines to a hyperbolic quadric in PG(3,q), q odd}, \emph{Discrete Mathematics} {\bf 343} (2020).
\item Roth, Julien and Upadhyay, Abhitosh: \href{https://link.springer.com/article/10.1007/s00025-019-1142-4?shared-article-renderer#citeas}{f-Biharmonic Submanifolds of Generalized Space Forms}, \emph{Results in Mathematics} {\bf 75} (2020).
\item Roth, Julien and Upadhyay, Abhitosh: \href{https://www.cambridge.org/core/journals/bulletin-of-the-australian-mathematical-society/article/on-almost-stable-cmc-hypersurfaces-in-manifolds-of-bounded-sectional-curvature/E0BF84D4D46C82F6D75CDCEAAF5DEE94}{On almost stable CMC hypersurfaces in manifolds of bounded sectional curvature}, \emph{Bulletin of the Australian Mathematical Society} {\bf 101} (2020), 333-338.
\end{enumerate}

\subsection{Accepted Papers}

\begin{enumerate}
\item M. Balodi, A. Banerjee, S. Ray: Cohomology of modules over H-categories and co-H-categories, to appear in \emph{Canadian Journal of Mathematics}.
\item M. Balodi, A. Banerjee, S. Ray: Entwined modules over linear categories and Galois extensions, to appear in \emph{Israel Journal of Mathematics (Accepted)}.
\item A. Banerjee: On Noetherian schemes over $(\mathcal C,\otimes,1)$ and the category of quasi-coherent sheaves, to appear in \emph{Indiana University Mathematics Journal}.
\item Ghosh, Subhajit: Total Variation Cutoff for the Transpose Top-2 with Random Shuffle, to appear in \emph{Journal of Theoretical Probability}.
\item Datta, B. and Gupta, S.: Semi-regular tilings of the hyperbolic plane \label{bdsg_2019}, to appear in \emph{Discrete Comput Geom}.
\item Martin Kreuzer and Dilip P. Patil: COMPUTATIONAL ASPECTS OF BURNSIDE RINGS PART II: IMPORTANT MAPS, to appear in \emph{Beiträge zur Algebra und Geometrie / Contributions to Algebra and Geometry}.
\item Bharali, Gautam and Maitra, Anwoy: A weak notion of visibility, a family of examples, and Wolff--Denjoy theorems, to appear in \emph{Annali della Scuola Normale Superiore di Pisa}.
\item Basu, R. and Ganguly, S. and Hegde, M. and Krishnapur, M.: Lower Deviations in β-ensembles and Law of Iterated Logarithm in Last Passage Percolation, to appear in \emph{Israel Journal of Mathematics}.
\item M. K. Ghosh and S. Pradhan: Zero-Sum Risk-Sensitive Stochastic Differential Games with Reflecting Diffusions in the Orthant, to appear in \emph{ESIAM: Control, Optimization and Calculus of Variations}.
\item M. K. Ghosh, K. Suresh Kumar, C. Pal and S. Pradhan: Nonzero-sum risk-sensitive stochastic differential games with discounted costs, Stochastic Analysis and Applications, to appear in \emph{Stochastic Analysis and Applications}.
\item Datta, Basudeb and Gupta, Subhojoy: Semi-regular tilings of the hyperbolic plane, to appear in \emph{Discrete and Computational Geometry}.
\item Thirupathi Gudi, Ramesh Ch. Sau: Finite Element Analysis of the Constrained Dirichlet Boundary Control Problem Governed by the Diffusion Problem, to appear in \emph{ESAIM: Control, Optim. Calc. Var}.
\item Pingali, Vamsi Pritham and Varolin, Dror: Nonuniformly flat affine algebraic hypersurfaces, to appear in \emph{Nagoya Mathematical Journal}.
\item Alvarez-Consul, Luis and Garcia-Fernandez, Mario  and García-Prada, Oscar and Pingali, Vamsi Pritham: Gravitating vortices and the Einstein–Bogomol’nyi equations, to appear in \emph{Mathematische Annalen}.
\item Datar, Ved and Jacob,Adam and Zhang, Yuguang: Adiabatic limits of anti-self-dual connections on collapsed K3 surfaces, to appear in \emph{Journal of differential geometry}.
\item Datar, Ved and Pingali, Vamsi Pritham: On coupled constant scalar curvature Kaehler metrics, to appear in \emph{Journal of symplectic geometry}.
\item Barbosa, Ezequiel and Santana, Farley and Upadhyay, Abhitosh: On stable CMC free-boundary surfaces in a strictly convex domain of a bi-invariant Lie group, to appear in \emph{International Journal of Mathematics}.
\item Choudhury, Projesh Nath and Sivakumar, K.C.: \href{https://doi.org/10.1080/03081087.2019.1602588}{Matrices with Positive Semidefinite Real Part}, to appear in \emph{Linear and Multilinear Algebra}.
\end{enumerate}

\subsection{Papers in Conference proceedings}

\begin{enumerate}
\item A. Banerjee: Spectra as universal objects in categories of supports in {\em Banach Center Publications(Quantum Dynamics Proceedings)} Ludwik Dabrowski, Ryszard Nest, Adam Skalski (ed.) , Banach Center Publications , IMPAN, Warsaw, Poland (2020), (accepted).
\item S. Ghara and G. Misra: Decomposition of the tensor product of two Hilbert modules in {\em Operator Theory, Operator Algebras and Their Interactions with Geometry and Topology Ronald G. Douglas Memorial Volume} Curto, R.E., Helton, W., Lin, H., Tang, X., Yang, R., Yu, G. (ed.) , Operator Theory: Advances and Applications , Birkhauser (2020).
\item Gadadhar Misra: Decomposition of the tensor product of two Hilbert modules in {\em Operator Theory, Operator Algebras and Their Interactions with Geometry and Topology Ronald G. Douglas Memorial Volume} R. Curto, W. Helton, H. Lin, X. Tang, R. Yang, G. Yu (ed.) , Operator Theory: Advances and Applications , Birkhauser (2020).
\end{enumerate}

\subsection{Books}

\subsection{Book Chapters}

\begin{enumerate}
\item Gupta, Subhojoy and Seshadri, Harish: Complex geometry of Teichmüller domains in {\em Handbook of Teichmüller theory} Papadopoulos, Athanase (ed.), IRMA Lect. Math. Theor. Phys, Eur. Math. Soc., Zürich (2020), 26.
\item Gupta, Subhojoy: Holomorphic quadratic differentials in Teichmüller theory in {\em Handbook of Teichmüller theory} Papadopoulos, Athanase (ed.), IRMA Lect. Math. Theor. Phys., Eur. Math. Soc., Zürich (2020), 36.
\end{enumerate}

\subsection{Preprints with links}

\begin{enumerate}
\item Khare, Apoorva: Sharp nonzero lower bounds for the Schur product theorem, submitted for publication.
\item Lübeck, Frank and Prasad, Dipendra and Ayyer, Arvind: A character relationship between symmetric group and hyperoctahedral group, preprint.
\item Ayyer, Arvind and Sinha, Shubham: The size of t-cores and hook lengths of random cells in random partitions, preprint.
\item Ayyer, Arvind: A simple symmetric exclusion process driven by an asymmetric tracer particle, preprint.
\item David Burns, Mahesh Kakde: On Weil-Etale Cohomology and Artin L-series, preprint.
\item Gupta, Purvi and Wawrzyniak, Chloe Urbanski: STABILITY OF THE HULL(S) OF AN $n$-SPHERE IN $\mathbb{C}^n$, submitted for publication.
\item Radhika Ganapathy: A Hecke algebra isomorphism over close local fields, preprint.
\item Miermont, Grégory and Sen, Sanchayan: On breadth-first constructions of scaling limits of random graphs and random unicellular maps, submitted for publication.
\item Gupta, Subhojoy and Mj, Mahan: Meromorphic projective structures, grafting and the monodromy map, preprint.
\item Gupta, Subhojoy: Monodromy groups of $CP^1$-structures on punctured surfaces, preprint.
\item Biswas, Indranil and Gupta, Subhojoy and Mj, Mahan and Whang, Junho Peter: Surface group representations in SL(2,C) with finite mapping class orbits, preprint.
\item Ganguly, Pritam and Thangavelu,Sundaram: On lacunary spherical maximal function on Heisenberg groups, preprint.
\item Bhattacharyya, Tirthankar and Kumar, Poornendu and Sau, Haripada: Distinguished Varieties Through the Berger--Coburn--Lebow Theorem, submitted for publication.
\item Bhattacharyya, T. and  Biswas, A. and Maitra, A.: An unbounded realization of the symmetrized bidisc, submitted for publication.
\item Pingali, Vamsi Pritham: The deformed Hermitian Yang-Mills equation on three-folds, preprint.
\item Garcia-Fernandez, Mario and Pingali, Vamsi Pritham and Yao, Chengjian: Gravitating vortices with positive curvature, preprint.
\item Arun Maiti: OPTIMAL INDEX GROWTH OF CLOSED GEODESICS, submitted for publication.
\item Behera, Kiran Kumar: \href{https://arxiv.org/abs/1909.12548}{Self-inversive polynomials from rational transformation of linear functionals}, submitted for publication.
\item Bera, Sudip: ENUMERATION OF WEIGHTED PATHS ON A DIGRAPH AND BLOCK HOOK DETERMINANT, submitted for publication.
\item Chattopahyay, Arup and Sarkar, Jaydeb and Sarkar, Srijan: \href{https://arxiv.org/abs/1812.05435}{Multiplicities, invariant subspaces and an additive formula.}, preprint.
\item Ghara, Soumitra and Kumar, Surjit and Pramanick, Paramita: $K$-homogeneous tuple of operators on bounded symmetric domains, preprint.
\item Hatui, Sumana and Singla Pooja: \href{https://arxiv.org/abs/1909.06589}{On Schur multiplier and Projective representations of Heisenberg groups}, submitted for publication.
\item Poria, Anirudha: \href{https://arxiv.org/abs/2004.04184}{Uncertainty Principles for the Fourier and the Short-Time Fourier Transforms}, submitted for publication.
\item Poria, Anirudha: \href{https://arxiv.org/abs/2005.14274}{Uncertainty principles for the Opdam-Cherednik transform on modulation spaces}, submitted for publication.
\item Quddus, Safdar: \href{https://arxiv.org/abs/2002.02192}{Equivariant Cartan Homotopy Formula for $DG$-algebra.}, preprint.
\item Toft, J. and Gumber, A. and Manna, R. and Ratnakumar, P. K.: \href{https://arxiv.org/abs/2004.02430v1}{Translation and modulation invariant Hilbert spaces}, submitted for publication.
\end{enumerate}

% from web page

\subsection{Publications (all kinds) from web page}

\begin{enumerate}
\item Aas, Erik and Ayyer, Arvind and Linusson, Svante and Potka,
Samu: The exact phase diagram for a semipermeable {TASEP} with
nonlocal boundary jumps, \emph{J. Phys. A} {\bf 52} (2019), 355001, 19.
\item Adhikari, Kartick and Kishore Reddy, Nanda and Ram Reddy,
Tulasi and Saha, Koushik: Determinantal point processes in the plane from products of
random matrices, \emph{Ann. Inst. Henri Poincar\'{e} Probab. Stat.} {\bf 52} (2016), 16--46.
\item Adhikari, Kartick and Reddy, Nanda Kishore: Hole probabilities for finite and infinite {G}inibre
ensembles, \emph{Int. Math. Res. Not. IMRN} {\bf } (2017), 6694--6730.
\item Adimurthi and Ratnakumar, P. K. and Sohani, Vijay Kumar: A {H}ardy-{S}obolev inequality for the twisted {L}aplacian, \emph{Proc. Roy. Soc. Edinburgh Sect. A} {\bf 147} (2017), 1--23.
\item Aiyappan, S. and Jose, Editha C. and Lomerio, Ivy Carol B. and
Nandakumaran, A. K.: Control problem on a rough circular domain and homogenization, \emph{Asymptot. Anal.} {\bf 115} (2019), 19--46.
\item Aiyappan, S. and Nandakumaran, A. K.: Optimal control problem in a domain with branched structure
and homogenization, \emph{Math. Methods Appl. Sci.} {\bf 40} (2017), 3173--3189.
\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi: Semi-linear optimal control problem on a smooth oscillating
domain, \emph{Commun. Contemp. Math.} {\bf 22} (2020), 1950029, 26.
\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi: Locally periodic unfolding operator for highly oscillating
rough domains, \emph{Ann. Mat. Pura Appl. (4)} {\bf 198} (2019), 1931--1954.
\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi: Generalization of unfolding operator for highly oscillating
smooth boundary domains and homogenization, \emph{Calc. Var. Partial Differential Equations} {\bf 57} (2018), Art. 86, 30.
\item Aiyappan, S. and Nandakumaran, A. K. and Sufian, Abu: Asymptotic analysis of a boundary optimal control problem on a
general branched structure, \emph{Math. Methods Appl. Sci.} {\bf 42} (2019), 6407--6434.
\item Aiyappan, S. and Sardar, Bidhan Chandra: Biharmonic equation in a highly oscillating domain and
homogenization of an associated control problem, \emph{Appl. Anal.} {\bf 98} (2019), 2783--2801.
\item Akutagawa, Kazuo and Endo, Hisaaki and Seshadri, Harish: A gap theorem for positive {E}instein metrics on the
four-sphere, \emph{Math. Ann.} {\bf 373} (2019), 1329--1339.
\item Anamby, Pramath and Das, Soumya: Distinguishing {H}ermitian cusp forms of degree 2 by a certain
subset of all {F}ourier coefficients, \emph{Publ. Mat.} {\bf 63} (2019), 307--341.
\item Anamby, Pramath and Das, Soumya: Bounds for the {P}etersson norms of the pullbacks of
{S}aito-{K}urokawa lifts, \emph{J. Number Theory} {\bf 191} (2018), 289--304.
\item Anand, Jatin and Bhattacharyya, Tirthankar and Srivastava,
Sachi: Some thoughts on composition operators on subspaces of the
{H}ardy space, \emph{Arch. Math. (Basel)} {\bf 114} (2020), 431--444.
\item Arunkumar, G. and Kus, Deniz and Venkatesh, R.: Root multiplicities for {B}orcherds algebras and graph
coloring, \emph{J. Algebra} {\bf 499} (2018), 538--569.
\item Ayyer, Arvind: Full current statistics for a disordered open exclusion
process, \emph{J. Phys. A} {\bf 49} (2016), 155003, 10.
\item Ayyer, Arvind: Squareness for the {M}onopole-{D}imer model, \emph{Ann. Comb.} {\bf 24} (2020), 237--255.
\item Ayyer, Arvind and Behrend, Roger E.: Factorization theorems for classical group characters, with
applications to alternating sign matrices and plane
partitions, \emph{J. Combin. Theory Ser. A} {\bf 165} (2019), 78--105.
\item Ayyer, Arvind and Behrend, Roger E. and Fischer, Ilse: Extreme diagonally and antidiagonally symmetric alternating
sign matrices of odd order, \emph{Adv. Math.} {\bf 367} (2020), 107125, 56.
\item Ayyer, Arvind and Bouttier, J\'{e}r\'{e}mie and Corteel, Sylvie and
Linusson, Svante and Nunzi, Fran\c{c}ois: Bumping sequences and multispecies juggling, \emph{Adv. in Appl. Math.} {\bf 98} (2018), 100--126.
\item Ayyer, Arvind and Finn, Caley and Roy, Dipankar: Matrix product solution of a left-permeable two-species
asymmetric exclusion process, \emph{Phys. Rev. E} {\bf 97} (2018), 012151, 10.
\item Ayyer, Arvind and Finn, Caley and Roy, Dipankar: The phase diagram for a multispecies left-permeable asymmetric
exclusion process, \emph{J. Stat. Phys.} {\bf 174} (2019), 605--621.
\item Ayyer, Arvind and Fischer, Ilse: Bijective proofs of skew {S}chur polynomial factorizations, \emph{J. Combin. Theory Ser. A} {\bf 174} (2020), 105241, 40.
\item Ayyer, Arvind and Josuat-Verg\`es, Matthieu and Ramassamy,
Sanjay: Extensions of partial cyclic orders and consecutive coordinate
polytopes, \emph{Ann. H. Lebesgue} {\bf 3} (2020), 275--297.
\item Ayyer, Arvind and Linusson, Svante: Correlations in the multispecies {TASEP} and a conjecture by
{L}am, \emph{Trans. Amer. Math. Soc.} {\bf 369} (2017), 1097--1125.
\item Ayyer, Arvind and Linusson, Svante: Reverse juggling processes, \emph{Random Structures Algorithms} {\bf 55} (2019), 56--72.
\item Ayyer, Arvind and Prasad, Amritanshu and Spallone, Steven: Odd partitions in {Y}oung's lattice, \emph{S\'{e}m. Lothar. Combin.} {\bf 75} (2018), Art. B75g, 13.
\item Ayyer, Arvind and Prasad, Amritanshu and Spallone, Steven: Representations of symmetric groups with non-trivial
determinant, \emph{J. Combin. Theory Ser. A} {\bf 150} (2017), 208--232.
\item Ayyer, Arvind and Ramassamy, Sanjay: The {H}ilbert-{G}alton board, \emph{ALEA Lat. Am. J. Probab. Math. Stat.} {\bf 15} (2018), 755--774.
\item Ayyer, Arvind and Schilling, Anne and Thi\'{e}ry, Nicolas M.: Spectral gap for random-to-random shuffling on linear
extensions, \emph{Exp. Math.} {\bf 26} (2017), 22--30.
\item Ayyer, Arvind and Singla, Pooja: Random motion on finite rings, {I}: commutative rings, \emph{Algebr. Represent. Theory} {\bf 23} (2020), 583--604.
\item Ayyer, Arvind and Sinha, Shubham: Random {$t$}-cores and hook lengths in random partitions, \emph{S\'{e}m. Lothar. Combin.} {\bf 84B} (2020), Art. 58, 11.
\item Ayyer, Arvind and Steinberg, Benjamin: Random walks on rings and modules, \emph{Algebr. Comb.} {\bf 3} (2020), 309--329.
\item B\"{o}cherer, Siegfried and Das, Soumya: Cuspidality and the growth of {F}ourier coefficients: small
weights, \emph{Math. Z.} {\bf 283} (2016), 539--553.
\item B\"{o}cherer, Siegfried and Das, Soumya: Cuspidality and the growth of {F}ourier coefficients of
modular forms, \emph{J. Reine Angew. Math.} {\bf 741} (2018), 161--178.
\item B\"{o}cherer, Siegfried and Das, Soumya: Petersson norms of not necessarily cuspidal {J}acobi modular
forms and applications, \emph{Adv. Math.} {\bf 336} (2018), 335--376.
\item Bagchi, Bhaskar and Burton, Benjamin A. and Datta, Basudeb and
Singh, Nitin and Spreer, Jonathan: Efficient algorithms to decide tightness, \emph{32nd {I}nternational {S}ymposium on {C}omputational
{G}eometry} {\bf 51} (2016), Art. 12, 15.
\item Bagchi, Bhaskar and Datta, Basudeb and Spreer, Jonathan: A characterization of tightly triangulated 3-manifolds, \emph{European J. Combin.} {\bf 61} (2017), 133--137.
\item Bagchi, Bhaskar and Datta, Basudeb and Spreer, Jonathan: Tight triangulations of closed 3-manifolds, \emph{European J. Combin.} {\bf 54} (2016), 103--120.
\item Bagchi, Sayan and Thangavelu, Sundaram: Weighted norm inequalities for {W}eyl multipliers and fourier
multipliers on the {H}eisenberg group, \emph{J. Anal. Math.} {\bf 136} (2018), 1--29.
\item Baklouti, Ali and Thangavelu, Sundaram: Hardy and {M}iyachi theorems for {H}eisenberg motion groups, \emph{Nagoya Math. J.} {\bf 229} (2018), 1--20.
\item Balakumar, G. P. and Borah, Diganta and Mahajan, Prachi and
Verma, Kaushal: Further remarks on the higher dimensional {S}uita conjecture, \emph{Ann. Polon. Math.} {\bf 125} (2020), 101--115.
\item Balakumar, G. P. and Borah, Diganta and Mahajan, Prachi and
Verma, Kaushal: Remarks on the higher dimensional {S}uita conjecture, \emph{Proc. Amer. Math. Soc.} {\bf 147} (2019), 3401--3411.
\item Balhara, Rakesh: Hardy's inequality for the fractional powers of the {G}rushin
operator, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 129} (2019), Art. 33, 25.
\item Balodi, Mamta and Banerjee, Abhishek and Naolekar, Anita: B{V}-operators and the secondary {H}ochschild complex, \emph{C. R. Math. Acad. Sci. Paris} {\bf 358} (2020), 1239--1258.
\item Balodi, Mamta and Banerjee, Abhishek and Ray, Samarpita: Cohomology of modules over {$H$}-categories and
co-{$H$}-categories, \emph{Canad. J. Math.} {\bf 72} (2020), 1352--1385.
\item Balodi, Mamta and Huang, Hua-Lin and Kumar, Shiv Datt: On the classification of finite quasi-quantum groups, \emph{Rev. Math. Phys.} {\bf 29} (2017), 1730003, 20.
\item Balodi, Mamta and Palcoux, Sebastien: On {B}oolean intervals of finite groups, \emph{J. Combin. Theory Ser. A} {\bf 157} (2018), 49--69.
\item Banerjee, Abhishek: On some spectral spaces associated to tensor triangulated
categories, \emph{Arch. Math. (Basel)} {\bf 108} (2017), 581--591.
\item Banerjee, Abhishek: On {A}uslander's formula and cohereditary torsion pairs, \emph{Commun. Contemp. Math.} {\bf 20} (2018), 1750071, 27.
\item Banerjee, Abhishek: Modular {H}ecke algebras over {M}\"{o}bius categories, \emph{J. Geom. Phys.} {\bf 131} (2018), 23--40.
\item Banerjee, Abhishek: Closure operators in abelian categories and spectral spaces, \emph{Theory Appl. Categ.} {\bf 32} (2017), Paper No. 20, 719--735.
\item Banerjee, Abhishek: Noetherian schemes over abelian symmetric monoidal categories, \emph{Internat. J. Math.} {\bf 28} (2017), 1750051, 34.
\item Banerjee, Abhishek: On differential torsion theories and rings with several
objects, \emph{Canad. Math. Bull.} {\bf 62} (2019), 703--714.
\item Banerjee, Abhishek: Une note sur les d\'erivations localement nilpotentes dans les
cat\'egories mono\"\i dales sym\'etriques, \emph{Bull. Belg. Math. Soc. Simon Stevin} {\bf 23} (2016), 151--160.
\item Banerjee, Abhishek: Realizations of pairs and {O}ka families in tensor
triangulated categories, \emph{Eur. J. Math.} {\bf 2} (2016), 760--797.
\item Banerjee, Abhishek: Homologie de {H}ochschild et espaces alg\'ebriques, \emph{Math. Scand.} {\bf 118} (2016), 33--42.
\item Banerjee, Abhishek: A topological {N}ullstellensatz for tensor-triangulated
categories, \emph{C. R. Math. Acad. Sci. Paris} {\bf 356} (2018), 365--375.
\item Banerjee, Abhishek: Some remarks on a theorem of {B}ergman, \emph{C. R. Math. Acad. Sci. Paris} {\bf 354} (2016), 665--670.
\item Banerjee, Abhishek: Monoidal semifilters and arrays of prime ideals, \emph{Fund. Math.} {\bf 237} (2017), 281--296.
\item Banerjee, Abhishek: Schemes over symmetric monoidal categories and torsion
theories, \emph{J. Pure Appl. Algebra} {\bf 220} (2016), 3017--3047.
\item Banerjee, Abhishek: Quasimodular {H}ecke algebras and {H}opf actions, \emph{J. Noncommut. Geom.} {\bf 12} (2018), 1041--1080.
\item Banerjee, Abhishek: Completions of monoid objects and descent results, \emph{J. Algebra} {\bf 507} (2018), 362--399.
\item Banerjee, Abhishek and Kour, Surjeet: {$(A,\delta)$}-modules, {H}ochschild homology and higher
derivations, \emph{Ann. Mat. Pura Appl. (4)} {\bf 198} (2019), 1781--1802.
\item Barbosa, Ezequiel and Santana, Farley and Upadhyay, Abhitosh: On stable {CMC} free-boundary surfaces in a strictly convex
domain of a bi-invariant {L}ie group, \emph{Internat. J. Math.} {\bf 31} (2020), 2050086, 9.
\item Basak, Biplab: An algorithmic approach to construct crystallizations of
3-manifolds from presentations of fundamental groups, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 126} (2016), 629--653.
\item Basak, Biplab and Casali, Maria Rita: Lower bounds for regular genus and gem-complexity of {PL}
4-manifolds, \emph{Forum Math.} {\bf 29} (2017), 761--773.
\item Basak, Biplab and Spreer, Jonathan: Simple crystallizations of 4-manifolds, \emph{Adv. Geom.} {\bf 16} (2016), 111--130.
\item Basak, Gopal K. and Ghosh, Mrinal K. and Mukherjee, Diganta: A mean-reverting stochastic model for the political business
cycle, \emph{Stoch. Anal. Appl.} {\bf 34} (2016), 96--116.
\item Basu, Arnab and Ghosh, Mrinal K.: Nonzero-sum risk-sensitive stochastic games on a countable
state space, \emph{Math. Oper. Res.} {\bf 43} (2018), 516--532.
\item Belton, Alexander and Guillot, Dominique and Khare, Apoorva
and Putinar, Mihai: Simultaneous kernels of matrix {H}adamard powers, \emph{Linear Algebra Appl.} {\bf 576} (2019), 142--157.
\item Benedetti, Carolina and Gonz\'{a}lez D'Le\'{o}n, Rafael S. and Hanusa,
Christopher R. H. and Harris, Pamela E. and Khare, Apoorva and
Morales, Alejandro H. and Yip, Martha: The volume of the caracol polytope, \emph{S\'{e}m. Lothar. Combin.} {\bf 80B} (2018), Art. 87, 12.
\item Benedetti, Carolina and Gonz\'{a}lez D'Le\'{o}n, Rafael S. and Hanusa,
Christopher R. H. and Harris, Pamela E. and Khare, Apoorva and
Morales, Alejandro H. and Yip, Martha: A combinatorial model for computing volumes of flow polytopes, \emph{Trans. Amer. Math. Soc.} {\bf 372} (2019), 3369--3404.
\item Bennett, Nathan and Iyer, Srikanth K. and Jammalamadaka, S.
Rao: Analysis of gamma and {W}eibull lifetime data under a general
censoring scheme and in the presence of covariates, \emph{Comm. Statist. Theory Methods} {\bf 46} (2017), 2277--2289.
\item Bera, Sayani and Pal, Ratna: Dynamics of semigroups of entire maps of {$\Bbb{C}^k$}, \emph{Complex Anal. Synerg.} {\bf 2} (2016), Paper No. 2, 22.
\item Bera, Sayani and Pal, Ratna and Verma, Kaushal: Examples of non-autonomous basins of attraction, \emph{Illinois J. Math.} {\bf 61} (2017), 531--567.
\item Bera, Sayani and Pal, Ratna and Verma, Kaushal: A rigidity theorem for {H}\'{e}non maps, \emph{Eur. J. Math.} {\bf 6} (2020), 508--532.
\item Bera, Sayani and Verma, Kaushal: Some aspects of shift-like automorphisms of {$\Bbb C^k$}, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 128} (2018), Art. 10, 48.
\item Bera, Sudip and Mukherjee, Sajal Kumar: Generalized power sum and {N}ewton-{G}irard identities, \emph{Graphs Combin.} {\bf 36} (2020), 1957--1964.
\item Bhamidi, Shankar and Dhara, Souvik and van der Hofstad, Remco
and Sen, Sanchayan: Universality for critical heavy-tailed network models: metric
structure of maximal components, \emph{Electron. J. Probab.} {\bf 25} (2020), Paper No. 47, 57.
\item Bhamidi, Shankar and Sen, Sanchayan: Geometry of the vacant set left by random walk on random
graphs, {W}right's constants, and critical random graphs with
prescribed degrees, \emph{Random Structures Algorithms} {\bf 56} (2020), 676--721.
\item Bharali, Gautam: Complex geodesics, their boundary regularity, and a
{H}ardy-{L}ittlewood-type lemma, \emph{Ann. Acad. Sci. Fenn. Math.} {\bf 41} (2016), 253--263.
\item Bharali, Gautam: On the growth of the {B}ergman metric near a point of infinite
type, \emph{J. Geom. Anal.} {\bf 30} (2020), 1238--1258.
\item Bharali, Gautam and Biswas, Indranil and Divakaran, Divakaran
and Janardhanan, Jaikrishnan: Proper holomorphic mappings onto symmetric products of a
{R}iemann surface, \emph{Doc. Math.} {\bf 23} (2018), 1291--1311.
\item Bharali, Gautam and Biswas, Indranil and Schumacher, Georg: A criterion for a degree-one holomorphic map to be a
biholomorphism, \emph{Complex Var. Elliptic Equ.} {\bf 62} (2017), 914--918.
\item Bharali, Gautam and Chandel, Vikramjeet Singh: Pick interpolation on the polydisc: small families of
sufficient kernels, \emph{Complex Anal. Oper. Theory} {\bf 13} (2019), 2069--2093.
\item Bharali, Gautam and Sridharan, Shrihari: Holomorphic correspondences related to finitely generated
rational semigroups, \emph{Internat. J. Math.} {\bf 28} (2017), 1750108, 25.
\item Bharali, Gautam and Sridharan, Shrihari: The dynamics of holomorphic correspondences of {$\Bbb P^1$}:
invariant measures and the normality set, \emph{Complex Var. Elliptic Equ.} {\bf 61} (2016), 1587--1613.
\item Bharali, Gautam and Zimmer, Andrew: Goldilocks domains, a weak notion of visibility, and
applications, \emph{Adv. Math.} {\bf 310} (2017), 377--425.
\item Bhatia, Manan: Moderate deviation and exit time estimates for stationary last
passage percolation, \emph{J. Stat. Phys.} {\bf 181} (2020), 1410--1432.
\item Bhattacharjee, M. and Eschmeier, J. and Keshari, Dinesh K. and
Sarkar, Jaydeb: Dilations, wandering subspaces, and inner functions, \emph{Linear Algebra Appl.} {\bf 523} (2017), 263--280.
\item Bhattacharjee, Monojit and Sarkar, Jaydeb: Operator positivity and analytic models of commuting tuples of
operators, \emph{Studia Math.} {\bf 232} (2016), 155--171.
\item Bhattacharyya, T. and Narayanan, E. K. and Sarkar, Jaydeb: Analytic model of doubly commuting contractions, \emph{Oper. Matrices} {\bf 11} (2017), 101--113.
\item Bhattacharyya, T. and Sau, H.: Explicit and unique construction of tetrablock unitary
dilation in a certain case, \emph{Complex Anal. Oper. Theory} {\bf 10} (2016), 749--768.
\item Bhattacharyya, Tirthankar and Biswas, Anindya and Chandel,
Vikramjeet Singh: On the {N}evanlinna problem: characterization of all
{S}chur-{A}gler class solutions affiliated with a given
kernel, \emph{Studia Math.} {\bf 255} (2020), 83--107.
\item Bhattacharyya, Tirthankar and Rao, T. S. S. R. K.: Survey of last ten years of work done in {I}ndia in some
selected areas of functional analysis and operator theory, \emph{Indian J. Pure Appl. Math.} {\bf 50} (2019), 599--617.
\item Bhattacharyya, Tirthankar and Sau, Haripada: {$\Gamma$}-unitaries, dilation and a natural example, \emph{Publ. Res. Inst. Math. Sci.} {\bf 53} (2017), 261--285.
\item Bhattacharyya, Tirthankar and Sau, Haripada: Holomorphic functions on the symmetrized bidisk---realization,
interpolation and extension, \emph{J. Funct. Anal.} {\bf 274} (2018), 504--524.
\item Bhimani, Divyang G. and Balhara, Rakesh and Thangavelu,
Sundaram: Hermite multipliers on modulation spaces, \emph{Analysis and partial differential equations: perspectives from
developing countries} {\bf 275} (2019), 42--64.
\item Bhosle, Usha N.: Hitchin pairs on reducible curves, \emph{Internat. J. Math.} {\bf 29} (2018), 1850015, 49.
\item Bhosle, Usha N.: On a conjecture of {B}utler, \emph{Analytic and algebraic geometry} {\bf } (2017), 29--48.
\item Bhosle, Usha N. and Biswas, Indranil: Moduli spaces of vector bundles on a singular rational ruled
surface, \emph{Geom. Dedicata} {\bf 180} (2016), 399--413.
\item Biswas, Indranil and D'Mello, Shane and Mukherjee, Ritwik and
Pingali, Vamsi P.: Rational cuspidal curves on del-{P}ezzo surfaces, \emph{J. Singul.} {\bf 17} (2018), 91--107.
\item Biswas, Indranil and Dumitrescu, Sorin and Gupta, Subhojoy: Branched projective structures on a {R}iemann surface and
logarithmic connections, \emph{Doc. Math.} {\bf 24} (2019), 2299--2337.
\item Biswas, Indranil and Pingali, Vamsi Pritham: A characterization of finite vector bundles on {G}auduchon
astheno-{K}\"{a}hler manifolds, \emph{\'{E}pijournal Geom. Alg\'{e}brique} {\bf 2} (2018), Art. 6, 13.
\item Biswas, Indranil and Pingali, Vamsi Pritham: Metric properties of parabolic ample bundles, \emph{Int. Math. Res. Not. IMRN} {\bf } (2020), 9336--9369.
\item Biswas, Indranil and Seshadri, Harish: Chern scalar curvature and symmetric products of compact
{R}iemann surfaces, \emph{Pure Appl. Math. Q.} {\bf 12} (2016), 463--471.
\item Biswas, Indranil and Seshadri, Harish: Quot schemes and {R}icci semipositivity, \emph{C. R. Math. Acad. Sci. Paris} {\bf 355} (2017), 577--581.
\item Biswas, Rahul and Dond, Asha K. and Gudi, Thirupathi: Edge patch-wise local projection stabilized nonconforming
{FEM} for the {O}seen problem, \emph{Comput. Methods Appl. Math.} {\bf 19} (2019), 189--214.
\item Biswas, Shibananda and Ghosh, Gargi and Misra, Gadadhar and
Shyam Roy, Subrata: On reducing submodules of {H}ilbert modules with
{${S}_n$}-invariant kernels, \emph{J. Funct. Anal.} {\bf 276} (2019), 751--784.
\item Bleher, F. M. and Chinburg, T. and Greenberg, R. and Kakde, M.
and Pappas, G. and Sharifi, R. and Taylor, M. J.: Higher {C}hern classes in {I}wasawa theory, \emph{Amer. J. Math.} {\bf 142} (2020), 627--682.
\item Boggarapu, Pradeep and Roncal, Luz and Thangavelu, Sundaram: Mixed norm estimates for the {C}es\`aro means associated with
{D}unkl-{H}ermite expansions, \emph{Trans. Amer. Math. Soc.} {\bf 369} (2017), 7021--7047.
\item Boggarapu, Pradeep and Roncal, Luz and Thangavelu, Sundaram: On extension problem, trace {H}ardy and {H}ardy's inequalities
for some fractional {L}aplacians, \emph{Commun. Pure Appl. Anal.} {\bf 18} (2019), 2575--2605.
\item Boggarapu, Pradeep and Thangavelu, S.: On the chaotic behavior of the {D}unkl heat semigroup on
weighted {$L^p$} spaces, \emph{Israel J. Math.} {\bf 217} (2017), 57--92.
\item Boggarapu, Pradeep and Thangavelu, S.: Mixed norm estimates for the {R}iesz transforms on {$SU(2)$}, \emph{Publ. Mat.} {\bf 60} (2016), 171--190.
\item Borah, Diganta and Haridas, Pranav and Verma, Kaushal: Comments on the {G}reen's function of a planar domain, \emph{Anal. Math. Phys.} {\bf 8} (2018), 383--414.
\item Brenner, Susanne C. and Gudi, Thirupathi and Porwal, Kamana
and Sung, Li-Yeng: A {M}orley finite element method for an elliptic distributed
optimal control problem with pointwise state and control
constraints, \emph{ESAIM Control Optim. Calc. Var.} {\bf 24} (2018), 1181--1206.
\item Burton, Benjamin A. and Datta, Basudeb and Singh, Nitin and
Spreer, Jonathan: A construction principle for tight and minimal triangulations
of manifolds, \emph{Exp. Math.} {\bf 27} (2018), 22--36.
\item Carstensen, Carsten and Dond, Asha K. and Rabus, Hella: Quasi-optimality of adaptive mixed {FEM}s for non-selfadjoint
indefinite second-order linear elliptic problems, \emph{Comput. Methods Appl. Math.} {\bf 19} (2019), 233--250.
\item Chakrabarti, A. and Mondal, Sumita Rani: On approximate numerical method of solution of {F}redholm
integral equations, \emph{Int. J. Appl. Math. Anal. Appl.} {\bf 11} (2016), 153--162.
\item Chakraborty, Atlanta and Chandru, Vijay and Rao, M. R.: A linear programming primer: from {F}ourier to {K}armarkar, \emph{Ann. Oper. Res.} {\bf 287} (2020), 593--616.
\item Chan, Chun Lam and Fernandes, Winston and Kashyap, Navin and
Krishnapur, Manjunath: Phase transitions for the uniform distribution in the pattern
maximum likelihood problem and its {B}ethe approximation, \emph{SIAM J. Discrete Math.} {\bf 31} (2017), 597--631.
\item Chandel, Vikramjeet Singh: The three-point {P}ick-{N}evanlinna interpolation problem on
the polydisc, \emph{Complex Var. Elliptic Equ.} {\bf 63} (2018), 1341--1352.
\item Chandel, Vikramjeet Singh: The 3-point spectral pick interpolation problem and an
application to holomorphic correspondences, \emph{J. Geom. Anal.} {\bf 30} (2020), 551--572.
\item Chas, Moira and Gadgil, Siddhartha: The extended {G}oldman bracket determines intersection numbers
for surfaces and orbifolds, \emph{Algebr. Geom. Topol.} {\bf 16} (2016), 2813--2838.
\item Chatterjee, Esha and Hassan, Sk. Sarif: On the asymptotic character of a generalized rational
difference equation, \emph{Discrete Contin. Dyn. Syst.} {\bf 38} (2018), 1707--1718.
\item Chattopadhyay, Arup and Sain, Debmalya and Senapati, Tanusri: A study of non-positive operators between real normed linear
spaces, \emph{Acta Sci. Math. (Szeged)} {\bf 86} (2020), 449--466.
\item Choudhary, A. and Martha, S. C. and Chakrabarti, A.: Modified method for the solution of dual trigonometric series
relations, \emph{Proc. Nat. Acad. Sci. India Sect. A} {\bf 90} (2020), 437--445.
\item Choudhury, Projesh Nath and Kannan, M. Rajesh and Sivakumar,
K. C.: A note on linear preservers of semipositive and minimally
semipositive matrices, \emph{Electron. J. Linear Algebra} {\bf 34} (2018), 687--694.
\item Choudhury, Projesh Nath and Tsatsomeros, Michael J.: Algorithmic detection and construction of {N}-matrices, \emph{Linear Algebra Appl.} {\bf 602} (2020), 46--56.
\item Chowdhury, Sudipto and Gudi, Thirupathi: A {$C^0$} interior penalty method for the {D}irichlet control
problem governed by biharmonic operator, \emph{J. Comput. Appl. Math.} {\bf 317} (2017), 290--306.
\item Chowdhury, Sudipto and Gudi, Thirupathi and Nandakumaran, A.
K.: Error bounds for a {D}irichlet boundary control problem based
on energy spaces, \emph{Math. Comp.} {\bf 86} (2017), 1103--1126.
\item Ciaurri, \'{O}scar and Roncal, Luz and Thangavelu, Sundaram: Hardy-type inequalities for fractional powers of the
{D}unkl-{H}ermite operator, \emph{Proc. Edinb. Math. Soc. (2)} {\bf 61} (2018), 513--544.
\item Cicuttin, Matteo and Ern, Alexandre and Gudi, Thirupathi: Hybrid high-order methods for the elliptic obstacle problem, \emph{J. Sci. Comput.} {\bf 83} (2020), Paper No. 8, 18.
\item Cui, Jintao and Gudi, Thirupathi: Finite element approximation of a free boundary plasma
problem, \emph{Adv. Comput. Math.} {\bf 43} (2017), 517--535.
\item Das, Soumya: Cuspidality and the growth of {F}ourier coefficients of
modular forms: a survey, \emph{Indian J. Pure Appl. Math.} {\bf 47} (2016), 9--22.
\item Das, Soumya and Ganguly, Satadal: A note on small gaps between nonzero {F}ourier coefficients of
cusp forms, \emph{Proc. Amer. Math. Soc.} {\bf 144} (2016), 2301--2305.
\item Das, Soumya and Jha, Abhash Kumar: Analytic properties of twisted real-analytic {H}ermitian
{K}lingen type {E}isenstein series and applications, \emph{Abh. Math. Semin. Univ. Hambg.} {\bf 89} (2019), 105--116.
\item Das, Soumya and Khan, Rizwanur: The third moment of symmetric square {$L$}-functions, \emph{Q. J. Math.} {\bf 69} (2018), 1063--1087.
\item Das, Soumya and Kohnen, Winfried: On sign changes of eigenvalues of {S}iegel cusp forms of genus
2 in prime powers, \emph{Acta Arith.} {\bf 183} (2018), 167--172.
\item Das, Soumya and Pal, Ritwik: The first negative eigenvalue of {Y}oshida lifts, \emph{Res. Number Theory} {\bf 5} (2019), Paper No. 20, 9.
\item Das, Soumya and Pal, Ritwik: Jacobi forms and differential operators: odd weights, \emph{J. Number Theory} {\bf 179} (2017), 113--125.
\item Datar, Ved V. and Pingali, Vamsi Pritham: On coupled constant scalar curvature {K}\"{a}hler metrics, \emph{J. Symplectic Geom.} {\bf 18} (2020), 961--994.
\item Datta, Basudeb and Maity, Dipendu: Correction to: {S}emi-equivelar and vertex-transitive maps on
the torus, \emph{Beitr. Algebra Geom.} {\bf 61} (2020), 187--188.
\item Datta, Basudeb and Maity, Dipendu: Semi-equivelar and vertex-transitive maps on the torus, \emph{Beitr. Algebra Geom.} {\bf 58} (2017), 617--634.
\item Datta, Basudeb and Maity, Dipendu: Semi-equivelar maps on the torus and the {K}lein bottle are
{A}rchimedean, \emph{Discrete Math.} {\bf 341} (2018), 3296--3309.
\item Datta, Basudeb and Murai, Satoshi: On stacked triangulated manifolds, \emph{Electron. J. Combin.} {\bf 24} (2017), Paper 4.12, 14.
\item Datta, Basudeb and Sarkar, Soumen: On equilibrium triangulations of quasitoric manifolds, \emph{Southeast Asian Bull. Math.} {\bf 44} (2020), 57-56.
\item Dhanya, R. and Prashanth, S. and Tiwari, Sweta and Sreenadh,
K.: Elliptic problems in {${\Bbb R}^N$} with critical and singular
discontinuous nonlinearities, \emph{Complex Var. Elliptic Equ.} {\bf 61} (2016), 1668--1688.
\item Dhanya, R. and Sarkar, Abhishek: Isolated singularities of polyharmonic operator in even
dimension, \emph{Complex Var. Elliptic Equ.} {\bf 61} (2016), 55--66.
\item Dhara, Souvik and van der Hofstad, Remco and van Leeuwaarden,
Johan S. H. and Sen, Sanchayan: Heavy-tailed configuration models at criticality, \emph{Ann. Inst. Henri Poincar\'{e} Probab. Stat.} {\bf 56} (2020), 1515--1558.
\item Dhillon, Gurbir and Khare, Apoorva: Faces of highest weight modules and the universal {W}eyl
polyhedron, \emph{Adv. Math.} {\bf 319} (2017), 111--152.
\item Dond, Asha K. and Gudi, Thirupathi: Patch-wise local projection stabilized finite element methods
for convection-diffusion-reaction problems, \emph{Numer. Methods Partial Differential Equations} {\bf 35} (2019), 638--663.
\item Dond, Asha K. and Gudi, Thirupathi and Nataraj, Neela: A nonconforming finite element approximation for optimal
control of an obstacle problem, \emph{Comput. Methods Appl. Math.} {\bf 16} (2016), 653--666.
\item Dond, Asha K. and Gudi, Thirupathi and Sau, Ramesh CH.: An error analysis of discontinuous finite element methods for
the optimal control problems governed by {S}tokes equation, \emph{Numer. Funct. Anal. Optim.} {\bf 40} (2019), 421--460.
\item Fern\'{o}s, Talia and Singla, Pooja: On images of real representations of special linear groups
over complete discrete valuation rings, \emph{Glasg. Math. J.} {\bf 58} (2016), 263--272.
\item Fritz, Tobias and Gadgil, Siddhartha and Khare, Apoorva and Nielsen, Pace P. and Silberman, Lior and Tao, Terence: Homogeneous length functions on groups, \emph{Algebra Number Theory} {\bf 12} (2018), 1773--1786.
\item Gaddam, Sharat and Gudi, Thirupathi: Inhomogeneous {D}irichlet boundary condition in the {\it a
posteriori} error control of the obstacle problem, \emph{Comput. Math. Appl.} {\bf 75} (2018), 2311--2327.
\item Gaddam, Sharat and Gudi, Thirupathi: Bubbles enriched quadratic finite element method for the
3{D}-elliptic obstacle problem, \emph{Comput. Methods Appl. Math.} {\bf 18} (2018), 223--236.
\item Gadgil, Siddhartha: Homogeneous {L}ength {F}unctions on {G}roups: {I}ntertwined
{C}omputer and {H}uman {P}roofs, \emph{J. Automat. Reason.} {\bf 64} (2020), 677--688.
\item Gadgil, Siddhartha and Kulkarni, Dheeraj: Relative symplectic caps, 4-genus and fibered knots, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 126} (2016), 261--275.
\item Garani, Shayan Srinivasa and Seshadri, Harish: An algorithmic approach to {S}outh {I}ndian classical music, \emph{J. Math. Music} {\bf 13} (2019), 107--134.
\item Gaussier, Herv\'{e} and Seshadri, Harish: Metric geometry of the {K}obayashi metric, \emph{Eur. J. Math.} {\bf 3} (2017), 1030--1044.
\item Gaussier, Herv\'{e} and Seshadri, Harish: Holomorphic motions and complex geometry, \emph{Proc. Amer. Math. Soc.} {\bf 147} (2019), 301--313.
\item Gaussier, Herv\'{e} and Seshadri, Harish: On the {G}romov hyperbolicity of convex domains in {$\Bbb
C^n$}, \emph{Comput. Methods Funct. Theory} {\bf 18} (2018), 617--641.
\item Gauttam, Sunil Kumar and Kumar, K. Suresh and Pal, Chandan: Risk-sensitive control of reflected diffusion processes on
orthrant, \emph{Pure Appl. Funct. Anal.} {\bf 2} (2017), 477--510.
\item Gayen, R. and Gupta, Sourav and Chakrabarti, A.: Approximate solution of the problem of scattering of surface
water waves by a partially immersed rigid plane vertical
barrier, \emph{Appl. Math. Lett.} {\bf 58} (2016), 19--25.
\item Ghara, Soumitra: The orbit of a bounded operator under the {M}\"{o}bius group
modulo similarity equivalence, \emph{Israel J. Math.} {\bf 238} (2020), 167--207.
\item Ghara, Soumitra and Kumar, Surjit: On the sum of two subnormal kernels, \emph{J. Math. Anal. Appl.} {\bf 469} (2019), 1015--1027.
\item Ghosh, Mrinal K. and Kumar, K. Suresh and Pal, Chandan: Zero-sum risk-sensitive stochastic games for continuous time
{M}arkov chains, \emph{Stoch. Anal. Appl.} {\bf 34} (2016), 835--851.
\item Ghosh, Mrinal K. and Pradhan, Somnath: Risk-sensitive stochastic differential games with reflecting
diffusions, \emph{Stoch. Anal. Appl.} {\bf 36} (2018), 1--27.
\item Ghosh, Mrinal Kanti and Pradhan, Somnath: Zero-sum risk-sensitive stochastic differential games with
reflecting diffusions in the orthant, \emph{ESAIM Control Optim. Calc. Var.} {\bf 26} (2020), Paper No. 114, 33.
\item Ghosh, Puja and Paul, Kallol and Sain, Debmalya: Symmetric properties of orthogonality of linear operators on
{$(\Bbb R^n,\|\cdot\|_1)$}, \emph{Novi Sad J. Math.} {\bf 47} (2017), 41--46.
\item Ghosh, Puja and Sain, Debmalya and Paul, Kallol: On symmetry of {B}irkhoff-{J}ames orthogonality of linear
operators, \emph{Adv. Oper. Theory} {\bf 2} (2017), 428--434.
\item Ghosh, Subhajit: Total variation cutoff for the transpose top-2 with random
shuffle, \emph{J. Theoret. Probab.} {\bf 33} (2020), 1832--1854.
\item Ghosh, Subhajit: Cutoff for the warp-transpose top with random shuffle, \emph{S\'{e}m. Lothar. Combin.} {\bf 84B} (2020), Art. 69, 12.
\item Ghosh, Subhroshekhar and Krishnapur, Manjunath and Peres,
Yuval: Continuum percolation for {G}aussian zeroes and {G}inibre
eigenvalues, \emph{Ann. Probab.} {\bf 44} (2016), 3357--3384.
\item Gudi, Thirupathi and Majumder, Papri: Crouzeix-{R}aviart finite element approximation for the
parabolic obstacle problem, \emph{Comput. Methods Appl. Math.} {\bf 20} (2020), 273--292.
\item Gudi, Thirupathi and Majumder, Papri: Convergence analysis of finite element method for a parabolic
obstacle problem, \emph{J. Comput. Appl. Math.} {\bf 357} (2019), 85--102.
\item Gudi, Thirupathi and Majumder, Papri: Conforming and discontinuous {G}alerkin {FEM} in space for
solving parabolic obstacle problem, \emph{Comput. Math. Appl.} {\bf 78} (2019), 3896--3915.
\item Gudi, Thirupathi and Porwal, Kamana: A {$C^0$} interior penalty method for a fourth-order
variational inequality of the second kind, \emph{Numer. Methods Partial Differential Equations} {\bf 32} (2016), 36--59.
\item Gudi, Thirupathi and Porwal, Kamana: A posteriori error estimates of discontinuous {G}alerkin
methods for the {S}ignorini problem, \emph{J. Comput. Appl. Math.} {\bf 292} (2016), 257--278.
\item Gudi, Thirupathi and Sau, Ramesh Ch.: Finite element analysis of the constrained {D}irichlet
boundary control problem governed by the diffusion problem, \emph{ESAIM Control Optim. Calc. Var.} {\bf 26} (2020), Paper No. 78, 19.
\item Gupta, Rajeev: An improvement on the bound for {$C_2(n)$}, \emph{Acta Sci. Math. (Szeged)} {\bf 83} (2017), 263--269.
\item Gupta, Rajeev and Kumar, Surjit and Trivedi, Shailesh: Unitary equivalence of operator-valued multishifts, \emph{J. Math. Anal. Appl.} {\bf 487} (2020), 124032, 23.
\item Gupta, Rajeev and Kumar, Surjit and Trivedi, Shailesh: Von {N}eumann's inequality for commuting operator-valued
multishifts, \emph{Proc. Amer. Math. Soc.} {\bf 147} (2019), 2599--2608.
\item Gupta, Rajeev and Misra, Gadadhar: The {C}arath\'{e}odory-{F}ej\'{e}r interpolation on the polydisc, \emph{Studia Math.} {\bf 254} (2020), 265--293.
\item Gupta, Rajeev and Ramiz Reza, Md.: Operator space structures on {$\ell^1(n)$}, \emph{Houston J. Math.} {\bf 44} (2018), 1205--1212.
\item Gupta, S. C.: The classical {S}tefan problem, \emph{} {\bf } (2018), xxiii+726.
\item Gupta, Subhojoy: Limits of harmonic maps and crowned hyperbolic surfaces, \emph{Trans. Amer. Math. Soc.} {\bf 372} (2019), 7573--7596.
\item Gupta, Subhojoy and Mj, Mahan: Monodromy representations of meromorphic projective
structures, \emph{Proc. Amer. Math. Soc.} {\bf 148} (2020), 2069--2078.
\item Gupta, Subhojoy and Seshadri, Harish: On domains biholomorphic to {T}eichm\"{u}ller spaces, \emph{Int. Math. Res. Not. IMRN} {\bf } (2020), 2542--2560.
\item Gupta, Subhojoy and Wolf, Michael: Meromorphic quadratic differentials and measured foliations on
a {R}iemann surface, \emph{Math. Ann.} {\bf 373} (2019), 73--118.
\item Gupta, Subhojoy and Wolf, Michael: Quadratic differentials, half-plane structures, and harmonic
maps to trees, \emph{Comment. Math. Helv.} {\bf 91} (2016), 317--356.
\item Gupta, Subhojoy and Wolf, Michael: Meromorphic quadratic differentials with complex residues and
spiralling foliations, \emph{In the tradition of {A}hlfors-{B}ers. {VII}} {\bf 696} (2017), 153--181.
\item Hatui, Sumana: Schur multipliers of special {$p$}-groups of rank 2, \emph{J. Group Theory} {\bf 23} (2020), 85--95.
\item Hazra, Somnath: Homogeneous 2-shifts, \emph{Complex Anal. Oper. Theory} {\bf 13} (2019), 1729--1763.
\item Iyer, Srikanth K.: The random connection model: connectivity, edge lengths, and
degree distributions, \emph{Random Structures Algorithms} {\bf 52} (2018), 283--300.
\item Iyer, Srikanth K. and Vaze, Rahul and Narasimha, Dheeraj: Autoregressive cascades on random networks, \emph{Phys. A} {\bf 447} (2016), 345--354.
\item Iyer, Srikanth K. and Yogeshwaran, D.: Thresholds for vanishing of `isolated' faces in random \v{C}ech
and {V}ietoris-{R}ips complexes, \emph{Ann. Inst. Henri Poincar\'{e} Probab. Stat.} {\bf 56} (2020), 1869--1897.
\item Iyer, Srikanth and Vaze, Rahul: Achieving nonzero information velocity in wireless networks, \emph{Ann. Appl. Probab.} {\bf 27} (2017), 48--64.
\item Jha, Abhash Kumar and Juyal, Abhishek and Pandey, Manish
Kumar: On simultaneous non-vanishing of twisted {$L$}-functions
associated to newforms on {$\Gamma_0(N)$}, \emph{J. Ramanujan Math. Soc.} {\bf 34} (2019), 245--252.
\item Jha, Abhash Kumar and Kumar, Arvind: Construction of cusp forms using {R}ankin-{C}ohen brackets, \emph{Analytic number theory, modular forms and {$q$}-hypergeometric
series} {\bf 221} (2017), 329--341.
\item Jha, Abhash Kumar and Sahu, Brundaban: Rankin-{C}ohen brackets on {J}acobi forms of several variables
and special values of certain {D}irichlet series, \emph{Int. J. Number Theory} {\bf 15} (2019), 925--933.
\item Jha, Abhash Kumar and Sahu, Brundaban: Differential operators on {J}acobi forms and special values of
certain {D}irichlet series, \emph{Automorphic forms and related topics} {\bf 732} (2019), 91--99.
\item Ji, Kui and Jiang, Chunlan and Keshari, Dinesh Kumar and
Misra, Gadadhar: Rigidity of the flag structure for a class of
{C}owen-{D}ouglas operators, \emph{J. Funct. Anal.} {\bf 272} (2017), 2899--2932.
\item Jiang, Chunlan and Ji, Kui and Misra, Gadadhar: Classification of quasi-homogeneous holomorphic curves and
operators in the {C}owen-{D}ouglas class, \emph{J. Funct. Anal.} {\bf 273} (2017), 2870--2915.
\item Kabiraj, Arpan: Center of the {G}oldman {L}ie algebra, \emph{Algebr. Geom. Topol.} {\bf 16} (2016), 2839--2849.
\item Kanukurthi, Bhavana and Bhavana Obbattu, Sai Lakshmi and
Sekar, Sruthi: Four-state non-malleable codes with explicit constant rate, \emph{J. Cryptology} {\bf 33} (2020), 1044--1079.
\item Kanukurthi, Bhavana and Bhavana Obbattu, Sai Lakshmi and
Sekar, Sruthi: Four-state non-malleable codes with explicit constant rate, \emph{Theory of cryptography. {P}art {II}} {\bf 10678} (2017), 344--375.
\item Kanukurthi, Bhavana and Bhavana Obbattu, Sai Lakshmi and
Sekar, Sruthi: Non-malleable randomness encoders and their applications, \emph{Advances in cryptology---{EUROCRYPT} 2018. {P}art {III}} {\bf 10822} (2018), 589--617.
\item Kashyap, G. A. R. S. R. K. and Bapat, D. and Das, D. and Gowaikar, R. D. and  Amritkar, R. E. and Rangarajan, G. and V. Ravindranath and G. Ambika: Synapse loss and progress of Alzheimer&rsquo;s disease - A network model, \emph{Scientific Reports} {\bf 9} (2019).
\item Kataria, Kuldeep Kumar and Vellaisamy, Palaniappan: On the convolution of {M}ittag-{L}effler distributions and its
applications to fractional point processes, \emph{Stoch. Anal. Appl.} {\bf 37} (2019), 115--122.
\item Kaur, Amandeep and Martha, S. C. and Chakrabarti, A.: An algebraic method of solution of a water wave scattering
problem involving an asymmetrical trench, \emph{Comput. Appl. Math.} {\bf 39} (2020), Paper No. 229, 19.
\item Khan, Mohammad Yasir Feroz and Biswas, Rahul and Sau, Ramesh
Chandra: Instability of {L}agrange interpolation on {$H^1(\Omega)\cap
C(\overline\Omega)$}, \emph{Math. Student} {\bf 88} (2019), 153--158.
\item Khare, Apoorva: Generalized nil-{C}oxeter algebras over discrete complex
reflection groups, \emph{Trans. Amer. Math. Soc.} {\bf 370} (2018), 2971--2999.
\item Khare, Apoorva and Rajaratnam, Bala: Probability inequalities and tail estimates for metric
semigroups, \emph{Adv. Oper. Theory} {\bf 5} (2020), 779--795.
\item Khare, Apoorva and Rajaratnam, Bala: The {H}offmann-{J}\o rgensen inequality in metric semigroups, \emph{Ann. Probab.} {\bf 45} (2017), 4101--4111.
\item Khare, Apoorva and Tikaradze, Akaki: A {C}arlitz--von {S}taudt type theorem for finite rings, \emph{Linear Algebra Appl.} {\bf 568} (2019), 106--126.
\item Khurana, Divya and Roy, Saikat and Sain, Debmalya: Symmetric points in spaces of linear operators between
{B}anach spaces, \emph{Acta Sci. Math. (Szeged)} {\bf 86} (2020), 617--634.
\item Kor\'{a}nyi, Adam and Misra, Gadadhar: Homogeneous {H}ermitian holomorphic vector bundles and the
{C}owen-{D}ouglas class over bounded symmetric domains, \emph{Adv. Math.} {\bf 351} (2019), 1105--1138.
\item Kor\'{a}nyi, Adam and Misra, Gadadhar: Homogeneous {H}ermitian holomorphic vector bundles and the
{C}owen-{D}ouglas class over bounded symmetric domains, \emph{C. R. Math. Acad. Sci. Paris} {\bf 354} (2016), 291--295.
\item Krakovski, Roi and Onn, Uri and Singla, Pooja: Regular characters of groups of type {$\ssf A_n$} over
discrete valuation rings, \emph{J. Algebra} {\bf 496} (2018), 116--137.
\item Kreuzer, Martin and Patil, Dilip P.: Computational aspects of {B}urnside rings, part {I}: the ring
structure, \emph{Beitr. Algebra Geom.} {\bf 58} (2017), 427--452.
\item Krishna, M. and Krishnapur, Manjunath: Persistence probabilities in centered, stationary, {G}aussian
processes in discrete time, \emph{Indian J. Pure Appl. Math.} {\bf 47} (2016), 183--194.
\item Krishnapur, Manjunath and Rider, Brian and Vir\'{a}g, B\'{a}lint: Universality of the stochastic {A}iry operator, \emph{Comm. Pure Appl. Math.} {\bf 69} (2016), 145--199.
\item Kundu, S. and Aggarwal, Manisha and Hazra, Somnath: Finitely chainable and totally bounded metric spaces:
equivalent characterizations, \emph{Topology Appl.} {\bf 216} (2017), 59--73.
\item Lin, X. and Kulkarni, P. and Bocci, F. and Schafer, N. and Roy, S. and Tsai, M-Y. and He, Y. and Chen, Y and Rajagopalan, K. and Mooney, S. and Zeng, Y. and Weninger, K. and Grishaev, A. and Onuchic, J. and Levine, H. and Wolynes, P. and Salgia, R. and Rangarajan, G. and Uversky, V. and Orban, J. and Jolly, M. K.: Structural and Dynamical Order of a Disordered Protein: Molecular Insights into Conformational Switching of PAGE4 at the Systems Level, \emph{Biomolecules} {\bf 9} (2019), 77.
\item Mahadevan, Rajesh and Nandakumaran, A. K. and Prakash, Ravi: Homogenization of an elliptic equation in a domain with
oscillating boundary with non-homogeneous non-linear boundary
conditions, \emph{Appl. Math. Optim.} {\bf 82} (2020), 245--278.
\item Mahajan, Prachi and Verma, Kaushal: A comparison of two biholomorphic invariants, \emph{Internat. J. Math.} {\bf 30} (2019), 1950012, 16.
\item Maiti, Arun: Quasi-vertex-transitive maps on the plane, \emph{Discrete Math.} {\bf 343} (2020), 111911, 7.
\item Maitra, Anwoy: On the continuous extension of {K}obayashi isometries, \emph{Proc. Amer. Math. Soc.} {\bf 148} (2020), 3437--3451.
\item Maitra, Anwoy: A form of {S}chwarz's lemma and a bound for the {K}obayashi
metric on convex domains, \emph{J. Math. Anal. Appl.} {\bf 484} (2020), 123694, 16.
\item Maity, Dipendu: Pentagonal maps on the torus and the plane, \emph{Beitr. Algebra Geom.} {\bf 60} (2019), 17--37.
\item Maity, Dipendu and Upadhyay, Ashish Kumar: On centrally symmetric manifolds, \emph{J. Ramanujan Math. Soc.} {\bf 34} (2019), 21--27.
\item Mal, Arpita and Paul, Kallol and Rao, T. S. S. R. K. and Sain,
Debmalya: Approximate {B}irkhoff-{J}ames orthogonality and smoothness in
the space of bounded linear operators, \emph{Monatsh. Math.} {\bf 190} (2019), 549--558.
\item Mal, Arpita and Sain, Debmalya and Paul, Kallol: On some geometric properties of operator spaces, \emph{Banach J. Math. Anal.} {\bf 13} (2019), 174--191.
\item Manchanda, Kaustubh and Bose, Amitabha and Ramaswamy,
Ramakrishna: Collective dynamics in heterogeneous networks of neuronal
cellular automata, \emph{Phys. A} {\bf 487} (2017), 111--124.
\item Mandal, S. and Sahoo, T. and Chakrabarti, A.: Characteristics of eigen-system for flexural gravity wave
problems, \emph{Geophys. Astrophys. Fluid Dyn.} {\bf 111} (2017), 249--281.
\item Manfio, F. and Turgay, N. C. and Upadhyay, A.: Biconservative submanifolds in {$\Bbb{S}^n\times\Bbb{R}$} and
{$\Bbb{H}^n\times\Bbb{R}$}, \emph{J. Geom. Anal.} {\bf 29} (2019), 283--298.
\item Manikandan, K. and Vishnu Priya, N. and Senthilvelan, M. and
Sankaranarayanan, R.: Deformation of dark solitons in a {$\Cal{PT}$}-invariant
variable coefficients nonlocal nonlinear {S}chr\"{o}dinger
equation, \emph{Chaos} {\bf 28} (2018), 083103, 12.
\item Misra, Gadadhar: Operators in the {C}owen-{D}ouglas class and related topics, \emph{Handbook of analytic operator theory} {\bf } (2019), 87--137.
\item Misra, Gadadhar and Pal, Avijit: Curvature inequalities for operators in the {C}owen-{D}ouglas
class and localization of the {W}allach set, \emph{J. Anal. Math.} {\bf 136} (2018), 31--54.
\item Misra, Gadadhar and Pal, Avijit and Varughese, Cherian: Contractivity and complete contractivity for finite
dimensional {B}anach spaces, \emph{J. Operator Theory} {\bf 82} (2019), 23--47.
\item Misra, Gadadhar and Reza, Md. Ramiz: Curvature inequalities and extremal operators, \emph{Illinois J. Math.} {\bf 63} (2019), 193--217.
\item Misra, Gadadhar and Upmeier, Harald: Homogeneous vector bundles and intertwining operators for
symmetric domains, \emph{Adv. Math.} {\bf 303} (2016), 1077--1121.
\item Mukherjee, Sajal Kumar and Bera, Sudip: Combinatorial proofs of the {N}ewton-{G}irard and
{C}hapman-{C}ostas-{S}antos identities, \emph{Discrete Math.} {\bf 342} (2019), 1577--1580.
\item Nandakumaran, A. K. and Datti, P. S. and George, Raju K.: Ordinary differential equations, \emph{} {\bf } (2017), xviii+328.
\item Nandakumaran, A. K. and Sili, Ali: Homogenization of a hyperbolic equation with highly
contrasting diffusivity coefficients, \emph{Differential Integral Equations} {\bf 29} (2016), 37--54.
\item Nandakumaran, Akambadath and Prakash, Ravi and Sardar, Bidhan
Chandra: Asymptotic analysis of {N}eumann periodic optimal boundary
control problem, \emph{Math. Methods Appl. Sci.} {\bf 39} (2016), 4354--4374.
\item Narayanan, E. K. and Singla, Pooja: On monomial representations of finitely generated nilpotent
groups, \emph{Comm. Algebra} {\bf 46} (2018), 2319--2331.
\item Nikolov, Nikolai and Verma, Kaushal: On the squeezing function and {F}ridman invariants, \emph{J. Geom. Anal.} {\bf 30} (2020), 1218--1225.
\item Pal, Chandan and Pradhan, Somnath: Risk sensitive control of pure jump processes on a general
state space, \emph{Stochastics} {\bf 91} (2019), 155--174.
\item Pal, Ratna and Verma, Kaushal: Ergodic properties of families of {H}\'{e}non maps, \emph{Ann. Polon. Math.} {\bf 121} (2018), 45--71.
\item Pal, Ritwik: On the signs of {F}ourier coefficients of {H}ilbert cusp
forms, \emph{Ramanujan J.} {\bf 53} (2020), 467--481.
\item Panda, Srikumar and Martha, S. C. and Chakrabarti, A.: An alternative approach to study nonlinear inviscid flow over
arbitrary bottom topography, \emph{Appl. Math. Comput.} {\bf 273} (2016), 165--177.
\item Patil, D. P.: Affine one dimensional domains which are not {E}uclidean
domains, \emph{Int. Rev. Pure Appl. Math.} {\bf 12} (2016), 45--50.
\item Paul, Kallol and Sain, Debmalya: Birkhoff-{J}ames orthogonality and its application in the
study of geometry of {B}anach space, \emph{Advanced topics in mathematical analysis} {\bf } (2019), 245--284.
\item Paul, Kallol and Sain, Debmalya and Ghosh, Puja: Symmetry of {B}irkhoff-{J}ames orthogonality of bounded linear
operators, \emph{Ulam type stability} {\bf } (2019), 331--344.
\item Paul, Kallol and Sain, Debmalya and Mal, Arpita: Approximate {B}irkhoff-{J}ames orthogonality in the space of
bounded linear operators, \emph{Linear Algebra Appl.} {\bf 537} (2018), 348--357.
\item Paul, Kallol and Sain, Debmalya and Mal, Arpita and Mandal,
Kalidas: Orthogonality of bounded linear operators on complex {B}anach
spaces, \emph{Adv. Oper. Theory} {\bf 3} (2018), 699--709.
\item Pingali, Vamsi P.: {$C^{2,\alpha}$} estimates and existence results for a
nonconcave {PDE}, \emph{Electron. J. Differential Equations} {\bf } (2016), Paper No. 168, 10.
\item Pingali, Vamsi P.: A note on the deformed {H}ermitian {Y}ang-{M}ills {PDE}, \emph{Complex Var. Elliptic Equ.} {\bf 64} (2019), 503--518.
\item Pingali, Vamsi Pritham: A vector bundle version of the {M}onge-{A}mp\`ere equation, \emph{Adv. Math.} {\bf 360} (2020), 106921, 40.
\item Pingali, Vamsi Pritham: A note on higher extremal metrics, \emph{Trans. Amer. Math. Soc.} {\bf 370} (2018), 6995--7010.
\item Pingali, Vamsi Pritham: Quillen metrics and perturbed equations, \emph{Lett. Math. Phys.} {\bf 110} (2020), 1861--1875.
\item Pingali, Vamsi Pritham: Representability of {C}hern-{W}eil forms, \emph{Math. Z.} {\bf 288} (2018), 629--641.
\item Poria, Anirudha: Positive weight function and classification of {$g$}-frames, \emph{Numer. Funct. Anal. Optim.} {\bf 41} (2020), 950--968.
\item Pradhan, Puspendu and Sahu, Bikramaditya: A characterization of the family of secant lines to a
hyperbolic quadric in {${\rm PG} (3,q)$}, {$q$} odd, \emph{Discrete Math.} {\bf 343} (2020), 112044, 8.
\item Prasad, Phoolan: Ray equations of a weak shock in a hyperbolic system of
conservation laws in multi-dimensions, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 126} (2016), 199--206.
\item Prasad, Phoolan: Propagation of multidimensional nonlinear waves and
kinematical conservation laws, \emph{} {\bf } (2017), xi+159.
\item Prasad, Phoolan: Kinematical conservation laws in a space of arbitrary
dimensions, \emph{Indian J. Pure Appl. Math.} {\bf 47} (2016), 641--653.
\item Pritham Pingali, Vamsi: Existence of coupled {K}\"{a}hler-{E}instein metrics using the
continuity method, \emph{Internat. J. Math.} {\bf 29} (2018), 1850041, 8.
\item Quddus, Safdar: Invariants of the {$\Bbb Z_2$} orbifolds of the {P}odle\'{s} two
spheres, \emph{J. Noncommut. Geom.} {\bf 13} (2019), 257--267.
\item Ram, Samrith: Counting zero kernel pairs over a finite field, \emph{Linear Algebra Appl.} {\bf 495} (2016), 1--10.
\item Ray, Anubhab and Roy, Saikat and Bagchi, Satya and Sain,
Debmalya: Extreme contractions on finite-dimensional polygonal {B}anach
spaces. {II}, \emph{J. Operator Theory} {\bf 84} (2020), 127--137.
\item Ray, Samarpita: Closure operations, continuous valuations on monoids and
spectral spaces, \emph{J. Algebra Appl.} {\bf 19} (2020), 2050006, 27.
\item Reddy, Nanda Kishore: Equality of {L}yapunov and stability exponents for products of
isotropic random matrices, \emph{Int. Math. Res. Not. IMRN} {\bf } (2019), 606--624.
\item Reza, Md. Ramiz: Curvature inequalities for operators in the {C}owen-{D}ouglas
class of a planar domain, \emph{Indiana Univ. Math. J.} {\bf 67} (2018), 1255--1279.
\item Richard, Thomas and Seshadri, Harish: Positive isotropic curvature and self-duality in dimension 4, \emph{Manuscripta Math.} {\bf 149} (2016), 443--457.
\item Roncal, Luz and Thangavelu, Sundaram: An extension problem and trace {H}ardy inequality for the
sub-{L}aplacian on {$H$}-type groups, \emph{Int. Math. Res. Not. IMRN} {\bf } (2020), 4238--4294.
\item Roncal, Luz and Thangavelu, Sundaram: Hardy's inequality for fractional powers of the sublaplacian
on the {H}eisenberg group, \emph{Adv. Math.} {\bf 302} (2016), 106--158.
\item Roth, Julien and Upadhyay, Abhitosh: On compact anisotropic {W}eingarten hypersurfaces in
{E}uclidean space, \emph{Arch. Math. (Basel)} {\bf 113} (2019), 213--224.
\item Roth, Julien and Upadhyay, Abhitosh: On almost stable {CMC} hypersurfaces in manifolds of bounded
sectional curvature, \emph{Bull. Aust. Math. Soc.} {\bf 101} (2020), 333--338.
\item Roy, Dipankar and Pandit, Rahul: One-dimensional {K}ardar-{P}arisi-{Z}hang and
{K}uramoto-{S}ivashinsky universality class: limit
distributions, \emph{Phys. Rev. E} {\bf 101} (2020), 030103(R), 6.
\item Roy, Krishanu and Venkatesh, R.: Maximal closed subroot systems of real affine root systems, \emph{Transform. Groups} {\bf 24} (2019), 1261--1308.
\item Sain, Debmalya: On extreme contractions and the norm attainment set of a
bounded linear operator, \emph{Ann. Funct. Anal.} {\bf 10} (2019), 135--143.
\item Sain, Debmalya: On a generalization of a complementary triangle inequality in
{H}ilbert spaces and {B}anach spaces, \emph{Indian J. Pure Appl. Math.} {\bf 51} (2020), 1815--1827.
\item Sain, Debmalya: On the norm attainment set of a bounded linear operator, \emph{J. Math. Anal. Appl.} {\bf 457} (2018), 67--76.
\item Sain, Debmalya: On the norm attainment set of a bounded linear operator and
semi-inner-products in normed spaces, \emph{Indian J. Pure Appl. Math.} {\bf 51} (2020), 179--186.
\item Sain, Debmalya: Smooth points in operator spaces and some
{B}ishop-{P}helps-{B}ollob\'{a}s type theorems in {B}anach spaces, \emph{Oper. Matrices} {\bf 13} (2019), 433--445.
\item Sain, Debmalya and Mal, Arpita and Paul, Kallol: Some remarks on {B}irkhoff-{J}ames orthogonality of linear
operators, \emph{Expo. Math.} {\bf 38} (2020), 138--147.
\item Sain, Debmalya and Paul, Kallol and Bhunia, Pintu and Bag,
Santanu: On the numerical index of polyhedral {B}anach spaces, \emph{Linear Algebra Appl.} {\bf 577} (2019), 121--133.
\item Sain, Debmalya and Paul, Kallol and Mal, Arpita: A complete characterization of {B}irkhoff-{J}ames
orthogonality in infinite dimensional normed space, \emph{J. Operator Theory} {\bf 80} (2018), 399--413.
\item Sain, Debmalya and Paul, Kallol and Mal, Arpita: On approximate {B}irkhoff-{J}ames orthogonality and normal
cones in a normed space, \emph{J. Convex Anal.} {\bf 26} (2019), 341--351.
\item Sain, Debmalya and Paul, Kallol and Mal, Arpita and Ray,
Anubhab: A complete characterization of smoothness in the space of
bounded linear operators, \emph{Linear Multilinear Algebra} {\bf 68} (2020), 2484--2494.
\item Sain, Debmalya and Paul, Kallol and Mandal, Kalidas: On two extremum problems related to the norm of a bounded
linear operator, \emph{Oper. Matrices} {\bf 13} (2019), 421--432.
\item Sain, Debmalya and Ray, Anubhab and Paul, Kallol: Extreme contractions on finite-dimensional polygonal {B}anach
spaces, \emph{J. Convex Anal.} {\bf 26} (2019), 877--885.
\item Sain, Debmalya and Tanaka, Ryotaro: Modular {B}irkhoff-{J}ames orthogonality in {$B(\Bbb{X,Y})$}
and {$K(\Bbb{X,Y})$}, \emph{Banach J. Math. Anal.} {\bf 14} (2020), 1347--1362.
\item Sanki, Bidyut: Filling of closed surfaces, \emph{J. Topol. Anal.} {\bf 10} (2018), 897--913.
\item Sanki, Bidyut and Gadgil, Siddhartha: Graphs of systoles on hyperbolic surfaces, \emph{J. Topol. Anal.} {\bf 11} (2019), 1--20.
\item Sarkar, Amar Deep and Verma, Kaushal: A submultiplicative property of the {C}arath\'{e}odory metric on
planar domains, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 130} (2020), Paper No. 35, 6.
\item Sarkar, Amar Deep and Verma, Kaushal: Boundary behaviour of some conformal invariants on planar
domains, \emph{Comput. Methods Funct. Theory} {\bf 20} (2020), 145--158.
\item Sen, Suparna: Segal-{B}argmann transform and {P}aley-{W}iener theorems on
{H}eisenberg motion groups, \emph{Adv. Pure Appl. Math.} {\bf 7} (2016), 13--28.
\item Senthil Raani, K. S.: {$L^p$}-{F}ourier asymptotics, {H}ardy-type inequality and
fractal measures, \emph{Monatsh. Math.} {\bf 184} (2017), 459--487.
\item Seshadri, Harish: Differential geometry in {I}ndia, \emph{Indian J. Pure Appl. Math.} {\bf 50} (2019), 795--799.
\item Seshadri, Harish and Verma, Kaushal: Some aspects of the automorphism groups of domains, \emph{Handbook of group actions. {V}ol. {III}} {\bf 40} (2018), 145--174.
\item Tanaka, Ryotaro and Sain, Debmalya: On symmetry of strong {B}irkhoff orthogonality in {$B(\Cal
H,\Cal K)$} and {$K(\Cal H,\Cal K)$}, \emph{Ann. Funct. Anal.} {\bf 11} (2020), 693--704.
\item Thangavelu, Sundaram: An analogue of {P}fannschmidt's theorem for the {H}eisenberg
group, \emph{J. Anal.} {\bf 26} (2018), 235--244.
\item Thangavelu, Sundaram and Naidu Dogga, Venku: {$L^p$}-{$L^2$} estimates for solutions of the wave equation
associated to the {G}rushin operator, \emph{Adv. Pure Appl. Math.} {\bf 9} (2018), 85--92.
\item Turgay, Nurettin Cenk and Upadhyay, Abhitosh: On biconservative hypersurfaces in 4-dimensional {R}iemannian
space forms, \emph{Math. Nachr.} {\bf 292} (2019), 905--921.
\item Urbina-Romero, Wilfredo: Gaussian harmonic analysis, \emph{} {\bf } (2019), xix+477.
\item Verma, Kaushal: Notes on the boundaries of quadrature domains, \emph{Anal. Math. Phys.} {\bf 9} (2019), 617--638.
\item Verma, Kaushal: Remembering {R}. {R}. {S}imha, \emph{Math. Newsl.} {\bf 27} (2017), 16--20.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan,
Govindan: On the role of four-wave mixing effect in the interactions
between nonlinear modes of coupled generalized nonlinear
{S}chr\"{o}dinger equation, \emph{Chaos} {\bf 29} (2019), 123135, 14.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan, Govindan
and Lakshmanan, M.: On symmetry preserving and symmetry broken bright, dark and
antidark soliton solutions of nonlocal nonlinear {S}chr\"{o}dinger
equation, \emph{Phys. Lett. A} {\bf 383} (2019), 15--26.
\item Yadav, Avinash Chand and Manchanda, Kaustubh and Ramaswamy,
Ramakrishna: Emergent organization in a model market, \emph{Phys. A} {\bf 482} (2017), 118--126.
\end{enumerate}

\subsection{Publications (all kinds) from web page with hyperlinks}

\begin{enumerate}
\item Aas, Erik and Ayyer, Arvind and Linusson, Svante and Potka,
Samu: \href{https://doi.org/10.1088/1751-8121/ab2e96}{The exact phase diagram for a semipermeable {TASEP} with
nonlocal boundary jumps}, \emph{J. Phys. A} {\bf 52} (2019), 355001, 19.
\item Adhikari, Kartick and Kishore Reddy, Nanda and Ram Reddy,
Tulasi and Saha, Koushik: \href{https://doi.org/10.1214/14-AIHP632}{Determinantal point processes in the plane from products of
random matrices}, \emph{Ann. Inst. Henri Poincar\'{e} Probab. Stat.} {\bf 52} (2016), 16--46.
\item Adhikari, Kartick and Reddy, Nanda Kishore: \href{https://doi.org/10.1093/imrn/rnw207}{Hole probabilities for finite and infinite {G}inibre
ensembles}, \emph{Int. Math. Res. Not. IMRN} {\bf } (2017), 6694--6730.
\item Adimurthi and Ratnakumar, P. K. and Sohani, Vijay Kumar: \href{https://doi.org/10.1017/S0308210516000081}{A {H}ardy-{S}obolev inequality for the twisted {L}aplacian}, \emph{Proc. Roy. Soc. Edinburgh Sect. A} {\bf 147} (2017), 1--23.
\item Aiyappan, S. and Jose, Editha C. and Lomerio, Ivy Carol B. and
Nandakumaran, A. K.: \href{https://doi.org/10.3233/asy-191526}{Control problem on a rough circular domain and homogenization}, \emph{Asymptot. Anal.} {\bf 115} (2019), 19--46.
\item Aiyappan, S. and Nandakumaran, A. K.: \href{https://doi.org/10.1002/mma.4231}{Optimal control problem in a domain with branched structure
and homogenization}, \emph{Math. Methods Appl. Sci.} {\bf 40} (2017), 3173--3189.
\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi: \href{https://doi.org/10.1142/S0219199719500299}{Semi-linear optimal control problem on a smooth oscillating
domain}, \emph{Commun. Contemp. Math.} {\bf 22} (2020), 1950029, 26.
\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi: \href{https://doi.org/10.1007/s10231-019-00848-7}{Locally periodic unfolding operator for highly oscillating
rough domains}, \emph{Ann. Mat. Pura Appl. (4)} {\bf 198} (2019), 1931--1954.
\item Aiyappan, S. and Nandakumaran, A. K. and Prakash, Ravi: \href{https://doi.org/10.1007/s00526-018-1354-6}{Generalization of unfolding operator for highly oscillating
smooth boundary domains and homogenization}, \emph{Calc. Var. Partial Differential Equations} {\bf 57} (2018), Art. 86, 30.
\item Aiyappan, S. and Nandakumaran, A. K. and Sufian, Abu: \href{https://doi.org/10.1002/mma.5748}{Asymptotic analysis of a boundary optimal control problem on a
general branched structure}, \emph{Math. Methods Appl. Sci.} {\bf 42} (2019), 6407--6434.
\item Aiyappan, S. and Sardar, Bidhan Chandra: \href{https://doi.org/10.1080/00036811.2018.1471207}{Biharmonic equation in a highly oscillating domain and
homogenization of an associated control problem}, \emph{Appl. Anal.} {\bf 98} (2019), 2783--2801.
\item Akutagawa, Kazuo and Endo, Hisaaki and Seshadri, Harish: \href{https://doi.org/10.1007/s00208-018-1749-x}{A gap theorem for positive {E}instein metrics on the
four-sphere}, \emph{Math. Ann.} {\bf 373} (2019), 1329--1339.
\item Anamby, Pramath and Das, Soumya: \href{https://doi.org/10.5565/PUBLMAT6311911}{Distinguishing {H}ermitian cusp forms of degree 2 by a certain
subset of all {F}ourier coefficients}, \emph{Publ. Mat.} {\bf 63} (2019), 307--341.
\item Anamby, Pramath and Das, Soumya: \href{https://doi.org/10.1016/j.jnt.2018.03.011}{Bounds for the {P}etersson norms of the pullbacks of
{S}aito-{K}urokawa lifts}, \emph{J. Number Theory} {\bf 191} (2018), 289--304.
\item Anand, Jatin and Bhattacharyya, Tirthankar and Srivastava,
Sachi: \href{https://doi.org/10.1007/s00013-019-01406-6}{Some thoughts on composition operators on subspaces of the
{H}ardy space}, \emph{Arch. Math. (Basel)} {\bf 114} (2020), 431--444.
\item Arunkumar, G. and Kus, Deniz and Venkatesh, R.: \href{https://doi.org/10.1016/j.jalgebra.2017.11.050}{Root multiplicities for {B}orcherds algebras and graph
coloring}, \emph{J. Algebra} {\bf 499} (2018), 538--569.
\item Ayyer, Arvind: \href{https://doi.org/10.1088/1751-8113/49/15/155003}{Full current statistics for a disordered open exclusion
process}, \emph{J. Phys. A} {\bf 49} (2016), 155003, 10.
\item Ayyer, Arvind: \href{https://doi.org/10.1007/s00026-019-00480-5}{Squareness for the {M}onopole-{D}imer model}, \emph{Ann. Comb.} {\bf 24} (2020), 237--255.
\item Ayyer, Arvind and Behrend, Roger E.: \href{https://doi.org/10.1016/j.jcta.2019.01.001}{Factorization theorems for classical group characters, with
applications to alternating sign matrices and plane
partitions}, \emph{J. Combin. Theory Ser. A} {\bf 165} (2019), 78--105.
\item Ayyer, Arvind and Behrend, Roger E. and Fischer, Ilse: \href{https://doi.org/10.1016/j.aim.2020.107125}{Extreme diagonally and antidiagonally symmetric alternating
sign matrices of odd order}, \emph{Adv. Math.} {\bf 367} (2020), 107125, 56.
\item Ayyer, Arvind and Bouttier, J\'{e}r\'{e}mie and Corteel, Sylvie and
Linusson, Svante and Nunzi, Fran\c{c}ois: \href{https://doi.org/10.1016/j.aam.2018.03.001}{Bumping sequences and multispecies juggling}, \emph{Adv. in Appl. Math.} {\bf 98} (2018), 100--126.
\item Ayyer, Arvind and Finn, Caley and Roy, Dipankar: \href{https://doi.org/10.1103/physreve.97.012151}{Matrix product solution of a left-permeable two-species
asymmetric exclusion process}, \emph{Phys. Rev. E} {\bf 97} (2018), 012151, 10.
\item Ayyer, Arvind and Finn, Caley and Roy, Dipankar: \href{https://doi.org/10.1007/s10955-018-2183-x}{The phase diagram for a multispecies left-permeable asymmetric
exclusion process}, \emph{J. Stat. Phys.} {\bf 174} (2019), 605--621.
\item Ayyer, Arvind and Fischer, Ilse: \href{https://doi.org/10.1016/j.jcta.2020.105241}{Bijective proofs of skew {S}chur polynomial factorizations}, \emph{J. Combin. Theory Ser. A} {\bf 174} (2020), 105241, 40.
\item Ayyer, Arvind and Josuat-Verg\`es, Matthieu and Ramassamy,
Sanjay: \href{https://doi.org/10.5802/ahl.33}{Extensions of partial cyclic orders and consecutive coordinate
polytopes}, \emph{Ann. H. Lebesgue} {\bf 3} (2020), 275--297.
\item Ayyer, Arvind and Linusson, Svante: \href{https://doi.org/10.1090/tran/6806}{Correlations in the multispecies {TASEP} and a conjecture by
{L}am}, \emph{Trans. Amer. Math. Soc.} {\bf 369} (2017), 1097--1125.
\item Ayyer, Arvind and Linusson, Svante: \href{https://doi.org/10.1002/rsa.20825}{Reverse juggling processes}, \emph{Random Structures Algorithms} {\bf 55} (2019), 56--72.
\item Ayyer, Arvind and Prasad, Amritanshu and Spallone, Steven: Odd partitions in {Y}oung's lattice, \emph{S\'{e}m. Lothar. Combin.} {\bf 75} (2018), Art. B75g, 13.
\item Ayyer, Arvind and Prasad, Amritanshu and Spallone, Steven: \href{https://doi.org/10.1016/j.jcta.2017.03.004}{Representations of symmetric groups with non-trivial
determinant}, \emph{J. Combin. Theory Ser. A} {\bf 150} (2017), 208--232.
\item Ayyer, Arvind and Ramassamy, Sanjay: \href{https://doi.org/10.30757/alea.v15-28}{The {H}ilbert-{G}alton board}, \emph{ALEA Lat. Am. J. Probab. Math. Stat.} {\bf 15} (2018), 755--774.
\item Ayyer, Arvind and Schilling, Anne and Thi\'{e}ry, Nicolas M.: \href{https://doi.org/10.1080/10586458.2015.1107868}{Spectral gap for random-to-random shuffling on linear
extensions}, \emph{Exp. Math.} {\bf 26} (2017), 22--30.
\item Ayyer, Arvind and Singla, Pooja: \href{https://doi.org/10.1007/s10468-019-09864-w}{Random motion on finite rings, {I}: commutative rings}, \emph{Algebr. Represent. Theory} {\bf 23} (2020), 583--604.
\item Ayyer, Arvind and Sinha, Shubham: Random {$t$}-cores and hook lengths in random partitions, \emph{S\'{e}m. Lothar. Combin.} {\bf 84B} (2020), Art. 58, 11.
\item Ayyer, Arvind and Steinberg, Benjamin: \href{https://doi.org/10.5802/alco.94}{Random walks on rings and modules}, \emph{Algebr. Comb.} {\bf 3} (2020), 309--329.
\item B\"{o}cherer, Siegfried and Das, Soumya: \href{https://doi.org/10.1007/s00209-015-1609-2}{Cuspidality and the growth of {F}ourier coefficients: small
weights}, \emph{Math. Z.} {\bf 283} (2016), 539--553.
\item B\"{o}cherer, Siegfried and Das, Soumya: \href{https://doi.org/10.1515/crelle-2015-0075}{Cuspidality and the growth of {F}ourier coefficients of
modular forms}, \emph{J. Reine Angew. Math.} {\bf 741} (2018), 161--178.
\item B\"{o}cherer, Siegfried and Das, Soumya: \href{https://doi.org/10.1016/j.aim.2018.07.033}{Petersson norms of not necessarily cuspidal {J}acobi modular
forms and applications}, \emph{Adv. Math.} {\bf 336} (2018), 335--376.
\item Bagchi, Bhaskar and Burton, Benjamin A. and Datta, Basudeb and
Singh, Nitin and Spreer, Jonathan: Efficient algorithms to decide tightness, \emph{32nd {I}nternational {S}ymposium on {C}omputational
{G}eometry} {\bf 51} (2016), Art. 12, 15.
\item Bagchi, Bhaskar and Datta, Basudeb and Spreer, Jonathan: \href{https://doi.org/10.1016/j.ejc.2016.10.005}{A characterization of tightly triangulated 3-manifolds}, \emph{European J. Combin.} {\bf 61} (2017), 133--137.
\item Bagchi, Bhaskar and Datta, Basudeb and Spreer, Jonathan: \href{https://doi.org/10.1016/j.ejc.2015.12.006}{Tight triangulations of closed 3-manifolds}, \emph{European J. Combin.} {\bf 54} (2016), 103--120.
\item Bagchi, Sayan and Thangavelu, Sundaram: \href{https://doi.org/10.1007/s11854-018-0053-8}{Weighted norm inequalities for {W}eyl multipliers and fourier
multipliers on the {H}eisenberg group}, \emph{J. Anal. Math.} {\bf 136} (2018), 1--29.
\item Baklouti, Ali and Thangavelu, Sundaram: \href{https://doi.org/10.1017/nmj.2016.58}{Hardy and {M}iyachi theorems for {H}eisenberg motion groups}, \emph{Nagoya Math. J.} {\bf 229} (2018), 1--20.
\item Balakumar, G. P. and Borah, Diganta and Mahajan, Prachi and
Verma, Kaushal: \href{https://doi.org/10.4064/ap200203-21-4}{Further remarks on the higher dimensional {S}uita conjecture}, \emph{Ann. Polon. Math.} {\bf 125} (2020), 101--115.
\item Balakumar, G. P. and Borah, Diganta and Mahajan, Prachi and
Verma, Kaushal: \href{https://doi.org/10.1090/proc/14421}{Remarks on the higher dimensional {S}uita conjecture}, \emph{Proc. Amer. Math. Soc.} {\bf 147} (2019), 3401--3411.
\item Balhara, Rakesh: \href{https://doi.org/10.1007/s12044-019-0471-2}{Hardy's inequality for the fractional powers of the {G}rushin
operator}, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 129} (2019), Art. 33, 25.
\item Balodi, Mamta and Banerjee, Abhishek and Naolekar, Anita: \href{https://doi.org/10.5802/crmath.157}{B{V}-operators and the secondary {H}ochschild complex}, \emph{C. R. Math. Acad. Sci. Paris} {\bf 358} (2020), 1239--1258.
\item Balodi, Mamta and Banerjee, Abhishek and Ray, Samarpita: \href{https://doi.org/10.4153/s0008414x19000403}{Cohomology of modules over {$H$}-categories and
co-{$H$}-categories}, \emph{Canad. J. Math.} {\bf 72} (2020), 1352--1385.
\item Balodi, Mamta and Huang, Hua-Lin and Kumar, Shiv Datt: \href{https://doi.org/10.1142/S0129055X17300035}{On the classification of finite quasi-quantum groups}, \emph{Rev. Math. Phys.} {\bf 29} (2017), 1730003, 20.
\item Balodi, Mamta and Palcoux, Sebastien: \href{https://doi.org/10.1016/j.jcta.2018.02.004}{On {B}oolean intervals of finite groups}, \emph{J. Combin. Theory Ser. A} {\bf 157} (2018), 49--69.
\item Banerjee, Abhishek: \href{https://doi.org/10.1007/s00013-017-1025-4}{On some spectral spaces associated to tensor triangulated
categories}, \emph{Arch. Math. (Basel)} {\bf 108} (2017), 581--591.
\item Banerjee, Abhishek: \href{https://doi.org/10.1142/S0219199717500717}{On {A}uslander's formula and cohereditary torsion pairs}, \emph{Commun. Contemp. Math.} {\bf 20} (2018), 1750071, 27.
\item Banerjee, Abhishek: \href{https://doi.org/10.1016/j.geomphys.2018.04.008}{Modular {H}ecke algebras over {M}\"{o}bius categories}, \emph{J. Geom. Phys.} {\bf 131} (2018), 23--40.
\item Banerjee, Abhishek: Closure operators in abelian categories and spectral spaces, \emph{Theory Appl. Categ.} {\bf 32} (2017), Paper No. 20, 719--735.
\item Banerjee, Abhishek: \href{https://doi.org/10.1142/S0129167X17500513}{Noetherian schemes over abelian symmetric monoidal categories}, \emph{Internat. J. Math.} {\bf 28} (2017), 1750051, 34.
\item Banerjee, Abhishek: \href{https://doi.org/10.4153/s0008439518000656}{On differential torsion theories and rings with several
objects}, \emph{Canad. Math. Bull.} {\bf 62} (2019), 703--714.
\item Banerjee, Abhishek: \href{http://projecteuclid.org/euclid.bbms/1457560860}{Une note sur les d\'erivations localement nilpotentes dans les
cat\'egories mono\"\i dales sym\'etriques}, \emph{Bull. Belg. Math. Soc. Simon Stevin} {\bf 23} (2016), 151--160.
\item Banerjee, Abhishek: \href{http://dx.doi.org/10.1007/s40879-016-0108-2}{Realizations of pairs and {O}ka families in tensor
triangulated categories}, \emph{Eur. J. Math.} {\bf 2} (2016), 760--797.
\item Banerjee, Abhishek: \href{http://dx.doi.org/10.7146/math.scand.a-23294}{Homologie de {H}ochschild et espaces alg\'ebriques}, \emph{Math. Scand.} {\bf 118} (2016), 33--42.
\item Banerjee, Abhishek: \href{https://doi.org/10.1016/j.crma.2018.02.012}{A topological {N}ullstellensatz for tensor-triangulated
categories}, \emph{C. R. Math. Acad. Sci. Paris} {\bf 356} (2018), 365--375.
\item Banerjee, Abhishek: \href{http://dx.doi.org/10.1016/j.crma.2016.05.005}{Some remarks on a theorem of {B}ergman}, \emph{C. R. Math. Acad. Sci. Paris} {\bf 354} (2016), 665--670.
\item Banerjee, Abhishek: \href{https://doi.org/10.4064/fm218-8-2016}{Monoidal semifilters and arrays of prime ideals}, \emph{Fund. Math.} {\bf 237} (2017), 281--296.
\item Banerjee, Abhishek: \href{http://dx.doi.org/10.1016/j.jpaa.2016.02.001}{Schemes over symmetric monoidal categories and torsion
theories}, \emph{J. Pure Appl. Algebra} {\bf 220} (2016), 3017--3047.
\item Banerjee, Abhishek: \href{https://doi.org/10.4171/JNCG/297}{Quasimodular {H}ecke algebras and {H}opf actions}, \emph{J. Noncommut. Geom.} {\bf 12} (2018), 1041--1080.
\item Banerjee, Abhishek: \href{https://doi.org/10.1016/j.jalgebra.2018.04.022}{Completions of monoid objects and descent results}, \emph{J. Algebra} {\bf 507} (2018), 362--399.
\item Banerjee, Abhishek and Kour, Surjeet: \href{https://doi.org/10.1007/s10231-019-00844-x}{{$(A,\delta)$}-modules, {H}ochschild homology and higher
derivations}, \emph{Ann. Mat. Pura Appl. (4)} {\bf 198} (2019), 1781--1802.
\item Barbosa, Ezequiel and Santana, Farley and Upadhyay, Abhitosh: \href{https://doi.org/10.1142/S0129167X2050086X}{On stable {CMC} free-boundary surfaces in a strictly convex
domain of a bi-invariant {L}ie group}, \emph{Internat. J. Math.} {\bf 31} (2020), 2050086, 9.
\item Basak, Biplab: \href{https://doi.org/10.1007/s12044-016-0302-7}{An algorithmic approach to construct crystallizations of
3-manifolds from presentations of fundamental groups}, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 126} (2016), 629--653.
\item Basak, Biplab and Casali, Maria Rita: \href{https://doi.org/10.1515/forum-2015-0080}{Lower bounds for regular genus and gem-complexity of {PL}
4-manifolds}, \emph{Forum Math.} {\bf 29} (2017), 761--773.
\item Basak, Biplab and Spreer, Jonathan: \href{https://doi.org/10.1515/advgeom-2015-0043}{Simple crystallizations of 4-manifolds}, \emph{Adv. Geom.} {\bf 16} (2016), 111--130.
\item Basak, Gopal K. and Ghosh, Mrinal K. and Mukherjee, Diganta: \href{https://doi.org/10.1080/07362994.2015.1106321}{A mean-reverting stochastic model for the political business
cycle}, \emph{Stoch. Anal. Appl.} {\bf 34} (2016), 96--116.
\item Basu, Arnab and Ghosh, Mrinal K.: \href{https://doi.org/10.1287/moor.2017.0870}{Nonzero-sum risk-sensitive stochastic games on a countable
state space}, \emph{Math. Oper. Res.} {\bf 43} (2018), 516--532.
\item Belton, Alexander and Guillot, Dominique and Khare, Apoorva
and Putinar, Mihai: \href{https://doi.org/10.1016/j.laa.2018.03.035}{Simultaneous kernels of matrix {H}adamard powers}, \emph{Linear Algebra Appl.} {\bf 576} (2019), 142--157.
\item Benedetti, Carolina and Gonz\'{a}lez D'Le\'{o}n, Rafael S. and Hanusa,
Christopher R. H. and Harris, Pamela E. and Khare, Apoorva and
Morales, Alejandro H. and Yip, Martha: The volume of the caracol polytope, \emph{S\'{e}m. Lothar. Combin.} {\bf 80B} (2018), Art. 87, 12.
\item Benedetti, Carolina and Gonz\'{a}lez D'Le\'{o}n, Rafael S. and Hanusa,
Christopher R. H. and Harris, Pamela E. and Khare, Apoorva and
Morales, Alejandro H. and Yip, Martha: \href{https://doi.org/10.1090/tran/7743}{A combinatorial model for computing volumes of flow polytopes}, \emph{Trans. Amer. Math. Soc.} {\bf 372} (2019), 3369--3404.
\item Bennett, Nathan and Iyer, Srikanth K. and Jammalamadaka, S.
Rao: \href{https://doi.org/10.1080/03610926.2015.1041981}{Analysis of gamma and {W}eibull lifetime data under a general
censoring scheme and in the presence of covariates}, \emph{Comm. Statist. Theory Methods} {\bf 46} (2017), 2277--2289.
\item Bera, Sayani and Pal, Ratna: \href{https://doi.org/10.1186/s40627-016-0006-x}{Dynamics of semigroups of entire maps of {$\Bbb{C}^k$}}, \emph{Complex Anal. Synerg.} {\bf 2} (2016), Paper No. 2, 22.
\item Bera, Sayani and Pal, Ratna and Verma, Kaushal: \href{https://projecteuclid.org/euclid.ijm/1534924839}{Examples of non-autonomous basins of attraction}, \emph{Illinois J. Math.} {\bf 61} (2017), 531--567.
\item Bera, Sayani and Pal, Ratna and Verma, Kaushal: \href{https://doi.org/10.1007/s40879-019-00326-7}{A rigidity theorem for {H}\'{e}non maps}, \emph{Eur. J. Math.} {\bf 6} (2020), 508--532.
\item Bera, Sayani and Verma, Kaushal: \href{https://doi.org/10.1007/s12044-018-0388-1}{Some aspects of shift-like automorphisms of {$\Bbb C^k$}}, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 128} (2018), Art. 10, 48.
\item Bera, Sudip and Mukherjee, Sajal Kumar: \href{https://doi.org/10.1007/s00373-020-02223-3}{Generalized power sum and {N}ewton-{G}irard identities}, \emph{Graphs Combin.} {\bf 36} (2020), 1957--1964.
\item Bhamidi, Shankar and Dhara, Souvik and van der Hofstad, Remco
and Sen, Sanchayan: \href{https://doi.org/10.1214/19-ejp408}{Universality for critical heavy-tailed network models: metric
structure of maximal components}, \emph{Electron. J. Probab.} {\bf 25} (2020), Paper No. 47, 57.
\item Bhamidi, Shankar and Sen, Sanchayan: \href{https://doi.org/10.1002/rsa.20880}{Geometry of the vacant set left by random walk on random
graphs, {W}right's constants, and critical random graphs with
prescribed degrees}, \emph{Random Structures Algorithms} {\bf 56} (2020), 676--721.
\item Bharali, Gautam: \href{https://doi.org/10.5186/aasfm.2016.4116}{Complex geodesics, their boundary regularity, and a
{H}ardy-{L}ittlewood-type lemma}, \emph{Ann. Acad. Sci. Fenn. Math.} {\bf 41} (2016), 253--263.
\item Bharali, Gautam: \href{https://doi.org/10.1007/s12220-019-00342-9}{On the growth of the {B}ergman metric near a point of infinite
type}, \emph{J. Geom. Anal.} {\bf 30} (2020), 1238--1258.
\item Bharali, Gautam and Biswas, Indranil and Divakaran, Divakaran
and Janardhanan, Jaikrishnan: Proper holomorphic mappings onto symmetric products of a
{R}iemann surface, \emph{Doc. Math.} {\bf 23} (2018), 1291--1311.
\item Bharali, Gautam and Biswas, Indranil and Schumacher, Georg: \href{https://doi.org/10.1080/17476933.2016.1252342}{A criterion for a degree-one holomorphic map to be a
biholomorphism}, \emph{Complex Var. Elliptic Equ.} {\bf 62} (2017), 914--918.
\item Bharali, Gautam and Chandel, Vikramjeet Singh: \href{https://doi.org/10.1007/s11785-017-0701-5}{Pick interpolation on the polydisc: small families of
sufficient kernels}, \emph{Complex Anal. Oper. Theory} {\bf 13} (2019), 2069--2093.
\item Bharali, Gautam and Sridharan, Shrihari: \href{https://doi.org/10.1142/S0129167X17501087}{Holomorphic correspondences related to finitely generated
rational semigroups}, \emph{Internat. J. Math.} {\bf 28} (2017), 1750108, 25.
\item Bharali, Gautam and Sridharan, Shrihari: \href{https://doi.org/10.1080/17476933.2016.1185419}{The dynamics of holomorphic correspondences of {$\Bbb P^1$}:
invariant measures and the normality set}, \emph{Complex Var. Elliptic Equ.} {\bf 61} (2016), 1587--1613.
\item Bharali, Gautam and Zimmer, Andrew: \href{https://doi.org/10.1016/j.aim.2017.02.005}{Goldilocks domains, a weak notion of visibility, and
applications}, \emph{Adv. Math.} {\bf 310} (2017), 377--425.
\item Bhatia, Manan: \href{https://doi.org/10.1007/s10955-020-02632-x}{Moderate deviation and exit time estimates for stationary last
passage percolation}, \emph{J. Stat. Phys.} {\bf 181} (2020), 1410--1432.
\item Bhattacharjee, M. and Eschmeier, J. and Keshari, Dinesh K. and
Sarkar, Jaydeb: \href{https://doi.org/10.1016/j.laa.2017.02.032}{Dilations, wandering subspaces, and inner functions}, \emph{Linear Algebra Appl.} {\bf 523} (2017), 263--280.
\item Bhattacharjee, Monojit and Sarkar, Jaydeb: Operator positivity and analytic models of commuting tuples of
operators, \emph{Studia Math.} {\bf 232} (2016), 155--171.
\item Bhattacharyya, T. and Narayanan, E. K. and Sarkar, Jaydeb: \href{https://doi.org/10.7153/oam-11-07}{Analytic model of doubly commuting contractions}, \emph{Oper. Matrices} {\bf 11} (2017), 101--113.
\item Bhattacharyya, T. and Sau, H.: \href{https://doi.org/10.1007/s11785-015-0472-9}{Explicit and unique construction of tetrablock unitary
dilation in a certain case}, \emph{Complex Anal. Oper. Theory} {\bf 10} (2016), 749--768.
\item Bhattacharyya, Tirthankar and Biswas, Anindya and Chandel,
Vikramjeet Singh: \href{https://doi.org/10.4064/sm190505-8-10}{On the {N}evanlinna problem: characterization of all
{S}chur-{A}gler class solutions affiliated with a given
kernel}, \emph{Studia Math.} {\bf 255} (2020), 83--107.
\item Bhattacharyya, Tirthankar and Rao, T. S. S. R. K.: \href{https://doi.org/10.1007/s13226-019-0345-4}{Survey of last ten years of work done in {I}ndia in some
selected areas of functional analysis and operator theory}, \emph{Indian J. Pure Appl. Math.} {\bf 50} (2019), 599--617.
\item Bhattacharyya, Tirthankar and Sau, Haripada: \href{https://doi.org/10.4171/PRIMS/53-2-2}{{$\Gamma$}-unitaries, dilation and a natural example}, \emph{Publ. Res. Inst. Math. Sci.} {\bf 53} (2017), 261--285.
\item Bhattacharyya, Tirthankar and Sau, Haripada: \href{https://doi.org/10.1016/j.jfa.2017.09.013}{Holomorphic functions on the symmetrized bidisk---realization,
interpolation and extension}, \emph{J. Funct. Anal.} {\bf 274} (2018), 504--524.
\item Bhimani, Divyang G. and Balhara, Rakesh and Thangavelu,
Sundaram: Hermite multipliers on modulation spaces, \emph{Analysis and partial differential equations: perspectives from
developing countries} {\bf 275} (2019), 42--64.
\item Bhosle, Usha N.: \href{https://doi.org/10.1142/S0129167X18500155}{Hitchin pairs on reducible curves}, \emph{Internat. J. Math.} {\bf 29} (2018), 1850015, 49.
\item Bhosle, Usha N.: On a conjecture of {B}utler, \emph{Analytic and algebraic geometry} {\bf } (2017), 29--48.
\item Bhosle, Usha N. and Biswas, Indranil: \href{https://doi.org/10.1007/s10711-015-0108-2}{Moduli spaces of vector bundles on a singular rational ruled
surface}, \emph{Geom. Dedicata} {\bf 180} (2016), 399--413.
\item Biswas, Indranil and D'Mello, Shane and Mukherjee, Ritwik and
Pingali, Vamsi P.: Rational cuspidal curves on del-{P}ezzo surfaces, \emph{J. Singul.} {\bf 17} (2018), 91--107.
\item Biswas, Indranil and Dumitrescu, Sorin and Gupta, Subhojoy: Branched projective structures on a {R}iemann surface and
logarithmic connections, \emph{Doc. Math.} {\bf 24} (2019), 2299--2337.
\item Biswas, Indranil and Pingali, Vamsi Pritham: A characterization of finite vector bundles on {G}auduchon
astheno-{K}\"{a}hler manifolds, \emph{\'{E}pijournal Geom. Alg\'{e}brique} {\bf 2} (2018), Art. 6, 13.
\item Biswas, Indranil and Pingali, Vamsi Pritham: \href{https://doi.org/10.1093/imrn/rny259}{Metric properties of parabolic ample bundles}, \emph{Int. Math. Res. Not. IMRN} {\bf } (2020), 9336--9369.
\item Biswas, Indranil and Seshadri, Harish: \href{https://doi.org/10.4310/PAMQ.2016.v12.n4.a1}{Chern scalar curvature and symmetric products of compact
{R}iemann surfaces}, \emph{Pure Appl. Math. Q.} {\bf 12} (2016), 463--471.
\item Biswas, Indranil and Seshadri, Harish: \href{https://doi.org/10.1016/j.crma.2017.03.012}{Quot schemes and {R}icci semipositivity}, \emph{C. R. Math. Acad. Sci. Paris} {\bf 355} (2017), 577--581.
\item Biswas, Rahul and Dond, Asha K. and Gudi, Thirupathi: \href{https://doi.org/10.1515/cmam-2018-0020}{Edge patch-wise local projection stabilized nonconforming
{FEM} for the {O}seen problem}, \emph{Comput. Methods Appl. Math.} {\bf 19} (2019), 189--214.
\item Biswas, Shibananda and Ghosh, Gargi and Misra, Gadadhar and
Shyam Roy, Subrata: \href{https://doi.org/10.1016/j.jfa.2018.10.025}{On reducing submodules of {H}ilbert modules with
{${S}_n$}-invariant kernels}, \emph{J. Funct. Anal.} {\bf 276} (2019), 751--784.
\item Bleher, F. M. and Chinburg, T. and Greenberg, R. and Kakde, M.
and Pappas, G. and Sharifi, R. and Taylor, M. J.: \href{https://doi.org/10.1353/ajm.2020.0017}{Higher {C}hern classes in {I}wasawa theory}, \emph{Amer. J. Math.} {\bf 142} (2020), 627--682.
\item Boggarapu, Pradeep and Roncal, Luz and Thangavelu, Sundaram: \href{https://doi.org/10.1090/tran/6861}{Mixed norm estimates for the {C}es\`aro means associated with
{D}unkl-{H}ermite expansions}, \emph{Trans. Amer. Math. Soc.} {\bf 369} (2017), 7021--7047.
\item Boggarapu, Pradeep and Roncal, Luz and Thangavelu, Sundaram: \href{https://doi.org/10.3934/cpaa.2019116}{On extension problem, trace {H}ardy and {H}ardy's inequalities
for some fractional {L}aplacians}, \emph{Commun. Pure Appl. Anal.} {\bf 18} (2019), 2575--2605.
\item Boggarapu, Pradeep and Thangavelu, S.: \href{https://doi.org/10.1007/s11856-017-1438-6}{On the chaotic behavior of the {D}unkl heat semigroup on
weighted {$L^p$} spaces}, \emph{Israel J. Math.} {\bf 217} (2017), 57--92.
\item Boggarapu, Pradeep and Thangavelu, S.: \href{http://projecteuclid.org/euclid.pm/1450818486}{Mixed norm estimates for the {R}iesz transforms on {$SU(2)$}}, \emph{Publ. Mat.} {\bf 60} (2016), 171--190.
\item Borah, Diganta and Haridas, Pranav and Verma, Kaushal: \href{https://doi.org/10.1007/s13324-017-0177-5}{Comments on the {G}reen's function of a planar domain}, \emph{Anal. Math. Phys.} {\bf 8} (2018), 383--414.
\item Brenner, Susanne C. and Gudi, Thirupathi and Porwal, Kamana
and Sung, Li-Yeng: \href{https://doi.org/10.1051/cocv/2017031}{A {M}orley finite element method for an elliptic distributed
optimal control problem with pointwise state and control
constraints}, \emph{ESAIM Control Optim. Calc. Var.} {\bf 24} (2018), 1181--1206.
\item Burton, Benjamin A. and Datta, Basudeb and Singh, Nitin and
Spreer, Jonathan: \href{https://doi.org/10.1080/10586458.2016.1212747}{A construction principle for tight and minimal triangulations
of manifolds}, \emph{Exp. Math.} {\bf 27} (2018), 22--36.
\item Carstensen, Carsten and Dond, Asha K. and Rabus, Hella: \href{https://doi.org/10.1515/cmam-2019-0034}{Quasi-optimality of adaptive mixed {FEM}s for non-selfadjoint
indefinite second-order linear elliptic problems}, \emph{Comput. Methods Appl. Math.} {\bf 19} (2019), 233--250.
\item Chakrabarti, A. and Mondal, Sumita Rani: On approximate numerical method of solution of {F}redholm
integral equations, \emph{Int. J. Appl. Math. Anal. Appl.} {\bf 11} (2016), 153--162.
\item Chakraborty, Atlanta and Chandru, Vijay and Rao, M. R.: \href{https://doi.org/10.1007/s10479-019-03186-2}{A linear programming primer: from {F}ourier to {K}armarkar}, \emph{Ann. Oper. Res.} {\bf 287} (2020), 593--616.
\item Chan, Chun Lam and Fernandes, Winston and Kashyap, Navin and
Krishnapur, Manjunath: \href{https://doi.org/10.1137/15M1028819}{Phase transitions for the uniform distribution in the pattern
maximum likelihood problem and its {B}ethe approximation}, \emph{SIAM J. Discrete Math.} {\bf 31} (2017), 597--631.
\item Chandel, Vikramjeet Singh: \href{https://doi.org/10.1080/17476933.2017.1370461}{The three-point {P}ick-{N}evanlinna interpolation problem on
the polydisc}, \emph{Complex Var. Elliptic Equ.} {\bf 63} (2018), 1341--1352.
\item Chandel, Vikramjeet Singh: \href{https://doi.org/10.1007/s12220-018-00123-w}{The 3-point spectral pick interpolation problem and an
application to holomorphic correspondences}, \emph{J. Geom. Anal.} {\bf 30} (2020), 551--572.
\item Chas, Moira and Gadgil, Siddhartha: \href{https://doi.org/10.2140/agt.2016.16.2813}{The extended {G}oldman bracket determines intersection numbers
for surfaces and orbifolds}, \emph{Algebr. Geom. Topol.} {\bf 16} (2016), 2813--2838.
\item Chatterjee, Esha and Hassan, Sk. Sarif: \href{https://doi.org/10.3934/dcds.2018070}{On the asymptotic character of a generalized rational
difference equation}, \emph{Discrete Contin. Dyn. Syst.} {\bf 38} (2018), 1707--1718.
\item Chattopadhyay, Arup and Sain, Debmalya and Senapati, Tanusri: \href{https://doi.org/10.14232/actasm-019-554-z}{A study of non-positive operators between real normed linear
spaces}, \emph{Acta Sci. Math. (Szeged)} {\bf 86} (2020), 449--466.
\item Choudhary, A. and Martha, S. C. and Chakrabarti, A.: \href{https://doi.org/10.1007/s40010-019-00600-z}{Modified method for the solution of dual trigonometric series
relations}, \emph{Proc. Nat. Acad. Sci. India Sect. A} {\bf 90} (2020), 437--445.
\item Choudhury, Projesh Nath and Kannan, M. Rajesh and Sivakumar,
K. C.: \href{https://doi.org/10.13001/1081-3810.3864}{A note on linear preservers of semipositive and minimally
semipositive matrices}, \emph{Electron. J. Linear Algebra} {\bf 34} (2018), 687--694.
\item Choudhury, Projesh Nath and Tsatsomeros, Michael J.: \href{https://doi.org/10.1016/j.laa.2020.04.028}{Algorithmic detection and construction of {N}-matrices}, \emph{Linear Algebra Appl.} {\bf 602} (2020), 46--56.
\item Chowdhury, Sudipto and Gudi, Thirupathi: \href{https://doi.org/10.1016/j.cam.2016.12.005}{A {$C^0$} interior penalty method for the {D}irichlet control
problem governed by biharmonic operator}, \emph{J. Comput. Appl. Math.} {\bf 317} (2017), 290--306.
\item Chowdhury, Sudipto and Gudi, Thirupathi and Nandakumaran, A.
K.: \href{https://doi.org/10.1090/mcom/3125}{Error bounds for a {D}irichlet boundary control problem based
on energy spaces}, \emph{Math. Comp.} {\bf 86} (2017), 1103--1126.
\item Ciaurri, \'{O}scar and Roncal, Luz and Thangavelu, Sundaram: \href{https://doi.org/10.1017/s0013091517000311}{Hardy-type inequalities for fractional powers of the
{D}unkl-{H}ermite operator}, \emph{Proc. Edinb. Math. Soc. (2)} {\bf 61} (2018), 513--544.
\item Cicuttin, Matteo and Ern, Alexandre and Gudi, Thirupathi: \href{https://doi.org/10.1007/s10915-020-01195-z}{Hybrid high-order methods for the elliptic obstacle problem}, \emph{J. Sci. Comput.} {\bf 83} (2020), Paper No. 8, 18.
\item Cui, Jintao and Gudi, Thirupathi: \href{https://doi.org/10.1007/s10444-016-9495-7}{Finite element approximation of a free boundary plasma
problem}, \emph{Adv. Comput. Math.} {\bf 43} (2017), 517--535.
\item Das, Soumya: \href{https://doi.org/10.1007/s13226-015-0165-0}{Cuspidality and the growth of {F}ourier coefficients of
modular forms: a survey}, \emph{Indian J. Pure Appl. Math.} {\bf 47} (2016), 9--22.
\item Das, Soumya and Ganguly, Satadal: \href{https://doi.org/10.1090/proc/12887}{A note on small gaps between nonzero {F}ourier coefficients of
cusp forms}, \emph{Proc. Amer. Math. Soc.} {\bf 144} (2016), 2301--2305.
\item Das, Soumya and Jha, Abhash Kumar: \href{https://doi.org/10.1007/s12188-019-00206-7}{Analytic properties of twisted real-analytic {H}ermitian
{K}lingen type {E}isenstein series and applications}, \emph{Abh. Math. Semin. Univ. Hambg.} {\bf 89} (2019), 105--116.
\item Das, Soumya and Khan, Rizwanur: \href{https://doi.org/10.1093/qmath/hay012}{The third moment of symmetric square {$L$}-functions}, \emph{Q. J. Math.} {\bf 69} (2018), 1063--1087.
\item Das, Soumya and Kohnen, Winfried: \href{https://doi.org/10.4064/aa170419-4-11}{On sign changes of eigenvalues of {S}iegel cusp forms of genus
2 in prime powers}, \emph{Acta Arith.} {\bf 183} (2018), 167--172.
\item Das, Soumya and Pal, Ritwik: \href{https://doi.org/10.1007/s40993-019-0158-x}{The first negative eigenvalue of {Y}oshida lifts}, \emph{Res. Number Theory} {\bf 5} (2019), Paper No. 20, 9.
\item Das, Soumya and Pal, Ritwik: \href{https://doi.org/10.1016/j.jnt.2017.03.021}{Jacobi forms and differential operators: odd weights}, \emph{J. Number Theory} {\bf 179} (2017), 113--125.
\item Datar, Ved V. and Pingali, Vamsi Pritham: \href{https://doi.org/10.4310/JSG.2020.v18.n4.a1}{On coupled constant scalar curvature {K}\"{a}hler metrics}, \emph{J. Symplectic Geom.} {\bf 18} (2020), 961--994.
\item Datta, Basudeb and Maity, Dipendu: \href{https://doi.org/10.1007/s13366-019-00458-7}{Correction to: {S}emi-equivelar and vertex-transitive maps on
the torus}, \emph{Beitr. Algebra Geom.} {\bf 61} (2020), 187--188.
\item Datta, Basudeb and Maity, Dipendu: \href{https://doi.org/10.1007/s13366-017-0332-z}{Semi-equivelar and vertex-transitive maps on the torus}, \emph{Beitr. Algebra Geom.} {\bf 58} (2017), 617--634.
\item Datta, Basudeb and Maity, Dipendu: \href{https://doi.org/10.1016/j.disc.2018.08.016}{Semi-equivelar maps on the torus and the {K}lein bottle are
{A}rchimedean}, \emph{Discrete Math.} {\bf 341} (2018), 3296--3309.
\item Datta, Basudeb and Murai, Satoshi: On stacked triangulated manifolds, \emph{Electron. J. Combin.} {\bf 24} (2017), Paper 4.12, 14.
\item Datta, Basudeb and Sarkar, Soumen: \href{https://doi.org/10.1007/jhep01(2020)036}{On equilibrium triangulations of quasitoric manifolds}, \emph{Southeast Asian Bull. Math.} {\bf 44} (2020), 57-56.
\item Dhanya, R. and Prashanth, S. and Tiwari, Sweta and Sreenadh,
K.: \href{https://doi.org/10.1080/17476933.2016.1198787}{Elliptic problems in {${\Bbb R}^N$} with critical and singular
discontinuous nonlinearities}, \emph{Complex Var. Elliptic Equ.} {\bf 61} (2016), 1668--1688.
\item Dhanya, R. and Sarkar, Abhishek: \href{https://doi.org/10.1080/17476933.2015.1055735}{Isolated singularities of polyharmonic operator in even
dimension}, \emph{Complex Var. Elliptic Equ.} {\bf 61} (2016), 55--66.
\item Dhara, Souvik and van der Hofstad, Remco and van Leeuwaarden,
Johan S. H. and Sen, Sanchayan: \href{https://doi.org/10.1214/19-AIHP980}{Heavy-tailed configuration models at criticality}, \emph{Ann. Inst. Henri Poincar\'{e} Probab. Stat.} {\bf 56} (2020), 1515--1558.
\item Dhillon, Gurbir and Khare, Apoorva: \href{https://doi.org/10.1016/j.aim.2017.08.005}{Faces of highest weight modules and the universal {W}eyl
polyhedron}, \emph{Adv. Math.} {\bf 319} (2017), 111--152.
\item Dond, Asha K. and Gudi, Thirupathi: \href{https://doi.org/10.1002/num.22317}{Patch-wise local projection stabilized finite element methods
for convection-diffusion-reaction problems}, \emph{Numer. Methods Partial Differential Equations} {\bf 35} (2019), 638--663.
\item Dond, Asha K. and Gudi, Thirupathi and Nataraj, Neela: \href{https://doi.org/10.1515/cmam-2016-0024}{A nonconforming finite element approximation for optimal
control of an obstacle problem}, \emph{Comput. Methods Appl. Math.} {\bf 16} (2016), 653--666.
\item Dond, Asha K. and Gudi, Thirupathi and Sau, Ramesh CH.: \href{https://doi.org/10.1080/01630563.2018.1538158}{An error analysis of discontinuous finite element methods for
the optimal control problems governed by {S}tokes equation}, \emph{Numer. Funct. Anal. Optim.} {\bf 40} (2019), 421--460.
\item Fern\'{o}s, Talia and Singla, Pooja: \href{https://doi.org/10.1017/S001708951500018X}{On images of real representations of special linear groups
over complete discrete valuation rings}, \emph{Glasg. Math. J.} {\bf 58} (2016), 263--272.
\item Fritz, Tobias and Gadgil, Siddhartha and Khare, Apoorva and Nielsen, Pace P. and Silberman, Lior and Tao, Terence: \href{https://doi.org/10.2140/ant.2018.12.1773}{Homogeneous length functions on groups}, \emph{Algebra Number Theory} {\bf 12} (2018), 1773--1786.
\item Gaddam, Sharat and Gudi, Thirupathi: \href{https://doi.org/10.1016/j.camwa.2017.12.010}{Inhomogeneous {D}irichlet boundary condition in the {\it a
posteriori} error control of the obstacle problem}, \emph{Comput. Math. Appl.} {\bf 75} (2018), 2311--2327.
\item Gaddam, Sharat and Gudi, Thirupathi: \href{https://doi.org/10.1515/cmam-2017-0018}{Bubbles enriched quadratic finite element method for the
3{D}-elliptic obstacle problem}, \emph{Comput. Methods Appl. Math.} {\bf 18} (2018), 223--236.
\item Gadgil, Siddhartha: \href{https://doi.org/10.1007/s10817-019-09523-1}{Homogeneous {L}ength {F}unctions on {G}roups: {I}ntertwined
{C}omputer and {H}uman {P}roofs}, \emph{J. Automat. Reason.} {\bf 64} (2020), 677--688.
\item Gadgil, Siddhartha and Kulkarni, Dheeraj: \href{https://doi.org/10.1007/s12044-016-0278-3}{Relative symplectic caps, 4-genus and fibered knots}, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 126} (2016), 261--275.
\item Garani, Shayan Srinivasa and Seshadri, Harish: \href{https://doi.org/10.1080/17459737.2019.1604845}{An algorithmic approach to {S}outh {I}ndian classical music}, \emph{J. Math. Music} {\bf 13} (2019), 107--134.
\item Gaussier, Herv\'{e} and Seshadri, Harish: \href{https://doi.org/10.1007/s40879-017-0177-x}{Metric geometry of the {K}obayashi metric}, \emph{Eur. J. Math.} {\bf 3} (2017), 1030--1044.
\item Gaussier, Herv\'{e} and Seshadri, Harish: \href{https://doi.org/10.1090/proc/14217}{Holomorphic motions and complex geometry}, \emph{Proc. Amer. Math. Soc.} {\bf 147} (2019), 301--313.
\item Gaussier, Herv\'{e} and Seshadri, Harish: \href{https://doi.org/10.1007/s40315-018-0243-5}{On the {G}romov hyperbolicity of convex domains in {$\Bbb
C^n$}}, \emph{Comput. Methods Funct. Theory} {\bf 18} (2018), 617--641.
\item Gauttam, Sunil Kumar and Kumar, K. Suresh and Pal, Chandan: Risk-sensitive control of reflected diffusion processes on
orthrant, \emph{Pure Appl. Funct. Anal.} {\bf 2} (2017), 477--510.
\item Gayen, R. and Gupta, Sourav and Chakrabarti, A.: \href{https://doi.org/10.1016/j.aml.2016.01.018}{Approximate solution of the problem of scattering of surface
water waves by a partially immersed rigid plane vertical
barrier}, \emph{Appl. Math. Lett.} {\bf 58} (2016), 19--25.
\item Ghara, Soumitra: \href{https://doi.org/10.1007/s11856-020-2016-x}{The orbit of a bounded operator under the {M}\"{o}bius group
modulo similarity equivalence}, \emph{Israel J. Math.} {\bf 238} (2020), 167--207.
\item Ghara, Soumitra and Kumar, Surjit: \href{https://doi.org/10.1016/j.jmaa.2018.09.048}{On the sum of two subnormal kernels}, \emph{J. Math. Anal. Appl.} {\bf 469} (2019), 1015--1027.
\item Ghosh, Mrinal K. and Kumar, K. Suresh and Pal, Chandan: \href{https://doi.org/10.1080/07362994.2016.1180995}{Zero-sum risk-sensitive stochastic games for continuous time
{M}arkov chains}, \emph{Stoch. Anal. Appl.} {\bf 34} (2016), 835--851.
\item Ghosh, Mrinal K. and Pradhan, Somnath: \href{https://doi.org/10.1080/07362994.2017.1356732}{Risk-sensitive stochastic differential games with reflecting
diffusions}, \emph{Stoch. Anal. Appl.} {\bf 36} (2018), 1--27.
\item Ghosh, Mrinal Kanti and Pradhan, Somnath: \href{https://doi.org/10.1051/cocv/2020029}{Zero-sum risk-sensitive stochastic differential games with
reflecting diffusions in the orthant}, \emph{ESAIM Control Optim. Calc. Var.} {\bf 26} (2020), Paper No. 114, 33.
\item Ghosh, Puja and Paul, Kallol and Sain, Debmalya: \href{https://doi.org/10.30755/nsjom.04671}{Symmetric properties of orthogonality of linear operators on
{$(\Bbb R^n,\|\cdot\|_1)$}}, \emph{Novi Sad J. Math.} {\bf 47} (2017), 41--46.
\item Ghosh, Puja and Sain, Debmalya and Paul, Kallol: \href{https://doi.org/10.22034/aot.1703-1137}{On symmetry of {B}irkhoff-{J}ames orthogonality of linear
operators}, \emph{Adv. Oper. Theory} {\bf 2} (2017), 428--434.
\item Ghosh, Subhajit: \href{https://doi.org/10.1007/s10959-019-00945-6}{Total variation cutoff for the transpose top-2 with random
shuffle}, \emph{J. Theoret. Probab.} {\bf 33} (2020), 1832--1854.
\item Ghosh, Subhajit: Cutoff for the warp-transpose top with random shuffle, \emph{S\'{e}m. Lothar. Combin.} {\bf 84B} (2020), Art. 69, 12.
\item Ghosh, Subhroshekhar and Krishnapur, Manjunath and Peres,
Yuval: \href{https://doi.org/10.1214/15-AOP1051}{Continuum percolation for {G}aussian zeroes and {G}inibre
eigenvalues}, \emph{Ann. Probab.} {\bf 44} (2016), 3357--3384.
\item Gudi, Thirupathi and Majumder, Papri: \href{https://doi.org/10.1515/cmam-2019-0057}{Crouzeix-{R}aviart finite element approximation for the
parabolic obstacle problem}, \emph{Comput. Methods Appl. Math.} {\bf 20} (2020), 273--292.
\item Gudi, Thirupathi and Majumder, Papri: \href{https://doi.org/10.1016/j.cam.2019.02.026}{Convergence analysis of finite element method for a parabolic
obstacle problem}, \emph{J. Comput. Appl. Math.} {\bf 357} (2019), 85--102.
\item Gudi, Thirupathi and Majumder, Papri: \href{https://doi.org/10.1016/j.camwa.2019.06.022}{Conforming and discontinuous {G}alerkin {FEM} in space for
solving parabolic obstacle problem}, \emph{Comput. Math. Appl.} {\bf 78} (2019), 3896--3915.
\item Gudi, Thirupathi and Porwal, Kamana: \href{https://doi.org/10.1002/num.21983}{A {$C^0$} interior penalty method for a fourth-order
variational inequality of the second kind}, \emph{Numer. Methods Partial Differential Equations} {\bf 32} (2016), 36--59.
\item Gudi, Thirupathi and Porwal, Kamana: \href{https://doi.org/10.1016/j.cam.2015.07.008}{A posteriori error estimates of discontinuous {G}alerkin
methods for the {S}ignorini problem}, \emph{J. Comput. Appl. Math.} {\bf 292} (2016), 257--278.
\item Gudi, Thirupathi and Sau, Ramesh Ch.: \href{https://doi.org/10.1051/cocv/2019068}{Finite element analysis of the constrained {D}irichlet
boundary control problem governed by the diffusion problem}, \emph{ESAIM Control Optim. Calc. Var.} {\bf 26} (2020), Paper No. 78, 19.
\item Gupta, Rajeev: \href{https://doi.org/10.14232/actasm-015-088-4}{An improvement on the bound for {$C_2(n)$}}, \emph{Acta Sci. Math. (Szeged)} {\bf 83} (2017), 263--269.
\item Gupta, Rajeev and Kumar, Surjit and Trivedi, Shailesh: \href{https://doi.org/10.1016/j.jmaa.2020.124032}{Unitary equivalence of operator-valued multishifts}, \emph{J. Math. Anal. Appl.} {\bf 487} (2020), 124032, 23.
\item Gupta, Rajeev and Kumar, Surjit and Trivedi, Shailesh: \href{https://doi.org/10.1090/proc/14410}{Von {N}eumann's inequality for commuting operator-valued
multishifts}, \emph{Proc. Amer. Math. Soc.} {\bf 147} (2019), 2599--2608.
\item Gupta, Rajeev and Misra, Gadadhar: \href{https://doi.org/10.4064/sm190314-18-8}{The {C}arath\'{e}odory-{F}ej\'{e}r interpolation on the polydisc}, \emph{Studia Math.} {\bf 254} (2020), 265--293.
\item Gupta, Rajeev and Ramiz Reza, Md.: Operator space structures on {$\ell^1(n)$}, \emph{Houston J. Math.} {\bf 44} (2018), 1205--1212.
\item Gupta, S. C.: The classical {S}tefan problem, \emph{} {\bf } (2018), xxiii+726.
\item Gupta, Subhojoy: \href{https://doi.org/10.1090/tran/7777}{Limits of harmonic maps and crowned hyperbolic surfaces}, \emph{Trans. Amer. Math. Soc.} {\bf 372} (2019), 7573--7596.
\item Gupta, Subhojoy and Mj, Mahan: \href{https://doi.org/10.1090/proc/14866}{Monodromy representations of meromorphic projective
structures}, \emph{Proc. Amer. Math. Soc.} {\bf 148} (2020), 2069--2078.
\item Gupta, Subhojoy and Seshadri, Harish: \href{https://doi.org/10.1093/imrn/rny204}{On domains biholomorphic to {T}eichm\"{u}ller spaces}, \emph{Int. Math. Res. Not. IMRN} {\bf } (2020), 2542--2560.
\item Gupta, Subhojoy and Wolf, Michael: \href{https://doi.org/10.1007/s00208-018-1674-z}{Meromorphic quadratic differentials and measured foliations on
a {R}iemann surface}, \emph{Math. Ann.} {\bf 373} (2019), 73--118.
\item Gupta, Subhojoy and Wolf, Michael: \href{https://doi.org/10.4171/CMH/388}{Quadratic differentials, half-plane structures, and harmonic
maps to trees}, \emph{Comment. Math. Helv.} {\bf 91} (2016), 317--356.
\item Gupta, Subhojoy and Wolf, Michael: \href{https://doi.org/10.1090/conm/696/14021}{Meromorphic quadratic differentials with complex residues and
spiralling foliations}, \emph{In the tradition of {A}hlfors-{B}ers. {VII}} {\bf 696} (2017), 153--181.
\item Hatui, Sumana: \href{https://doi.org/10.1515/jgth-2019-0045}{Schur multipliers of special {$p$}-groups of rank 2}, \emph{J. Group Theory} {\bf 23} (2020), 85--95.
\item Hazra, Somnath: \href{https://doi.org/10.1007/s11785-018-0822-5}{Homogeneous 2-shifts}, \emph{Complex Anal. Oper. Theory} {\bf 13} (2019), 1729--1763.
\item Iyer, Srikanth K.: \href{https://doi.org/10.1002/rsa.20741}{The random connection model: connectivity, edge lengths, and
degree distributions}, \emph{Random Structures Algorithms} {\bf 52} (2018), 283--300.
\item Iyer, Srikanth K. and Vaze, Rahul and Narasimha, Dheeraj: \href{https://doi.org/10.1016/j.physa.2015.12.030}{Autoregressive cascades on random networks}, \emph{Phys. A} {\bf 447} (2016), 345--354.
\item Iyer, Srikanth K. and Yogeshwaran, D.: \href{https://doi.org/10.1214/19-AIHP1020}{Thresholds for vanishing of `isolated' faces in random \v{C}ech
and {V}ietoris-{R}ips complexes}, \emph{Ann. Inst. Henri Poincar\'{e} Probab. Stat.} {\bf 56} (2020), 1869--1897.
\item Iyer, Srikanth and Vaze, Rahul: \href{https://doi.org/10.1214/16-AAP1196}{Achieving nonzero information velocity in wireless networks}, \emph{Ann. Appl. Probab.} {\bf 27} (2017), 48--64.
\item Jha, Abhash Kumar and Juyal, Abhishek and Pandey, Manish
Kumar: On simultaneous non-vanishing of twisted {$L$}-functions
associated to newforms on {$\Gamma_0(N)$}, \emph{J. Ramanujan Math. Soc.} {\bf 34} (2019), 245--252.
\item Jha, Abhash Kumar and Kumar, Arvind: Construction of cusp forms using {R}ankin-{C}ohen brackets, \emph{Analytic number theory, modular forms and {$q$}-hypergeometric
series} {\bf 221} (2017), 329--341.
\item Jha, Abhash Kumar and Sahu, Brundaban: \href{https://doi.org/10.1142/S1793042119500490}{Rankin-{C}ohen brackets on {J}acobi forms of several variables
and special values of certain {D}irichlet series}, \emph{Int. J. Number Theory} {\bf 15} (2019), 925--933.
\item Jha, Abhash Kumar and Sahu, Brundaban: \href{https://doi.org/10.1090/conm/732/14793}{Differential operators on {J}acobi forms and special values of
certain {D}irichlet series}, \emph{Automorphic forms and related topics} {\bf 732} (2019), 91--99.
\item Ji, Kui and Jiang, Chunlan and Keshari, Dinesh Kumar and
Misra, Gadadhar: \href{https://doi.org/10.1016/j.jfa.2016.12.019}{Rigidity of the flag structure for a class of
{C}owen-{D}ouglas operators}, \emph{J. Funct. Anal.} {\bf 272} (2017), 2899--2932.
\item Jiang, Chunlan and Ji, Kui and Misra, Gadadhar: \href{https://doi.org/10.1016/j.jfa.2017.06.015}{Classification of quasi-homogeneous holomorphic curves and
operators in the {C}owen-{D}ouglas class}, \emph{J. Funct. Anal.} {\bf 273} (2017), 2870--2915.
\item Kabiraj, Arpan: \href{https://doi.org/10.2140/agt.2016.16.2839}{Center of the {G}oldman {L}ie algebra}, \emph{Algebr. Geom. Topol.} {\bf 16} (2016), 2839--2849.
\item Kanukurthi, Bhavana and Bhavana Obbattu, Sai Lakshmi and
Sekar, Sruthi: \href{https://doi.org/10.1007/s00145-019-09339-7}{Four-state non-malleable codes with explicit constant rate}, \emph{J. Cryptology} {\bf 33} (2020), 1044--1079.
\item Kanukurthi, Bhavana and Bhavana Obbattu, Sai Lakshmi and
Sekar, Sruthi: Four-state non-malleable codes with explicit constant rate, \emph{Theory of cryptography. {P}art {II}} {\bf 10678} (2017), 344--375.
\item Kanukurthi, Bhavana and Bhavana Obbattu, Sai Lakshmi and
Sekar, Sruthi: Non-malleable randomness encoders and their applications, \emph{Advances in cryptology---{EUROCRYPT} 2018. {P}art {III}} {\bf 10822} (2018), 589--617.
\item Kashyap, G. A. R. S. R. K. and Bapat, D. and Das, D. and Gowaikar, R. D. and  Amritkar, R. E. and Rangarajan, G. and V. Ravindranath and G. Ambika: Synapse loss and progress of Alzheimer&rsquo;s disease - A network model, \emph{Scientific Reports} {\bf 9} (2019).
\item Kataria, Kuldeep Kumar and Vellaisamy, Palaniappan: \href{https://doi.org/10.1080/07362994.2018.1538803}{On the convolution of {M}ittag-{L}effler distributions and its
applications to fractional point processes}, \emph{Stoch. Anal. Appl.} {\bf 37} (2019), 115--122.
\item Kaur, Amandeep and Martha, S. C. and Chakrabarti, A.: \href{https://doi.org/10.1007/s40314-020-01255-y}{An algebraic method of solution of a water wave scattering
problem involving an asymmetrical trench}, \emph{Comput. Appl. Math.} {\bf 39} (2020), Paper No. 229, 19.
\item Khan, Mohammad Yasir Feroz and Biswas, Rahul and Sau, Ramesh
Chandra: Instability of {L}agrange interpolation on {$H^1(\Omega)\cap
C(\overline\Omega)$}, \emph{Math. Student} {\bf 88} (2019), 153--158.
\item Khare, Apoorva: \href{https://doi.org/10.1090/tran/7304}{Generalized nil-{C}oxeter algebras over discrete complex
reflection groups}, \emph{Trans. Amer. Math. Soc.} {\bf 370} (2018), 2971--2999.
\item Khare, Apoorva and Rajaratnam, Bala: \href{https://doi.org/10.1007/s43036-020-00048-8}{Probability inequalities and tail estimates for metric
semigroups}, \emph{Adv. Oper. Theory} {\bf 5} (2020), 779--795.
\item Khare, Apoorva and Rajaratnam, Bala: \href{https://doi.org/10.1214/16-AOP1160}{The {H}offmann-{J}\o rgensen inequality in metric semigroups}, \emph{Ann. Probab.} {\bf 45} (2017), 4101--4111.
\item Khare, Apoorva and Tikaradze, Akaki: \href{https://doi.org/10.1016/j.laa.2018.05.035}{A {C}arlitz--von {S}taudt type theorem for finite rings}, \emph{Linear Algebra Appl.} {\bf 568} (2019), 106--126.
\item Khurana, Divya and Roy, Saikat and Sain, Debmalya: \href{https://doi.org/10.14232/actasm-020-420-6}{Symmetric points in spaces of linear operators between
{B}anach spaces}, \emph{Acta Sci. Math. (Szeged)} {\bf 86} (2020), 617--634.
\item Kor\'{a}nyi, Adam and Misra, Gadadhar: \href{https://doi.org/10.1016/j.aim.2019.05.018}{Homogeneous {H}ermitian holomorphic vector bundles and the
{C}owen-{D}ouglas class over bounded symmetric domains}, \emph{Adv. Math.} {\bf 351} (2019), 1105--1138.
\item Kor\'{a}nyi, Adam and Misra, Gadadhar: \href{https://doi.org/10.1016/j.crma.2015.11.002}{Homogeneous {H}ermitian holomorphic vector bundles and the
{C}owen-{D}ouglas class over bounded symmetric domains}, \emph{C. R. Math. Acad. Sci. Paris} {\bf 354} (2016), 291--295.
\item Krakovski, Roi and Onn, Uri and Singla, Pooja: \href{https://doi.org/10.1016/j.jalgebra.2017.10.018}{Regular characters of groups of type {$\ssf A_n$} over
discrete valuation rings}, \emph{J. Algebra} {\bf 496} (2018), 116--137.
\item Kreuzer, Martin and Patil, Dilip P.: \href{https://doi.org/10.1007/s13366-016-0324-4}{Computational aspects of {B}urnside rings, part {I}: the ring
structure}, \emph{Beitr. Algebra Geom.} {\bf 58} (2017), 427--452.
\item Krishna, M. and Krishnapur, Manjunath: \href{https://doi.org/10.1007/s13226-016-0183-6}{Persistence probabilities in centered, stationary, {G}aussian
processes in discrete time}, \emph{Indian J. Pure Appl. Math.} {\bf 47} (2016), 183--194.
\item Krishnapur, Manjunath and Rider, Brian and Vir\'{a}g, B\'{a}lint: \href{https://doi.org/10.1002/cpa.21573}{Universality of the stochastic {A}iry operator}, \emph{Comm. Pure Appl. Math.} {\bf 69} (2016), 145--199.
\item Kundu, S. and Aggarwal, Manisha and Hazra, Somnath: \href{https://doi.org/10.1016/j.topol.2016.11.008}{Finitely chainable and totally bounded metric spaces:
equivalent characterizations}, \emph{Topology Appl.} {\bf 216} (2017), 59--73.
\item Lin, X. and Kulkarni, P. and Bocci, F. and Schafer, N. and Roy, S. and Tsai, M-Y. and He, Y. and Chen, Y and Rajagopalan, K. and Mooney, S. and Zeng, Y. and Weninger, K. and Grishaev, A. and Onuchic, J. and Levine, H. and Wolynes, P. and Salgia, R. and Rangarajan, G. and Uversky, V. and Orban, J. and Jolly, M. K.: Structural and Dynamical Order of a Disordered Protein: Molecular Insights into Conformational Switching of PAGE4 at the Systems Level, \emph{Biomolecules} {\bf 9} (2019), 77.
\item Mahadevan, Rajesh and Nandakumaran, A. K. and Prakash, Ravi: \href{https://doi.org/10.1007/s00245-018-9499-4}{Homogenization of an elliptic equation in a domain with
oscillating boundary with non-homogeneous non-linear boundary
conditions}, \emph{Appl. Math. Optim.} {\bf 82} (2020), 245--278.
\item Mahajan, Prachi and Verma, Kaushal: \href{https://doi.org/10.1142/S0129167X19500125}{A comparison of two biholomorphic invariants}, \emph{Internat. J. Math.} {\bf 30} (2019), 1950012, 16.
\item Maiti, Arun: \href{https://doi.org/10.1016/j.disc.2020.111911}{Quasi-vertex-transitive maps on the plane}, \emph{Discrete Math.} {\bf 343} (2020), 111911, 7.
\item Maitra, Anwoy: \href{https://doi.org/10.1090/proc/15038}{On the continuous extension of {K}obayashi isometries}, \emph{Proc. Amer. Math. Soc.} {\bf 148} (2020), 3437--3451.
\item Maitra, Anwoy: \href{https://doi.org/10.1016/j.jmaa.2019.123694}{A form of {S}chwarz's lemma and a bound for the {K}obayashi
metric on convex domains}, \emph{J. Math. Anal. Appl.} {\bf 484} (2020), 123694, 16.
\item Maity, Dipendu: \href{https://doi.org/10.1007/s13366-018-0405-7}{Pentagonal maps on the torus and the plane}, \emph{Beitr. Algebra Geom.} {\bf 60} (2019), 17--37.
\item Maity, Dipendu and Upadhyay, Ashish Kumar: On centrally symmetric manifolds, \emph{J. Ramanujan Math. Soc.} {\bf 34} (2019), 21--27.
\item Mal, Arpita and Paul, Kallol and Rao, T. S. S. R. K. and Sain,
Debmalya: \href{https://doi.org/10.1007/s00605-019-01289-3}{Approximate {B}irkhoff-{J}ames orthogonality and smoothness in
the space of bounded linear operators}, \emph{Monatsh. Math.} {\bf 190} (2019), 549--558.
\item Mal, Arpita and Sain, Debmalya and Paul, Kallol: \href{https://doi.org/10.1215/17358787-2018-0021}{On some geometric properties of operator spaces}, \emph{Banach J. Math. Anal.} {\bf 13} (2019), 174--191.
\item Manchanda, Kaustubh and Bose, Amitabha and Ramaswamy,
Ramakrishna: \href{https://doi.org/10.1016/j.physa.2017.06.021}{Collective dynamics in heterogeneous networks of neuronal
cellular automata}, \emph{Phys. A} {\bf 487} (2017), 111--124.
\item Mandal, S. and Sahoo, T. and Chakrabarti, A.: \href{https://doi.org/10.1080/03091929.2017.1318129}{Characteristics of eigen-system for flexural gravity wave
problems}, \emph{Geophys. Astrophys. Fluid Dyn.} {\bf 111} (2017), 249--281.
\item Manfio, F. and Turgay, N. C. and Upadhyay, A.: \href{https://doi.org/10.1007/s12220-018-9990-9}{Biconservative submanifolds in {$\Bbb{S}^n\times\Bbb{R}$} and
{$\Bbb{H}^n\times\Bbb{R}$}}, \emph{J. Geom. Anal.} {\bf 29} (2019), 283--298.
\item Manikandan, K. and Vishnu Priya, N. and Senthilvelan, M. and
Sankaranarayanan, R.: \href{https://doi.org/10.1063/1.5039901}{Deformation of dark solitons in a {$\Cal{PT}$}-invariant
variable coefficients nonlocal nonlinear {S}chr\"{o}dinger
equation}, \emph{Chaos} {\bf 28} (2018), 083103, 12.
\item Misra, Gadadhar: Operators in the {C}owen-{D}ouglas class and related topics, \emph{Handbook of analytic operator theory} {\bf } (2019), 87--137.
\item Misra, Gadadhar and Pal, Avijit: \href{https://doi.org/10.1007/s11854-018-0054-7}{Curvature inequalities for operators in the {C}owen-{D}ouglas
class and localization of the {W}allach set}, \emph{J. Anal. Math.} {\bf 136} (2018), 31--54.
\item Misra, Gadadhar and Pal, Avijit and Varughese, Cherian: \href{https://doi.org/10.7900/jot.2018jun13.2225}{Contractivity and complete contractivity for finite
dimensional {B}anach spaces}, \emph{J. Operator Theory} {\bf 82} (2019), 23--47.
\item Misra, Gadadhar and Reza, Md. Ramiz: \href{https://doi.org/10.1215/00192082-7768711}{Curvature inequalities and extremal operators}, \emph{Illinois J. Math.} {\bf 63} (2019), 193--217.
\item Misra, Gadadhar and Upmeier, Harald: \href{https://doi.org/10.1016/j.aim.2016.08.036}{Homogeneous vector bundles and intertwining operators for
symmetric domains}, \emph{Adv. Math.} {\bf 303} (2016), 1077--1121.
\item Mukherjee, Sajal Kumar and Bera, Sudip: \href{https://doi.org/10.1016/j.disc.2019.02.013}{Combinatorial proofs of the {N}ewton-{G}irard and
{C}hapman-{C}ostas-{S}antos identities}, \emph{Discrete Math.} {\bf 342} (2019), 1577--1580.
\item Nandakumaran, A. K. and Datti, P. S. and George, Raju K.: \href{https://doi.org/10.1017/9781108236843}{Ordinary differential equations}, \emph{} {\bf } (2017), xviii+328.
\item Nandakumaran, A. K. and Sili, Ali: \href{http://projecteuclid.org/euclid.die/1448323252}{Homogenization of a hyperbolic equation with highly
contrasting diffusivity coefficients}, \emph{Differential Integral Equations} {\bf 29} (2016), 37--54.
\item Nandakumaran, Akambadath and Prakash, Ravi and Sardar, Bidhan
Chandra: \href{https://doi.org/10.1002/mma.3865}{Asymptotic analysis of {N}eumann periodic optimal boundary
control problem}, \emph{Math. Methods Appl. Sci.} {\bf 39} (2016), 4354--4374.
\item Narayanan, E. K. and Singla, Pooja: \href{https://doi.org/10.1080/00927872.2017.1378894}{On monomial representations of finitely generated nilpotent
groups}, \emph{Comm. Algebra} {\bf 46} (2018), 2319--2331.
\item Nikolov, Nikolai and Verma, Kaushal: \href{https://doi.org/10.1007/s12220-019-00237-9}{On the squeezing function and {F}ridman invariants}, \emph{J. Geom. Anal.} {\bf 30} (2020), 1218--1225.
\item Pal, Chandan and Pradhan, Somnath: \href{https://doi.org/10.1080/17442508.2018.1521413}{Risk sensitive control of pure jump processes on a general
state space}, \emph{Stochastics} {\bf 91} (2019), 155--174.
\item Pal, Ratna and Verma, Kaushal: \href{https://doi.org/10.4064/ap170928-2-4}{Ergodic properties of families of {H}\'{e}non maps}, \emph{Ann. Polon. Math.} {\bf 121} (2018), 45--71.
\item Pal, Ritwik: \href{https://doi.org/10.1007/s11139-019-00206-4}{On the signs of {F}ourier coefficients of {H}ilbert cusp
forms}, \emph{Ramanujan J.} {\bf 53} (2020), 467--481.
\item Panda, Srikumar and Martha, S. C. and Chakrabarti, A.: \href{https://doi.org/10.1016/j.amc.2015.09.086}{An alternative approach to study nonlinear inviscid flow over
arbitrary bottom topography}, \emph{Appl. Math. Comput.} {\bf 273} (2016), 165--177.
\item Patil, D. P.: Affine one dimensional domains which are not {E}uclidean
domains, \emph{Int. Rev. Pure Appl. Math.} {\bf 12} (2016), 45--50.
\item Paul, Kallol and Sain, Debmalya: Birkhoff-{J}ames orthogonality and its application in the
study of geometry of {B}anach space, \emph{Advanced topics in mathematical analysis} {\bf } (2019), 245--284.
\item Paul, Kallol and Sain, Debmalya and Ghosh, Puja: Symmetry of {B}irkhoff-{J}ames orthogonality of bounded linear
operators, \emph{Ulam type stability} {\bf } (2019), 331--344.
\item Paul, Kallol and Sain, Debmalya and Mal, Arpita: \href{https://doi.org/10.1016/j.laa.2017.10.008}{Approximate {B}irkhoff-{J}ames orthogonality in the space of
bounded linear operators}, \emph{Linear Algebra Appl.} {\bf 537} (2018), 348--357.
\item Paul, Kallol and Sain, Debmalya and Mal, Arpita and Mandal,
Kalidas: \href{https://doi.org/10.15352/aot.1712-1268}{Orthogonality of bounded linear operators on complex {B}anach
spaces}, \emph{Adv. Oper. Theory} {\bf 3} (2018), 699--709.
\item Pingali, Vamsi P.: {$C^{2,\alpha}$} estimates and existence results for a
nonconcave {PDE}, \emph{Electron. J. Differential Equations} {\bf } (2016), Paper No. 168, 10.
\item Pingali, Vamsi P.: \href{https://doi.org/10.1080/17476933.2018.1454914}{A note on the deformed {H}ermitian {Y}ang-{M}ills {PDE}}, \emph{Complex Var. Elliptic Equ.} {\bf 64} (2019), 503--518.
\item Pingali, Vamsi Pritham: \href{https://doi.org/10.1016/j.aim.2019.106921}{A vector bundle version of the {M}onge-{A}mp\`ere equation}, \emph{Adv. Math.} {\bf 360} (2020), 106921, 40.
\item Pingali, Vamsi Pritham: \href{https://doi.org/10.1090/tran/7416}{A note on higher extremal metrics}, \emph{Trans. Amer. Math. Soc.} {\bf 370} (2018), 6995--7010.
\item Pingali, Vamsi Pritham: \href{https://doi.org/10.1007/s11005-020-01279-9}{Quillen metrics and perturbed equations}, \emph{Lett. Math. Phys.} {\bf 110} (2020), 1861--1875.
\item Pingali, Vamsi Pritham: \href{https://doi.org/10.1007/s00209-017-1903-2}{Representability of {C}hern-{W}eil forms}, \emph{Math. Z.} {\bf 288} (2018), 629--641.
\item Poria, Anirudha: \href{https://doi.org/10.1080/01630563.2020.1728771}{Positive weight function and classification of {$g$}-frames}, \emph{Numer. Funct. Anal. Optim.} {\bf 41} (2020), 950--968.
\item Pradhan, Puspendu and Sahu, Bikramaditya: \href{https://doi.org/10.1016/j.disc.2020.112044}{A characterization of the family of secant lines to a
hyperbolic quadric in {${\rm PG} (3,q)$}, {$q$} odd}, \emph{Discrete Math.} {\bf 343} (2020), 112044, 8.
\item Prasad, Phoolan: \href{https://doi.org/10.1007/s12044-016-0275-6}{Ray equations of a weak shock in a hyperbolic system of
conservation laws in multi-dimensions}, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 126} (2016), 199--206.
\item Prasad, Phoolan: \href{https://doi.org/10.1007/978-981-10-7581-0}{Propagation of multidimensional nonlinear waves and
kinematical conservation laws}, \emph{} {\bf } (2017), xi+159.
\item Prasad, Phoolan: \href{https://doi.org/10.1007/s13226-016-0197-0}{Kinematical conservation laws in a space of arbitrary
dimensions}, \emph{Indian J. Pure Appl. Math.} {\bf 47} (2016), 641--653.
\item Pritham Pingali, Vamsi: \href{https://doi.org/10.1142/S0129167X18500416}{Existence of coupled {K}\"{a}hler-{E}instein metrics using the
continuity method}, \emph{Internat. J. Math.} {\bf 29} (2018), 1850041, 8.
\item Quddus, Safdar: \href{https://doi.org/10.4171/JNCG/320}{Invariants of the {$\Bbb Z_2$} orbifolds of the {P}odle\'{s} two
spheres}, \emph{J. Noncommut. Geom.} {\bf 13} (2019), 257--267.
\item Ram, Samrith: \href{https://doi.org/10.1016/j.laa.2016.01.029}{Counting zero kernel pairs over a finite field}, \emph{Linear Algebra Appl.} {\bf 495} (2016), 1--10.
\item Ray, Anubhab and Roy, Saikat and Bagchi, Satya and Sain,
Debmalya: \href{https://doi.org/10.7900/jot}{Extreme contractions on finite-dimensional polygonal {B}anach
spaces. {II}}, \emph{J. Operator Theory} {\bf 84} (2020), 127--137.
\item Ray, Samarpita: \href{https://doi.org/10.1142/S0219498820500061}{Closure operations, continuous valuations on monoids and
spectral spaces}, \emph{J. Algebra Appl.} {\bf 19} (2020), 2050006, 27.
\item Reddy, Nanda Kishore: \href{https://doi.org/10.1093/imrn/rnx134}{Equality of {L}yapunov and stability exponents for products of
isotropic random matrices}, \emph{Int. Math. Res. Not. IMRN} {\bf } (2019), 606--624.
\item Reza, Md. Ramiz: \href{https://doi.org/10.1512/iumj.2018.67.7320}{Curvature inequalities for operators in the {C}owen-{D}ouglas
class of a planar domain}, \emph{Indiana Univ. Math. J.} {\bf 67} (2018), 1255--1279.
\item Richard, Thomas and Seshadri, Harish: \href{https://doi.org/10.1007/s00229-015-0790-2}{Positive isotropic curvature and self-duality in dimension 4}, \emph{Manuscripta Math.} {\bf 149} (2016), 443--457.
\item Roncal, Luz and Thangavelu, Sundaram: \href{https://doi.org/10.1093/imrn/rny137}{An extension problem and trace {H}ardy inequality for the
sub-{L}aplacian on {$H$}-type groups}, \emph{Int. Math. Res. Not. IMRN} {\bf } (2020), 4238--4294.
\item Roncal, Luz and Thangavelu, Sundaram: \href{https://doi.org/10.1016/j.aim.2016.07.010}{Hardy's inequality for fractional powers of the sublaplacian
on the {H}eisenberg group}, \emph{Adv. Math.} {\bf 302} (2016), 106--158.
\item Roth, Julien and Upadhyay, Abhitosh: \href{https://doi.org/10.1007/s00013-019-01315-8}{On compact anisotropic {W}eingarten hypersurfaces in
{E}uclidean space}, \emph{Arch. Math. (Basel)} {\bf 113} (2019), 213--224.
\item Roth, Julien and Upadhyay, Abhitosh: \href{https://doi.org/10.1017/s0004972719000935}{On almost stable {CMC} hypersurfaces in manifolds of bounded
sectional curvature}, \emph{Bull. Aust. Math. Soc.} {\bf 101} (2020), 333--338.
\item Roy, Dipankar and Pandit, Rahul: One-dimensional {K}ardar-{P}arisi-{Z}hang and
{K}uramoto-{S}ivashinsky universality class: limit
distributions, \emph{Phys. Rev. E} {\bf 101} (2020), 030103(R), 6.
\item Roy, Krishanu and Venkatesh, R.: \href{https://doi.org/10.1007/s00031-018-9510-9}{Maximal closed subroot systems of real affine root systems}, \emph{Transform. Groups} {\bf 24} (2019), 1261--1308.
\item Sain, Debmalya: \href{https://doi.org/10.1215/20088752-2018-0014}{On extreme contractions and the norm attainment set of a
bounded linear operator}, \emph{Ann. Funct. Anal.} {\bf 10} (2019), 135--143.
\item Sain, Debmalya: \href{https://doi.org/10.1007/s13226-020-0498-1}{On a generalization of a complementary triangle inequality in
{H}ilbert spaces and {B}anach spaces}, \emph{Indian J. Pure Appl. Math.} {\bf 51} (2020), 1815--1827.
\item Sain, Debmalya: \href{https://doi.org/10.1016/j.jmaa.2017.07.070}{On the norm attainment set of a bounded linear operator}, \emph{J. Math. Anal. Appl.} {\bf 457} (2018), 67--76.
\item Sain, Debmalya: \href{https://doi.org/10.1007/s13226-020-0393-9}{On the norm attainment set of a bounded linear operator and
semi-inner-products in normed spaces}, \emph{Indian J. Pure Appl. Math.} {\bf 51} (2020), 179--186.
\item Sain, Debmalya: \href{https://doi.org/10.7153/oam-2019-13-32}{Smooth points in operator spaces and some
{B}ishop-{P}helps-{B}ollob\'{a}s type theorems in {B}anach spaces}, \emph{Oper. Matrices} {\bf 13} (2019), 433--445.
\item Sain, Debmalya and Mal, Arpita and Paul, Kallol: \href{https://doi.org/10.1016/j.exmath.2019.01.001}{Some remarks on {B}irkhoff-{J}ames orthogonality of linear
operators}, \emph{Expo. Math.} {\bf 38} (2020), 138--147.
\item Sain, Debmalya and Paul, Kallol and Bhunia, Pintu and Bag,
Santanu: \href{https://doi.org/10.1016/j.laa.2019.04.024}{On the numerical index of polyhedral {B}anach spaces}, \emph{Linear Algebra Appl.} {\bf 577} (2019), 121--133.
\item Sain, Debmalya and Paul, Kallol and Mal, Arpita: A complete characterization of {B}irkhoff-{J}ames
orthogonality in infinite dimensional normed space, \emph{J. Operator Theory} {\bf 80} (2018), 399--413.
\item Sain, Debmalya and Paul, Kallol and Mal, Arpita: On approximate {B}irkhoff-{J}ames orthogonality and normal
cones in a normed space, \emph{J. Convex Anal.} {\bf 26} (2019), 341--351.
\item Sain, Debmalya and Paul, Kallol and Mal, Arpita and Ray,
Anubhab: \href{https://doi.org/10.1080/03081087.2019.1586824}{A complete characterization of smoothness in the space of
bounded linear operators}, \emph{Linear Multilinear Algebra} {\bf 68} (2020), 2484--2494.
\item Sain, Debmalya and Paul, Kallol and Mandal, Kalidas: \href{https://doi.org/10.7153/oam-2019-13-31}{On two extremum problems related to the norm of a bounded
linear operator}, \emph{Oper. Matrices} {\bf 13} (2019), 421--432.
\item Sain, Debmalya and Ray, Anubhab and Paul, Kallol: Extreme contractions on finite-dimensional polygonal {B}anach
spaces, \emph{J. Convex Anal.} {\bf 26} (2019), 877--885.
\item Sain, Debmalya and Tanaka, Ryotaro: \href{https://doi.org/10.1007/s43037-020-00064-z}{Modular {B}irkhoff-{J}ames orthogonality in {$B(\Bbb{X,Y})$}
and {$K(\Bbb{X,Y})$}}, \emph{Banach J. Math. Anal.} {\bf 14} (2020), 1347--1362.
\item Sanki, Bidyut: \href{https://doi.org/10.1142/S1793525318500309}{Filling of closed surfaces}, \emph{J. Topol. Anal.} {\bf 10} (2018), 897--913.
\item Sanki, Bidyut and Gadgil, Siddhartha: \href{https://doi.org/10.1142/S1793525319500018}{Graphs of systoles on hyperbolic surfaces}, \emph{J. Topol. Anal.} {\bf 11} (2019), 1--20.
\item Sarkar, Amar Deep and Verma, Kaushal: \href{https://doi.org/10.1007/s12044-020-00565-9}{A submultiplicative property of the {C}arath\'{e}odory metric on
planar domains}, \emph{Proc. Indian Acad. Sci. Math. Sci.} {\bf 130} (2020), Paper No. 35, 6.
\item Sarkar, Amar Deep and Verma, Kaushal: \href{https://doi.org/10.1007/s40315-020-00303-2}{Boundary behaviour of some conformal invariants on planar
domains}, \emph{Comput. Methods Funct. Theory} {\bf 20} (2020), 145--158.
\item Sen, Suparna: \href{https://doi.org/10.1515/apam-2015-0010}{Segal-{B}argmann transform and {P}aley-{W}iener theorems on
{H}eisenberg motion groups}, \emph{Adv. Pure Appl. Math.} {\bf 7} (2016), 13--28.
\item Senthil Raani, K. S.: \href{https://doi.org/10.1007/s00605-017-1081-7}{{$L^p$}-{F}ourier asymptotics, {H}ardy-type inequality and
fractal measures}, \emph{Monatsh. Math.} {\bf 184} (2017), 459--487.
\item Seshadri, Harish: \href{https://doi.org/10.1007/s13226-019-0355-2}{Differential geometry in {I}ndia}, \emph{Indian J. Pure Appl. Math.} {\bf 50} (2019), 795--799.
\item Seshadri, Harish and Verma, Kaushal: Some aspects of the automorphism groups of domains, \emph{Handbook of group actions. {V}ol. {III}} {\bf 40} (2018), 145--174.
\item Tanaka, Ryotaro and Sain, Debmalya: \href{https://doi.org/10.1007/s43034-019-00048-7}{On symmetry of strong {B}irkhoff orthogonality in {$B(\Cal
H,\Cal K)$} and {$K(\Cal H,\Cal K)$}}, \emph{Ann. Funct. Anal.} {\bf 11} (2020), 693--704.
\item Thangavelu, Sundaram: \href{https://doi.org/10.1007/s41478-018-0147-9}{An analogue of {P}fannschmidt's theorem for the {H}eisenberg
group}, \emph{J. Anal.} {\bf 26} (2018), 235--244.
\item Thangavelu, Sundaram and Naidu Dogga, Venku: \href{https://doi.org/10.1515/apam-2017-0042}{{$L^p$}-{$L^2$} estimates for solutions of the wave equation
associated to the {G}rushin operator}, \emph{Adv. Pure Appl. Math.} {\bf 9} (2018), 85--92.
\item Turgay, Nurettin Cenk and Upadhyay, Abhitosh: \href{https://doi.org/10.1002/mana.201700328}{On biconservative hypersurfaces in 4-dimensional {R}iemannian
space forms}, \emph{Math. Nachr.} {\bf 292} (2019), 905--921.
\item Urbina-Romero, Wilfredo: \href{https://doi.org/10.1007/978-3-030-05597-4}{Gaussian harmonic analysis}, \emph{} {\bf } (2019), xix+477.
\item Verma, Kaushal: \href{https://doi.org/10.1007/s13324-018-0221-0}{Notes on the boundaries of quadrature domains}, \emph{Anal. Math. Phys.} {\bf 9} (2019), 617--638.
\item Verma, Kaushal: Remembering {R}. {R}. {S}imha, \emph{Math. Newsl.} {\bf 27} (2017), 16--20.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan,
Govindan: \href{https://doi.org/10.1063/1.5121245}{On the role of four-wave mixing effect in the interactions
between nonlinear modes of coupled generalized nonlinear
{S}chr\"{o}dinger equation}, \emph{Chaos} {\bf 29} (2019), 123135, 14.
\item Vishnu Priya, N. and Senthilvelan, M. and Rangarajan, Govindan
and Lakshmanan, M.: \href{https://doi.org/10.1016/j.physleta.2018.10.011}{On symmetry preserving and symmetry broken bright, dark and
antidark soliton solutions of nonlocal nonlinear {S}chr\"{o}dinger
equation}, \emph{Phys. Lett. A} {\bf 383} (2019), 15--26.
\item Yadav, Avinash Chand and Manchanda, Kaustubh and Ramaswamy,
Ramakrishna: \href{https://doi.org/10.1016/j.physa.2017.04.029}{Emergent organization in a model market}, \emph{Phys. A} {\bf 482} (2017), 118--126.
\end{enumerate}